### 2015-11-16 paper.tex
Getting all the equations in place

\documentclass[11pt, oneside]{article}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{ptex2tex, minted}
\usepackage{xcolor}
\usepackage{cancel}
\usepackage{caption}
\usepackage{subfigure}
\usepackage{algorithm}
\usepackage{algpseudocode}

\subfiglabelskip=0pt

\definecolor{gray}{gray}{0.97}
\colorlet{commentcolour}{green!50!black}
\colorlet{stringcolour}{red!60!black}
\colorlet{keywordcolour}{magenta!90!black}
\colorlet{exceptioncolour}{yellow!50!red}
\colorlet{commandcolour}{blue!60!black}
\colorlet{numpycolour}{blue!60!green}
\colorlet{literatecolour}{magenta!90!black}
\colorlet{promptcolour}{green!50!black}
\colorlet{specmethodcolour}{violet}
\colorlet{indendifiercolour}{green!70!white}

%\newcommand{\codetitlestyle}[1]{\small\textit{#1}\hspace{0.1cm}}
\newcommand{\belowtitleskip}{2pt}%\smallskipamount}
%\newcommand{\captionposition}{t}

%\newcommand{\mmo}[1]{\emph{#1}}

%\renewcommand{\ttdefault}{pcr}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[top=3cm,bottom=4cm,left=3cm,right=3.2cm,asymmetric]{geometry}

\lstset{
numbers=none,
aboveskip=1ex,
belowskip=1ex,
basicstyle=\ttfamily\footnotesize,
}

\lstdefinestyle{pythonstyle}{
%%\lstset{
%%keepspaces=true,
language=python,
showtabs=true,
tab=,
tabsize=2,
basicstyle=\ttfamily\footnotesize,%\setstretch{.5},
stringstyle=\color{stringcolour},
showstringspaces=false,
alsoletter={1234567890},
otherkeywords={\ , \}, \{, \%, \&, \|},
keywordstyle=\color{keywordcolour}\bfseries,
emph={and,break,class,continue,def,yield,del,elif ,else,%
except,exec,finally,for,from,global,if,import,in,%
lambda,not,or,pass,print,raise,return,try,while,assert},
emphstyle=\color{blue}\bfseries,
emph={[2]True, False, None},
emphstyle=[2]\color{keywordcolour},
emph={[3]object,type,isinstance,copy,deepcopy,zip,enumerate,reversed,list,len,dict,tuple,xrange,append,execfile,real,imag,reduce,str,repr},
emphstyle=[3]\color{commandcolour},
emph={Exception,NameError,IndexError,SyntaxError,TypeError,ValueError,OverflowError,ZeroDivisionError},
emphstyle=\color{exceptioncolour}\bfseries,
%upquote=true,
morestring=[s]{"""}{"""},
morestring=[s]{'''}{'''},
commentstyle=\color{commentcolour}\slshape,
emph={[4]1, 2, 3, 4, 5, 6, 7, 8, 9, 0,ode, fsolve, sqrt, exp, sin, cos, arccos, pi, array, norm, solve,float,complex, dot, arange, isscalar, max, sum, flatten, shape, reshape, find, any, all, abs, plot, linspace, legend, quad, polyval,polyfit, hstack,vector, concatenate,vstack,column_stack,empty,zeros,ones,rand,vander,grid,pcolor,eig,eigs,eigvals,svd,qr,tan,det,logspace,roll,min,mean,cumsum,cumprod,diff,vectorize,lstsq,cla,eye,xlabel,ylabel,squeeze},
emphstyle=[4]\color{commandcolour},
emph={[5]__init__,__add__,__mul__,__div__,__sub__,__call__,__getitem__,__setitem__,__eq__,__ne__,__nonzero__,__rmul__,__radd__,__repr__,__str__,__get__,__truediv__,__pow__,__name__,__future__,__all__},
emphstyle=[5]\color{specmethodcolour},
emph={[6]assert,range,yield},
emphstyle=[6]\color{keywordcolour}\bfseries,
emph={[7]def, return, and, print},
emphstyle=[7]\color{black}\bfseries,
% emph={[7]self},
% emphstyle=[7]\bfseries,
literate=*%
%{:}{{\literatecolour:}}{1}%
%{=}{{\literatecolour=}}{1}%
%{-}{{\literatecolour-}}{1}%
%{+}{{\literatecolour+}}{1}%
%{*}{{\literatecolour*}}{1}%
{/}{{\literatecolour/}}{1}%
{!}{{\literatecolour!}}{1}%
%{(}{{\literatecolour(}}{1}%
%{)}{{\literatecolour)}}{1}%
%{[}{{\literatecolour[}}{1}%
%{]}{{\literatecolour]}}{1}%
{<}{{\literatecolour<}}{1}%
{>}{{\literatecolour>}}{1}%
{>>>}{{\textcolor{promptcolour}{>>>}}}{1}%
,%
breaklines=true,
breakatwhitespace= true,
xleftmargin=\framemargin,
xrightmargin=\framemargin,
aboveskip=1ex,
belowskip=1ex,
frame=trbl, %trbl
numbers=none,
%frameround=tttt,
rulecolor=\color{black!40},
%framexleftmargin=\framemargin,
%framextopmargin=.1ex,
%framexbottommargin=.1ex,
%framexrightmargin=\framemargin,
%framexleftmargin=1mm, framextopmargin=1mm, frame=shadowbox, rulesepcolor=\color{blue},#1
%frame=tb,
%backgroundcolor=\color{yellow!10}
backgroundcolor=\color{gray}
%}
}

\newcommand{\N}[1]{\check{#1}}
\newcommand{\D}[1]{\overline{#1}}

\newcommand{\inpyth}{\lstinline[style=pythonstyle, basicstyle=\ttfamily]} %[]%

\lstnewenvironment{inpython}[1][]{
\lstset{style=pythonstyle, frame=trbl, belowcaptionskip=\belowtitleskip}
}{}

\newcommand{\includecode}[2][py]{\lstinputlisting[caption=#2,label=list:#2,style=pythonstyle,
float=!htpb]{#2}}

\newcommand{\mmo}[1]{({\bf mmo comment:} \emph{#1})}
\bibliographystyle{plain}

\title{Assessment of a Shen-Fourier turbulent channel flow solver for large-scale simulations}
\author{Mikael Mortensen and Diako Darian}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
\section*{Abstract}

\section{Introduction}

\section{Preliminaries}
The Navier-Stokes equations used to describe turbulent flow in a doubly periodic channel are
\begin{align}
 \frac{\partial \bm{u}}{\partial t} + (\bm{u} \cdot \nabla) \bm{u}   &= \nu \nabla^2 \bm{u} - \nabla{p} + \bm{f},  \label{eq:NS} \\
 \nabla \cdot \bm{u} &= 0,
\end{align}
where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ and $t$ are position and time,   $p(\bm{x}, t)$ is the instantaneous pressure normalized by a constant density and $\bm{f} = (0, f_y, 0)$ is a constant forcing vector. The computational domain is $\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$ with channel walls located at $x=\pm 1$ such that no-slip applies at $ \bm{u}(\pm 1, y, z, t) = 0$. The walls are spanning the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions.

The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall normal direction. Three different sets of basis functions and function spaces are relevant for the wall normal direction
\begin{align}
&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_0, \phi_1,\ldots, \phi_{N_x}\} \label{eq:Tk}\\
& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_0, \D{\phi}_1, \ldots, \D{\phi}_{N_x-2} \} \label{eq:phiD}\\
& \N{\phi}_k(x) = T_k(x) - \left(\frac{k}{k+2}\right)^2 T_{k+2}(x), & \N{V}_{N_x} &= \text{span} \{\N{\phi}_1, \N{\phi}_2, \ldots, \N{\phi}_{N_x-2} \} \label{eq:phiN} 
\end{align}
where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind. The basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were suggested by Shen, and satisfy, respectively, the boundary conditions $\D{\phi}_k(\pm 1) = 0$ and $\N{\phi}'_k(\pm 1)=0$. The last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: \int_{-1}^{1}v/\sqrt{1-x^2} dx = 0, \, v'(\pm 1) = 0 \}$.

Three-dimensional basis functions that are periodic in $y$ and $z$ directions, can now be defined as
\begin{align}
  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
\end{align}
where $\imath=\sqrt{-1}$ and 
\begin{align}
K_N = \Big\{\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / &(l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
 &(l, m, n) \in  [0, \ldots, N_x] \times [-\frac{N_y}{2},\ldots, \frac{N_y}{2}-1] \times[-\frac{N_z}{2},\ldots, \frac{N_z}{2}-1] \Big\}. \label{eq:wavenumbermesh}
\end{align}
$\D{K}_N$ and $\N{K}_N$ are both equal to $K_N$ except from in index $l$ that ranges from $[0, 1, \ldots, N_x-2]$ and $[1, 2, \ldots, N_x-2]$ for $\D{K}_N$ and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). The computational mesh has in real space been created using $N = (N_x, N_y, N_z)$ intervals, where the two periodic directions use uniform intervals. The computational mesh is given as
\begin{align}
  X_N = \Big\{ \bm{x} \in \mathcal{R}^3/&(x_i, y_j, z_k) \in \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
  &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\} \label{eq:Xn}
\end{align}
where $x_i = h(i)$ represents either
\begin{equation}
 h(i) = \begin{cases}
   \cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Lobatto points}, \\
   \cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Chebyshev points}. \\
 \end{cases}
\end{equation}

For the Navier Stokes equations we look for solutions for the velocity components of the form
\begin{equation}
u(\bm{x}, t) = \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{u}_{\bm{k}}(t) \D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_sol}
\end{equation}
where $\hat{u}_{\bm{k}}(t) = \hat{u}(l, {m}, {n}, t)$ are the expansion coefficients for the velocity component in $x$-direction and the scaling by $N_y$ and $N_z$ is merely for convenience and compliance with the inverse discrete Fourier transform. Similar expressions apply for the remaining two velocity components. Note that from now on we will simply use the notation $\hat{u}$ for $\hat{u}_{\bm{k}}(t)$, when it is possible to simplify without loss of clarity. Likewise we will simply use $u$ for $u(\bm{x}, t)$.

If the expansion coefficients $\hat{u}$  are known for the entire wavenumber mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_sol}) may be evaluated on the mesh (\ref{eq:Xn}) using fast Fourier and fast Shen inverse transforms as
\begin{align}
u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n \underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l \hat{u}(l,m,n)\D{\phi}_l(x_i)}_{\mathcal{S}_x^{-1}} e^{\imath \underline{m} y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , \quad \forall \, \bm{x} \in X_N, \notag \\
  u &= \D{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y (\D{\mathcal{S}}^{-1}_x (\hat{u}))), \quad \forall \, \bm{x} \in X_N.  \label{eq:ifft} 
\end{align}
Here $\D{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represents inverse Fourier transforms along directions $y$ and $z$ respectively, and the inverse Shen transform $\D{\mathcal{S}}_{x}^{-1}$ is performed along the wall normal $x$ direction. Note that the transforms are applied sequentially to the entire computational mesh. The inverse Shen transform may be computed using fast Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, \ref{eq:phiN}). The transforms are slightly different for the three bases and ${\mathcal{S}}, \D{\mathcal{S}}$ and $\N{\mathcal{S}}$ are used to distinguish between them with obvious notation. Python implementations of forward and inverse Shen transforms and scalar products are given in the Appendix.

A three-dimensional scalar product in the weighted $[L^2_w(\Omega)]^3$ space is defined as
\begin{align}
 \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, \notag \\
   &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}\,dx_wdydz,
\end{align}
where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction, and we have used $dx_w = dx/\sqrt{1-x^2}$. In this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, where quadrature is employed for integration. As such the scalar product corresponds to sums that may be computed using fast transforms
\begin{align}
 \left< u, \D{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} }_{\mathcal{F}_m} \D{\phi}_{l}(x_i) w_i}_{\D{\mathcal{S}}_l},   \notag \\
  &=  h \D{\mathcal{S}}(u) = h \D{\mathcal{S}}_{l} (\mathcal{F}_{{m}} (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \D{K}_N. \label{eq:sst1}
\end{align}
Here $h\D{\mathcal{S}}$ denotes the complete three dimensional scalar product and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and $\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and $y$-directions, respectively, and $\D{\mathcal{S}}_{l}(\cdot) = (\cdot, \D{\phi}_l)_w$ is the Shen scalar product in the $x$-direction. The weights, $w_i$, required for the Shen transforms are defined as
\begin{equation}
 w_i = \begin{cases}
       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for Gauss-Lobatto},\\
       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for Gauss-Chebyshev},      
 \end{cases}
\end{equation}
where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.

The scalar product  $h\D{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a complete transform. To find a transformation between physical $u$ and spectral $\hat{u}$ we make use of Eq. (\ref{eq:u_sol}) directly and use the discrete orthogonality of the Fourier basis functions
\begin{align}
\left<u, \D{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< \sum_{\bm{m}=(q,r,s) \in \D{K}_N} \hat{u}_{\bm{m}} \D{\psi}_{\bm{m}}, \D{\psi}_{\bm{k}} \right>_w, \notag \\
           &= h \sum_{q=0}^{N_x-2} \sum_{i=0}^{N_x} \D{\phi}_{q}(x_i) \D{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
           &= h \sum_{p=0}^{N_x-2} \D{B}_{lq} \hat{u}(q, {m}, {n}, t), \label{eq:sst2}
\end{align}
where the scalar product $\D{B}_{lq} = (\D{\phi}_q, \D{\phi}_l)_w = \sum_{i=0}^{N_x} \D{\phi}_{q}(x_i) \D{\phi}_{l}(x_i) w_i$ can be assembled to a tridiagonal mass matrix. The complete transformation is now obtained by setting Eq. (\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. Moving to matrix notation we get
\begin{align}
h\sum_{q=0}^{N_x-2} \D{B}_{lq} \hat{u}(q, {m}, {n}, t) &= h\D{\mathcal{S}}(u)(l, m, n, t), \notag \\
 \hat{u}_{\bm{k}}(t) &= \D{\mathcal{T}}(u) =  \D{B}^{-1}\D{\mathcal{S}}(u), \quad \forall \, \bm{k} \in \D{K}_N.
\end{align}
where $\D{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = \D{\mathcal{T}}^{-1}(\D{\mathcal{T}}(u))$. Note that since $\D{B}$ assembles to a tridiagonal matrix the solution can be obtained very fast ($O(N_x)$) using one forward elimination followed by a backwards substitution (Thomas algorithm). The complete transformation thus requires a fast Chebyshev transform and a fast linear algebraic solve. Similar transforms $\mathcal{T}$ and $\N{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and (\ref{eq:phiN}), using mass matrices $B_{kj}=(T_j, T_k)_w$ and $\N{B}_{kj}=(\N{\phi}_j, \N{\phi}_j)_w$ and scalar products $\mathcal{S}_k(\cdot) = (\cdot, T_k)_w$ and $\N{\mathcal{S}}_k(\cdot) = (\cdot, \N{\phi}_k)_w$.

\section{The incremental pressure correction scheme}

The Navier Stokes equations are discretized in time using a second order incremental pressure correction scheme with Crank-Nicolson on diffusion and explicit Adams Bashforth on the nonlinear convection. The incremental pressure correction scheme (IPCS) that is used consists of a tentative momentum equation followed by a pressure correction. The two steps may be performed iteratively. The equations are iterated forward in time in equal size time steps $\triangle t$ from initial conditions $u(\bm{x}, 0) = u^0(\bm{x})$ at $t=0$, such that $t_n = n\triangle t$, $n \in \mathcal{Z}$. For turbulent channel flows initial conditions are simply a perturbed state used to get a turbulent flow going.

The IPCS attempts to find velocity vector $\bm{u}^{n}$ on time step $t_n$  and pressure midway between $t_n$ and $t_{n-1}$, $p^{n-1/2}$, given the solutions on all previous time steps. The three steps that are required solved on each time step are
\begin{align}
i)&\,\, \frac{\bm{u}^{*}-\bm{u}^{n-1}}{\triangle t} + \bm{N}^{n-1/2}   = \nu \nabla^2 \bm{u}^{n-1/2} - \nabla{p}^{n-3/2} + \bm{f}, \quad \bm{u}^*=0 \text{ for } x=\pm1   \label{eq:NSCN} \\
ii)&\,\,\begin{cases}
  \frac{\bm{u}^{n}-\bm{u}^*}{\triangle t} &= -\nabla p^* \\
  \nabla \cdot \bm{u}^{n} &= 0, \quad \bm{u}^n\cdot \bm{n} = 0,\text{ for } x=\pm1
  \end{cases} \label{eq:momupdate}\\
  iii)&\,\,p^{n-1/2}=p^{n-3/2}+p^* \label{eq:pupdate} 
\end{align}
Here $\bm{u}^{n-1/2} = 0.5(\bm{u}^*+\bm{u}^{n-1})$, $\bm{u}^*$ is a tentative velocity, $p^*$ is a pressure correction, and the explicit convection is computed as $\bm{N}^{n-1/2}=1.5(\bm{u}^{n-1}\cdot \nabla) \bm{u}^{n-1} - 0.5(\bm{u}^{n-2}\cdot \nabla) \bm{u}^{n-2}$. The second step is computed by solving a Poisson equation for the pressure correction
\begin{equation}
  \nabla^2p^* = \frac{\nabla \cdot \bm{u}^*}{\triangle t}
\end{equation}
and subsequently updating the velocity through (\ref{eq:momupdate}).

All three velocity components use the Dirichlet basis (\ref{eq:u_sol}), whereas the pressure is searched for in $\N{V}_N$
\begin{equation}
p(\bm{x}, t) = \frac{1}{N_yN_z}\sum_{\bm{k} \in \N{K}_N} \hat{p}_{\bm{k}}(t) \N{\psi}_{\bm{k}}(\bm{x}), \label{eq:p_sol}
\end{equation}

The spectral Galerkin method becomes:

1) Find $\bm{u}^*$ in $\D{V}_N^3$ by solving
\begin{multline}
 \frac{1}{\triangle t}\left<\bm{u}^*-\bm{u}^{n-1}, \bm{\D{\psi}}\right>_w + \left<\bm{N}^{n-1/2}, \bm{\D{\psi}} \right>_w   = \nu \left<\nabla^2 \bm{u}^{n-1/2}, \bm{\D{\psi}} \right>_w \\ - \left<\nabla{p}^{n-3/2}, \bm{\D{\psi}} \right>_w + \left<\bm{f}, \bm{\D{\psi}} \right>_w \,\,  \forall \, \bm{\D{\psi}} \in \D{V}_N^3. \label{eq:tentative}
\end{multline}

2) Compute the pressure correction $p^*\in \N{V}_N$ by solving
\begin{equation}
\left< \nabla^2 p^*, \N{\psi} \right>_w = \frac{1}{\triangle t} \left<\nabla \cdot \bm{u}^*, \N{\psi} \right>_w\,\, \forall \, \N{\psi} \in \N{V}_N. \label{eq:pcorr}
\end{equation}

3) Update the velocity, $\bm{u}^n \in \D{V}_N^3$, and pressure, $p^{n-1/2} \in \N{V}_N$, by solving
\begin{align}
\left< \bm{u}^n, \bm{\D{\psi}}\right>_w &= \left< \bm{u}^*, \bm{\D{\psi}}\right>_w - \triangle t \left< \nabla p^*, \bm{\D{\psi}} \right>_w \,\,  &\forall \, \bm{\D{\psi}} \in \D{V}_N^3, \label{eq:u_update} \\
\left<p^{n-1/2}, \N{\psi}\right>_w &= \left<p^{n-3/2}, \N{\psi}\right>_w + \left<p^*, \N{\psi}\right>_w \,\, &\forall \, \N{\psi} \in \N{V}_N,\label{eq:pres_update}
\end{align}
where the pressure update simply resorts to $\hat{p}^{n-1/2} = \hat{p}^{n-3/2} + \hat{p}^* \, \forall \, \bm{k}\in \N{K}_N$.

\section{Implementation}
The vector valued equations (\ref{eq:tentative}) and (\ref{eq:u_update}) are solved in a segregated manner, one component at the time. We use $u_i$ and $\hat{u}_i$ for component $i$ of the real and transformed velocity vector. The inner products in Eqs. (\ref{eq:tentative}, \ref{eq:pcorr}, \ref{eq:u_update}, \ref{eq:pres_update}) lead to a range of new matrices
\begin{align}
 \left< u_i^* - u_i^{n-1}, \D{\psi}\right>_w &= h\D{B}\left({\hat{u}}_i^*-{\hat{u}}_i^{n-1} \right) \quad &i \in (0,1,2), \\ 
 \left<\nabla^2 {u}_i^{n-1/2}, \D{\psi}\right>_w &= -h  \left( \D{A} +(\underline{m}^2+\underline{n}^2)\D{B}\right) \hat{u}_i^{n-1/2} \quad &i \in (0,1,2), \\ 
 \left< f_i, \D{\psi}\right>_w &= h\D{\mathcal{S}}(f_i) &i \in (0,1,2), \\
 \left<N_i^{n-1/2}, \D{\psi} \right>_w &= h\D{\mathcal{S}}(N_i^{n-1/2}) &i \in (0,1,2), \\  
 \left< \nabla p^{n-3/2}, \bm{\D{\psi}} \right>_w &= h\left[\grave{C},\, \imath \underline{m} \grave{B},\, \imath \underline{n} \grave{B} \right] \hat{p}^{n-3/2}, \\ 
 \left< \nabla^2 p^*, \N{\psi} \right>_w &= -h\left(\N{A} +(\underline{m}^2 + \underline{n}^2)\N{B} \right) \hat{p}^*, \\ 
 \left<\nabla \cdot \bm{u}^*, \N{\psi} \right>_w &= h\left( \tilde{C} \hat{u}^* + \imath \underline{m} \tilde{B} \hat{v}^* + \imath \underline{n} \tilde{B} \hat{w}^* \right), 
\end{align}
where $\D{A}_{kj} = -\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w, \N{A}_{kj} = -\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w,  \grave{C}_{kj} = \left(\N{\phi}^{'}_j, \D{\phi}_k \right)_w, \tilde{C}_{kj} = (\D{\phi}^{'}_j, \N{\phi}_k)_w, \N{B}_{kj} = \left( \N{\phi}_j, \N{\phi}_k \right)_w, \grave{B}_{kj} = (\N{\phi}_j, \D{\phi}_k)_w$ and $\tilde{B}_{kj} = (\D{\phi}_j, \N{\phi}_k)_w$ are all more or less sparse matrices. Arranging the governing equations on matrix form we obtain for the three tentative velocity components in (\ref{eq:tentative})
\begin{align}
\D{H} \hat{u}^* &= \left(\frac{4}{\nu \triangle t} \D{B}-\D{H}\right) \hat{u}^{n-1} - \left[\D{\mathcal{S}}(N_x^{n-1/2}) + \grave{C} \hat{p}^{n-3/2} + \D{\mathcal{S}}(f_x)\right]\frac{2}{\nu}, \label{eq:tentativeUA}\\
\D{H} \hat{v}^* &=  \left(\frac{4}{\nu \triangle t} \D{B}-\D{H}\right)\hat{v}^{n-1} - \left[\D{\mathcal{S}}(N_y^{n-1/2}) + \imath \underline{m} \grave{B} \hat{p}^{n-3/2} + \D{\mathcal{S}}(f_y)\right]\frac{2}{\nu}, \label{eq:tentativeVA}\\
\D{H} \hat{w}^* &= \left(\frac{4}{\nu \triangle t} \D{B}-\D{H}\right) \hat{w}^{n-1} - \left[\D{\mathcal{S}}(N_z^{n-1/2}) + \imath \underline{n} \grave{B} \hat{p}^{n-3/2} + \D{\mathcal{S}}(f_z)\right]\frac{2}{\nu},\label{eq:tentativeWA}
\end{align}
where $\D{H}=\D{A} +(\underline{m}^2 + \underline{n}^2 + \frac{2}{\nu \triangle t})\D{B}$ is a matrix consisting of one subdiagonal, and where every second diagonal is zero. As such, we may split the matrix into two smaller (and decoupled) matrices for odd and even coefficients $H^e_{k,j} = H_{2k,2j}$ and $H^o_{k,j} = H_{2k+1,2j+1}$, where the two matrices $H^{e,o}$ are of type upper Hessenberg, with only 4 distinct values on each row. This structure of $H$ allows us to perform a very efficient tailored LU decomposition, as shown in Algorithm (\ref{alg:lu}), and a very efficient tailored solve through Algorithm (\ref{alg:lusolve}). Regarding Algorithm (\ref{alg:lusolve}), we note that the ability to decouple the system into odd and even coefficients leads to two subsystem that may be trivially solved simultaneously in two threads by letting one thread call LUODDEVEN with even coefficients, and the second thread call it for odd.

The linear system arising for the pressure correction equation is similar to the tentative velocities
\begin{equation}
\N{H}\hat{p}^* = -\frac{1}{\triangle t}\left( \tilde{C} \hat{u}^* + \imath \underline{m} \tilde{B} \hat{v}^* + \imath \underline{n} \tilde{B} \hat{w}^* \right), \label{eq:pressureA}
\end{equation}
where $\N{H}=\N{A} +(\underline{m}^2 + \underline{n}^2)\N{B}$ is of the same structure as $\D{H}$ and as such  should be factorizable in the same way. However, entries of the matrix $\N{H}_{kj}$ contain the column index $j^2$ (see $\N{A}$), and as such the values on each row of the matrix $\N{H}_{kj}$ are not constant for $j \ge k+4$ and the LU decomposition of $\N{H}$ will require a full upper $U$ matrix. To circumvent this we define $\N{H}_{kj} = \underline{H}_{kl}J_{lj}$, where $J_{kj}= j^2\delta_{kj}$ (no summation implied). We then set $w_k=J_{kj}\hat{p}_j$, use $b_k$ for the right hand side of (\ref{eq:pressureA}) and solve
\begin{align}
  \underline{H}_{kj}w_j &= b_k &\forall \, k \in 1, 2, \ldots, N_x-2, \\
  \hat{p}_k &= J_{kj}^{-1}w_j &\forall \, k \in 1, 2, \ldots, N_x-2.
\end{align}
Here the first system $\underline{H}_{kj} w_j = b_k$ can be solved using the same LU factorization as the $\D{H}$ matrix, since $\underline{H}_{kj}=\underline{H}_{k, k+4}\, \forall \, j=k+6, k+8, \ldots$ and there are only 4 distinct values on each row. The implementation is the same as for $\D{H}$, except that for the matrix $\underline{H}$ the indices start at 1 instead of zero.

The velocity update is computed on matrix form as
\begin{align}
  \hat{u}^n &= \hat{u}^* - \triangle t \D{B}^{-1} \left( \grave{C} \hat{p}^* \right),\\
  \hat{v}^n &= \hat{v}^* - \triangle t \D{B}^{-1} \left( \imath \underline{m} \grave{B} \hat{p}^* \right),\\ 
  \hat{w}^n &= \hat{w}^* - \triangle t \D{B}^{-1} \left( \imath \underline{n} \grave{B} \hat{p}^* \right).
\end{align}
using a tridiagonal solver.

The convection term is computed in standard form with a pseudo-spectral method that requires a fast scalar product
\begin{equation}
  \D{\mathcal{S}}(N_i^{n-1}) = \D{\mathcal{S}}(\bm{u}^{n-1} \cdot \nabla u_i^{n-1}) \quad \forall \, i \in (0, 1, 2).
\end{equation}
We obtain the gradients of $u_i$ in real space by projecting $\partial u/\partial x$ to $\D{V}_{N_x}$ (boundary condition $\partial u/\partial x(\pm 1)=0$ follows from continuity) and $\partial v/\partial x$ and $\partial w/\partial x$ to $V_{N_x}$. 
\begin{align}
  \frac{\partial u}{\partial x} &= \D{\mathcal{T}}^{-1} (\D{B}^{-1}(\D{C}\hat{u})), \\
  \frac{\partial v}{\partial x} &= \mathcal{T}^{-1} (B^{-1}(\acute{C}\hat{v})), \\
  \frac{\partial w}{\partial x} &= \mathcal{T}^{-1} (B^{-1}(\acute{C}\hat{w})),  
\end{align}
where $\D{C}_{kj} = (\D{\phi}_j^{'}, \D{\phi}_k)_w$,  $\acute{C_{kj}} = (\D{\phi}_j^{'}, T_k)_w$ and $B=(T_j, T_k)_w$. Partial derivatives with respect to periodic directions are computed as
\begin{align}
 \frac{\partial u_i}{\partial y} &= \D{\mathcal{T}}^{-1}(\imath \underline{m} \hat{u}_i) &\forall \, i \in (0, 1, 2), \\
 \frac{\partial u_i}{\partial w} &= \D{\mathcal{T}}^{-1}(\imath \underline{n} \hat{u}_i) &\forall \, i \in (0, 1, 2).
\end{align}
Dealiasing is implemented using the 2/3-rule.

We finally note that the right hand sides of all the linear systems that need to solved can be assembled fast using tailored matrix vector products taking advantage of the fact that there are at most three distinct diagonals in any matrix. This concludes the implementation required to solve the Navier-Stokes equations by the incremental pressure correction method and the spectral Shen-Fourier Galerkin method. We note that all terms in assembly and solve may be performed using fast routines, at worst scaling like $O(N_x \log_2 N_x)$.

\section{Serial scaling}

\section{Parallel scaling}

\begin{algorithm}
\caption{LU factorization of matrix $H_{k,j}$, that may be split into two decoupled upper Hessenberg matrices $H^e_{k,j} = H_{2k,2j}$ and $H^o_{k,j} = H_{2k+1,2j+1}$ for even and odd coefficients. Return LU factorization for even and odd coefficients, such that $L^eU^e=H^e$ and $L^oU^o=H^o$. Note that $L^{e}$ and $L^{o}$ each have one single nonzero diagonal in addition to the unity main diagonal, and as such only a vector is stored. Each $U^{e}$ and $U^{o}$ has three distinct diagonals and $U^{e}_{kj}=U^{e}_{k, k+2} \,\forall j > k+2, j< M^e$ and $U^{o}_{kj}=U^{o}_{k, k+2} \,\forall j > k+2, j< M^o$. As such, a sparse three-diagonal storage may be used for $U^e$ and $U^o$.}
\label{alg:lu}
\begin{algorithmic}[1]
\Procedure{LUoddeven}{}
  \State \textbf{Input:} $\{H\}_{k,j=0}^{N-2}$
  \State \textit{integer} $M^e \gets (N-2)/2$
  \State \textit{integer} $M^o \gets (N-3)/2$
  \State ${U}^e_{0j} \gets {H}_{0,2j}, \, \forall j=0,1,2 $
  \State ${U}^o_{0j} \gets {H}_{1,2j+1}, \, \forall j=0,1,2 $
  \For{$k=1, 2,  \ldots, M^e-2$}
    \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
    \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall j=k, k+1, k+2$
    \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
    \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, \forall j=k, k+1, k+2$
  \EndFor
  \State $k \gets M^e-1$
  \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
  \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall j=k, k+1$
  \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
  \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, \forall j=k, k+1$
  \State $k \gets M^e$
  \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
  \State ${U}^e_{kk} \gets {H}_{2k,2k} - {L}^e_{k}{U}^e_{k-1,k}$
  \If{$M^o == M^e$ }
    \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
    \State ${U}^o_{kk} \gets {H}_{2k+1,2k+1} - {L}^o_{k}{U}^o_{k-1,k}$

  \EndIf  
  \State \Return $\{L^e_k\}_{k=0}^{M^e},\{U^e\}_{k,j=0}^{M^e}, \{L^o_k\}_{k=0}^{M^o},\{U^o\}_{k,j=0}^{M^o}$
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Solve of $H_{kj}u_j = b_k$, where $H$ has previously been LU factorized for odd even coefficients using Algorithm (\ref{alg:lu}).}
\label{alg:lusolve}
\begin{algorithmic}[1]
\Procedure{LUoddeven}{}
  \State \textbf{Input:} $\{U_{kj}\}_{k,j=0}^{M}, \{L_k\}_{k=0}^{M}, \{b_k\}_{k=0}^{M}$ 
  \State // Solve $Ly=b$  
  \State $y_0 \gets b_0$
  \For{$k = 1$ \textbf{to} $M$}
    \State $y_k \gets b_k - L_ky_{k-1}$
  \EndFor
  \State // Solve $Uu=y$
  \State double $s = 0.0$     
  \State $u_M = y_M / U_{MM}$    
  \For{$i = M-1$ \textbf{to} $0$ \textbf{step} $-1$}
     \State $u_i \gets y_i - U_{i, i+1}\cdot u_{i+1}$
     \If{$i < M-1$}
       \State $s \gets s + u_{i+2}$
       \State $u_i \gets u_i - s\cdot U_{i,i+2}$            
     \EndIf
     \State $u_i \gets  u_i / U_{i,i}$
   \EndFor
  \State \Return $\{u_k\}_{k=0}^{M}$  
\EndProcedure
\end{algorithmic}
\begin{algorithmic}[1]
\Procedure{LUSolve}{}
  \State \textbf{Input: $\{u_k\}_{k=0}^{N-2}, \{b_k\}_{k=0}^{N-2}, \{U_{kj}^{e}\}_{k,j=0}^{M^{e}}, \{U_{kj}^{o}\}_{k,j=0}^{M^{o}}, \{L^{e}_k\}_{k=0}^{M^{e}}, \{L^{o}_k\}_{k=0}^{M^{o}}$}
  \State $\{u_{2k}\}_{k=0}^{M^e} \gets $ LUoddeven($U^{e}, L^e, \{b_{2k}\}_{k=0}^{M^e}$)
  \State $\{u_{2k+1}\}_{k=0}^{M^o} \gets $ LUoddeven($U^{o}, L^o, \{b_{2k+1}\}_{k=0}^{M^o}$)

\EndProcedure
\end{algorithmic}

\end{algorithm}

\section{Appendices}
\subsection*{Fast Shen transforms}
The fast forward and inverse Chebyshev transforms are used by the Shen transforms. The code below computes both scalar products and complete transforms for both

\begin{python}
from numpy import *
from scipy.fftpack import dct

def fct(f, transform="GL"):
    """Fast Chebyshev transform."""
    N = f.shape[0]
    f_hat = zeros(N)
    if transform == "GL":
        fk[:] = dct(fj, type=2, axis=0)            
        fk /= N
        fk[0] /= 2

    elif transform == "GC":
        fk[:] = dct(fj, type=1, axis=0)/(N-1)
        fk[0] /= 2
        fk[-1] /= 2

    return fk

def ifct(fk, transform="GL"):
    """Inverse fast Chebyshev transform."""
    if transform == "GL":
        f = 0.5*dct(fk, type=3, axis=0)
        f += 0.5*fk[0]

    elif transform == "GC":
        f = 0.5*dct(fk, type=1, axis=0)
        f += 0.5*fk[0]
        f[::2] += 0.5*fk[-1]
        f[1::2] -= 0.5*fk[-1]

    return f

def fastChebScalar(f, transform="GL"):
    """Fast Chebyshev scalar product."""
    N = f.shape[0]
    if transform == "GL":
        fk = dct(fj, type=2, axis=0)*pi/(2*N)

    elif transform == "GC":
        fk = dct(fj, type=1, axis=0)*pi/(2*(N-1))
    return fk

def fastShenScalar(f, transform="GL"):
    """Fast Shen scalar product"""                
    fk = fastChebScalar(f, transform)
    fk[:-2] -= fk[2:]
    return fk

def ifst(fk):
    """Fast inverse Shen transform
    """
    N = len(fj)
    wk = np.zeros(N, dtype=fk.dtype)        
    wk[:-2] = fk[:-2] 
    wk[2:] -= fk[:-2] 
    return ifct(wk)

def fst(fj, transform="GL"):
    """Fast Shen transform"""
    fk = fastShenScalar(fj, transform)

    N = fj.shape[0]
    if transform == "GL":
        ck = np.ones(N-2); ck[0] = 2

    elif transform == "GC":
        ck = np.ones(N-2); ck[0] = 2; ck[-1] = 2

    a = np.ones(N-4)*(-np.pi/2)
    b = np.pi/2*(ck+1)
    c = a.copy()
    fk[:-2] = TDMA_1D(a, b, c, fk[:-2])

    return fk

def fastShenScalar(f, transform="GL"):
    """Fast Shen scalar product Neumann basis"""        
    k  = wavenumbers(f.shape)
    fk = fastChebScalar(f, transform)
    fk[:-2] -= ((k/(k+2))**2) * fk[2:]
    return fk

def ifst(fk):
    """Fast inverse Shen scalar transform Neumann basis"""
    k = self.wavenumbers(fk.shape[0])
    wk = np.zeros(fk.shape[0])    
    wk[1:-2] = fk[1:-2]
    wk[3:] -= (k[1:]/(k[1:]+2))**2*fk[1:-2]
    f = ifct(wk, transform)
    return f

def fst(f, transform="GL"):
    """Fast Shen transform Neumann basis"""
    fk = fastShenScalar(f, transform)
    N = fj.shape[0]
    k = wavenumbers(N)
    ck = np.ones(N-3)
    if transform == "GC": ck[-1] = 2
    a = np.ones(N-5)*(-np.pi/2)*(k[1:-2]/(k[1:-2]+2))**2
    b = np.pi/2*(1+ck*(k[1:]/(k[1:]+2))**4)
    c = a.copy()
    fk[1:-2] = TDMA_1D(a, b, c, fk[1:-2])
    return fk

# Cython implementation of TDMA    
import numpy as np
cimport cython
cimport numpy as np
#cython: boundscheck=False
#cython: wraparound=False

ctypedef fused T:
    np.float64_t
    np.complex128_t

def TDMA_1D(np.ndarray[np.float64_t, ndim=1] a, 
            np.ndarray[np.float64_t, ndim=1] b, 
            np.ndarray[np.float64_t, ndim=1] c, 
            np.ndarray[T, ndim=1] d):
    cdef:
        unsigned int n = b.shape[0]
        unsigned int m = a.shape[0]
        unsigned int k = n - m
        int i

    for i in range(m):
        d[i + k] -= d[i] * a[i] / b[i]
        b[i + k] -= c[i] * a[i] / b[i]
    for i in range(m - 1, -1, -1):
        d[i] -= d[i + k] * c[i] / b[i + k]
    for i in range(n):
        d[i] /= b[i]

    return d

\end{python}
\subsection*{Matrices}

\begin{equation}
 \{\D{A}_{kj}\}_{k,j=0}^{N_x-2} = -\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w = \begin{cases}
 2\pi(k+1)(k+2), &j=k,\\
 4\pi(k+1), & j=k+2, k+4, k+6, \ldots, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\N{A}_{kj}\}_{k,j=1}^{N_x-2} = -\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w = \begin{cases}
 2\pi j^2(k+1)/(k+2), &j=k,\\
 4\pi j^2(k+1)/(k+2)^2, & j=k+2, k+4, k+6, \ldots, \\
 0, &j>k, \text{ or } k+j \text{ odd}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\D{C}_{kj}\}_{k,j=0}^{N_x-2} = \left(\D{\phi}^{'}_j, \D{\phi}_k \right)_w = \begin{cases}
 \pi(k+1), &j=k+1,\\
 -\pi(k+1), & j=k-1, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\grave{C}_{kj}\}_{k=0,j=1}^{N_x-2} = \left(\N{\phi}^{'}_j, \D{\phi}_k \right)_w = \begin{cases}
 \pi(k+1), &j=k+1,\\
 -\pi(k-1)^2/(k+1) , & j=k-1, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\tilde{C}_{kj}\}_{k=1,j=0}^{N_x-2} = (\D{\phi}^{'}_j, \N{\phi}_k)_w = \begin{cases}
 -\pi(k+1), &j=k-1,\\
 -\pi(2-\left(\frac{k}{k+2}\right)^2(k+3)) , & j=k+1, \\
 -2 \pi(1-\left(\frac{k}{k+2}\right)^2), &j=k+3, k+5, k+7, \ldots,\\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
  \{\acute{C}_{kj}\}_{k,j=0}^{N_x-2, N_x} = (\D{\phi}^{'}_j, T_k)_w = \begin{cases}
    -\pi (k+1), &j=k-1, \\
    -2 \pi,   &j = k+1, k+3, k+5, \ldots, \\
    0, &\text{otherwise}
  \end{cases}
\end{equation}

\begin{equation}
  \{\D{B}_{kj}\}_{j,k=0}^{N_x-2} = (\D{\phi}_j, \D{\phi}_k)_w = \begin{cases}
  \frac{c_k+c_{k+2}}{2}\pi,& k = j \\
  -\frac{\pi}{2},& j=k\pm 2 \\
  0,& \text{otherwise}
\end{cases}
\end{equation}

\begin{equation}
 \{\tilde{B}_{kj}\}_{k=1, j=0}^{N_x-2} = (\D{\phi}_j, \N{\phi}_k)_w = \begin{cases}
 -\frac{\pi}{2}, &j=k-2,\\
 -\frac{\pi}{2}\left(1+ {c_{k+2}}\left(\frac{k}{k+2}\right)^2 \right), &j=k,\\
 -\frac{\pi}{2} k^2/(k+2)^2 , & j=k+2, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\grave{B}_{kj}\}_{k=0, j=1}^{N_x-2} = (\N{\phi}_j, \D{\phi}_k)_w = \begin{cases}
 -\frac{\pi}{2} (k-2)^2/k^2 , & j=k-2, \\
 -\frac{\pi}{2}\left(1+ {c_{k+2}} \left(\frac{k}{k+2}\right)^2 \right), &j=k,\\
 -\frac{\pi}{2}, &j=k-2,\\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\N{B}_{kj}\}_{k,j=1}^{N_x-2} = (\N{\phi}_j, \N{\phi}_k)_w = \begin{cases}
 -\frac{\pi}{2} \left(\frac{k-2}{k}\right)^2, &j=k-2,\\
 -\frac{\pi}{2}\left(1+ {c_{k+2}}\left(\frac{k}{k+2}\right)^4 \right), &j=k,\\
 -\frac{\pi}{2} k^2/(k+2)^2 , & j=k+2, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{B_{kj}\}_{k,j=0}^{N_x} = (T_k, T_j)_w = \frac{c_k \pi}{2} \delta_{kj}
\end{equation}

For Gauss-Lobatto points we have $c_0=c_{N_x}=2$ and $c_k=1$ for $0<k<N_x$. For Gauss-Chebyshev points we have $c_0=2$ and $c_k=1$ for $k>0$. Note that for an $L^2_w$ scalar product we would have $c_{N_x}=1$ as well, but for the $l^2_w$ scalar product used here we get $c_{N_x}=2$. This disagreement, that follows from inexact quadrature at the highest mode, is not reported by Shen, but is reported by, e.g., Kopriva.

\end{document}

### 2016-01-29 paper.tex
Move to KMM solver

@@
 \section{Preliminaries}
-The Navier-Stokes equations used to describe turbulent flow in a doubly periodic channel are
+The Navier-Stokes equations used to describe turbulent flow in a doubly 
+periodic channel can be written as
 \begin{align}
- \frac{\partial \bm{u}}{\partial t} + (\bm{u} \cdot \nabla) \bm{u}   &= \nu \nabla^2 \bm{u} - \nabla{p} + \bm{f},  \label{eq:NS} \\
- \nabla \cdot \bm{u} &= 0,
+ \frac{\partial \bm{u}}{\partial t}   &= \bm{\mathcal{H}} + \nu 
+ \nabla^2 \bm{u} - \nabla{p}, \notag \\
+ \nabla \cdot \bm{u} &= 0, \label{eq:NS}
 \end{align}
-where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ and $t$ are position and time,   $p(\bm{x}, t)$ is the instantaneous pressure normalized by a constant density and $\bm{f} = (0, f_y, 0)$ is a constant forcing vector.
-The computational domain is $\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$ with channel walls located at $x=\pm 1$ such that no-slip applies at $ \bm{u}(\pm 1, y, z, t) = 0$.
-The walls are spanning the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions.
+where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ 
+and $t$ are position and time, $\bm{\mathcal{H}}(\bm{x}, t)$ represents the 
+nonlinear convection and the mean pressure gradient (the driving force), $\nu$ 
+is a constant dynamic viscosity and $p$ is the instantaneous pressure 
+fluctuation normalized by a constant density. The 
+computational domain is $\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$ with 
+channel walls located at $x=\pm 1$ such that no-slip applies at $ \bm{u}(\pm 1, 
+y, z, t) = 0$. The walls are spanning the $y-z$ plane and the equations are 
+periodic in the $y$ and $z$ directions.
 
@@
 & \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_0, \D{\phi}_1, \ldots, \D{\phi}_{N_x-2} \} \label{eq:phiD}\\
-& \N{\phi}_k(x) = T_k(x) - \left(\frac{k}{k+2}\right)^2 T_{k+2}(x), & \N{V}_{N_x} &= \text{span} \{\N{\phi}_1, \N{\phi}_2, \ldots, \N{\phi}_{N_x-2} \} \label{eq:phiN} 
+& \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
+\frac{k+1}{k+3} T_{k+4}, & \N{V}_{N_x} &= \text{span} \{\N{\phi}_0, \N{\phi}_1, 
+\ldots, \N{\phi}_{N_x-4} \} \label{eq:phiN} 
 \end{align}
-where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind.
-The basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were suggested by Shen, and satisfy, respectively, the boundary conditions $\D{\phi}_k(\pm 1) = 0$ and $\N{\phi}'_k(\pm 1)=0$.
-The last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: \int_{-1}^{1}v/\sqrt{1-x^2} dx = 0, \, v'(\pm 1) = 0 \}$.
+where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind. The 
+basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were 
+suggested by Shen, and satisfy, respectively, the boundary conditions 
+$\D{\phi}_k(\pm 1) = 0$, $\N{\phi}_k(\pm 1)=0$ and $\N{\phi}'_k(\pm 1)=0$. The 
+last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in 
+V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1) = v'(\pm 
+1) = 0 \}$.
 
-Three-dimensional basis functions that are periodic in $y$ and $z$ directions, can now be defined as
+Three-dimensional basis functions and function spaces that are periodic in $y$ 
+and $z$ directions, can now be defined as
 \begin{align}
@@
 \end{align}
-$\D{K}_N$ and $\N{K}_N$ are both equal to $K_N$ except from in index $l$ that ranges from $[0, 1, \ldots, N_x-2]$ and $[1, 2, \ldots, N_x-2]$ for $\D{K}_N$ and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}).
-The computational mesh has in real space been created using $N = (N_x, N_y, N_z)$ intervals, where the two periodic directions use uniform intervals.
-The computational mesh is given as
+$\D{K}_N$ and $\N{K}_N$ are both equal to $K_N$ except from index $l$ that 
+ranges from $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
+and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). The 
+computational mesh has in real space been created using $N = (N_x, N_y, N_z)$ 
+intervals, where the two periodic directions use uniform intervals. The 
+computational mesh is given as
 \begin{align}
@@
 
-For the Navier Stokes equations we look for solutions for the velocity components of the form
-\begin{equation}
-u(\bm{x}, t) = \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{u}_{\bm{k}}(t) \D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_sol}
-\end{equation}
-where $\hat{u}_{\bm{k}}(t) = \hat{u}(l, {m}, {n}, t)$ are the expansion coefficients for the velocity component in $x$-direction and the scaling by $N_y$ and $N_z$ is merely for convenience and compliance with the inverse discrete Fourier transform.
-Similar expressions apply for the remaining two velocity components.
-Note that from now on we will simply use the notation $\hat{u}$ for $\hat{u}_{\bm{k}}(t)$, when it is possible to simplify without loss of clarity.
+For the Navier Stokes equations we look for solutions of the velocity 
+components of the form
+\begin{align}
+u(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \N{K}_N} \hat{u}_{\bm{k}}(t) 
+\N{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_solx} \\
+v(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{v}_{\bm{k}}(t) 
+\D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_soly} \\
+w(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{w}_{\bm{k}}(t) 
+\D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_solz}
+\end{align}
+where $\hat{u}_{\bm{k}}(t) = \hat{u}(l, {m}, {n}, t)$ are the expansion 
+coefficients for the velocity component in $x$-direction (and similar for the 
+other two components) and the scaling by $N_y$ and $N_z$ is merely for 
+convenience 
+and compliance with the definition used for the inverse discrete Fourier 
+transform. Note that from now on we will simply use the notation $\hat{u}$ for 
+$\hat{u}_{\bm{k}}(t)$, when it is possible to simplify without loss of clarity. 
 Likewise we will simply use $u$ for $u(\bm{x}, t)$.
 
-If the expansion coefficients $\hat{u}$  are known for the entire wavenumber mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_sol}) may be evaluated on the mesh (\ref{eq:Xn}) using fast Fourier and fast Shen inverse transforms as
+If the expansion coefficients $\hat{u}$  are known for the entire wavenumber 
+mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_solx}) may be 
+evaluated on the mesh (\ref{eq:Xn}) using fast Fourier and fast Shen inverse 
+transforms as
 \begin{align}
-u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n \underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l \hat{u}(l,m,n)\D{\phi}_l(x_i)}_{\mathcal{S}_x^{-1}} e^{\imath \underline{m} y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , \quad \forall \, \bm{x} \in X_N, \notag \\
-  u &= \D{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y (\D{\mathcal{S}}^{-1}_x (\hat{u}))), \quad \forall \, \bm{x} \in X_N.  \label{eq:ifft} 
+u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n 
+\underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l 
+\hat{u}(l,m,n)\N{\phi}_l(x_i)}_{\N{\mathcal{S}}_x^{-1}} e^{\imath \underline{m} 
+y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , 
+\quad \forall \, \bm{x} \in X_N, \notag \\
+  u &= \N{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y 
+  (\N{\mathcal{S}}^{-1}_x (\hat{u}))), \quad \forall \, \bm{x} \in X_N.  
+  \label{eq:ifft} 
 \end{align}
-Here $\D{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represents inverse Fourier transforms along directions $y$ and $z$ respectively, and the inverse Shen transform $\D{\mathcal{S}}_{x}^{-1}$ is performed along the wall normal $x$ direction.
-Note that the transforms are applied sequentially to the entire computational mesh.
-The inverse Shen transform may be computed using fast Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, \ref{eq:phiN}).
-The transforms are slightly different for the three bases and ${\mathcal{S}}, \D{\mathcal{S}}$ and $\N{\mathcal{S}}$ are used to distinguish between them with obvious notation.
-Python implementations of forward and inverse Shen transforms and scalar products are given in the Appendix.
+Here $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
+transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represents inverse 
+Fourier transforms along directions $y$ and $z$ respectively, and the inverse 
+Shen transform $\N{\mathcal{S}}_{x}^{-1}$ is performed along the wall normal 
+$x$ direction. Note that the transforms are applied sequentially to the entire 
+computational mesh. The inverse Shen transform may be computed using fast 
+Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, 
+\ref{eq:phiN}). The transforms are slightly different for the three bases and 
+${\mathcal{S}}, \D{\mathcal{S}}$ and $\N{\mathcal{S}}$ are used to distinguish 
+between them with obvious notation. Python implementations of forward and 
+inverse Shen transforms and scalar products are given in the Appendix.
 
@@
 \begin{align}
- \left< u, \D{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} }_{\mathcal{F}_m} \D{\phi}_{l}(x_i) w_i}_{\D{\mathcal{S}}_l},   \notag \\
-  &=  h \D{\mathcal{S}}(u) = h \D{\mathcal{S}}_{l} (\mathcal{F}_{{m}} (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \D{K}_N. \label{eq:sst1}
+ \left< u, \N{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i 
+ \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath 
+ \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} 
+ }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) w_i}_{\N{\mathcal{S}}_l},   \notag \\
+  &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
+  (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
 \end{align}
-Here $h\D{\mathcal{S}}$ denotes the complete three dimensional scalar product and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant.
-$\mathcal{F}_{{n}}$ and $\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and $y$-directions, respectively, and $\D{\mathcal{S}}_{l}(\cdot) = (\cdot, \D{\phi}_l)_w$ is the Shen scalar product in the $x$-direction.
-The weights, $w_i$, required for the Shen transforms are defined as
+Here $h\N{\mathcal{S}}$ denotes the complete three dimensional scalar product 
+and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
+$\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and 
+$y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
+\N{\phi}_l)_w$ is the Shen scalar product in the $x$-direction for the 
+$\N{V}_N$ space. The weights, $w_i$, required for the Shen transforms are 
+defined as
 \begin{equation}
  w_i = \begin{cases}
-       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for Gauss-Lobatto},\\
-       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for Gauss-Chebyshev},      
+       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
+       Gauss-Lobatto points},\\
+       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for 
+       Gauss-Chebyshev points},      
  \end{cases}
@@
 
-The scalar product  $h\D{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a complete transform.
-To find a transformation between physical $u$ and spectral $\hat{u}$ we make use of Eq.
-(\ref{eq:u_sol}) directly and use the discrete orthogonality of the Fourier basis functions
+The scalar product  $h\N{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a 
+complete transform. To find a transformation between physical $u$ and spectral 
+$\hat{u}$ we make use of Eq. (\ref{eq:u_solx}) directly and use the discrete 
+orthogonality of the Fourier basis functions
 \begin{align}
-\left<u, \D{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< \sum_{\bm{m}=(q,r,s) \in \D{K}_N} \hat{u}_{\bm{m}} \D{\psi}_{\bm{m}}, \D{\psi}_{\bm{k}} \right>_w, \notag \\
-           &= h \sum_{q=0}^{N_x-2} \sum_{i=0}^{N_x} \D{\phi}_{q}(x_i) \D{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
-           &= h \sum_{p=0}^{N_x-2} \D{B}_{lq} \hat{u}(q, {m}, {n}, t), \label{eq:sst2}
+\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
+\sum_{\bm{m}=(q,r,s) \in \N{K}_N} \hat{u}_{\bm{m}} \N{\psi}_{\bm{m}}, 
+\N{\psi}_{\bm{k}} \right>_w, \notag \\
+           &= h \sum_{q=0}^{N_x-4} \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) 
+           \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
+           &= h \sum_{p=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t), 
+           \label{eq:sst2}
 \end{align}
-where the scalar product $\D{B}_{lq} = (\D{\phi}_q, \D{\phi}_l)_w = \sum_{i=0}^{N_x} \D{\phi}_{q}(x_i) \D{\phi}_{l}(x_i) w_i$ can be assembled to a tridiagonal mass matrix.
-The complete transformation is now obtained by setting Eq.
-(\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$.
-Moving to matrix notation we get
+where the scalar product $\N{B}_{lq} = (\N{\phi}_q, \N{\phi}_l)_w = 
+\sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
+pentadiagonal mass matrix. Likewise, for the remaining two velocity components, 
+the corresponding mass matrix, $\D{B}_{lq} = (\D{\phi}_q, \D{\phi}_l)_w$, is 
+tridiagonal. The complete transformation is now obtained by setting Eq. 
+(\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. Moving to 
+matrix notation we get
 \begin{align}
-h\sum_{q=0}^{N_x-2} \D{B}_{lq} \hat{u}(q, {m}, {n}, t) &= h\D{\mathcal{S}}(u)(l, m, n, t), \notag \\
- \hat{u}_{\bm{k}}(t) &= \D{\mathcal{T}}(u) =  \D{B}^{-1}\D{\mathcal{S}}(u), \quad \forall \, \bm{k} \in \D{K}_N.
+h\sum_{q=0}^{N_x-4} \N{B}_{lq}\, \hat{u}(q, {m}, {n}, t) &= 
+h\N{\mathcal{S}}(u)(l, m, n, t), \notag \\
+ \hat{u}_{\bm{k}}(t) &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
+ \quad \forall \, \bm{k} \in \N{K}_N.
 \end{align}
-where $\D{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = \D{\mathcal{T}}^{-1}(\D{\mathcal{T}}(u))$.
-Note that since $\D{B}$ assembles to a tridiagonal matrix the solution can be obtained very fast ($O(N_x)$) using one forward elimination followed by a backwards substitution (Thomas algorithm).
-The complete transformation thus requires a fast Chebyshev transform and a fast linear algebraic solve.
-Similar transforms $\mathcal{T}$ and $\N{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and (\ref{eq:phiN}), using mass matrices $B_{kj}=(T_j, T_k)_w$ and $\N{B}_{kj}=(\N{\phi}_j, \N{\phi}_j)_w$ and scalar products $\mathcal{S}_k(\cdot) = (\cdot, T_k)_w$ and $\N{\mathcal{S}}_k(\cdot) = (\cdot, \N{\phi}_k)_w$.
+where $\N{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = 
+\N{\mathcal{T}}^{-1}(\N{\mathcal{T}}(u))$. Note that since $\N{B}$ assembles to 
+a pentadiagonal matrix the solution ($\N{B}^{-1}$) can be obtained very fast 
+and 
+the complete 
+transformation thus requires a fast Chebyshev transform ($O(N_x \log N_x)$) and 
+a fast linear algebraic solve ($O(N_x)$). Similar transforms $\mathcal{T}$ 
+and $\D{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and 
+(\ref{eq:phiD}), using mass matrices $B_{lq}=(T_q, T_l)_w$ (diagonal) and 
+$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_w$ (tridiagonal) and scalar products 
+$\mathcal{S}_l(\cdot) = (\cdot, T_l)_w$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
+\D{\phi}_l)_w$.
 
-\section{The incremental pressure correction scheme}
+\section{Discretization of Navier Stokes equations}
+\label{sec:discretizationNS}
+The Navier Stokes equations (\ref{eq:NS}) are solved using a scheme 
+proposed by Kim, Moin and Moser (1987). This scheme is developed by taking the 
+Laplacian of the wall normal momentum equation and the curl of the momentum 
+equation. Following elimination of the pressure, the equations to solve are 
+\begin{align}
+\frac{\partial}{\partial t} \nabla^2 u &= h_u + \nu \nabla^4 u, 
+\label{eq:biharmonic} \\
+\frac{\partial g}{\partial t} &= h_g + \nu \nabla^2 g, \label{eq:g} \\
+f + \frac{\partial u}{\partial x} &= 0, \label{eq:f}
+\end{align}
+where
+\begin{align}
+f &= \frac{\partial v}{\partial y} + \frac{\partial w}{\partial z}, \\
+g &= \frac{\partial w}{\partial y} - \frac{\partial v}{\partial z}, \\
+h_u = -\frac{\partial}{\partial x} &\left( \frac{\partial 
+\mathcal{H}_y}{\partial y} + \frac{\partial \mathcal{H}_z}{\partial z} \right) 
++ \left(\frac{\partial^2}{\partial y^2} + \frac{\partial^2}{\partial z^2} 
+\right) \mathcal{H}_x ,
+\\
+h_g &= \frac{\partial \mathcal{H}_z}{\partial y} - \frac{\partial 
+\mathcal{H}_y}{\partial z}.
+\end{align}
+The two remaining velocity components are computed from the definitions of $f$ 
+and $g$. The biharmonic equation (\ref{eq:biharmonic}) is solved with boundary 
+conditions are $u(\pm 1) = u'(\pm 1) = 0$, where the Neumann condition follows 
+from the continuity equation. The boundary conditions for Eq. (\ref{eq:g}) are 
+$g(\pm 1) = 0$. The variational formulation becomes: Find ${u} \in 
+\N{V}_N$, ${g} \in \D{V}_N$ and ${f} \in \D{V}_N$ such that
+\begin{align}
+	\frac{\partial }{\partial t} \left< \nabla^2 u, \N{\psi}\right>_w &= 
+	\left<h_u, \N{\psi} \right>_w + \nu \left<\nabla^4u, \N{\psi}\right>_w 
+	&\forall \N{\psi} \in \N{V}_N, \label{eq:u1} \\
+	\frac{\partial}{\partial t}\left<g, \D{\psi}\right>_w &= \left<h_g, 
+	\D{\psi}\right>_w + \nu 
+	\left<\nabla^2 g, \D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N ,
+	\label{eq:g1} \\
+	\left<f, \D{\psi}\right>_w &= \left<\frac{\partial u}{\partial x}, 
+	\D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N. \label{eq:f1}
+\end{align}
+The two remaining velocity components are computed by projection to the 
+Dirichlet space $\D{V}_N$: Find 
+${v}, {w}$ in $\D{V}_N \times \D{V}_N$ such that
+\begin{align}
+\left<f, \D{\psi}\right>_w &= \left<\frac{\partial v}{\partial y} + 
+\frac{\partial w}{\partial z}, \D{\psi}\right>_w, \, &\forall \D{\psi} \in 
+\D{V}_N \label{eq:f2} \\
+\left<g, \D{\psi}\right>_w &= \left<\frac{\partial w}{\partial y}  - 
+\frac{\partial v}{\partial z}, \D{\psi}\right>_w, \, &\forall \D{\psi} \in 
+\D{V}_N, \label{eq:g2}
+\end{align}
+which simplifies considerably because all the derivatives are in periodic 
+directions. Written in spectral space Eqs. (\ref{eq:f2}) and (\ref{eq:g2}) 
+become simply
+\begin{align}
+\hat{f}_{\bm{k}} &= \imath \underline{m}\, \hat{v}_{\bm{k}} + \imath 
+\underline{n}\, \hat{w}_{\bm{k}} &\forall \bm{k} \in \D{K}_N^0, \label{eq:f3} \\
+\hat{g}_{\bm{k}} &= \imath \underline{m}\, \hat{w}_{\bm{k}} - \imath 
+\underline{n}\, \hat{v}_{\bm{k}} & \forall \bm{k} \in \D{K}_N^0, \label{eq:g3}
+\end{align}
+where $\D{K}_N^0$ is used to denote that these equations can be solved for all 
+wavenumbers except $m=n=0$. For $m=n=0$ we solve instead the 
+corresponding momentum equations in $y$ and $z$ directions:
+\begin{align}
+\frac{\partial }{\partial t} \D{B}_{lq}\hat{v}(q, 0, 0, t) &= 
+\D{B}_{lq}\hat{\mathcal{H}}_y(q, 0, 0, t) + \nu \D{A}_{lq} \hat{v}(q, 0, 0, t) 
+& \forall\, l \in [0, 1, 
+\ldots, N_x-2], \label{eq:v0}\\
+\frac{\partial }{\partial t} \D{B}_{lq}\hat{w}(q, 0, 0, t) &= 
+\D{B}_{lq}\hat{\mathcal{H}}_z(q, 0, 0, t) + \nu \D{A}_{lq} \hat{w}(q, 0, 0, t) 
+& \forall\, l \in [0, 1, \ldots, N_x-2].\label{eq:w0}
+\end{align}
 
-The Navier Stokes equations are discretized in time using a second order incremental pressure correction scheme with Crank-Nicolson on diffusion and explicit Adams Bashforth on the nonlinear convection.
-The incremental pressure correction scheme (IPCS) that is used consists of a tentative momentum equation followed by a pressure correction.
-The two steps may be performed iteratively.
-The equations are iterated forward in time in equal size time steps $\triangle t$ from initial conditions $u(\bm{x}, 0) = u^0(\bm{x})$ at $t=0$, such that $t_n = n\triangle t$, $n \in \mathcal{Z}$.
-For turbulent channel flows initial conditions are simply a perturbed state used to get a turbulent flow going.
+Equations (\ref{eq:u1}-\ref{eq:f1}) are also solved in spectral space. 
+Inserting for (\ref{eq:u_solx}) and a similar expansion for $g$ in $\D{V}_N$, 
+the required inner products become
+\begin{align}
+\left<\nabla^4u, \N{\psi}_{\bm{k}}\right>_w &= h\left[ \left( 
+\N{\phi}_q^{''''}, 
+\N{\phi}_l\right)_w -2(\underline{m}^2+\underline{n}^2) \left( \N{\phi}_q^{''}, 
+\N{\phi}_l\right)_w + (\underline{m}^2+\underline{n}^2)^2\left( \N{\phi}_q, 
+\N{\phi}_l\right)_w  \right] \hat{u}_q, \\
+\left< \nabla^2 u, \N{\psi}_{\bm{k}}\right>_w &= h\left[\left( \N{\phi}_q^{''}, 
+\N{\phi}_l\right)_w - (\underline{m}^2+\underline{n}^2)\left( \N{\phi}_q, 
+\N{\phi}_l \right)_w \right] \hat{u}_q, \\
+\left< \nabla^2 g, \D{\psi}_{\bm{k}}\right>_w &= h\left[\left( \D{\phi}_q^{''}, 
+\D{\phi}_l\right)_w - (\underline{m}^2+\underline{n}^2)\left( \D{\phi}_q, 
+\D{\phi}_l \right)_w \right] \hat{g}_q, \\
+\left<\frac{\partial u}{\partial x}, \D{\psi}_{\bm{k}}\right>_w &=
+h\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q, \\
+\left<h_u, \N{\psi} \right>_w &= h\left[-(\underline{m}^2+\underline{n}^2) 
+\N{\mathcal{S}}(\mathcal{H}_x) - \imath \underline{m}\left(\D{\phi}_q, 
+\N{\phi}_l \right)_w \hat{\mathcal{H}}_y - \imath 
+\underline{n}\left(\D{\phi}_q, \N{\phi}_l \right)_w \hat{\mathcal{H}}_z\right], 
+\\
+\left< h_g, \D{\psi} \right>_w &= h\left[ \imath \underline{m}\, \D{B} 
+\hat{\mathcal{H}}_z - \imath \underline{n}\, \D{B} \hat{\mathcal{H}}_y 
+\right]
+\end{align}
+where for brevity in notation (as before) it is understood that the matrices 
+act along the first dimension of the transformed variables, i.e.,  $ 
+\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q$ is short for the matrix 
+vector product $ \sum_{q=0}^{N_x-4}\left( \N{\phi}_q^{'}, \D{\phi}_l\right)_w 
+\hat{u}(q, {m}, {n}, t)$. The inner products are used to set up linear systems 
+of equations for the inhomogeneous wall normal direction. All inner products 
+$\left(\cdot, \cdot\right)_w$ can be represented by sparse matrices.
+%\begin{align}
+%\frac{\partial }{\partial t} \left( \N{\psi}_j^{''}, \N{\psi}_k\right)_w 
+%\hat{u}_j &=  \left(h_u, \N{\psi}_k \right)_w + \nu \left(\N{\psi}^{''''}_j, 
+%\N{\psi}_k\right)_w \hat{u}_j, \, &\forall \N{\psi}_k \in \N{V}_N, \\
+%\frac{\partial}{\partial t}\left(\D{\psi}_j, \D{\psi}_k\right)_w \hat{g}_j &= 
+%\left(h_g, \D{\psi}_k\right)_w + \nu \left(\D{\psi}_j^{''}, 
+%\D{\psi}_k\right)_w 
+%\hat{g}_j, \, &\forall \D{\psi}_k \in \D{V}_N \\
+%\left(\D{\psi}_j, \D{\psi}_k\right)_w \hat{f}_j &= \left(\N{\psi}_j^{'}, 
+%\D{\psi}_k\right)_w \hat{u}_j, \, &\forall \D{\psi}_k \in \D{V}_N.
+%\end{align}
+The mass matrices, $\N{A}_{lq} = \left( \N{\phi}_q^{''}, \N{\phi}_l\right)_w 
+$ and $\N{L}_{lq} = \left(\N{\phi}^{''''}_q, \N{\phi}_l\right)_w$ are given by 
+Shen, whereas $\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w$ is given in the 
+appendix.
 
-The IPCS attempts to find velocity vector $\bm{u}^{n}$ on time step $t_n$  and pressure midway between $t_n$ and $t_{n-1}$, $p^{n-1/2}$, given the solutions on all previous time steps.
-The three steps that are required solved on each time step are
+It remains to discretize the equations in time. Kim, Moin and Mansour use 
+Crank-Nicolson for the diffusive terms and Adams-Bashforth for the nonlinear 
+convective terms, whereas a third order Runge-Kutta method is used for the 
+nonlinear term by Moser, Kim and Mansour (1999). We have implemented both 
+approaches, but for simplicity only the first one is described here. 
+Discretization in time is performed using a constant time step $\triangle t$, 
+such that the time is represented discretely as $t_n = n \triangle t$. After 
+discretizing in time, equations (\ref{eq:u1}) and (\ref{eq:g1}) are transformed 
+to a matrix form that can be solved for $\hat{u}^{n+1} (= 
+\hat{u}_{\bm{k}}(t_{n+1}))$ and $\hat{g}^{n+1}$
 \begin{align}
-i)&\,\, \frac{\bm{u}^{*}-\bm{u}^{n-1}}{\triangle t} + \bm{N}^{n-1/2}   = \nu \nabla^2 \bm{u}^{n-1/2} - \nabla{p}^{n-3/2} + \bm{f}, \quad \bm{u}^*=0 \text{ for } x=\pm1   \label{eq:NSCN} \\
-ii)&\,\,\begin{cases}
-  \frac{\bm{u}^{n}-\bm{u}^*}{\triangle t} &= -\nabla p^* \\
-  \nabla \cdot \bm{u}^{n} &= 0, \quad \bm{u}^n\cdot \bm{n} = 0,\text{ for } x=\pm1
-  \end{cases} \label{eq:momupdate}\\
-  iii)&\,\,p^{n-1/2}=p^{n-3/2}+p^* \label{eq:pupdate} 
+\N{H}\hat{u}^{n+1} & = \left(2\N{A} - 2\underline{z}^2\N{B} - \N{H} 
+\right)\hat{u}^{n} + \N{\mathcal{S}}(h_v^{n+1/2}) \triangle t, 
+\label{eq:ufin}\\ 
+\D{H} \hat{g}^{n+1} &= \left(2 \D{B}-\D{H}\right) 
+\hat{g}^{n} + \D{\mathcal{S}}(h_g^{n+1/2}) \triangle t. \label{eq:gfin}
 \end{align}
-Here $\bm{u}^{n-1/2} = 0.5(\bm{u}^*+\bm{u}^{n-1})$, $\bm{u}^*$ is a tentative velocity, $p^*$ is a pressure correction, and the explicit convection is computed as $\bm{N}^{n-1/2}=1.5(\bm{u}^{n-1}\cdot \nabla) \bm{u}^{n-1} - 0.5(\bm{u}^{n-2}\cdot \nabla) \bm{u}^{n-2}$.
-The second step is computed by solving a Poisson equation for the pressure correction
-\begin{equation}
-  \nabla^2p^* = \frac{\nabla \cdot \bm{u}^*}{\triangle t}
-\end{equation}
-and subsequently updating the velocity through (\ref{eq:momupdate}).
+Here $\N{H} = -\frac{\nu \triangle t}{2}\N{L} + \left( 1 + \nu \triangle t 
+\underline{z}^2 \right) \N{A} - \frac{2\underline{z}^2 + \nu \triangle t 
+\underline{z}^4}{2} \N{B}$, $\D{H} = -\frac{\nu \triangle t}{2}\D{A} + (1 + 
+\frac{\nu \triangle t \underline{z}^2}{2}) \D{B}$ and $\underline{z}^2 = 
+\underline{m}^2 + \underline{n}^2$. The convective terms are computed with 
+a forward Adams-Bashforth projection, such that $\N{\mathcal{S}}(h_v^{n+1/2}) = 
+\N{\mathcal{S}}(1.5h_v^{n} - 0.5 h_v^{n-1})$ and $\D{\mathcal{S}}(h_g^{n+1/2}) 
+= \D{\mathcal{S}}(1.5h_g^{n} - 0.5 h_g^{n-1})$.
 
-All three velocity components use the Dirichlet basis (\ref{eq:u_sol}), whereas the pressure is searched for in $\N{V}_N$
-\begin{equation}
-p(\bm{x}, t) = \frac{1}{N_yN_z}\sum_{\bm{k} \in \N{K}_N} \hat{p}_{\bm{k}}(t) \N{\psi}_{\bm{k}}(\bm{x}), \label{eq:p_sol}
-\end{equation}
-
-The spectral Galerkin method becomes:
-
-1) Find $\bm{u}^*$ in $\D{V}_N^3$ by solving
-\begin{multline}
- \frac{1}{\triangle t}\left<\bm{u}^*-\bm{u}^{n-1}, \bm{\D{\psi}}\right>_w + \left<\bm{N}^{n-1/2}, \bm{\D{\psi}} \right>_w   = \nu \left<\nabla^2 \bm{u}^{n-1/2}, \bm{\D{\psi}} \right>_w \\ - \left<\nabla{p}^{n-3/2}, \bm{\D{\psi}} \right>_w + \left<\bm{f}, \bm{\D{\psi}} \right>_w \,\,  \forall \, \bm{\D{\psi}} \in \D{V}_N^3.
-\label{eq:tentative}
-\end{multline}
-
-2) Compute the pressure correction $p^*\in \N{V}_N$ by solving
-\begin{equation}
-\left< \nabla^2 p^*, \N{\psi} \right>_w = \frac{1}{\triangle t} \left<\nabla \cdot \bm{u}^*, \N{\psi} \right>_w\,\, \forall \, \N{\psi} \in \N{V}_N. \label{eq:pcorr}
-\end{equation}
-
-3) Update the velocity, $\bm{u}^n \in \D{V}_N^3$, and pressure, $p^{n-1/2} \in \N{V}_N$, by solving
+Equations (\ref{eq:v0}) and (\ref{eq:w0}) are also discretized in time using 
+Crank-Nicolson and Adams-Bashforth, and we obtain
 \begin{align}
-\left< \bm{u}^n, \bm{\D{\psi}}\right>_w &= \left< \bm{u}^*, \bm{\D{\psi}}\right>_w - \triangle t \left< \nabla p^*, \bm{\D{\psi}} \right>_w \,\,  &\forall \, \bm{\D{\psi}} \in \D{V}_N^3, \label{eq:u_update} \\
-\left<p^{n-1/2}, \N{\psi}\right>_w &= \left<p^{n-3/2}, \N{\psi}\right>_w + \left<p^*, \N{\psi}\right>_w \,\, &\forall \, \N{\psi} \in \N{V}_N,\label{eq:pres_update}
+\left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{v}_0^{n+1} &= 
+\left(\frac{\nu \triangle t}{2}\D{A} + \D{B} 
+\right)\hat{v}^{n}_0 + \D{B}\hat{\mathcal{H}}_{y,0}^{n+1/2}, \label{eq:v00} \\
+\left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{w}_0^{n+1} &= 
+\left(\frac{\nu \triangle t}{2}\D{A} + \D{B} 
+\right)\hat{w}^{n}_0 + \D{B}\hat{\mathcal{H}}_{z,0}^{n+1/2}, \label{eq:w00}
 \end{align}
-where the pressure update simply resorts to $\hat{p}^{n-1/2} = \hat{p}^{n-3/2} + \hat{p}^* \, \forall \, \bm{k}\in \N{K}_N$.
+where we have used notation $\hat{v}^{n}_0 = \hat{v}(l, 0, 0, t_n)$, 
+$\hat{\mathcal{H}}^{n}_{y,0} = \hat{\mathcal{H}}_y(l, 0, 0, t_n)$ and similar 
+for $\hat{w}^n_0$ and $\hat{\mathcal{H}}^n_{z,0}$.
 
 \section{Implementation}
-The vector valued equations (\ref{eq:tentative}) and (\ref{eq:u_update}) are solved in a segregated manner, one component at the time.
-We use $u_i$ and $\hat{u}_i$ for component $i$ of the real and transformed velocity vector.
-The inner products in Eqs.
-(\ref{eq:tentative}, \ref{eq:pcorr}, \ref{eq:u_update}, \ref{eq:pres_update}) lead to a range of new matrices
-\begin{align}
- \left< u_i^* - u_i^{n-1}, \D{\psi}\right>_w &= h\D{B}\left({\hat{u}}_i^*-{\hat{u}}_i^{n-1} \right) \quad &i \in (0,1,2), \\ 
- \left<\nabla^2 {u}_i^{n-1/2}, \D{\psi}\right>_w &= -h  \left( \D{A} +(\underline{m}^2+\underline{n}^2)\D{B}\right) \hat{u}_i^{n-1/2} \quad &i \in (0,1,2), \\ 
- \left< f_i, \D{\psi}\right>_w &= h\D{\mathcal{S}}(f_i) &i \in (0,1,2), \\
- \left<N_i^{n-1/2}, \D{\psi} \right>_w &= h\D{\mathcal{S}}(N_i^{n-1/2}) &i \in (0,1,2), \\  
- \left< \nabla p^{n-3/2}, \bm{\D{\psi}} \right>_w &= h\left[\grave{C},\, \imath \underline{m} \grave{B},\, \imath \underline{n} \grave{B} \right] \hat{p}^{n-3/2}, \\ 
- \left< \nabla^2 p^*, \N{\psi} \right>_w &= -h\left(\N{A} +(\underline{m}^2 + \underline{n}^2)\N{B} \right) \hat{p}^*, \\ 
- \left<\nabla \cdot \bm{u}^*, \N{\psi} \right>_w &= h\left( \tilde{C} \hat{u}^* + \imath \underline{m} \tilde{B} \hat{v}^* + \imath \underline{n} \tilde{B} \hat{w}^* \right), 
-\end{align}
-where $\D{A}_{kj} = -\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w, \N{A}_{kj} = -\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w,  \grave{C}_{kj} = \left(\N{\phi}^{'}_j, \D{\phi}_k \right)_w, \tilde{C}_{kj} = (\D{\phi}^{'}_j, \N{\phi}_k)_w, \N{B}_{kj} = \left( \N{\phi}_j, \N{\phi}_k \right)_w, \grave{B}_{kj} = (\N{\phi}_j, \D{\phi}_k)_w$ and $\tilde{B}_{kj} = (\D{\phi}_j, \N{\phi}_k)_w$ are all more or less sparse matrices.
-Arranging the governing equations on matrix form we obtain for the three tentative velocity components in (\ref{eq:tentative})
-\begin{align}
-\D{H} \hat{u}^* &= \left(\frac{4}{\nu \triangle t} \D{B}-\D{H}\right) \hat{u}^{n-1} - \left[\D{\mathcal{S}}(N_x^{n-1/2}) + \grave{C} \hat{p}^{n-3/2} + \D{\mathcal{S}}(f_x)\right]\frac{2}{\nu}, \label{eq:tentativeUA}\\
-\D{H} \hat{v}^* &=  \left(\frac{4}{\nu \triangle t} \D{B}-\D{H}\right)\hat{v}^{n-1} - \left[\D{\mathcal{S}}(N_y^{n-1/2}) + \imath \underline{m} \grave{B} \hat{p}^{n-3/2} + \D{\mathcal{S}}(f_y)\right]\frac{2}{\nu}, \label{eq:tentativeVA}\\
-\D{H} \hat{w}^* &= \left(\frac{4}{\nu \triangle t} \D{B}-\D{H}\right) \hat{w}^{n-1} - \left[\D{\mathcal{S}}(N_z^{n-1/2}) + \imath \underline{n} \grave{B} \hat{p}^{n-3/2} + \D{\mathcal{S}}(f_z)\right]\frac{2}{\nu},\label{eq:tentativeWA}
-\end{align}
-where $\D{H}=\D{A} +(\underline{m}^2 + \underline{n}^2 + \frac{2}{\nu \triangle t})\D{B}$ is a matrix consisting of one subdiagonal, and where every second diagonal is zero.
-As such, we may split the matrix into two smaller (and decoupled) matrices for odd and even coefficients $H^e_{k,j} = H_{2k,2j}$ and $H^o_{k,j} = H_{2k+1,2j+1}$, where the two matrices $H^{e,o}$ are of type upper Hessenberg, with only 4 distinct values on each row.
-This structure of $H$ allows us to perform a very efficient tailored LU decomposition, as shown in Algorithm (\ref{alg:lu}), and a very efficient tailored solve through Algorithm (\ref{alg:lusolve}).
-Regarding Algorithm (\ref{alg:lusolve}), we note that the ability to decouple the system into odd and even coefficients leads to two subsystem that may be trivially solved simultaneously in two threads by letting one thread call LUODDEVEN with even coefficients, and the second thread call it for odd.
 
-The linear system arising for the pressure correction equation is similar to the tentative velocities
-\begin{equation}
-\N{H}\hat{p}^* = -\frac{1}{\triangle t}\left( \tilde{C} \hat{u}^* + \imath \underline{m} \tilde{B} \hat{v}^* + \imath \underline{n} \tilde{B} \hat{w}^* \right), \label{eq:pressureA}
-\end{equation}
-where $\N{H}=\N{A} +(\underline{m}^2 + \underline{n}^2)\N{B}$ is of the same structure as $\D{H}$ and as such  should be factorizable in the same way.
-However, entries of the matrix $\N{H}_{kj}$ contain the column index $j^2$ (see $\N{A}$), and as such the values on each row of the matrix $\N{H}_{kj}$ are not constant for $j \ge k+4$ and the LU decomposition of $\N{H}$ will require a full upper $U$ matrix.
-To circumvent this we define $\N{H}_{kj} = \underline{H}_{kl}J_{lj}$, where $J_{kj}= j^2\delta_{kj}$ (no summation implied).
-We then set $w_k=J_{kj}\hat{p}_j$, use $b_k$ for the right hand side of (\ref{eq:pressureA}) and solve
-\begin{align}
-  \underline{H}_{kj}w_j &= b_k &\forall \, k \in 1, 2, \ldots, N_x-2, \\
-  \hat{p}_k &= J_{kj}^{-1}w_j &\forall \, k \in 1, 2, \ldots, N_x-2.
-\end{align}
-Here the first system $\underline{H}_{kj} w_j = b_k$ can be solved using the same LU factorization as the $\D{H}$ matrix, since $\underline{H}_{kj}=\underline{H}_{k, k+4}\, \forall \, j=k+6, k+8, \ldots$ and there are only 4 distinct values on each row.
-The implementation is the same as for $\D{H}$, except that for the matrix $\underline{H}$ the indices start at 1 instead of zero.
-
-The velocity update is computed on matrix form as
-\begin{align}
-  \hat{u}^n &= \hat{u}^* - \triangle t \D{B}^{-1} \left( \grave{C} \hat{p}^* \right),\\
-  \hat{v}^n &= \hat{v}^* - \triangle t \D{B}^{-1} \left( \imath \underline{m} \grave{B} \hat{p}^* \right),\\ 
-  \hat{w}^n &= \hat{w}^* - \triangle t \D{B}^{-1} \left( \imath \underline{n} \grave{B} \hat{p}^* \right).
-\end{align}
-using a tridiagonal solver.
-
-The convection term is computed in standard form with a pseudo-spectral method that requires a fast scalar product
-\begin{equation}
-  \D{\mathcal{S}}(N_i^{n-1}) = \D{\mathcal{S}}(\bm{u}^{n-1} \cdot \nabla u_i^{n-1}) \quad \forall \, i \in (0, 1, 2).
-\end{equation}
-We obtain the gradients of $u_i$ in real space by projecting $\partial u/\partial x$ to $\D{V}_{N_x}$ (boundary condition $\partial u/\partial x(\pm 1)=0$ follows from continuity) and $\partial v/\partial x$ and $\partial w/\partial x$ to $V_{N_x}$.
-\begin{align}
-  \frac{\partial u}{\partial x} &= \D{\mathcal{T}}^{-1} (\D{B}^{-1}(\D{C}\hat{u})), \\
-  \frac{\partial v}{\partial x} &= \mathcal{T}^{-1} (B^{-1}(\acute{C}\hat{v})), \\
-  \frac{\partial w}{\partial x} &= \mathcal{T}^{-1} (B^{-1}(\acute{C}\hat{w})),  
-\end{align}
-where $\D{C}_{kj} = (\D{\phi}_j^{'}, \D{\phi}_k)_w$,  $\acute{C_{kj}} = (\D{\phi}_j^{'}, T_k)_w$ and $B=(T_j, T_k)_w$. Partial derivatives with respect to periodic directions are computed as
-\begin{align}
- \frac{\partial u_i}{\partial y} &= \D{\mathcal{T}}^{-1}(\imath \underline{m} \hat{u}_i) &\forall \, i \in (0, 1, 2), \\
- \frac{\partial u_i}{\partial w} &= \D{\mathcal{T}}^{-1}(\imath \underline{n} \hat{u}_i) &\forall \, i \in (0, 1, 2).
-\end{align}
-Dealiasing is implemented using the 2/3-rule.
-
-We finally note that the right hand sides of all the linear systems that need to solved can be assembled fast using tailored matrix vector products taking advantage of the fact that there are at most three distinct diagonals in any matrix.
-This concludes the implementation required to solve the Navier-Stokes equations by the incremental pressure correction method and the spectral Shen-Fourier Galerkin method.
-We note that all terms in assembly and solve may be performed using fast routines, at worst scaling like $O(N_x \log_2 N_x)$.
-
-\section{Serial scaling}
-
-\section{Parallel scaling}
+The solution procedure for the numerical method described in 
+Sec.~\ref{sec:discretizationNS} is given in Algorithm \ref{alg:NS}
 
 \begin{algorithm}
-\caption{LU factorization of matrix $H_{k,j}$, that may be split into two decoupled upper Hessenberg matrices $H^e_{k,j} = H_{2k,2j}$ and $H^o_{k,j} = H_{2k+1,2j+1}$ for even and odd coefficients.
-Return LU factorization for even and odd coefficients, such that $L^eU^e=H^e$ and $L^oU^o=H^o$.
-Note that $L^{e}$ and $L^{o}$ each have one single nonzero diagonal in addition to the unity main diagonal, and as such only a vector is stored.
-Each $U^{e}$ and $U^{o}$ has three distinct diagonals and $U^{e}_{kj}=U^{e}_{k, k+2} \,\forall j > k+2, j< M^e$ and $U^{o}_{kj}=U^{o}_{k, k+2} \,\forall j > k+2, j< M^o$.
-As such, a sparse three-diagonal storage may be used for $U^e$ and $U^o$.}
+	\caption{Solution procedure}
+    \label{alg:NS}
+    \begin{algorithmic}[1]
+    	\State \text{Initialize} $\bm{u}(\bm{x}, 0)$ and compute 
+    	$\hat{g}_{\bm{k}}(0)$
+		\State \text{Assemble} $\N{H}, \D{H}$
+		\State $t=0$
+    	\While{$t<T$}
+	    	\State $t \gets t+dt$
+	    	\State \text{Compute nonlinear convection} $\bm{\mathcal{H}}$
+		    \State \text{Solve Eq.}(\ref{eq:ufin}) for 
+		    $\hat{u}^{n+1}_{\bm{k}} \, \forall \,\bm{k} \in \N{K}_N$
+		    \State \text{Solve Eqs.}(\ref{eq:gfin}) for 
+		    $\hat{g}^{n+1}_{\bm{k}} \, \forall\, \bm{k} \in \D{K}_N$
+		    \State \text{Solve Eqs.}(\ref{eq:f3}, \ref{eq:g3}) for  
+		    $\hat{v}_{\bm{k}}, \hat{w}_{\bm{k}}\, \forall\, \bm{k} 
+		    \in \D{K}_N^0$
+		    \State \text{Solve Eqs.}(\ref{eq:v00}, \ref{eq:w00}) for 
+		    $\hat{v}^{n+1}(l, 0, 0, t_{n+1}), \hat{w}^{n+1}(l, 0, 0, t_{n+1})\, 
+		    \forall \, l \in [0, 1, \ldots, N_x-2] $
+		    \State \text{Update to new time step}
+	    \EndWhile
+
+    \end{algorithmic}
+\end{algorithm}
+
+The major computational cost is found in computing the nonlinear convection and 
+in solving for Eqs.~(\ref{eq:ufin}, \ref{eq:gfin}). The nonlinear convection 
+term $\bm{\hat{\mathcal{H}}}$ is computed using a pseudo-spectral method with 
+the 3/2-rule for de-aliasing in the periodic directions (ref Canuto). The 
+transforms are carried out using fast Fourier and fast Chebyshev transforms 
+that are all of order $O(N\log N)$ per one-dimensional transform of size 
+N. The Shen transforms make use of the Chebyshev transforms that are using 
+discrete cosine transforms. The forward scalar product is shown in Algorithm 
+\ref{alg:fst}.
+\begin{algorithm}
+	\caption{Forward scalar product for all spaces $V_N, \D{V}_N, \N{V}_N$. 
+	Here "dct" is the discrete cosine transform from SciPy.}
+	\label{alg:fst}
+	\begin{algorithmic}[1]
+		\State \textbf{Input:} $\{f\}_{j=0}^{N}$
+		\State \textbf{Input:} points
+		\State \textbf{Input:} space
+		\If{ points = "Gauss-Lobatto"}
+			\State $f_k \gets \text{dct}(f_j, type=2, axis=0)\pi/(2 N)\, 
+			\forall \, \, k \in 
+			[0, 1, \ldots, N]$
+		\Else{ points = "Gauss-Chebyshev"}
+		    \State $f_k \gets \text{dct}(f_j, type=1, axis=0)\pi/(2 (N-1))\, 
+		    \forall \, \, k 
+		    \in [0, 1, \ldots, N]$
+	    \EndIf
+	    \If{ space = $\D{V}_N$}
+	        \State $f_k \gets f_k - f_{k+2} \, \forall k \in [0, 1, \ldots, 
+	        N-2]$
+	    \ElsIf{ space = $\N{V}_N$}	                
+	        \State $f_k \gets f_k - 2(k+2)/(k+3)f_{k+2} + (k+1)/(k+3)f_{k+4} 
+	        \,\, 
+	        \forall \, k \in [0, 1, \ldots, N-4]$
+	    \Else{}
+	        \State pass
+	    \EndIf
+\State \Return $f_k$    
+	\end{algorithmic}
+\end{algorithm}
+
+We will now describe very efficient direct solvers for Eqs.~(\ref{eq:ufin}, 
+\ref{eq:gfin}) that are of order $O(N)$. The solvers are generated using clever 
+direct LU decomposition and only a few diagonals need to be stored for each 
+final lower (L) and upper (U) matrix.
+
+To solve Eq. (\ref{eq:gfin}) we perform an LU decomposition of the coefficient 
+matrix $\D{H}$. $\D{H}$ consists of one subdiagonal, and every 
+second upper diagonal is zero. As such, we may split the matrix into two 
+smaller (and decoupled) matrices for odd and even coefficients $\D{H}^e_{k,j} = 
+\D{H}_{2k,2j}$ and $\D{H}^o_{k,j} = \D{H}_{2k+1,2j+1}$, where the two matrices 
+$\D{H}^{e,o}$ are of type upper 
+Hessenberg, with only 4 distinct values on each row. This structure of $\D{H}$ 
+allows us to perform a very efficient tailored LU decomposition, as shown in 
+Algorithm (\ref{alg:lu}), and a very efficient tailored solve through 
+Algorithm (\ref{alg:lusolve}). The efficiency follows from the fact that all 
+$U^{o,e}_{kj} = U^{o,e}_{kk+2} \, \forall \, j > k+2$, i.e., the upper 
+tridiagonal matrix has a maximum of three distinct values for each row. 
+Regarding Algorithm 
+(\ref{alg:lusolve}), we note that the ability to decouple the system into odd 
+and even coefficients leads to two subsystem that may be trivially solved 
+simultaneously in two threads by 
+letting one thread call LUODDEVEN with even coefficients, and the second thread 
+call it for odd.
+
+\begin{algorithm}
+\caption{LU factorization of matrix $H_{k,j}$, that may be split into two 
+decoupled upper Hessenberg matrices $H^e_{k,j} = H_{2k,2j}$ and $H^o_{k,j} = 
+H_{2k+1,2j+1}$ for even and odd coefficients. Return LU factorization for even 
+and odd coefficients, such that $L^eU^e=H^e$ and $L^oU^o=H^o$. Note that 
+$L^{e}$ and $L^{o}$ each have one single nonzero diagonal in addition to the 
+unity main diagonal, and as such only a vector is stored. Each $U^{e}$ and 
+$U^{o}$ has three distinct diagonals and $U^{e}_{kj}=U^{e}_{k, k+2} \,\forall 
+j 
+> k+2, j< M^e$ and $U^{o}_{kj}=U^{o}_{k, k+2} \,\forall j > k+2, j< M^o$. As 
+such, a sparse three-diagonal storage may be used for $U^e$ and $U^o$.}
 \label{alg:lu}
@@
     \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
-    \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall j=k, k+1, k+2$
+    \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall 
+j=k, k+1, k+2$
     \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
-    \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, \forall j=k, k+1, k+2$
+    \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, 
+\forall j=k, k+1, k+2$
   \EndFor
@@
   \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
-  \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall j=k, k+1$
+  \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall 
+j=k, k+1$
   \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
-  \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, \forall j=k, k+1$
+  \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, 
+\forall j=k, k+1$
   \State $k \gets M^e$
@@
   \EndIf  
-  \State \Return $\{L^e_k\}_{k=0}^{M^e},\{U^e\}_{k,j=0}^{M^e}, \{L^o_k\}_{k=0}^{M^o},\{U^o\}_{k,j=0}^{M^o}$
+  \State \Return $\{L^e_k\}_{k=0}^{M^e},\{U^e\}_{k,j=0}^{M^e}, 
+\{L^o_k\}_{k=0}^{M^o},\{U^o\}_{k,j=0}^{M^o}$
 \EndProcedure
@@
 \begin{algorithm}
-\caption{Solve of $H_{kj}u_j = b_k$, where $H$ has previously been LU factorized for odd even coefficients using Algorithm (\ref{alg:lu}).}
+\caption{Solve of $H_{kj}u_j = b_k$, where $H$ has previously been LU 
+factorized for odd even coefficients using Algorithm (\ref{alg:lu}).}
 \label{alg:lusolve}
 \begin{algorithmic}[1]
-\Procedure{LUoddeven}{}
-  \State \textbf{Input:} $\{U_{kj}\}_{k,j=0}^{M}, \{L_k\}_{k=0}^{M}, \{b_k\}_{k=0}^{M}$ 
-  \State // Solve $Ly=b$  
+\Procedure{LUSolveoddeven}{}
+  \State \textbf{Input:} $\{U_{kj}\}_{k,j=0}^{M}, \{L_k\}_{k=0}^{M}, 
+\{b_k\}_{k=0}^{M}$ 
+  \State // Solve $Ly=b$ by forward elimination 
   \State $y_0 \gets b_0$
@@
   \EndFor
-  \State // Solve $Uu=y$
+  \State // Solve $Uu=y$ with back substitution
   \State double $s = 0.0$     
@@
 \Procedure{LUSolve}{}
-  \State \textbf{Input: $\{u_k\}_{k=0}^{N-2}, \{b_k\}_{k=0}^{N-2}, \{U_{kj}^{e}\}_{k,j=0}^{M^{e}}, \{U_{kj}^{o}\}_{k,j=0}^{M^{o}}, \{L^{e}_k\}_{k=0}^{M^{e}}, \{L^{o}_k\}_{k=0}^{M^{o}}$}
-  \State $\{u_{2k}\}_{k=0}^{M^e} \gets $ LUoddeven($U^{e}, L^e, \{b_{2k}\}_{k=0}^{M^e}$)
-  \State $\{u_{2k+1}\}_{k=0}^{M^o} \gets $ LUoddeven($U^{o}, L^o, \{b_{2k+1}\}_{k=0}^{M^o}$)
+  \State \textbf{Input: $\{u_k\}_{k=0}^{N-2}, \{b_k\}_{k=0}^{N-2}, 
+\{U_{kj}^{e}\}_{k,j=0}^{M^{e}}, \{U_{kj}^{o}\}_{k,j=0}^{M^{o}}, 
+\{L^{e}_k\}_{k=0}^{M^{e}}, \{L^{o}_k\}_{k=0}^{M^{o}}$}
+  \State $\{u_{2k}\}_{k=0}^{M^e} \gets $ LUSolveoddeven($U^{e}, L^e, 
+\{b_{2k}\}_{k=0}^{M^e}$)
+  \State $\{u_{2k+1}\}_{k=0}^{M^o} \gets $ LUSolveoddeven($U^{o}, L^o, 
+\{b_{2k+1}\}_{k=0}^{M^o}$)
 
@@
 \end{algorithm}
+
+%\section{The incremental pressure correction scheme}
+%
+%The Navier Stokes equations are discretized in time using a second order 
+%incremental pressure correction scheme with Crank-Nicolson on diffusion and 
+%explicit Adams Bashforth on the nonlinear convection. The incremental pressure 
+%correction scheme (IPCS) that is used consists of a tentative momentum 
+%equation 
+%followed by a pressure correction. The two steps may be performed iteratively. 
+%The equations are iterated forward in time in equal size time steps $\triangle 
+%t$ from initial conditions $u(\bm{x}, 0) = u^0(\bm{x})$ at $t=0$, such that 
+%$t_n = n\triangle t$, $n \in \mathcal{Z}$. For turbulent channel flows initial 
+%conditions are simply a perturbed state used to get a turbulent flow going. 
+%
+%The IPCS attempts to find velocity vector $\bm{u}^{n}$ on time step $t_n$  and 
+%pressure midway between $t_n$ and $t_{n-1}$, $p^{n-1/2}$, given the solutions 
+%on all previous time steps. The three steps that are required solved on each 
+%time step are
+%\begin{align}
+%i)&\,\, \frac{\bm{u}^{*}-\bm{u}^{n-1}}{\triangle t} + \bm{N}^{n-1/2}   = \nu 
+%\nabla^2 \bm{u}^{n-1/2} - \nabla{p}^{n-3/2} + \bm{f}, \quad \bm{u}^*=0 \text{ 
+%for } x=\pm1   \label{eq:NSCN} \\
+%ii)&\,\,\begin{cases}
+%  \frac{\bm{u}^{n}-\bm{u}^*}{\triangle t} &= -\nabla p^* \\
+%  \nabla \cdot \bm{u}^{n} &= 0, \quad \bm{u}^n\cdot \bm{n} = 0,\text{ for } 
+%x=\pm1
+%  \end{cases} \label{eq:momupdate}\\
+%  iii)&\,\,p^{n-1/2}=p^{n-3/2}+p^* \label{eq:pupdate} 
+%\end{align}
+%Here $\bm{u}^{n-1/2} = 0.5(\bm{u}^*+\bm{u}^{n-1})$, $\bm{u}^*$ is a tentative 
+%velocity, $p^*$ is a pressure correction, and the explicit convection is 
+%computed as $\bm{N}^{n-1/2}=1.5(\bm{u}^{n-1}\cdot \nabla) \bm{u}^{n-1} - 
+%0.5(\bm{u}^{n-2}\cdot \nabla) \bm{u}^{n-2}$. The second step is computed by 
+%solving a Poisson equation for the pressure correction
+%\begin{equation}
+%  \nabla^2p^* = \frac{\nabla \cdot \bm{u}^*}{\triangle t}
+%\end{equation}
+%and subsequently updating the velocity through (\ref{eq:momupdate}).
+%
+%All three velocity components use the Dirichlet basis (\ref{eq:u_sol}), 
+%whereas the pressure is searched for in $\N{V}_N$
+%\begin{equation}
+%p(\bm{x}, t) = \frac{1}{N_yN_z}\sum_{\bm{k} \in \N{K}_N} \hat{p}_{\bm{k}}(t) 
+%\N{\psi}_{\bm{k}}(\bm{x}), \label{eq:p_sol}
+%\end{equation}
+% 
+%The spectral Galerkin method becomes: 
+%
+%1) Find $\bm{u}^*$ in $\D{V}_N^3$ by solving
+%\begin{multline}
+% \frac{1}{\triangle t}\left<\bm{u}^*-\bm{u}^{n-1}, \bm{\D{\psi}}\right>_w + 
+%\left<\bm{N}^{n-1/2}, \bm{\D{\psi}} \right>_w   = \nu \left<\nabla^2 
+%\bm{u}^{n-1/2}, \bm{\D{\psi}} \right>_w \\ - \left<\nabla{p}^{n-3/2}, 
+%\bm{\D{\psi}} \right>_w + \left<\bm{f}, \bm{\D{\psi}} \right>_w \,\,  \forall 
+%\, \bm{\D{\psi}} \in \D{V}_N^3. \label{eq:tentative}
+%\end{multline}
+%
+%2) Compute the pressure correction $p^*\in \N{V}_N$ by solving
+%\begin{equation}
+%\left< \nabla^2 p^*, \N{\psi} \right>_w = \frac{1}{\triangle t} \left<\nabla 
+%\cdot \bm{u}^*, \N{\psi} \right>_w\,\, \forall \, \N{\psi} \in \N{V}_N. 
+%\label{eq:pcorr}
+%\end{equation}
+%
+%3) Update the velocity, $\bm{u}^n \in \D{V}_N^3$, and pressure, $p^{n-1/2} \in 
+%\N{V}_N$, by solving
+%\begin{align}
+%\left< \bm{u}^n, \bm{\D{\psi}}\right>_w &= \left< \bm{u}^*, 
+%\bm{\D{\psi}}\right>_w - \triangle t \left< \nabla p^*, \bm{\D{\psi}} 
+%\right>_w 
+%\,\,  &\forall \, \bm{\D{\psi}} \in \D{V}_N^3, \label{eq:u_update} \\
+%\left<p^{n-1/2}, \N{\psi}\right>_w &= \left<p^{n-3/2}, \N{\psi}\right>_w + 
+%\left<p^*, \N{\psi}\right>_w \,\, &\forall \, \N{\psi} \in 
+%\N{V}_N,\label{eq:pres_update}
+%\end{align}
+%where the pressure update simply resorts to $\hat{p}^{n-1/2} = \hat{p}^{n-3/2} 
+%+ \hat{p}^* \, \forall \, \bm{k}\in \N{K}_N$.
+%
+%\section{Implementation}
+%The vector valued equations (\ref{eq:tentative}) and (\ref{eq:u_update}) are 
+%solved in a segregated manner, one component at the time. We use $u_i$ and 
+%$\hat{u}_i$ for component $i$ of the real and transformed velocity vector. The 
+%inner products in Eqs. (\ref{eq:tentative}, \ref{eq:pcorr}, \ref{eq:u_update}, 
+%\ref{eq:pres_update}) lead to a range of new matrices
+%\begin{align}
+% \left< u_i^* - u_i^{n-1}, \D{\psi}\right>_w &= 
+%h\D{B}\left({\hat{u}}_i^*-{\hat{u}}_i^{n-1} \right) \quad &i \in (0,1,2), \\ 
+% \left<\nabla^2 {u}_i^{n-1/2}, \D{\psi}\right>_w &= -h  \left( \D{A} 
+%+(\underline{m}^2+\underline{n}^2)\D{B}\right) \hat{u}_i^{n-1/2} \quad &i \in 
+%(0,1,2), \\ 
+% \left< f_i, \D{\psi}\right>_w &= h\D{\mathcal{S}}(f_i) &i \in (0,1,2), \\
+% \left<N_i^{n-1/2}, \D{\psi} \right>_w &= h\D{\mathcal{S}}(N_i^{n-1/2}) &i \in 
+%(0,1,2), \\  
+% \left< \nabla p^{n-3/2}, \bm{\D{\psi}} \right>_w &= h\left[\grave{C},\, 
+%\imath \underline{m} \grave{B},\, \imath \underline{n} \grave{B} \right] 
+%\hat{p}^{n-3/2}, \\ 
+% \left< \nabla^2 p^*, \N{\psi} \right>_w &= -h\left(\N{A} +(\underline{m}^2 + 
+%\underline{n}^2)\N{B} \right) \hat{p}^*, \\ 
+% \left<\nabla \cdot \bm{u}^*, \N{\psi} \right>_w &= h\left( \tilde{C} 
+%\hat{u}^* + \imath \underline{m} \tilde{B} \hat{v}^* + \imath \underline{n} 
+%\tilde{B} \hat{w}^* \right), 
+%\end{align}
+%where $\D{A}_{kj} = -\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w, 
+%\N{A}_{kj} = -\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w,  \grave{C}_{kj} 
+%= \left(\N{\phi}^{'}_j, \D{\phi}_k \right)_w, \tilde{C}_{kj} = 
+%(\D{\phi}^{'}_j, 
+%\N{\phi}_k)_w, \N{B}_{kj} = \left( \N{\phi}_j, \N{\phi}_k \right)_w, 
+%\grave{B}_{kj} = (\N{\phi}_j, \D{\phi}_k)_w$ and $\tilde{B}_{kj} = 
+%(\D{\phi}_j, 
+%\N{\phi}_k)_w$ are all more or less sparse matrices. Arranging the governing 
+%equations on matrix form we obtain for the three tentative velocity components 
+%in (\ref{eq:tentative})
+%\begin{align}
+%\D{H} \hat{u}^* &= \left(\frac{4}{\nu \triangle t} \D{B}-\D{H}\right) 
+%\hat{u}^{n-1} - \left[\D{\mathcal{S}}(N_x^{n-1/2}) + \grave{C} \hat{p}^{n-3/2} 
+%+ \D{\mathcal{S}}(f_x)\right]\frac{2}{\nu}, \label{eq:tentativeUA}\\
+%\D{H} \hat{v}^* &=  \left(\frac{4}{\nu \triangle t} 
+%\D{B}-\D{H}\right)\hat{v}^{n-1} - \left[\D{\mathcal{S}}(N_y^{n-1/2}) + \imath 
+%\underline{m} \grave{B} \hat{p}^{n-3/2} + 
+%\D{\mathcal{S}}(f_y)\right]\frac{2}{\nu}, \label{eq:tentativeVA}\\
+%\D{H} \hat{w}^* &= \left(\frac{4}{\nu \triangle t} \D{B}-\D{H}\right) 
+%\hat{w}^{n-1} - \left[\D{\mathcal{S}}(N_z^{n-1/2}) + \imath \underline{n} 
+%\grave{B} \hat{p}^{n-3/2} + 
+%\D{\mathcal{S}}(f_z)\right]\frac{2}{\nu},\label{eq:tentativeWA}
+%\end{align}
+%where $\D{H}=\D{A} +(\underline{m}^2 + \underline{n}^2 + \frac{2}{\nu 
+%\triangle t})\D{B}$ is a matrix consisting of one subdiagonal, and where every 
+%second diagonal is zero. As such, we may split the matrix into two smaller 
+%(and 
+%decoupled) matrices for odd and even coefficients $H^e_{k,j} = H_{2k,2j}$ and 
+%$H^o_{k,j} = H_{2k+1,2j+1}$, where the two matrices $H^{e,o}$ are of type 
+%upper 
+%Hessenberg, with only 4 distinct values on each row. This structure of $H$ 
+%allows us to perform a very efficient tailored LU decomposition, as shown in 
+%Algorithm (\ref{alg:lu}), and a very efficient tailored solve through 
+%Algorithm 
+%(\ref{alg:lusolve}). Regarding Algorithm (\ref{alg:lusolve}), we note that the 
+%ability to decouple the system into odd and even coefficients leads to two 
+%subsystem that may be trivially solved simultaneously in two threads by 
+%letting 
+%one thread call LUODDEVEN with even coefficients, and the second thread call 
+%it 
+%for odd. 
+%
+%The linear system arising for the pressure correction equation is similar to 
+%the tentative velocities
+%\begin{equation}
+%\N{H}\hat{p}^* = -\frac{1}{\triangle t}\left( \tilde{C} \hat{u}^* + \imath 
+%\underline{m} \tilde{B} \hat{v}^* + \imath \underline{n} \tilde{B} \hat{w}^* 
+%\right), \label{eq:pressureA}
+%\end{equation}
+%where $\N{H}=\N{A} +(\underline{m}^2 + \underline{n}^2)\N{B}$ is of the same 
+%structure as $\D{H}$ and as such  should be factorizable in the same way. 
+%However, entries of the matrix $\N{H}_{kj}$ contain the column index $j^2$ 
+%(see 
+%$\N{A}$), and as such the values on each row of the matrix $\N{H}_{kj}$ are 
+%not 
+%constant for $j \ge k+4$ and the LU decomposition of $\N{H}$ will require a 
+%full upper $U$ matrix. To circumvent this we define $\N{H}_{kj} = 
+%\underline{H}_{kl}J_{lj}$, where $J_{kj}= j^2\delta_{kj}$ (no summation 
+%implied). We then set $w_k=J_{kj}\hat{p}_j$, use $b_k$ for the right hand side 
+%of (\ref{eq:pressureA}) and solve
+%\begin{align}
+%  \underline{H}_{kj}w_j &= b_k &\forall \, k \in 1, 2, \ldots, N_x-2, \\
+%  \hat{p}_k &= J_{kj}^{-1}w_j &\forall \, k \in 1, 2, \ldots, N_x-2.
+%\end{align}
+%Here the first system $\underline{H}_{kj} w_j = b_k$ can be solved using the 
+%same LU factorization as the $\D{H}$ matrix, since 
+%$\underline{H}_{kj}=\underline{H}_{k, k+4}\, \forall \, j=k+6, k+8, \ldots$ 
+%and 
+%there are only 4 distinct values on each row. The implementation is the same 
+%as 
+%for $\D{H}$, except that for the matrix $\underline{H}$ the indices start at 1 
+%instead of zero.
+%
+%The velocity update is computed on matrix form as
+%\begin{align}
+%  \hat{u}^n &= \hat{u}^* - \triangle t \D{B}^{-1} \left( \grave{C} \hat{p}^* 
+%\right),\\
+%  \hat{v}^n &= \hat{v}^* - \triangle t \D{B}^{-1} \left( \imath \underline{m} 
+%\grave{B} \hat{p}^* \right),\\ 
+%  \hat{w}^n &= \hat{w}^* - \triangle t \D{B}^{-1} \left( \imath \underline{n} 
+%\grave{B} \hat{p}^* \right).
+%\end{align}
+%using a tridiagonal solver.
+%
+%The convection term is computed in standard form with a pseudo-spectral method 
+%that requires a fast scalar product
+%\begin{equation}
+%  \D{\mathcal{S}}(N_i^{n-1}) = \D{\mathcal{S}}(\bm{u}^{n-1} \cdot \nabla 
+%u_i^{n-1}) \quad \forall \, i \in (0, 1, 2).
+%\end{equation}
+%We obtain the gradients of $u_i$ in real space by projecting $\partial 
+%u/\partial x$ to $\D{V}_{N_x}$ (boundary condition $\partial u/\partial x(\pm 
+%1)=0$ follows from continuity) and $\partial v/\partial x$ and $\partial 
+%w/\partial x$ to $V_{N_x}$. 
+%\begin{align}
+%  \frac{\partial u}{\partial x} &= \D{\mathcal{T}}^{-1} 
+%(\D{B}^{-1}(\D{C}\hat{u})), \\
+%  \frac{\partial v}{\partial x} &= \mathcal{T}^{-1} 
+%(B^{-1}(\acute{C}\hat{v})), \\
+%  \frac{\partial w}{\partial x} &= \mathcal{T}^{-1} 
+%(B^{-1}(\acute{C}\hat{w})),  
+%\end{align}
+%where $\D{C}_{kj} = (\D{\phi}_j^{'}, \D{\phi}_k)_w$,  $\acute{C_{kj}} = 
+%(\D{\phi}_j^{'}, T_k)_w$ and $B=(T_j, T_k)_w$. Partial derivatives with 
+%respect 
+%to periodic directions are computed as
+%\begin{align}
+% \frac{\partial u_i}{\partial y} &= \D{\mathcal{T}}^{-1}(\imath \underline{m} 
+%\hat{u}_i) &\forall \, i \in (0, 1, 2), \\
+% \frac{\partial u_i}{\partial w} &= \D{\mathcal{T}}^{-1}(\imath \underline{n} 
+%\hat{u}_i) &\forall \, i \in (0, 1, 2).
+%\end{align}
+%Dealiasing is implemented using the 2/3-rule.
+%
+%We finally note that the right hand sides of all the linear systems that need 
+%to solved can be assembled fast using tailored matrix vector products taking 
+%advantage of the fact that there are at most three distinct diagonals in any 
+%matrix. This concludes the implementation required to solve the Navier-Stokes 
+%equations by the incremental pressure correction method and the spectral 
+%Shen-Fourier Galerkin method. We note that all terms in assembly and solve may 
+%be performed using fast routines, at worst scaling like $O(N_x \log_2 N_x)$. 
+%
+%\section{Serial scaling}
+%
+%\section{Parallel scaling}
+%
+%
+%
 

### 2016-01-29 paper.tex
Triangle t

@@
 \left(\frac{\nu \triangle t}{2}\D{A} + \D{B} 
-\right)\hat{v}^{n}_0 + \D{B}\hat{\mathcal{H}}_{y,0}^{n+1/2}, \label{eq:v00} \\
+\right)\hat{v}^{n}_0 + \triangle t\D{B}\hat{\mathcal{H}}_{y,0}^{n+1/2} , 
+\label{eq:v00} \\
 \left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{w}_0^{n+1} &= 
 \left(\frac{\nu \triangle t}{2}\D{A} + \D{B} 
-\right)\hat{w}^{n}_0 + \D{B}\hat{\mathcal{H}}_{z,0}^{n+1/2}, \label{eq:w00}
+\right)\hat{w}^{n}_0 + \triangle t\D{B}\hat{\mathcal{H}}_{z,0}^{n+1/2}, 
+\label{eq:w00}
 \end{align}
@@
 and even coefficients leads to two subsystem that may be trivially solved 
-simultaneously in two threads by 
-letting one thread call LUODDEVEN with even coefficients, and the second thread 
-call it for odd.
+simultaneously in two threads by letting one thread call LUSolveoddeven with 
+even coefficients, and the second thread call it for odd. For optimal 
+performance, the odd and even coefficient would then need to be stored 
+contiguously in computer memory, and not alternately, which would be the normal 
+way of storing the coefficients.
 
@@
 \end{algorithmic}
-
 \end{algorithm}
+
+The biharmonic operator $\N{H}$ is more challenging to solve efficiently, but 
+it is still possible to find an algorithm that is $O(N)$ for $N$ unknowns. Let 
+first $\N{H}_kj$ be decoupled into odd $\N{H}^o_{k,j}= \N{H}_{2k,2k}$ and even 
+$\N{H}^e_{k,j}= \N{H}_{2k+1,2k+1}$ coefficients. Both systems can be solve with 
+the same algorithm, so we show only the details for the even coefficients. Note 
+that $\N{H}$ is the sum of three matrices $\N{L}, \N{A}$ and $\N{B}$ and can be 
+written
+\begin{equation}
+\N{H} = c_0\N{L} + c_1\N{A} + c_2 \N{B},
+\end{equation}
+where $c_0, c_1$ and $c_2$ are constants. The matrix $\N{A}$ contains only 
+three nonzero diagonals at $k-2, k, k+2$,  whereas matrix $\N{B}$ contains five 
+nonzero diagonals at $k-4, k-2, k, k+2, k+4$. The $\N{L}$ matrix is given as
+\begin{align}
+ \N{L}_{kk} &= -4(k+1)(k+2)^2 \\
+ \N{L}_{kj} &= p(k)q(j) + r(k)s(j), \quad j = k+2, k+4, \ldots N_x-4
+\end{align}
+where 
+\begin{align}
+p(k) &= 8 k (k+1)(k+2)(k+4)\pi, &q(j) = \frac{1}{j+3} \\
+r(k) &= 24(k+1)(k+2)\pi, &s(j) =  \frac{(j+2)^2}{(j+3)}
+\end{align} 
+For each row in $\N{H}_{kj}$ we have that $\N{H}_{kj}=c_0\N{L}_{kj}\, \forall j 
+> k+4$. A direct LU decomposition of $\N{H}^e$ leads to a lower triangluar 
+matrix $l^e_{kj}$ with two nonzero diagonals and an upper triangular matrix 
+$U^e_{kj}$ that can be written as
+\begin{equation}
+\N{U}^e_{kj} = a^e(k)q(2j) + b^e(k)s(2j), \quad j > k+2
+\end{equation}
+where $a^e(k)$ and $b^e(k)$ are two new vectors that can be pre-computed 
+recursively as
+\begin{align}
+a^e[0] &= p(0),\\
+b^e[0] &= r(0), \\
+a^e[1] &= p(2) - l^e_{10}a[0]\\
+b^e[1] &= r(2) - l^e_{10}b[0]\\
+a^e(k) &= p(2k) - l^e_{k,k-1} a^e(k-1) - l^e_{k,k-2}a^e(k-2) \quad k = 2, 3, 
+\ldots, (N_x-4)/2, \\
+b^e(k) &= r(2k) - l^e_{k,k-1} b^e(k-1) - l^e_{k,k-2}b(^ek-2)\quad k = 2, 3, 
+\ldots, (N_x-4)/2.
+\end{align}
+In other words, we need the three closest upper diagonals of $U$ plus $a$ and 
+$b$. In other words, the total storage requirement is $7N$. Having completed 
+the LU decomposition with minimal storage, the solution of the system can be 
+obtained very quickly with one forward elimination and a back solve. The back 
+solve is very fast since there is only one new addition to the sum for each row
 

### 2016-02-08 paper.tex
Getting all LU algorithms in there

--- 2016-01-29 paper.tex
@@
 The Navier-Stokes equations used to describe turbulent flow in a doubly 
-periodic channel can be written as
+periodic channel can be written in rotational form as
 \begin{align}
  \frac{\partial \bm{u}}{\partial t}   &= \bm{\mathcal{H}} + \nu 
- \nabla^2 \bm{u} - \nabla{p}, \notag \\
+ \nabla^2 \bm{u} - \nabla{P}, \notag \\
  \nabla \cdot \bm{u} &= 0, \label{eq:NS}
@@
 where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ 
-and $t$ are position and time, $\bm{\mathcal{H}}(\bm{x}, t)$ represents the 
-nonlinear convection and the mean pressure gradient (the driving force), $\nu$ 
-is a constant dynamic viscosity and $p$ is the instantaneous pressure 
-fluctuation normalized by a constant density. The 
-computational domain is $\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$ with 
-channel walls located at $x=\pm 1$ such that no-slip applies at $ \bm{u}(\pm 1, 
-y, z, t) = 0$. The walls are spanning the $y-z$ plane and the equations are 
-periodic in the $y$ and $z$ directions.
+and $t$ are position and time and $\bm{\mathcal{H}}(\bm{x}, t) = \bm{u}\times 
+\bm{\omega}$ (where $\bm{\omega} = \nabla \times \bm{u}$)  represents the 
+nonlinear convection.  The constant dynamic viscosity is denoted as $\nu$ and 
+$P$ is a pressure modified to account for the kinetic energy, i.e., $P = p + 
+\bm{u} \cdot \bm{u}/2$, where $p$ the instantaneous pressure 
+normalized by a constant density. The computational domain is $\Omega=[-1, 
+1]\times [0, L_y] \times [0, L_z]$ with channel walls located at $x=\pm 1$ such 
+that no-slip applies at $ \bm{u}(\pm 1, y, z, t) = 0$. The walls are spanning 
+the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions 
+with periodic lengths $L_y$ and $L_z$, respectively.
 
@@
 $\D{K}_N$ and $\N{K}_N$ are both equal to $K_N$ except from index $l$ that 
-ranges from $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
-and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). The 
-computational mesh has in real space been created using $N = (N_x, N_y, N_z)$ 
+ranges $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
+and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). We use 
+notation $K_N^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
+dimensions of $K_N$.
+
+The computational mesh has in real space been created using $N = (N_x, N_y, 
+N_z)$ 
 intervals, where the two periodic directions use uniform intervals. The 
@@
 ${\mathcal{S}}, \D{\mathcal{S}}$ and $\N{\mathcal{S}}$ are used to distinguish 
-between them with obvious notation. Python implementations of forward and 
-inverse Shen transforms and scalar products are given in the Appendix.
+between them with obvious notation.
 
@@
 $y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
-\N{\phi}_l)_w$ is the Shen scalar product in the $x$-direction for the 
-$\N{V}_N$ space. The weights, $w_i$, required for the Shen transforms are 
+\N{\phi}_l)_w$ is the Shen scalar product in the $x$-direction. The weights, 
+$w_i$, required for the Shen transforms are 
 defined as
@@
 \end{align}
-where the scalar product $\N{B}_{lq} = (\N{\phi}_q, \N{\phi}_l)_w = 
+where $\N{B}_{lq} = (\N{\phi}_q, \N{\phi}_l)_w = 
 \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
-pentadiagonal mass matrix. Likewise, for the remaining two velocity components, 
-the corresponding mass matrix, $\D{B}_{lq} = (\D{\phi}_q, \D{\phi}_l)_w$, is 
-tridiagonal. The complete transformation is now obtained by setting Eq. 
-(\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. Moving to 
-matrix notation we get
+pentadiagonal mass matrix. The complete transformation is now obtained by 
+setting Eq. (\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. 
+Moving to matrix notation we get
 \begin{align}
@@
  \hat{u}_{\bm{k}}(t) &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
- \quad \forall \, \bm{k} \in \N{K}_N.
+ \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:fstB}
 \end{align}
@@
 and $g$. The biharmonic equation (\ref{eq:biharmonic}) is solved with boundary 
-conditions are $u(\pm 1) = u'(\pm 1) = 0$, where the Neumann condition follows 
+conditions are $u(\pm 1) = u'(\pm 1) = 0$, where the Neumann conditions follow 
 from the continuity equation. The boundary conditions for Eq. (\ref{eq:g}) are 
@@
 \left<f, \D{\psi}\right>_w &= \left<\frac{\partial v}{\partial y} + 
-\frac{\partial w}{\partial z}, \D{\psi}\right>_w, \, &\forall \D{\psi} \in 
-\D{V}_N \label{eq:f2} \\
+\frac{\partial w}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
+\D{V}_N, \label{eq:f2} \\
 \left<g, \D{\psi}\right>_w &= \left<\frac{\partial w}{\partial y}  - 
-\frac{\partial v}{\partial z}, \D{\psi}\right>_w, \, &\forall \D{\psi} \in 
+\frac{\partial v}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
 \D{V}_N, \label{eq:g2}
@@
 \left<h_u, \N{\psi} \right>_w &= h\left[-(\underline{m}^2+\underline{n}^2) 
-\N{\mathcal{S}}(\mathcal{H}_x) - \imath \underline{m}\left(\D{\phi}_q, 
-\N{\phi}_l \right)_w \hat{\mathcal{H}}_y - \imath 
-\underline{n}\left(\D{\phi}_q, \N{\phi}_l \right)_w \hat{\mathcal{H}}_z\right], 
+\N{\mathcal{S}}(\mathcal{H}_x) - \imath \underline{m}\left(\D{\phi}_q^{'}, 
+\N{\phi}_l \right)_w \hat{\mathcal{H}}_{y, q} - \imath 
+\underline{n}\left(\D{\phi}_q^{'}, \N{\phi}_l \right)_w 
+\hat{\mathcal{H}}_{z, q}\right], 
 \\
 \left< h_g, \D{\psi} \right>_w &= h\left[ \imath \underline{m}\, \D{B} 
-\hat{\mathcal{H}}_z - \imath \underline{n}\, \D{B} \hat{\mathcal{H}}_y 
-\right]
+\hat{\mathcal{H}}_z - \imath \underline{n}\, \D{B} \hat{\mathcal{H}}_{y} 
+\right],
 \end{align}
@@
 %\end{align}
-The mass matrices, $\N{A}_{lq} = \left( \N{\phi}_q^{''}, \N{\phi}_l\right)_w 
-$ and $\N{L}_{lq} = \left(\N{\phi}^{''''}_q, \N{\phi}_l\right)_w$ are given by 
-Shen, whereas $\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w$ is given in the 
+The mass matrices, stiffness matrices $\N{A}_{lq} = \left( \N{\phi}_q^{''}, 
+\N{\phi}_l\right)_w, 
+\D{A}_{lq} = \left( \D{\phi}_q^{''}, \D{\phi}_l\right)_w 
+$ and $\N{Q}_{lq} = \left(\N{\phi}^{''''}_q, \N{\phi}_l\right)_w$ are given by 
+Shen, whereas $\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w$ and 
+$\left(\D{\phi}_q^{'}, \N{\phi}_l\right)_w$ are given in the 
 appendix.
@@
 Discretization in time is performed using a constant time step $\triangle t$, 
-such that the time is represented discretely as $t_n = n \triangle t$. After 
+such that the time is represented discretely as $t_n = n \triangle t, \, n \in 
+\mathcal{Z}^{\ge}$. After 
 discretizing in time, equations (\ref{eq:u1}) and (\ref{eq:g1}) are transformed 
-to a matrix form that can be solved for $\hat{u}^{n+1} (= 
-\hat{u}_{\bm{k}}(t_{n+1}))$ and $\hat{g}^{n+1}$
+to a matrix form that can be solved for $\hat{u}^{n+1}$ and $\hat{g}^{n+1}$ 
+given wavenumbers $m$ and $n$
 \begin{align}
 \N{H}\hat{u}^{n+1} & = \left(2\N{A} - 2\underline{z}^2\N{B} - \N{H} 
-\right)\hat{u}^{n} + \N{\mathcal{S}}(h_v^{n+1/2}) \triangle t, 
+\right)\hat{u}^{n} + \N{\mathcal{S}}(h_v^{n+1/2}) \triangle t & 
+\forall \, {\bm{k}} \in \N{K}_N, 
 \label{eq:ufin}\\ 
 \D{H} \hat{g}^{n+1} &= \left(2 \D{B}-\D{H}\right) 
-\hat{g}^{n} + \D{\mathcal{S}}(h_g^{n+1/2}) \triangle t. \label{eq:gfin}
-\end{align}
-Here $\N{H} = -\frac{\nu \triangle t}{2}\N{L} + \left( 1 + \nu \triangle t 
-\underline{z}^2 \right) \N{A} - \frac{2\underline{z}^2 + \nu \triangle t 
-\underline{z}^4}{2} \N{B}$, $\D{H} = -\frac{\nu \triangle t}{2}\D{A} + (1 + 
-\frac{\nu \triangle t \underline{z}^2}{2}) \D{B}$ and $\underline{z}^2 = 
-\underline{m}^2 + \underline{n}^2$. The convective terms are computed with 
+\hat{g}^{n} + \D{\mathcal{S}}(h_g^{n+1/2}) \triangle t &\forall \, 
+{\bm{k}} \in 
+\D{K}_N. \label{eq:gfin}
+\end{align}
+The convective terms are computed with 
 a forward Adams-Bashforth projection, such that $\N{\mathcal{S}}(h_v^{n+1/2}) = 
 \N{\mathcal{S}}(1.5h_v^{n} - 0.5 h_v^{n-1})$ and $\D{\mathcal{S}}(h_g^{n+1/2}) 
-= \D{\mathcal{S}}(1.5h_g^{n} - 0.5 h_g^{n-1})$.
-
-Equations (\ref{eq:v0}) and (\ref{eq:w0}) are also discretized in time using 
+= \D{\mathcal{S}}(1.5h_g^{n} - 0.5 h_g^{n-1})$ and the coefficient matrices are 
+given as
+\begin{align}
+\N{H}_{lq}(m, n) &= -\frac{\nu \triangle t}{2}\N{Q}_{lq} + \left( 1 + \nu 
+\triangle t 
+\underline{z}^2 \right) \N{A}_{lq} - \frac{2\underline{z}^2 + \nu \triangle t 
+\underline{z}^4}{2} \N{B}_{lq} &\forall m, n \in K_N^p, \\
+\D{H}_{lq}(m, n) &= -\frac{\nu \triangle 
+t}{2}\D{A}_{lq} + (1 + \frac{\nu \triangle t \underline{z}^2}{2}) \D{B}_{lq} 
+&\forall m, n \in K_N^p,
+\end{align}
+where $\underline{z}^2 = \underline{m}^2 + \underline{n}^2$. Equations 
+(\ref{eq:v0}) and (\ref{eq:w0}) are also discretized in time using 
 Crank-Nicolson and Adams-Bashforth, and we obtain
@@
 \begin{algorithm}
-	\caption{Solution procedure}
+	\caption{Solution procedure for Navier Stokes equations}
     \label{alg:NS}
@@
     	\State \text{Initialize} $\bm{u}(\bm{x}, 0)$ and compute 
-    	$\hat{g}_{\bm{k}}(0)$
-		\State \text{Assemble} $\N{H}, \D{H}$
-		\State $t=0$
+    	$\hat{g}_{\bm{k}}(0)$ and $\hat{u}_{\bm{k}}(0)$
+		\State \text{Compute LU decomposition of} $\N{H}(m,n), \D{H}(m,n) 
+		\,\forall \, m, n \in K_N^p$
+		\State $t=0; n=0$
     	\While{$t<T$}
-	    	\State $t \gets t+dt$
-	    	\State \text{Compute nonlinear convection} $\bm{\mathcal{H}}$
+	    	\State \text{Compute nonlinear convection} $\bm{\mathcal{H}}^n$
+	    	\State \text{Compute right hand sides of Eqs.~}(\ref{eq:ufin}), 
+	    	(\ref{eq:gfin})
 		    \State \text{Solve Eq.}(\ref{eq:ufin}) for 
@@
 		    \State \text{Solve Eqs.}(\ref{eq:v00}, \ref{eq:w00}) for 
-		    $\hat{v}^{n+1}(l, 0, 0, t_{n+1}), \hat{w}^{n+1}(l, 0, 0, t_{n+1})\, 
+		    $\hat{v}_0^{n+1}, \hat{w}_0^{n+1}\, 
 		    \forall \, l \in [0, 1, \ldots, N_x-2] $
+		    \State $n \gets n + 1$
+		    \State $t \gets t+\triangle t$
 		    \State \text{Update to new time step}
@@
 
-The major computational cost is found in computing the nonlinear convection and 
-in solving for Eqs.~(\ref{eq:ufin}, \ref{eq:gfin}). The nonlinear convection 
+The major computational cost comes from  computing the nonlinear convection 
+and solving for Eqs.~(\ref{eq:ufin}, \ref{eq:gfin}). The nonlinear 
+convection 
 term $\bm{\hat{\mathcal{H}}}$ is computed using a pseudo-spectral method with 
@@
 N. The Shen transforms make use of the Chebyshev transforms that are using 
-discrete cosine transforms. The forward scalar product is shown in Algorithm 
-\ref{alg:fst}.
+discrete cosine transforms. The forward scalar product is implemented as shown 
+in Algorithm \ref{alg:fst}. A complete transform is obtained through a scalar 
+product followed by solving the linear system involving the mass 
+matrix, just like shown in Eq. (\ref{eq:fstB}) for space $\N{V}_N$ and similar 
+for $\D{V}_N$ and $V_N$. Since the mass matrices are diagonal, tridiagonal and 
+pentadiagonal for $V_N, \D{V}_N$ and $\N{V}_N$ respectively, solving the linear 
+system is $O(N)$ and the complete transform still scales like $O(N\log N)$. 
 \begin{algorithm}
@@
 	\begin{algorithmic}[1]
-		\State \textbf{Input:} $\{f\}_{j=0}^{N}$
+	\Procedure{fastScalar}{}
+		\State \textbf{Input:} $\{f_j\}_{j=0}^{N}$
 		\State \textbf{Input:} points
@@
 		\If{ points = "Gauss-Lobatto"}
-			\State $f_k \gets \text{dct}(f_j, type=2, axis=0)\pi/(2 N)\, 
-			\forall \, \, k \in 
-			[0, 1, \ldots, N]$
+			\State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=2}, 
+			\text{axis=0})\frac{\pi}{2 N}$
 		\Else{ points = "Gauss-Chebyshev"}
-		    \State $f_k \gets \text{dct}(f_j, type=1, axis=0)\pi/(2 (N-1))\, 
-		    \forall \, \, k 
-		    \in [0, 1, \ldots, N]$
+		    \State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=1}, 
+		    \text{axis=0}) \frac{\pi}{2 (N-1)}$
 	    \EndIf
 	    \If{ space = $\D{V}_N$}
-	        \State $f_k \gets f_k - f_{k+2} \, \forall k \in [0, 1, \ldots, 
-	        N-2]$
+	        \For{$k=0$ to $N-2$}
+		        \State $\hat{f}_k \gets w_k - w_{k+2}$
+	        \EndFor
+			\State \Return $\{\hat{f}_k\}_{k=0}^{N-2}$	        
 	    \ElsIf{ space = $\N{V}_N$}	                
-	        \State $f_k \gets f_k - 2(k+2)/(k+3)f_{k+2} + (k+1)/(k+3)f_{k+4} 
-	        \,\, 
-	        \forall \, k \in [0, 1, \ldots, N-4]$
+  	        \For{$k=0$ to $N-4$}
+       	        \State $\hat{f}_k \gets w_k - \frac{2(k+2)}{k+3} w_{k+2} + 
+       	        \frac{k+1}{k+3}w_{k+4}$
+   	        \EndFor
+   	        \State \Return $\{\hat{f}_k\}_{k=0}^{N-4}$
    	    \Else{}
-	        \State pass
+   	        \For{$k = 0$ to $N$}
+     	        \State $f_k \gets \frac{2}{\pi} w_k $
+     	    \EndFor
+     	    \State $f_0 \gets f_0 / 2$
+     	    \If{ points = "Gauss-Lobatto"}
+	     	\State $f_N \gets f_N / 2$	     	
      	    \EndIf
-\State \Return $f_k$    
+			\State \Return $\{f_k\}_{k=0}^{N}$    
+	    \EndIf
+	\EndProcedure
 \end{algorithmic}
@@
 direct LU decomposition and only a few diagonals need to be stored for each 
-final lower (L) and upper (U) matrix.
+final lower (L) and upper (U) triangular matrix.
 
@@
 Algorithm (\ref{alg:lu}), and a very efficient tailored solve through 
-Algorithm (\ref{alg:lusolve}). The efficiency follows from the fact that all 
+Algorithm (\ref{alg:lusolve}). The efficiency follows from the fact that 
 $U^{o,e}_{kj} = U^{o,e}_{kk+2} \, \forall \, j > k+2$, i.e., the upper 
-tridiagonal matrix has a maximum of three distinct values for each row. 
+triangular matrix has a maximum of three distinct values for each row, which 
+allows for a very fast solve through back substitution. 
 Regarding Algorithm 
@@
   \EndFor
+
   \State $k \gets M^e-1$
@@
 \end{algorithmic}
+
 \begin{algorithmic}[1]
@@
 \{b_{2k+1}\}_{k=0}^{M^o}$)
-
+  \State \Return $\{u_k\}_{k=0}^{N-2}$
 \EndProcedure
 \end{algorithmic}
-
+\end{algorithm}
+
+The biharmonic operator $\N{H}$ is more challenging to solve efficiently, but 
+it is still possible to find an algorithm that is $O(N)$ for a system of $N$ 
+unknowns. Note 
+that $\N{H}$ is the sum of three matrices $\N{Q}, \N{A}$ and $\N{B}$ and can be 
+written as
+\begin{equation}
+\N{H} = c_0\N{Q} + c_1\N{A} + c_2 \N{B},
+\end{equation}
+where $c_0, c_1$ and $c_2$ are constants. The matrix $\N{A}_{kj}$ contains only 
+three nonzero diagonals at $ j = k-2, k, k+2$,  whereas matrix $\N{B}_{kj}$ 
+contains five 
+nonzero diagonals at $ j = k-4, k-2, k, k+2, k+4$. The nonzero elements of the 
+$\N{Q}_{kj}$ matrix are given as
+\begin{align}
+ \N{Q}_{kk} &= -4(k+1)(k+2)^2 \\
+ \N{Q}_{kj} &= p_kq_j + r_ks_j, \quad j = k+2, k+4, \ldots N_x-4
+\end{align}
+where 
+\begin{align}
+p_k &= 8 k (k+1)(k+2)(k+4)\pi, &q_j = \frac{1}{j+3} \label{eq:pk} \\
+r_k &= 24(k+1)(k+2)\pi, &s_j =  \frac{(j+2)^2}{(j+3)} \label{eq:rk}
+\end{align} 
+For each row in $\N{H}_{kj}$ we have that $\N{H}_{kj}=c_0\N{Q}_{kj}\, \forall j 
+> k+4$. A direct LU decomposition of $\N{H}$ can be performed as shown in Alg. 
+\ref{alg:lu_biharmonic}, which leads to a lower triangular 
+matrix $\N{L}_{kj}$ with two nonzero diagonals at $j=k-2$ and $k-4$ plus a 
+unity 
+main diagonal. The upper triangular matrix $\N{U}_{kj}$ contains three distinct 
+diagonals at $j=k, k+2$ and $k+4$. The remaining part of $\N{U}_{kj}$ can be 
+written as
+\begin{equation}
+\N{U}_{kj} = a_k q_j + b_k s_j, \quad j = k+6, k+8, \ldots, N_x-4 \label{eq:ab}
+\end{equation}
+where $\bm{a}$ and $\bm{b}$ are two new vectors that can be computed 
+recursively as shown in Alg. (\ref{alg:ab}). If $\bm{a}$ and $\bm{b}$ are 
+pre-computed 
+the total storage requirement for the complete LU decomposition is no more than 
+$7N$ and the solution of the complete system can be obtained very quickly with 
+one forward elimination and a back solve. The back 
+solve is very fast because of the recursive formulation (\ref{eq:ab}), leading 
+to a formula like (valid for $k \le N_x-8$)
+\begin{equation}
+u_k = \left(y_k - \N{U}_{k, k+2} u_{k+2} - \N{U}_{k, k+4} u_{k+4} - 
+a_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} q_j u_j - 
+b_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} 
+s_j u_j\right)/\N{U}_{kk},
+\end{equation}
+that only requires 
+one new addition to the row-sums for each back-traversed
+row, see Alg. (\ref{alg:SolveBiharmonic}).
+
+\begin{algorithm}
+	\caption{LU decomposition of biharmonic operator $\N{H}$. Current algorithm 
+	is using dense storage of $U_{kj}$, but no more than the 3 diagonals need 
+	to be stored.}
+	\label{alg:lu_biharmonic}
+	\begin{algorithmic}[1]
+		\Procedure{LUbiharmonic}{}
+			\State \textbf{Input:} $\{\N{H}_{kj}\}_{k, j = 0}^{N}$
+	        \State $U_{0j} = \N{H}_{0j} \quad \forall \, j \in 0, 2, \ldots, N$
+	        \State $U_{1j} = \N{H}_{1j} \quad \forall \, j \in 1, 3, \ldots, N$
+	        \State $L_{20} = \N{H}_{20}/U_{00}$
+	        \State $L_{31} = \N{H}_{31}/U_{11}$
+	        \State $U_{2j} = L_{20} U_{0j} \quad \forall \, j \in 2, 4, \ldots, 
+	        N$
+	        \State $U_{3j} = L_{31} U_{1j} \quad \forall \, j \in 3, 5, \ldots, 
+	        N$
+	        \For{$k = 4, 5, \ldots, N$}
+	            \State $L_{k,k-2} = {U}_{k,k-2}/{U}_{k-2, k-2}$
+		        \State $U_{kj} = {U}_{kj} - L_{k, k-2} U_{k-2, j} \quad 
+		        \forall j = k, k+2, \ldots, N$
+		        \If{$k <= N-2$}
+		        \State $L_{k+2,k-2} = \N{H}_{k+2, k-2} / U_{k-2, k-2}$
+		        \State $U_{k+2,j} = \N{H}_{k+2,j} - L_{k+2, k-2} U_{k-2, j} 
+		        \quad \forall j = k-2, k, \ldots, N$
+		        \EndIf
+	        \EndFor
+	        \State \Return $\{L_{kj}\}_{k, j = 0}^{N}$, $\{U_{kj}\}_{k, j = 
+	        0}^{N}$
+		\EndProcedure	
+	\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+	\caption{Recursive formula to compute $\{a_k\}_{k=0}^{N_x-4}, 
+	\{b_k\}_{k=0}^{N_x-4}$ in Eq. (\ref{eq:ab}). Note that the lower triangular 
+	matrix $\{L_{kj}\}_{k,j=0}^{N_x-4}$ contains only two nonzero diagonals and 
+	the parameters $p_k$ and $r_k$ are given in Eqs. (\ref{eq:pk}) and 
+	(\ref{eq:rk}) respectively. }
+    \label{alg:ab}
+	\begin{algorithmic}[1]
+		\Procedure{ReduceAB}{}
+		\State \textbf{Input:} $ \{L_{kj}\}_{k,j=0}^{N_x-4}$
+		\For{$k=0,1$}
+		     \State $a_k = p_k$
+		     \State $b_k = r_k$
+		\EndFor
+		\For{$k=2,3$}
+		    \State $a_k = p_k - L_{k,k-2}\,a_{k-2}$
+		    \State $b_k = r_k - L_{k, k-2}\,b_{k-2}$
+		\EndFor
+		\For{$k=4, 5, \ldots, N_x-4$}
+     		\State $a_k = p_k - L_{k,k-2}\,a_{k-2} - L_{k, k-4} \,a_{k-4}$
+	    	\State $b_k = r_k - L_{k, k-2}\,b_{k-2}- L_{k, k-4} \,b_{k-4}$
+		\EndFor
+		\State \Return $\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$
+        \EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+	\caption{Solve biharmonic Eq. (\ref{eq:ufin}) with pre-computed $L$, $U$ 
+	matrices. The parameters $q_j$ and $s_j$ are given in Eqs. (\ref{eq:pk}) 
+	and 
+	(\ref{eq:rk}) respectively. }
+	\label{alg:SolveBiharmonic}
+	\begin{algorithmic}[1]
+		\Procedure{LUSolveBiharmonic}{}
+		  \State \textbf{Input:} $\{U_{kj}\}_{k,j=0}^{N}, 
+		  \{L_{kj}\}_{k,j=0}^{N}, 
+		  \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N}, \{b_k\}_{k=0}^{N}$ 
+		  \State // Solve $Ly=f$ by forward elimination 
+		  \State $y_0 \gets f_0$
+		  \State $y_1 \gets f_1$
+		  \State $y_2 \gets f_2 - L_{20}y_{0}$
+		  \State $y_3 \gets f_3 - L_{31}y_{1}$
+		  \For{$k = 4$ \textbf{to} $N$}
+		  \State $y_k \gets f_k - L_{k,k-2}y_{k-2} - L_{k, k-4}y_{k-4}$
+		  \EndFor
+		  \State // Solve $Uu=y$ with back substitution                    
+		  \State $u_N = y_N / U_{NN}$
+	      \State $u_{N-1} = y_{N-1} / U_{N-1,N-1}$
+	      \State $u_{N-2} = \left(y_{N-2} - U_{N-2, N}y_N\right)/U_{N-2,N-2}$
+	      \State $u_{N-3} = \left(y_{N-3} - U_{N-3, 
+	      N-1}y_{N-1}\right)/U_{N-3,N-3}$
+
+	      \State $q^o = q^e = s^o = s^e = 0.0$
+		  \For{$k = N-4$ \textbf{to} $0$ \textbf{step} $-1$}
+		  \State $u_k \gets y_k - U_{k, k+2} u_{k+2} - U_{k, k+4} 
+		  u_{k+4} $ 
+		  \If{$k < N-5$}
+		  \State $j=k+6$
+		  \If{$k$ is odd}
+		    \State $q^o \gets q^o + u_{j}q_{j}$
+		    \State $s^o \gets s^o + u_{j}s_{j}$
+		    \State $u_k \gets u_k - c_0 q^o a_k - c_0 s^o b_k $	    
+		  \Else
+		    \State $q^e \gets q^e + u_{j}q_{j}$
+		    \State $s^e \gets s^e + u_{j}s_{j}$	  
+		    \State $u_k \gets u_k - c_0 q^e a_k - c_0 s^e b_k $
+		  \EndIf
+		  \EndIf
+		  \State $u_k \gets  u_k / U_{kk}$
+		  \EndFor
+		  \State \Return $\{u_k\}_{k=0}^{N}$
+
+		\EndProcedure
+	\end{algorithmic}
 \end{algorithm}
@@
 \section{Appendices}
-\subsection*{Fast Shen transforms}
-The fast forward and inverse Chebyshev transforms are used by the Shen transforms. The code below computes both scalar products and complete transforms for both
-
-\begin{python}
-from numpy import *
-from scipy.fftpack import dct
-
-def fct(f, transform="GL"):
-    """Fast Chebyshev transform."""
-    N = f.shape[0]
-    f_hat = zeros(N)
-    if transform == "GL":
-        fk[:] = dct(fj, type=2, axis=0)            
-        fk /= N
-        fk[0] /= 2
-
-    elif transform == "GC":
-        fk[:] = dct(fj, type=1, axis=0)/(N-1)
-        fk[0] /= 2
-        fk[-1] /= 2
-
-    return fk
-
-def ifct(fk, transform="GL"):
-    """Inverse fast Chebyshev transform."""
-    if transform == "GL":
-        f = 0.5*dct(fk, type=3, axis=0)
-        f += 0.5*fk[0]
-
-    elif transform == "GC":
-        f = 0.5*dct(fk, type=1, axis=0)
-        f += 0.5*fk[0]
-        f[::2] += 0.5*fk[-1]
-        f[1::2] -= 0.5*fk[-1]
-
-    return f
-
-def fastChebScalar(f, transform="GL"):
-    """Fast Chebyshev scalar product."""
-    N = f.shape[0]
-    if transform == "GL":
-        fk = dct(fj, type=2, axis=0)*pi/(2*N)
-
-    elif transform == "GC":
-        fk = dct(fj, type=1, axis=0)*pi/(2*(N-1))
-    return fk
-
-def fastShenScalar(f, transform="GL"):
-    """Fast Shen scalar product"""                
-    fk = fastChebScalar(f, transform)
-    fk[:-2] -= fk[2:]
-    return fk
-
-def ifst(fk):
-    """Fast inverse Shen transform
-    """
-    N = len(fj)
-    wk = np.zeros(N, dtype=fk.dtype)        
-    wk[:-2] = fk[:-2] 
-    wk[2:] -= fk[:-2] 
-    return ifct(wk)
-
-def fst(fj, transform="GL"):
-    """Fast Shen transform"""
-    fk = fastShenScalar(fj, transform)
-
-    N = fj.shape[0]
-    if transform == "GL":
-        ck = np.ones(N-2); ck[0] = 2
-
-    elif transform == "GC":
-        ck = np.ones(N-2); ck[0] = 2; ck[-1] = 2
-
-    a = np.ones(N-4)*(-np.pi/2)
-    b = np.pi/2*(ck+1)
-    c = a.copy()
-    fk[:-2] = TDMA_1D(a, b, c, fk[:-2])
-
-    return fk
-
-def fastShenScalar(f, transform="GL"):
-    """Fast Shen scalar product Neumann basis"""        
-    k  = wavenumbers(f.shape)
-    fk = fastChebScalar(f, transform)
-    fk[:-2] -= ((k/(k+2))**2) * fk[2:]
-    return fk
-
-def ifst(fk):
-    """Fast inverse Shen scalar transform Neumann basis"""
-    k = self.wavenumbers(fk.shape[0])
-    wk = np.zeros(fk.shape[0])    
-    wk[1:-2] = fk[1:-2]
-    wk[3:] -= (k[1:]/(k[1:]+2))**2*fk[1:-2]
-    f = ifct(wk, transform)
-    return f
-
-def fst(f, transform="GL"):
-    """Fast Shen transform Neumann basis"""
-    fk = fastShenScalar(f, transform)
-    N = fj.shape[0]
-    k = wavenumbers(N)
-    ck = np.ones(N-3)
-    if transform == "GC": ck[-1] = 2
-    a = np.ones(N-5)*(-np.pi/2)*(k[1:-2]/(k[1:-2]+2))**2
-    b = np.pi/2*(1+ck*(k[1:]/(k[1:]+2))**4)
-    c = a.copy()
-    fk[1:-2] = TDMA_1D(a, b, c, fk[1:-2])
-    return fk
-
-# Cython implementation of TDMA    
-import numpy as np
-cimport cython
-cimport numpy as np
-#cython: boundscheck=False
-#cython: wraparound=False
-
-ctypedef fused T:
-    np.float64_t
-    np.complex128_t
-
-def TDMA_1D(np.ndarray[np.float64_t, ndim=1] a, 
-            np.ndarray[np.float64_t, ndim=1] b, 
-            np.ndarray[np.float64_t, ndim=1] c, 
-            np.ndarray[T, ndim=1] d):
-    cdef:
-        unsigned int n = b.shape[0]
-        unsigned int m = a.shape[0]
-        unsigned int k = n - m
-        int i
-
-    for i in range(m):
-        d[i + k] -= d[i] * a[i] / b[i]
-        b[i + k] -= c[i] * a[i] / b[i]
-    for i in range(m - 1, -1, -1):
-        d[i] -= d[i + k] * c[i] / b[i + k]
-    for i in range(n):
-        d[i] /= b[i]
-
-    return d
-
-\end{python}
+%\subsection*{Fast Shen transforms}
+%The fast forward and inverse Chebyshev transforms are used by the Shen 
+%transforms. The code below computes both scalar products and complete 
+%transforms for both 
+%
+%\begin{python}
+%from numpy import *
+%from scipy.fftpack import dct
+%
+%def fct(f, transform="GL"):
+%    """Fast Chebyshev transform."""
+%    N = f.shape[0]
+%    f_hat = zeros(N)
+%    if transform == "GL":
+%        fk[:] = dct(fj, type=2, axis=0)            
+%        fk /= N
+%        fk[0] /= 2
+%                
+%    elif transform == "GC":
+%        fk[:] = dct(fj, type=1, axis=0)/(N-1)
+%        fk[0] /= 2
+%        fk[-1] /= 2
+%            
+%    return fk
+%
+%def ifct(fk, transform="GL"):
+%    """Inverse fast Chebyshev transform."""
+%    if transform == "GL":
+%        f = 0.5*dct(fk, type=3, axis=0)
+%        f += 0.5*fk[0]
+%    
+%    elif transform == "GC":
+%        f = 0.5*dct(fk, type=1, axis=0)
+%        f += 0.5*fk[0]
+%        f[::2] += 0.5*fk[-1]
+%        f[1::2] -= 0.5*fk[-1]
+%
+%    return f
+%
+%def fastChebScalar(f, transform="GL"):
+%    """Fast Chebyshev scalar product."""
+%    N = f.shape[0]
+%    if transform == "GL":
+%        fk = dct(fj, type=2, axis=0)*pi/(2*N)
+%        
+%    elif transform == "GC":
+%        fk = dct(fj, type=1, axis=0)*pi/(2*(N-1))
+%    return fk
+%
+%def fastShenScalar(f, transform="GL"):
+%    """Fast Shen scalar product"""                
+%    fk = fastChebScalar(f, transform)
+%    fk[:-2] -= fk[2:]
+%    return fk
+%
+%def ifst(fk):
+%    """Fast inverse Shen transform
+%    """
+%    N = len(fj)
+%    wk = np.zeros(N, dtype=fk.dtype)        
+%    wk[:-2] = fk[:-2] 
+%    wk[2:] -= fk[:-2] 
+%    return ifct(wk)
+%    
+%
+%def fst(fj, transform="GL"):
+%    """Fast Shen transform"""
+%    fk = fastShenScalar(fj, transform)
+%    
+%    N = fj.shape[0]
+%    if transform == "GL":
+%        ck = np.ones(N-2); ck[0] = 2
+%        
+%    elif transform == "GC":
+%        ck = np.ones(N-2); ck[0] = 2; ck[-1] = 2
+%        
+%    a = np.ones(N-4)*(-np.pi/2)
+%    b = np.pi/2*(ck+1)
+%    c = a.copy()
+%    fk[:-2] = TDMA_1D(a, b, c, fk[:-2])
+%        
+%    return fk
+%
+%def fastShenScalar(f, transform="GL"):
+%    """Fast Shen scalar product Neumann basis"""        
+%    k  = wavenumbers(f.shape)
+%    fk = fastChebScalar(f, transform)
+%    fk[:-2] -= ((k/(k+2))**2) * fk[2:]
+%    return fk
+%
+%def ifst(fk):
+%    """Fast inverse Shen scalar transform Neumann basis"""
+%    k = self.wavenumbers(fk.shape[0])
+%    wk = np.zeros(fk.shape[0])    
+%    wk[1:-2] = fk[1:-2]
+%    wk[3:] -= (k[1:]/(k[1:]+2))**2*fk[1:-2]
+%    f = ifct(wk, transform)
+%    return f
+%        
+%def fst(f, transform="GL"):
+%    """Fast Shen transform Neumann basis"""
+%    fk = fastShenScalar(f, transform)
+%    N = fj.shape[0]
+%    k = wavenumbers(N)
+%    ck = np.ones(N-3)
+%    if transform == "GC": ck[-1] = 2
+%    a = np.ones(N-5)*(-np.pi/2)*(k[1:-2]/(k[1:-2]+2))**2
+%    b = np.pi/2*(1+ck*(k[1:]/(k[1:]+2))**4)
+%    c = a.copy()
+%    fk[1:-2] = TDMA_1D(a, b, c, fk[1:-2])
+%    return fk
+%
+%# Cython implementation of TDMA    
+%import numpy as np
+%cimport cython
+%cimport numpy as np
+%#cython: boundscheck=False
+%#cython: wraparound=False
+%
+%ctypedef fused T:
+%    np.float64_t
+%    np.complex128_t
+%
+%def TDMA_1D(np.ndarray[np.float64_t, ndim=1] a, 
+%            np.ndarray[np.float64_t, ndim=1] b, 
+%            np.ndarray[np.float64_t, ndim=1] c, 
+%            np.ndarray[T, ndim=1] d):
+%    cdef:
+%        unsigned int n = b.shape[0]
+%        unsigned int m = a.shape[0]
+%        unsigned int k = n - m
+%        int i
+%        
+%    for i in range(m):
+%        d[i + k] -= d[i] * a[i] / b[i]
+%        b[i + k] -= c[i] * a[i] / b[i]
+%    for i in range(m - 1, -1, -1):
+%        d[i] -= d[i + k] * c[i] / b[i + k]
+%    for i in range(n):
+%        d[i] /= b[i]
+%        
+%    return d
+%    
+%
+%\end{python}
 \subsection*{Matrices}
@@
 \begin{equation}
- \{\D{C}_{kj}\}_{k,j=0}^{N_x-2} = \left(\D{\phi}^{'}_j, \D{\phi}_k \right)_w = \begin{cases}
- \pi(k+1), &j=k+1,\\
- -\pi(k+1), & j=k-1, \\
+ \{\N{C}_{kj}\}_{k,j=0}^{N_x-4, N_x-2} = \left(\D{\phi}^{'}_j, \N{\phi}_k 
+ \right)_w = \begin{cases}
+ -\pi(k+1), &j=k-1,\\
+ 2\pi(k+1), & j=k+1, \\
+ -\pi(k+1), & j=k+3, \\
  0, &\text{otherwise}.
@@
 \begin{equation}
- \{\grave{C}_{kj}\}_{k=0,j=1}^{N_x-2} = \left(\N{\phi}^{'}_j, \D{\phi}_k \right)_w = \begin{cases}
- \pi(k+1), &j=k+1,\\
- -\pi(k-1)^2/(k+1) , & j=k-1, \\
+ \{\D{C}_{kj}\}_{k,j=0}^{N_x-2, N_x-4} = \left(\N{\phi}^{'}_j, \D{\phi}_k 
+ \right)_w = \begin{cases}
+ \pi (k-2)(k+1)/k, &j=k-3,\\
+ -2 \pi(k+1)^2/(k+2), & j=k-1, \\
+ \pi(k+1), & j=k+1, \\
  0, &\text{otherwise}.
@@
 \begin{equation}
- \{\tilde{C}_{kj}\}_{k=1,j=0}^{N_x-2} = (\D{\phi}^{'}_j, \N{\phi}_k)_w = \begin{cases}
- -\pi(k+1), &j=k-1,\\
- -\pi(2-\left(\frac{k}{k+2}\right)^2(k+3)) , & j=k+1, \\
- -2 \pi(1-\left(\frac{k}{k+2}\right)^2), &j=k+3, k+5, k+7, \ldots,\\
- 0, &\text{otherwise}.
- \end{cases}
-\end{equation}
-
-\begin{equation}
-  \{\acute{C}_{kj}\}_{k,j=0}^{N_x-2, N_x} = (\D{\phi}^{'}_j, T_k)_w = \begin{cases}
-    -\pi (k+1), &j=k-1, \\
-    -2 \pi,   &j = k+1, k+3, k+5, \ldots, \\
-    0, &\text{otherwise}
-  \end{cases}
-\end{equation}
-
-\begin{equation}
-  \{\D{B}_{kj}\}_{j,k=0}^{N_x-2} = (\D{\phi}_j, \D{\phi}_k)_w = \begin{cases}
-  \frac{c_k+c_{k+2}}{2}\pi,& k = j \\
-  -\frac{\pi}{2},& j=k\pm 2 \\
-  0,& \text{otherwise}
-\end{cases}
-\end{equation}
-
-\begin{equation}
- \{\tilde{B}_{kj}\}_{k=1, j=0}^{N_x-2} = (\D{\phi}_j, \N{\phi}_k)_w = \begin{cases}
- -\frac{\pi}{2}, &j=k-2,\\
- -\frac{\pi}{2}\left(1+ {c_{k+2}}\left(\frac{k}{k+2}\right)^2 \right), &j=k,\\
- -\frac{\pi}{2} k^2/(k+2)^2 , & j=k+2, \\
- 0, &\text{otherwise}.
- \end{cases}
-\end{equation}
-
-\begin{equation}
- \{\grave{B}_{kj}\}_{k=0, j=1}^{N_x-2} = (\N{\phi}_j, \D{\phi}_k)_w = \begin{cases}
- -\frac{\pi}{2} (k-2)^2/k^2 , & j=k-2, \\
- -\frac{\pi}{2}\left(1+ {c_{k+2}} \left(\frac{k}{k+2}\right)^2 \right), &j=k,\\
- -\frac{\pi}{2}, &j=k-2,\\
- 0, &\text{otherwise}.
- \end{cases}
-\end{equation}
-
-\begin{equation}
- \{\N{B}_{kj}\}_{k,j=1}^{N_x-2} = (\N{\phi}_j, \N{\phi}_k)_w = \begin{cases}
- -\frac{\pi}{2} \left(\frac{k-2}{k}\right)^2, &j=k-2,\\
- -\frac{\pi}{2}\left(1+ {c_{k+2}}\left(\frac{k}{k+2}\right)^4 \right), &j=k,\\
- -\frac{\pi}{2} k^2/(k+2)^2 , & j=k+2, \\
+ \{\N{B}_{kj}\}_{k,j=0}^{N_x-4} = \{\N{B}_{jk}\}_{j,k=0}^{N_x-4} = (\N{\phi}_j, 
+ \N{\phi}_k)_w = \begin{cases}
+ \frac{\pi}{2} \left(c_k + 4 \left(\frac{k+2}{k+3} \right)^2 + c_{k+4} 
+ \left(\frac{k+1}{k+3}\right)^2    \right), &j=k,\\
+ -\pi \left( \frac{k+2}{k+3} + \frac{(k+4)(k+1)}{(k+5)(k+3)} \right), &j=k+2,\\
+ \frac{\pi}{2} \frac{k+1}{k+3} , & j=k+4, \\
  0, &\text{otherwise}.

### 2016-03-17 paper.tex
Work on getting all algorithms in there

@@
 K_N = \Big\{\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / &(l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
- &(l, m, n) \in  [0, \ldots, N_x] \times [-\frac{N_y}{2},\ldots, \frac{N_y}{2}-1] \times[-\frac{N_z}{2},\ldots, \frac{N_z}{2}-1] \Big\}. \label{eq:wavenumbermesh}
-\end{align}
-$\D{K}_N$ and $\N{K}_N$ are both equal to $K_N$ except from index $l$ that 
+ &(l, m, n) \in  [0, \ldots, N_x] \times [-\frac{N_y}{2},\ldots, 
+ \frac{N_y}{2}-1] \times[-\frac{N_z}{2},\ldots, \frac{N_z}{2}-1] \Big\} 
+ \label{eq:wavenumbermesh}
+\end{align}
+is the wavenumber mesh for $V_N$. The other two wavenumber meshes $\D{K}_N$ and 
+$\N{K}_N$ are both equal to $K_N$ except from in index $l$ that 
 ranges $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
@@
 
-The computational mesh has in real space been created using $N = (N_x, N_y, 
-N_z)$ 
-intervals, where the two periodic directions use uniform intervals. The 
+The computational mesh in real space can be created using $N = (N_x, N_y, 
+N_z)$ intervals, where the two periodic directions use uniform intervals. The 
 computational mesh is given as
@@
 Here $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
-transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represents inverse 
+transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represent inverse 
 Fourier transforms along directions $y$ and $z$ respectively, and the inverse 
@@
            \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
-           &= h \sum_{p=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t), 
+           &= h \sum_{q=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t), 
            \label{eq:sst2}
@@
 \end{align}
-The convective terms are computed with 
-a forward Adams-Bashforth projection, such that $\N{\mathcal{S}}(h_v^{n+1/2}) = 
-\N{\mathcal{S}}(1.5h_v^{n} - 0.5 h_v^{n-1})$ and $\D{\mathcal{S}}(h_g^{n+1/2}) 
-= \D{\mathcal{S}}(1.5h_g^{n} - 0.5 h_g^{n-1})$ and the coefficient matrices are 
-given as
+The coefficient matrices are given as
 \begin{align}
@@
 \end{align}
-where $\underline{z}^2 = \underline{m}^2 + \underline{n}^2$. Equations 
+where $\underline{z}^2 = \underline{m}^2 + \underline{n}^2$. The convective 
+terms are computed with a forward Adams-Bashforth projection, such that 
+$\N{\mathcal{S}}(h_v^{n+1/2}) = 
+\N{\mathcal{S}}(1.5h_v^{n} - 0.5 h_v^{n-1})$ and $\D{\mathcal{S}}(h_g^{n+1/2}) 
+= \D{\mathcal{S}}(1.5h_g^{n} - 0.5 h_g^{n-1})$. Equations 
 (\ref{eq:v0}) and (\ref{eq:w0}) are also discretized in time using 
@@
 nonzero diagonals at $ j = k-4, k-2, k, k+2, k+4$. The nonzero elements of the 
-$\N{Q}_{kj}$ matrix are given as
+upper triangular $\N{Q}_{kj}$ matrix are given as
 \begin{align}
@@
 \begin{align}
-p_k &= 8 k (k+1)(k+2)(k+4)\pi, &q_j = \frac{1}{j+3} \label{eq:pk} \\
-r_k &= 24(k+1)(k+2)\pi, &s_j =  \frac{(j+2)^2}{(j+3)} \label{eq:rk}
+p_k &= 8 k (k+1)(k+2)(k+4)\pi, &q_j = \frac{1}{j+3}, \label{eq:pk} \\
+r_k &= 24(k+1)(k+2)\pi, &s_j =  \frac{(j+2)^2}{(j+3)}. \label{eq:rk}
 \end{align} 
@@
 one forward elimination and a back solve. The back 
-solve is very fast because of the recursive formulation (\ref{eq:ab}), leading 
-to a formula like (valid for $k \le N_x-8$)
+solve is very fast because of the formulation (\ref{eq:ab}), leading 
+to a formula for the backwards substitution like (valid for $k \le N_x-8$)
 \begin{equation}
@@
 	is using dense storage of $U_{kj}$, but no more than the 3 diagonals need 
-	to be stored.}
+	to be stored eventually.}
 	\label{alg:lu_biharmonic}

### 2016-09-20 paper.tex
Added MHD
Added a new section containing MHD discretization, and extended the appendix by including matrices that are used in the MHD section.
Some speed results

@@
 \underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l 
-\hat{u}(l,m,n)\N{\phi}_l(x_i)}_{\N{\mathcal{S}}_x^{-1}} e^{\imath \underline{m} 
+\hat{u}(l,m,n,t)\N{\phi}_l(x_i)}_{\N{\mathcal{S}}_x^{-1}} e^{\imath 
+\underline{m} 
 y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , 
@@
 Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, 
-\ref{eq:phiN}). The transforms are slightly different for the three bases and 
-${\mathcal{S}}, \D{\mathcal{S}}$ and $\N{\mathcal{S}}$ are used to distinguish 
-between them with obvious notation.
+\ref{eq:phiN}). The inverse transforms are slightly different for the three 
+bases and 
+${\mathcal{S}}^{-1}, \D{\mathcal{S}}^{-1}$ and $\N{\mathcal{S}}^{-1}$ are used 
+to distinguish between them with obvious notation.
 
@@
 \begin{align}
- \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, \notag \\
-   &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}\,dx_wdydz,
-\end{align}
-where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction, and we have used $dx_w = dx/\sqrt{1-x^2}$.
-In this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, where quadrature is employed for integration.
-As such the scalar product corresponds to sums that may be computed using fast transforms
+ \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, 
+\end{align}
+where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for 
+periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction. In 
+this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, 
+where quadrature is employed for integration. As such the scalar product 
+corresponds to sums that may be computed using fast transforms. Using 
+$\N{\psi}_{\bm{k}}$ for $v$ we obtain
 \begin{align}
@@
 $y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
-\N{\phi}_l)_w$ is the Shen scalar product in the $x$-direction. The weights, 
-$w_i$, required for the Shen transforms are 
+\N{\phi}_l)_w$ is used to represent the Shen scalar product in the 
+$x$-direction. The weights, $w_i$, required for the Shen transforms are 
 defined as
@@
 \end{equation}
-where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.
+where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.\footnote{Note that 
+$c_{N_x}=1$ for the continuous scalar product $L^2_w(\Omega)$ (see, e.g., Eq. 
+(2.7) of Shen2), but for the discrete $l^2_w(\Omega)$ it follows that 
+$c_{N_x}=2$ due to inexact quadrature for the highest wavenumber.}
 
@@
 \begin{align}
+%\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
+%\sum_{\bm{m}=(q,r,s) \in \N{K}_N} \hat{u}_{\bm{m}} \N{\psi}_{\bm{m}}, 
+%\N{\psi}_{\bm{k}} \right>_w, \notag \\
 \left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
-\sum_{\bm{m}=(q,r,s) \in \N{K}_N} \hat{u}_{\bm{m}} \N{\psi}_{\bm{m}}, 
-\N{\psi}_{\bm{k}} \right>_w, \notag \\
+\sum_{(q,r,s) \in \N{K}_N} \hat{u}(q,r,s,t) \N{\psi}(q,r,s), 
+\N{\psi}(l,m,n,t) \right>_w, \notag \\
            &= h \sum_{q=0}^{N_x-4} \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) 
@@
  \hat{u}_{\bm{k}}(t) &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
- \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:fstB}
+ \quad \forall \, \bm{k} \in \N{K}_N, \label{eq:fstB}
 \end{align}
@@
 and $g$. The biharmonic equation (\ref{eq:biharmonic}) is solved with boundary 
-conditions are $u(\pm 1) = u'(\pm 1) = 0$, where the Neumann conditions follow 
+conditions $u(\pm 1) = u'(\pm 1) = 0$, where the Neumann conditions follow 
 from the continuity equation. The boundary conditions for Eq. (\ref{eq:g}) are 
@@
 Dirichlet space $\D{V}_N$: Find 
-${v}, {w}$ in $\D{V}_N \times \D{V}_N$ such that
+${v} \in \D{V}_N$ and $w \in \D{V}_N$ such that
 \begin{align}
@@
 wavenumbers except $m=n=0$. For $m=n=0$ we solve instead the 
-corresponding momentum equations in $y$ and $z$ directions:
+corresponding transformed momentum equations in $y$ and $z$ directions:
 \begin{align}
@@
 Inserting for (\ref{eq:u_solx}) and a similar expansion for $g$ in $\D{V}_N$, 
-the required inner products become
+the inner products required to solve Eqs. (\ref{eq:u1}-\ref{eq:f1}) are
 \begin{align}
@@
 h\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q, \\
-\left<h_u, \N{\psi} \right>_w &= h\left[-(\underline{m}^2+\underline{n}^2) 
+\left<h_u, \N{\psi}_{\bm{k}} \right>_w &= 
+h\left[-(\underline{m}^2+\underline{n}^2) 
 \N{\mathcal{S}}(\mathcal{H}_x) - \imath \underline{m}\left(\D{\phi}_q^{'}, 
@@
 \\
-\left< h_g, \D{\psi} \right>_w &= h\left[ \imath \underline{m}\, \D{B} 
+\left< h_g, \D{\psi}_{\bm{k}} \right>_w &= h\left[ \imath \underline{m}\, \D{B} 
 \hat{\mathcal{H}}_z - \imath \underline{n}\, \D{B} \hat{\mathcal{H}}_{y} 
@@
 \underline{z}^2 \right) \N{A}_{lq} - \frac{2\underline{z}^2 + \nu \triangle t 
-\underline{z}^4}{2} \N{B}_{lq} &\forall m, n \in K_N^p, \\
+\underline{z}^4}{2} \N{B}_{lq} &\forall m, n \in K_N^p, 
+\label{eq:Biharmonic_matrix} \\
 \D{H}_{lq}(m, n) &= -\frac{\nu \triangle 
 t}{2}\D{A}_{lq} + (1 + \frac{\nu \triangle t \underline{z}^2}{2}) \D{B}_{lq} 
-&\forall m, n \in K_N^p,
+&\forall m, n \in K_N^p, \label{eq:Helmholtz_matrix}
 \end{align}
@@
 \section{Implementation}
-
+\label{sec:implementation}
 The solution procedure for the numerical method described in 
@@
 
-We will now describe very efficient direct solvers for Eqs.~(\ref{eq:ufin}, 
-\ref{eq:gfin}) that are of order $O(N)$. The solvers are generated using clever 
-direct LU decomposition and only a few diagonals need to be stored for each 
-final lower (L) and upper (U) triangular matrix.
-
-To solve Eq. (\ref{eq:gfin}) we perform an LU decomposition of the coefficient 
-matrix $\D{H}$. $\D{H}$ consists of one subdiagonal, and every 
-second upper diagonal is zero. As such, we may split the matrix into two 
-smaller (and decoupled) matrices for odd and even coefficients $\D{H}^e_{k,j} = 
-\D{H}_{2k,2j}$ and $\D{H}^o_{k,j} = \D{H}_{2k+1,2j+1}$, where the two matrices 
-$\D{H}^{e,o}$ are of type upper 
-Hessenberg, with only 4 distinct values on each row. This structure of $\D{H}$ 
-allows us to perform a very efficient tailored LU decomposition, as shown in 
-Algorithm (\ref{alg:lu}), and a very efficient tailored solve through 
-Algorithm (\ref{alg:lusolve}). The efficiency follows from the fact that 
-$U^{o,e}_{kj} = U^{o,e}_{kk+2} \, \forall \, j > k+2$, i.e., the upper 
-triangular matrix has a maximum of three distinct values for each row, which 
-allows for a very fast solve through back substitution. 
-Regarding Algorithm 
-(\ref{alg:lusolve}), we note that the ability to decouple the system into odd 
-and even coefficients leads to two subsystem that may be trivially solved 
-simultaneously in two threads by letting one thread call LUSolveoddeven with 
-even coefficients, and the second thread call it for odd. For optimal 
+Shen \cite{shen1995} states that it is possible to solve Eq.~(\ref{eq:gfin}) 
+with 
+essentially the same number of operations as a pentadiagonal solver. We note 
+that the matrix 
+$\D{H}$ is of type upper Hessenberg and a solution
+procedure is simply to perform a direct LU decomposition (se, e.g., 
+\cite{stewart98}), which leads to a 
+lower triangular  
+matrix $L$ with only one subdiagonal and an upper triangular matrix $U$ which 
+is dense. However, each row in $U$ 
+contains at most three distinct values at $U_{k,k}, U_{k,k+2}$ and $U_{k,k+4}$, 
+and then $U_{k,j} = U_{k,k+4}\, \forall \, j = k+6, k+8, \ldots N$. 
+Consequently 
+only three diagonals in $U$ need storage and the back solve can be traversed 
+very efficiently in $O(N)$. Note also that the system could be decoupled into 
+odd and even coefficients, which leads to two subsystem that may be trivially 
+solved 
+simultaneously in two threads. For optimal 
 performance, the odd and even coefficient would then need to be stored 
@@
 
-\begin{algorithm}
-\caption{LU factorization of matrix $H_{k,j}$, that may be split into two 
-decoupled upper Hessenberg matrices $H^e_{k,j} = H_{2k,2j}$ and $H^o_{k,j} = 
-H_{2k+1,2j+1}$ for even and odd coefficients. Return LU factorization for even 
-and odd coefficients, such that $L^eU^e=H^e$ and $L^oU^o=H^o$. Note that 
-$L^{e}$ and $L^{o}$ each have one single nonzero diagonal in addition to the 
-unity main diagonal, and as such only a vector is stored. Each $U^{e}$ and 
-$U^{o}$ has three distinct diagonals and $U^{e}_{kj}=U^{e}_{k, k+2} \,\forall 
-j 
-> k+2, j< M^e$ and $U^{o}_{kj}=U^{o}_{k, k+2} \,\forall j > k+2, j< M^o$. As 
-such, a sparse three-diagonal storage may be used for $U^e$ and $U^o$.}
-\label{alg:lu}
-\begin{algorithmic}[1]
-\Procedure{LUoddeven}{}
-  \State \textbf{Input:} $\{H\}_{k,j=0}^{N-2}$
-  \State \textit{integer} $M^e \gets (N-2)/2$
-  \State \textit{integer} $M^o \gets (N-3)/2$
-  \State ${U}^e_{0j} \gets {H}_{0,2j}, \, \forall j=0,1,2 $
-  \State ${U}^o_{0j} \gets {H}_{1,2j+1}, \, \forall j=0,1,2 $
-  \For{$k=1, 2,  \ldots, M^e-2$}
-    \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
-    \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall 
-j=k, k+1, k+2$
-    \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
-    \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, 
-\forall j=k, k+1, k+2$
-  \EndFor
-
-  \State $k \gets M^e-1$
-  \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
-  \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall 
-j=k, k+1$
-  \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
-  \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, 
-\forall j=k, k+1$
-  \State $k \gets M^e$
-  \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
-  \State ${U}^e_{kk} \gets {H}_{2k,2k} - {L}^e_{k}{U}^e_{k-1,k}$
-  \If{$M^o == M^e$ }
-    \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
-    \State ${U}^o_{kk} \gets {H}_{2k+1,2k+1} - {L}^o_{k}{U}^o_{k-1,k}$
-
-  \EndIf  
-  \State \Return $\{L^e_k\}_{k=0}^{M^e},\{U^e\}_{k,j=0}^{M^e}, 
-\{L^o_k\}_{k=0}^{M^o},\{U^o\}_{k,j=0}^{M^o}$
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{Solve of $H_{kj}u_j = b_k$, where $H$ has previously been LU 
-factorized for odd even coefficients using Algorithm (\ref{alg:lu}).}
-\label{alg:lusolve}
-\begin{algorithmic}[1]
-\Procedure{LUSolveoddeven}{}
-  \State \textbf{Input:} $\{U_{kj}\}_{k,j=0}^{M}, \{L_k\}_{k=0}^{M}, 
-\{b_k\}_{k=0}^{M}$ 
-  \State // Solve $Ly=b$ by forward elimination 
-  \State $y_0 \gets b_0$
-  \For{$k = 1$ \textbf{to} $M$}
-    \State $y_k \gets b_k - L_ky_{k-1}$
-  \EndFor
-  \State // Solve $Uu=y$ with back substitution
-  \State double $s = 0.0$     
-  \State $u_M = y_M / U_{MM}$    
-  \For{$i = M-1$ \textbf{to} $0$ \textbf{step} $-1$}
-     \State $u_i \gets y_i - U_{i, i+1}\cdot u_{i+1}$
-     \If{$i < M-1$}
-       \State $s \gets s + u_{i+2}$
-       \State $u_i \gets u_i - s\cdot U_{i,i+2}$            
-     \EndIf
-     \State $u_i \gets  u_i / U_{i,i}$
-   \EndFor
-  \State \Return $\{u_k\}_{k=0}^{M}$  
-\EndProcedure
-\end{algorithmic}
-
-\begin{algorithmic}[1]
-\Procedure{LUSolve}{}
-  \State \textbf{Input: $\{u_k\}_{k=0}^{N-2}, \{b_k\}_{k=0}^{N-2}, 
-\{U_{kj}^{e}\}_{k,j=0}^{M^{e}}, \{U_{kj}^{o}\}_{k,j=0}^{M^{o}}, 
-\{L^{e}_k\}_{k=0}^{M^{e}}, \{L^{o}_k\}_{k=0}^{M^{o}}$}
-  \State $\{u_{2k}\}_{k=0}^{M^e} \gets $ LUSolveoddeven($U^{e}, L^e, 
-\{b_{2k}\}_{k=0}^{M^e}$)
-  \State $\{u_{2k+1}\}_{k=0}^{M^o} \gets $ LUSolveoddeven($U^{o}, L^o, 
-\{b_{2k+1}\}_{k=0}^{M^o}$)
-  \State \Return $\{u_k\}_{k=0}^{N-2}$
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-The biharmonic operator $\N{H}$ is more challenging to solve efficiently, but 
+%We will now describe very efficient direct solvers for Eqs.~(\ref{eq:ufin}, 
+%\ref{eq:gfin}) that are of order $O(N)$. The solvers are generated using 
+%clever 
+%direct LU decomposition and only a few diagonals need to be stored for each 
+%final lower (L) and upper (U) triangular matrix.
+
+%To solve Eq. (\ref{eq:gfin}) we perform an LU decomposition of the coefficient 
+%matrix $\D{H}$. $\D{H}$ consists of one subdiagonal, and every 
+%second upper diagonal is zero. As such, we may split the matrix into two 
+%smaller (and decoupled) matrices for odd and even coefficients $\D{H}^e_{k,j} 
+%= 
+%\D{H}_{2k,2j}$ and $\D{H}^o_{k,j} = \D{H}_{2k+1,2j+1}$, where the two matrices 
+%$\D{H}^{e,o}$ are of type upper 
+%Hessenberg, with only 4 distinct values on each row. This structure of $\D{H}$ 
+%allows us to perform a very efficient tailored LU decomposition, as shown in 
+%Algorithm (\ref{alg:lu}), and a very efficient tailored solve through 
+%Algorithm (\ref{alg:lusolve}). The efficiency follows from the fact that 
+%$U^{o,e}_{kj} = U^{o,e}_{kk+2} \, \forall \, j > k+2$, i.e., the upper 
+%triangular matrix has a maximum of three distinct values for each row, which 
+%allows for a very fast solve through back substitution. 
+%Regarding Algorithm 
+%(\ref{alg:lusolve}), we note that the ability to decouple the system into odd 
+%and even coefficients leads to two subsystem that may be trivially solved 
+%simultaneously in two threads by letting one thread call LUSolveoddeven with 
+%even coefficients, and the second thread call it for odd. For optimal 
+%performance, the odd and even coefficient would then need to be stored 
+%contiguously in computer memory, and not alternately, which would be the 
+%normal 
+%way of storing the coefficients.
+
+%\begin{algorithm}
+%\caption{LU factorization of matrix $H_{k,j}$, that may be split into two 
+%decoupled upper Hessenberg matrices $H^e_{k,j} = H_{2k,2j}$ and $H^o_{k,j} = 
+%H_{2k+1,2j+1}$ for even and odd coefficients. Return LU factorization for even 
+%and odd coefficients, such that $L^eU^e=H^e$ and $L^oU^o=H^o$. Note that 
+%$L^{e}$ and $L^{o}$ each have one single nonzero diagonal in addition to the 
+%unity main diagonal, and as such only a vector is stored. Each $U^{e}$ and 
+%$U^{o}$ has three distinct diagonals and $U^{e}_{kj}=U^{e}_{k, k+2} \,\forall 
+%j 
+%> k+2, j< M^e$ and $U^{o}_{kj}=U^{o}_{k, k+2} \,\forall j > k+2, j< M^o$. As 
+%such, a sparse three-diagonal storage may be used for $U^e$ and $U^o$.}
+%\label{alg:lu}
+%\begin{algorithmic}[1]
+%\Procedure{LUoddeven}{}
+%  \State \textbf{Input:} $\{H\}_{k,j=0}^{N-2}$
+%  \State \textit{integer} $M^e \gets (N-2)/2$
+%  \State \textit{integer} $M^o \gets (N-3)/2$
+%  \State ${U}^e_{0j} \gets {H}_{0,2j}, \, \forall j=0,1,2 $
+%  \State ${U}^o_{0j} \gets {H}_{1,2j+1}, \, \forall j=0,1,2 $
+%  \For{$k=1, 2,  \ldots, M^e-2$}
+%    \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
+%    \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall 
+%j=k, k+1, k+2$
+%    \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
+%    \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, 
+%\forall j=k, k+1, k+2$
+%  \EndFor
+%  
+%  \State $k \gets M^e-1$
+%  \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
+%  \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall 
+%j=k, k+1$
+%  \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
+%  \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, 
+%\forall j=k, k+1$
+%  \State $k \gets M^e$
+%  \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
+%  \State ${U}^e_{kk} \gets {H}_{2k,2k} - {L}^e_{k}{U}^e_{k-1,k}$
+%  \If{$M^o == M^e$ }
+%    \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
+%    \State ${U}^o_{kk} \gets {H}_{2k+1,2k+1} - {L}^o_{k}{U}^o_{k-1,k}$
+%
+%  \EndIf  
+%  \State \Return $\{L^e_k\}_{k=0}^{M^e},\{U^e\}_{k,j=0}^{M^e}, 
+%\{L^o_k\}_{k=0}^{M^o},\{U^o\}_{k,j=0}^{M^o}$
+%\EndProcedure
+%\end{algorithmic}
+%\end{algorithm}
+%
+%\begin{algorithm}
+%\caption{Solve of $H_{kj}u_j = b_k$, where $H$ has previously been LU 
+%factorized for odd even coefficients using Algorithm (\ref{alg:lu}).}
+%\label{alg:lusolve}
+%\begin{algorithmic}[1]
+%\Procedure{LUSolveoddeven}{}
+%  \State \textbf{Input:} $\{U_{kj}\}_{k,j=0}^{M}, \{L_k\}_{k=0}^{M}, 
+%\{b_k\}_{k=0}^{M}$ 
+%  \State // Solve $Ly=b$ by forward elimination 
+%  \State $y_0 \gets b_0$
+%  \For{$k = 1$ \textbf{to} $M$}
+%    \State $y_k \gets b_k - L_ky_{k-1}$
+%  \EndFor
+%  \State // Solve $Uu=y$ with back substitution
+%  \State double $s = 0.0$     
+%  \State $u_M = y_M / U_{MM}$    
+%  \For{$i = M-1$ \textbf{to} $0$ \textbf{step} $-1$}
+%     \State $u_i \gets y_i - U_{i, i+1}\cdot u_{i+1}$
+%     \If{$i < M-1$}
+%       \State $s \gets s + u_{i+2}$
+%       \State $u_i \gets u_i - s\cdot U_{i,i+2}$            
+%     \EndIf
+%     \State $u_i \gets  u_i / U_{i,i}$
+%   \EndFor
+%  \State \Return $\{u_k\}_{k=0}^{M}$  
+%\EndProcedure
+%\end{algorithmic}
+%
+%\begin{algorithmic}[1]
+%\Procedure{LUSolve}{}
+%  \State \textbf{Input: $\{u_k\}_{k=0}^{N-2}, \{b_k\}_{k=0}^{N-2}, 
+%\{U_{kj}^{e}\}_{k,j=0}^{M^{e}}, \{U_{kj}^{o}\}_{k,j=0}^{M^{o}}, 
+%\{L^{e}_k\}_{k=0}^{M^{e}}, \{L^{o}_k\}_{k=0}^{M^{o}}$}
+%  \State $\{u_{2k}\}_{k=0}^{M^e} \gets $ LUSolveoddeven($U^{e}, L^e, 
+%\{b_{2k}\}_{k=0}^{M^e}$)
+%  \State $\{u_{2k+1}\}_{k=0}^{M^o} \gets $ LUSolveoddeven($U^{o}, L^o, 
+%\{b_{2k+1}\}_{k=0}^{M^o}$)
+%  \State \Return $\{u_k\}_{k=0}^{N-2}$
+%\EndProcedure
+%\end{algorithmic}
+%\end{algorithm}
+
+The biharmonic matrix $\N{H}$ in Eq.~(\ref{eq:Biharmonic_matrix}) is more 
+challenging to solve for efficiently, but 
 it is still possible to find an algorithm that is $O(N)$ for a system of $N$ 
@@
 \begin{align}
- \N{Q}_{kk} &= -4(k+1)(k+2)^2 \\
- \N{Q}_{kj} &= p_kq_j + r_ks_j, \quad j = k+2, k+4, \ldots N_x-4
+ \N{Q}_{kk} &= -4(k+1)(k+2)^2, \\
+ \N{Q}_{kj} &= p_kq_j + r_ks_j, \quad j = k+2, k+4, \ldots, N_x-4,
 \end{align}
@@
 \begin{align}
-p_k &= 8 k (k+1)(k+2)(k+4)\pi, &q_j = \frac{1}{j+3}, \label{eq:pk} \\
-r_k &= 24(k+1)(k+2)\pi, &s_j =  \frac{(j+2)^2}{(j+3)}. \label{eq:rk}
+p_k &= 8 k (k+1)(k+2)(k+4)\pi, &q_j &= \frac{1}{j+3}, \label{eq:pk} \\
+r_k &= 24(k+1)(k+2)\pi, &s_j &=  \frac{(j+2)^2}{(j+3)}. \label{eq:rk}
 \end{align} 
 For each row in $\N{H}_{kj}$ we have that $\N{H}_{kj}=c_0\N{Q}_{kj}\, \forall j 
-> k+4$. A direct LU decomposition of $\N{H}$ can be performed as shown in Alg. 
+> k+4$. A straight forward direct LU decomposition of $\N{H}$ can be performed 
+as shown in Alg. 
 \ref{alg:lu_biharmonic}, which leads to a lower triangular 
@@
 unity 
-main diagonal. The upper triangular matrix $\N{U}_{kj}$ contains three distinct 
-diagonals at $j=k, k+2$ and $k+4$. The remaining part of $\N{U}_{kj}$ can be 
-written as
+main diagonal. The upper triangular matrix $\N{U}_{kj}$ is dense and as such a 
+show stopper for an efficient back solve. However, we note that $\N{U}_{kj}$ 
+contains only three distinct 
+diagonals at $j=k, k+2$ and $k+4$, whereas the remaining part of $\N{U}_{kj}$ 
+can be expressed as
 \begin{equation}
-\N{U}_{kj} = a_k q_j + b_k s_j, \quad j = k+6, k+8, \ldots, N_x-4 \label{eq:ab}
+\N{U}_{kj} = c_0(a_k q_j + b_k s_j), \quad j = k+6, k+8, \ldots, N_x-4, 
+\label{eq:ab}
 \end{equation}
 where $\bm{a}$ and $\bm{b}$ are two new vectors that can be computed 
-recursively as shown in Alg. (\ref{alg:ab}). If $\bm{a}$ and $\bm{b}$ are 
-pre-computed 
+recursively from $\N{L}$ 
+\begin{align}
+a_k &= p_k, \, &\forall \, k= 0, 1, \notag \\
+b_k &= r_k, \, &\forall \, k= 0, 1, \notag\\
+a_k &= p_k - \N{L}_{k,k-2}a_{k-2}, \, &\forall \, k = 2, 3, \notag\\
+b_k &= r_k - \N{L}_{k,k-2}b_{k-2}, \, &\forall \, k = 2, 3, \notag\\
+a_k &= p_k - \N{L}_{k,k-2}a_{k-2} - \N{L}_{k,k-4}a_{k-4}, \, &\forall \, k = 4, 
+5, 
+\ldots, N, \notag\\
+b_k &= r_k - \N{L}_{k,k-2}b_{k-2} - \N{L}_{k,k-4}b_{k-4}, \, &\forall \, k = 4, 
+5, 
+\ldots, N.
+\end{align}
+If $\bm{a}$ and $\bm{b}$ 
+are pre-computed 
 the total storage requirement for the complete LU decomposition is no more than 
-$7N$ and the solution of the complete system can be obtained very quickly with 
-one forward elimination and a back solve. The back 
-solve is very fast because of the formulation (\ref{eq:ab}), leading 
+$7N$, i.e., two diagonals for $\N{L}$, three diagonals for $\N{U}$ plus 
+$\bm{a}$ and 
+$\bm{b}$. The solution of the complete system can be obtained very quickly with 
+one simple forward elimination and a back solve, where the back 
+solve is very fast ($O(N)$) because of the formulation (\ref{eq:ab}), leading 
 to a formula for the backwards substitution like (valid for $k \le N_x-8$)
@@
 u_k = \left(y_k - \N{U}_{k, k+2} u_{k+2} - \N{U}_{k, k+4} u_{k+4} - 
-a_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} q_j u_j - 
-b_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} 
+c_0 a_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} q_j u_j - 
+c_0 b_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} 
 s_j u_j\right)/\N{U}_{kk},
@@
 	\caption{LU decomposition of biharmonic operator $\N{H}$. Current algorithm 
-	is using dense storage of $U_{kj}$, but no more than the 3 diagonals need 
+	is using dense storage of $\N{U}_{kj}$, but no more than the 3 diagonals 
+	need 
 	to be stored eventually.}
@@
 			\State \textbf{Input:} $\{\N{H}_{kj}\}_{k, j = 0}^{N}$
-	        \State $U_{0j} = \N{H}_{0j} \quad \forall \, j \in 0, 2, \ldots, N$
-	        \State $U_{1j} = \N{H}_{1j} \quad \forall \, j \in 1, 3, \ldots, N$
-	        \State $L_{20} = \N{H}_{20}/U_{00}$
-	        \State $L_{31} = \N{H}_{31}/U_{11}$
-	        \State $U_{2j} = L_{20} U_{0j} \quad \forall \, j \in 2, 4, \ldots, 
-	        N$
-	        \State $U_{3j} = L_{31} U_{1j} \quad \forall \, j \in 3, 5, \ldots, 
-	        N$
+	        \State $\N{U}_{0j} = \N{H}_{0j} \quad \forall \, j \in 0, 2, 
+	        \ldots, N$
+	        \State $\N{U}_{1j} = \N{H}_{1j} \quad \forall \, j \in 1, 3, 
+	        \ldots, N$
+	        \State $\N{L}_{20} = \N{H}_{20}/\N{U}_{00}$
+	        \State $\N{L}_{31} = \N{H}_{31}/\N{U}_{11}$
+	        \State $\N{U}_{2j} = \N{H}_{2j} - \N{L}_{20} \N{U}_{0j} \quad 
+	        \forall \, j \in 2, 4, \ldots, N$
+	        \State $\N{U}_{3j} = \N{H}_{3j} - \N{L}_{31} \N{U}_{1j} \quad 
+	        \forall \, j \in 
+	        3, 5, \ldots, N$
 	        \For{$k = 4, 5, \ldots, N$}
-	            \State $L_{k,k-2} = {U}_{k,k-2}/{U}_{k-2, k-2}$
-		        \State $U_{kj} = {U}_{kj} - L_{k, k-2} U_{k-2, j} \quad 
+		        \State $\N{L}_{k,k-4} = \N{H}_{k, k-4} / \N{U}_{k-4, k-4}$
+		        \State $\N{U}_{k,j} = \N{H}_{k,j} - \N{L}_{k, k-4} 
+		        \N{U}_{k-4, j} 
+		        \quad \forall j = k-2, k, \ldots, N$
+	            \State $\N{L}_{k,k-2} = {\N{U}}_{k,k-2}/{\N{U}}_{k-2, k-2}$
+		        \State $\N{U}_{kj} = {\N{U}}_{kj} - \N{L}_{k, k-2} \N{U}_{k-2, 
+		        j} \quad 
 		        \forall j = k, k+2, \ldots, N$
-		        \If{$k <= N-2$}
-		        \State $L_{k+2,k-2} = \N{H}_{k+2, k-2} / U_{k-2, k-2}$
-		        \State $U_{k+2,j} = \N{H}_{k+2,j} - L_{k+2, k-2} U_{k-2, j} 
-		        \quad \forall j = k-2, k, \ldots, N$
-		        \EndIf
 	        \EndFor
-	        \State \Return $\{L_{kj}\}_{k, j = 0}^{N}$, $\{U_{kj}\}_{k, j = 
+	        \State \Return $\{\N{L}_{kj}\}_{k, j = 0}^{N}$, $\{\N{U}_{kj}\}_{k, 
+	        j = 
 	        0}^{N}$
@@
 
+%\begin{algorithm}
+%	\caption{Recursive formula to compute $\{a_k\}_{k=0}^{N_x-4}, 
+%	\{b_k\}_{k=0}^{N_x-4}$ in Eq. (\ref{eq:ab}). Note that the lower triangular 
+%	matrix $\{L_{kj}\}_{k,j=0}^{N_x-4}$ contains only two nonzero diagonals and 
+%	the parameters $p_k$ and $r_k$ are given in Eqs. (\ref{eq:pk}) and 
+%	(\ref{eq:rk}) respectively. }
+%    \label{alg:ab}
+%	\begin{algorithmic}[1]
+%		\Procedure{ReduceAB}{}
+%		\State \textbf{Input:} $ \{\N{L}_{kj}\}_{k,j=0}^{N_x-4}, c_0$
+%		\For{$k=0,1$}
+%		     \State $a_k \gets p_k$
+%		     \State $b_k \gets r_k$
+%		\EndFor
+%		\For{$k=2,3$}
+%		    \State $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2}$
+%		    \State $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}$
+%		\EndFor
+%		\For{$k=4, 5, \ldots, N_x-4$}
+%     		\State $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2} - \N{L}_{k, k-4} 
+%\,a_{k-4}$
+%	    	\State $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}- \N{L}_{k, k-4} 
+%\,b_{k-4}$
+%		\EndFor
+%		\For{$k=0, 1, \ldots, N_x-4$}
+%		    \State $a_k \gets a_k\cdot c_0$
+%		    \State $b_k \gets b_k \cdot c_0$
+%		\EndFor
+%		\State \Return $\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$
+%        \EndProcedure
+%\end{algorithmic}
+%\end{algorithm}
+
 \begin{algorithm}
-	\caption{Recursive formula to compute $\{a_k\}_{k=0}^{N_x-4}, 
-	\{b_k\}_{k=0}^{N_x-4}$ in Eq. (\ref{eq:ab}). Note that the lower triangular 
-	matrix $\{L_{kj}\}_{k,j=0}^{N_x-4}$ contains only two nonzero diagonals and 
-	the parameters $p_k$ and $r_k$ are given in Eqs. (\ref{eq:pk}) and 
-	(\ref{eq:rk}) respectively. }
-    \label{alg:ab}
-	\begin{algorithmic}[1]
-		\Procedure{ReduceAB}{}
-		\State \textbf{Input:} $ \{L_{kj}\}_{k,j=0}^{N_x-4}$
-		\For{$k=0,1$}
-		     \State $a_k = p_k$
-		     \State $b_k = r_k$
-		\EndFor
-		\For{$k=2,3$}
-		    \State $a_k = p_k - L_{k,k-2}\,a_{k-2}$
-		    \State $b_k = r_k - L_{k, k-2}\,b_{k-2}$
-		\EndFor
-		\For{$k=4, 5, \ldots, N_x-4$}
-     		\State $a_k = p_k - L_{k,k-2}\,a_{k-2} - L_{k, k-4} \,a_{k-4}$
-	    	\State $b_k = r_k - L_{k, k-2}\,b_{k-2}- L_{k, k-4} \,b_{k-4}$
-		\EndFor
-		\State \Return $\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$
-        \EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-	\caption{Solve biharmonic Eq. (\ref{eq:ufin}) with pre-computed $L$, $U$ 
+	\caption{Solve biharmonic Eq. (\ref{eq:ufin}) with pre-computed $\N{L}$, 
+	$\N{U}$ 
 	matrices. The parameters $q_j$ and $s_j$ are given in Eqs. (\ref{eq:pk}) 
@@
 		\Procedure{LUSolveBiharmonic}{}
-		  \State \textbf{Input:} $\{U_{kj}\}_{k,j=0}^{N}, 
-		  \{L_{kj}\}_{k,j=0}^{N}, 
+		  \State \textbf{Input:} $\{\N{U}_{kj}\}_{k,j=0}^{N}, 
+		  \{\N{L}_{kj}\}_{k,j=0}^{N}, 
 		  \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N}, \{b_k\}_{k=0}^{N}$ 
-		  \State // Solve $Ly=f$ by forward elimination 
+		  \State // Solve $\N{L}_{kj}y_j=f_k$ by forward elimination 
 		  \State $y_0 \gets f_0$
 		  \State $y_1 \gets f_1$
-		  \State $y_2 \gets f_2 - L_{20}y_{0}$
-		  \State $y_3 \gets f_3 - L_{31}y_{1}$
+		  \State $y_2 \gets f_2 - \N{L}_{20}y_{0}$
+		  \State $y_3 \gets f_3 - \N{L}_{31}y_{1}$
 		  \For{$k = 4$ \textbf{to} $N$}
-		  \State $y_k \gets f_k - L_{k,k-2}y_{k-2} - L_{k, k-4}y_{k-4}$
+		  \State $y_k \gets f_k - \N{L}_{k,k-2}y_{k-2} - \N{L}_{k, k-4}y_{k-4}$
 		  \EndFor
-		  \State // Solve $Uu=y$ with back substitution                    
-		  \State $u_N = y_N / U_{NN}$
-	      \State $u_{N-1} = y_{N-1} / U_{N-1,N-1}$
-	      \State $u_{N-2} = \left(y_{N-2} - U_{N-2, N}y_N\right)/U_{N-2,N-2}$
-	      \State $u_{N-3} = \left(y_{N-3} - U_{N-3, 
-	      N-1}y_{N-1}\right)/U_{N-3,N-3}$
-
-	      \State $q^o = q^e = s^o = s^e = 0.0$
+		  \State // Solve $\N{U}_{kj}u_j=y_k$ with back 
+		  substitution                    
+		  \State $u_N = y_N / \N{U}_{NN}$
+	      \State $u_{N-1} = y_{N-1} / \N{U}_{N-1,N-1}$
+	      \State $u_{N-2} = \left(y_{N-2} - \N{U}_{N-2, 
+	      N}y_N\right)/\N{U}_{N-2,N-2}$
+	      \State $u_{N-3} = \left(y_{N-3} - \N{U}_{N-3, 
+	      N-1}y_{N-1}\right)/\N{U}_{N-3,N-3}$
+
+	      \State $q^o = q^e = s^o = s^e = 0$
 		  \For{$k = N-4$ \textbf{to} $0$ \textbf{step} $-1$}
-		  \State $u_k \gets y_k - U_{k, k+2} u_{k+2} - U_{k, k+4} 
+		  \State $u_k \gets y_k - \N{U}_{k, k+2} u_{k+2} - \N{U}_{k, k+4} 
 		  u_{k+4} $ 
@@
 		    \State $s^o \gets s^o + u_{j}s_{j}$
-		    \State $u_k \gets u_k - c_0 q^o a_k - c_0 s^o b_k $	    
+		    \State $u_k \gets u_k - q^o a_k - s^o b_k $	    
 		  \Else
@@
 		    \State $s^e \gets s^e + u_{j}s_{j}$	  
-		    \State $u_k \gets u_k - c_0 q^e a_k - c_0 s^e b_k $
+		    \State $u_k \gets u_k - q^e a_k - s^e b_k $
 		  \EndIf
 		  \EndIf
-		  \State $u_k \gets  u_k / U_{kk}$
+		  \State $u_k \gets  u_k / \N{U}_{kk}$
 		  \EndFor
@@
 \end{algorithm}
+
+\section{Verification}
+We have described a Navier-Stokes solver for channel flows that scales no worse 
+than the fast transforms used to go between spectral and real space. All the 
+matrices that make up the linear systems can be inverted using sparse direct LU 
+decompositions that scales like $O(N)$. To verify and to further show that we 
+can use these algorithms for large scale simulations, we will first study some 
+simple 1D problems.
+
+We first consider the roundoff error of the matrix decomposition methods 
+described in Sec.~\ref{sec:implementation}. The roundoff error may be 
+approximated 
+by the following approach: Let $\{u_i\}_{i=0}^{N-4}$ be a uniformly distributed 
+random vector. Compute the matrix vector product $\bm{f}=\N{H}\bm{u}$ and 
+then solve $\N{H} \bm{v} = \bm{f}$, using Alg. \ref{alg:SolveBiharmonic}, for 
+$\{v_i\}_{i=0}^{i=N-4}$. The roundoff error of the matrix decomposition 
+algorithm may then be found as $\max_{0 \ge i \ge N-4}(|u_i-v_i|)/\max_{0 \ge i 
+\ge N-4}|u_i|$. For this experiment we assume that $\triangle t=0.0001$ 
+and $\nu=1/1200$, that are reasonable parameters for a turbulent channel flow 
+simulation at $\text{Re}_{\tau}=1200$. Three different wavenumbers 
+$z=(2, 200, 5600)$ are 
+also chosen as representative for simulations that are using up to 8000 
+wavenumbers in each periodic direction, which is approximately the resolution 
+of the largest known channel simulations, performed by Lee and 
+Moser~\cite{leemoser15}. The mesh in the wall normal direction 
+is varied from 64 to 4096 points, which covers almost three times higher 
+resolution than used by Lee and Moser \cite{leemoser15}. Results shown in the 
+first three rows of Table \ref{tab:roundoff} indicate that 
+roundoff errors are not very pronounced, even for problems as large as 
+$N_x=4096$. Further, the roundoff errors increase only slowly with increasing 
+the wavenumbers. Corresponding results are also computed for the Helmholtz 
+matrix (\ref{eq:Helmholtz_matrix}) and 
+results are shown in the last three rows of Table \ref{tab:roundoff}. For 
+the Helmholtz matrix the roundoff errors are seen to be insignificant.
+
+\begin{table}
+	\caption{Roundoff error of the matrix decomposition method. The first three 
+	rows show results for Alg. (\ref{alg:SolveBiharmonic}) and matrix $\N{H}(z, 
+	\triangle t, \nu)$, whereas the last three rows show results for matrix 
+	$\D{H}(z, \triangle t, \nu)$. Parameters 
+	are chosen as $\nu=1/1200$ and $\triangle t=0.0001$ and the numbers shown 
+	are for an average of 100 runs. }
+	\label{tab:roundoff}
+\begin{tabular}{cccccccc}
+	   \multicolumn{8}{c}{$N_x$} \\
+	   \hline
+	$z$ & 64 & 128 & 256 & 512 & 1024 & 2048 & 4096 \\
+	\hline
+0 & 2.66e-15 & 6.71e-15 & 1.73e-14 & 4.64e-14 & 1.18e-13 & 3.30e-13 & 9.21e-13 
+\\
+200 & 8.39e-15 & 1.78e-14 & 5.12e-14 & 1.56e-13 & 4.47e-13 & 1.18e-12 & 
+3.45e-12 \\
+5600 & 7.09e-15 & 2.08e-14 & 5.65e-14 & 1.43e-13 & 4.38e-13 & 1.01e-12 & 
+2.61e-12 \\
+		\hline
+0 & 3.29e-15 & 6.07e-15 & 6.05e-15 & 6.50e-15 & 5.63e-15 & 5.43e-15 & 5.13e-15 
+\\
+200 & 3.03e-15 & 6.30e-15 & 6.14e-15 & 5.87e-15 & 5.67e-15 & 5.50e-15 & 
+5.39e-15 \\
+5600 & 3.35e-15 & 7.78e-15 & 8.86e-15 & 7.94e-15 & 7.18e-15 & 7.20e-15 & 
+7.03e-15
+
+\end{tabular}
+\end{table}
+
+To show that the solvers are scaling with size, we investigate the execution time 
+for solving $\N{H} \bm{v} = \bm{f}$ and $\D{H} \bm{v} = \bm{f}$ on a complete 
+wavenumber mesh of size $(N_x,64,64)$. The numbers have been produced on a MacBook Pro laptop with the 2.8 GHz intel Core i7 processor. The average time for one single 
+solve is shown in Table~\ref{tab:timings}, where the numbers in parenthesis are used to indicate scaling (unity would be perfectly $O(N)$). The scaling is seen to be close to linear, and variations 
+are most likely due to how well datastructures fit in the cache. Naturally, the timings are not 100$\%$ reproducible, but the ability to solve large scale problems is evident.
+\begin{table}
+	\centering
+	\caption{Execution time (in seconds) for one solve  of the Helmholtz 
+	(\ref{eq:gfin}) and Biharmonic 
+	(\ref{eq:ufin}) problems. The numbers in parenthesis are computed as $t_n / t_{n-1} / 2$, where $t_n$ is the current solvers timing in row $n$.}
+	\label{tab:timings}
+	\begin{tabular}{ccc}
+%		$N_x$ & Biharmonic & Helmholtz \\
+%		\hline
+%64 & 6.49e-07 (*) & 4.64e-07 (*) \\ 
+%128 & 1.40e-06 (1.08) & 9.43e-07 (1.02) \\ 
+%256 & 2.96e-06 (1.06) & 1.93e-06 (1.02) \\ 
+%512 & 6.09e-06 (1.03) & 4.01e-06 (1.04) \\ 
+%1024 & 1.49e-05 (1.22) & 9.83e-06 (1.23) \\ 
+%2048 & 2.53e-05 (0.85) & 1.64e-05 (0.83) \\ 
+%4096 & 6.15e-05 (1.22) & 4.09e-05 (1.25) \\ 
+%8192 & 1.25e-04 (1.02) & 8.18e-05 (1.00) 
+Nx & Biharmonic & Helmholtz \\ 
+\hline
+64 & 8.09e-07 (0.00) & 4.51e-07 (0.00) \\ 
+128 & 1.57e-06 (0.97) & 9.41e-07 (1.04) \\ 
+256 & 2.76e-06 (0.88) & 1.97e-06 (1.05) \\ 
+512 & 4.96e-06 (0.90) & 4.22e-06 (1.07) \\ 
+1024 & 1.18e-05 (1.18) & 8.40e-06 (1.00) \\ 
+2048 & 2.37e-05 (1.01) & 1.79e-05 (1.07) \\ 
+4096 & 4.82e-05 (1.02) & 3.78e-05 (1.05) \\ 
+8192 & 9.72e-05 (1.01) & 7.91e-05 (1.05)
+	\end{tabular}
+\end{table}
 
@@
 
+\bibliography{bib.bib}
+
 \section{Appendices}

### 2016-12-06 paper.tex
Huge commit

@@
-\documentclass[11pt, oneside]{article}
+\documentclass[11pt, oneside]{elsarticle}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{graphicx}
-\usepackage{ptex2tex, minted}
 \usepackage{xcolor}
@@
 
-\definecolor{gray}{gray}{0.97}
-\colorlet{commentcolour}{green!50!black}
-\colorlet{stringcolour}{red!60!black}
-\colorlet{keywordcolour}{magenta!90!black}
-\colorlet{exceptioncolour}{yellow!50!red}
-\colorlet{commandcolour}{blue!60!black}
-\colorlet{numpycolour}{blue!60!green}
-\colorlet{literatecolour}{magenta!90!black}
-\colorlet{promptcolour}{green!50!black}
-\colorlet{specmethodcolour}{violet}
-\colorlet{indendifiercolour}{green!70!white}
-
 %\newcommand{\codetitlestyle}[1]{\small\textit{#1}\hspace{0.1cm}}
@@
 
-\lstset{
-numbers=none,
-aboveskip=1ex,
-belowskip=1ex,
-basicstyle=\ttfamily\footnotesize,
-}
-
-\lstdefinestyle{pythonstyle}{
-%%\lstset{
-%%keepspaces=true,
-language=python,
-showtabs=true,
-tab=,
-tabsize=2,
-basicstyle=\ttfamily\footnotesize,%\setstretch{.5},
-stringstyle=\color{stringcolour},
-showstringspaces=false,
-alsoletter={1234567890},
-otherkeywords={\ , \}, \{, \%, \&, \|},
-keywordstyle=\color{keywordcolour}\bfseries,
-emph={and,break,class,continue,def,yield,del,elif ,else,%
-except,exec,finally,for,from,global,if,import,in,%
-lambda,not,or,pass,print,raise,return,try,while,assert},
-emphstyle=\color{blue}\bfseries,
-emph={[2]True, False, None},
-emphstyle=[2]\color{keywordcolour},
-emph={[3]object,type,isinstance,copy,deepcopy,zip,enumerate,reversed,list,len,dict,tuple,xrange,append,execfile,real,imag,reduce,str,repr},
-emphstyle=[3]\color{commandcolour},
-emph={Exception,NameError,IndexError,SyntaxError,TypeError,ValueError,OverflowError,ZeroDivisionError},
-emphstyle=\color{exceptioncolour}\bfseries,
-%upquote=true,
-morestring=[s]{"""}{"""},
-morestring=[s]{'''}{'''},
-commentstyle=\color{commentcolour}\slshape,
-emph={[4]1, 2, 3, 4, 5, 6, 7, 8, 9, 0,ode, fsolve, sqrt, exp, sin, cos, arccos, pi, array, norm, solve,float,complex, dot, arange, isscalar, max, sum, flatten, shape, reshape, find, any, all, abs, plot, linspace, legend, quad, polyval,polyfit, hstack,vector, concatenate,vstack,column_stack,empty,zeros,ones,rand,vander,grid,pcolor,eig,eigs,eigvals,svd,qr,tan,det,logspace,roll,min,mean,cumsum,cumprod,diff,vectorize,lstsq,cla,eye,xlabel,ylabel,squeeze},
-emphstyle=[4]\color{commandcolour},
-emph={[5]__init__,__add__,__mul__,__div__,__sub__,__call__,__getitem__,__setitem__,__eq__,__ne__,__nonzero__,__rmul__,__radd__,__repr__,__str__,__get__,__truediv__,__pow__,__name__,__future__,__all__},
-emphstyle=[5]\color{specmethodcolour},
-emph={[6]assert,range,yield},
-emphstyle=[6]\color{keywordcolour}\bfseries,
-emph={[7]def, return, and, print},
-emphstyle=[7]\color{black}\bfseries,
-% emph={[7]self},
-% emphstyle=[7]\bfseries,
-literate=*%
-%{:}{{\literatecolour:}}{1}%
-%{=}{{\literatecolour=}}{1}%
-%{-}{{\literatecolour-}}{1}%
-%{+}{{\literatecolour+}}{1}%
-%{*}{{\literatecolour*}}{1}%
-{/}{{\literatecolour/}}{1}%
-{!}{{\literatecolour!}}{1}%
-%{(}{{\literatecolour(}}{1}%
-%{)}{{\literatecolour)}}{1}%
-%{[}{{\literatecolour[}}{1}%
-%{]}{{\literatecolour]}}{1}%
-{<}{{\literatecolour<}}{1}%
-{>}{{\literatecolour>}}{1}%
-{>>>}{{\textcolor{promptcolour}{>>>}}}{1}%
-,%
-breaklines=true,
-breakatwhitespace= true,
-xleftmargin=\framemargin,
-xrightmargin=\framemargin,
-aboveskip=1ex,
-belowskip=1ex,
-frame=trbl, %trbl
-numbers=none,
-%frameround=tttt,
-rulecolor=\color{black!40},
-%framexleftmargin=\framemargin,
-%framextopmargin=.1ex,
-%framexbottommargin=.1ex,
-%framexrightmargin=\framemargin,
-%framexleftmargin=1mm, framextopmargin=1mm, frame=shadowbox, rulesepcolor=\color{blue},#1
-%frame=tb,
-%backgroundcolor=\color{yellow!10}
-backgroundcolor=\color{gray}
-%}
-}
-
 \newcommand{\N}[1]{\check{#1}}
@@
 
-\newcommand{\inpyth}{\lstinline[style=pythonstyle, basicstyle=\ttfamily]} %[]%
-
-\lstnewenvironment{inpython}[1][]{
-\lstset{style=pythonstyle, frame=trbl, belowcaptionskip=\belowtitleskip}
-}{}
-
 \newcommand{\includecode}[2][py]{\lstinputlisting[caption=#2,label=list:#2,style=pythonstyle,
@@
 
+\begin{document}
+\begin{frontmatter}
+
 \title{Assessment of a Shen-Fourier turbulent channel flow solver for large-scale simulations}
-\author{Mikael Mortensen and Diako Darian}
+\author[mmo]{Mikael Mortensen}
+\ead{mikaem@math.uio.no}
 %\date{}							% Activate to display a given date or no date
-
-\begin{document}
-\maketitle
-\section*{Abstract}
-
+\address[mmo]{Department of Mathematics, Division of Mechanics, University of Oslo}
+
+\begin{abstract}
+A fully (pseudo-)spectral solver for direct numerical simulations of large-scale turbulent channel flows is described.
+The solver utilizes the Chebyshev base functions suggested by J.
+Shen [SIAM J.
+Sci.
+Comput., 16, 1, 1995], that lead to stable and robust numerics, even at very large scale.
+New algorithms for direct solution of the linear systems are devised, and we show that the solver is accurate and applicable for very large simulations.
+Specifically, the solver is shown to reproduce first order statistics of a channel flow at $Re_{\tau}=2000$ by Hoyas and Jim\'{e}nez [Phys.
+Fluids, 18(1), 2006], which is higher than has been achieved in a channel by a fully spectral method before.
+\end{abstract}
+\begin{keyword}
+DNS \sep Fourier \sep Chebyshev
+\end{keyword}
+
+\end{frontmatter}
 \section{Introduction}
+Direct Numerical Simulations (DNS) of turbulent flows is a very important research tool, utilized across a range of scientific communities.
+DNS is used extensively to validate statistical models, and to further our understanding of complex mechanisms taking place inside turbulent flows.
+One of the many advantages of DNS is that it provides all information about a flow, and quantities that can be very hard to study experimentally, like velocity-pressure interactions, are trivially extracted from a DNS.
+In this regard, DNS both complements and extends the knowledge we are able to extract from experiments.
+
+The most commonly known DNS use simple geometries, because turbulence physics may then most easily be isolated and studied.
+Isotropic and homogeneous turbulence are usually studied in triply periodic domains, which allows a spectral Fourier decomposition in all three spatial directions.
+The equations are solved in spectral space and nonlinearities are usually handled with pseudo-spectral techniques.
+An important feature is that the pressure can be trivially eliminated, which simplifies the Navier-Stokes equations substantially.
+
+Inhomogeneous, wall bounded, flows, successfully studied with DNS are, e.g., the pipe flow, free shear layers and channel flow between two parallel plates (both Couette and Poiseuille).
+In this paper we will constrain ourselves to the pressure driven Poiseuille channel flow.
+In a channel there are two periodic directions that can be handled with Fourier expansions, whereas the wall normal direction is non-periodic and requires a different type of discretization.
+There are many challenges associated with this inhomogeneity not faced by the pure Fourier solvers, but the first problem at hand is the discretization.
+The first DNS channel solvers typically used a Chebyshev expansion for the wall-normal direction and, as such, were still able to obtain spectral accuracy in all three spatial directions.
+A Chebyshev-tau technique, that utilize the recurrence relations of the Chebyshev polynomials, is then used to approximate derivatives, and the coefficient matrices that appear (tridiagonal) can be inverted directly and efficiently.
+A downside to the Chebyshev-tau method is usually quoted \cite{canuto1988} as numerical stability and roundoff errors, caused by the recurrence relation, and severe condition numbers of the coefficient matrices.
+For Chebyshev-tau methods the condition numbers have been reported to grow with size as $\mathcal{O}(N^8)$, for a discretization using N points in the wall-normal direction.
+Discouraged by these numbers, all major recent channel flow simulations have found other, non-spectral, ways of discretizing the non-periodic direction.
+
+The largest known channel simulations to date have been performed by Lee and Moser \cite{leemoser15}, where for $Re_{\tau}\approx 5200$ they used a computational box of resolution $[10240 \times 1530 \times 7680]$ for the streamwise, wall-normal and spanwise directions, respectively.
+Lee and Moser used seventh-order B-splines for the wall-normal direction.
+Other simulations of similar magnitude are performed by Hoyas and Jim\'{e}nez \cite{hoyas06, hoyas08}, Lozano-Duran and Jim\'{e}nez \cite{Lozano2014} and Bernardini, Pirozzoli and Orlandi \cite{bernardini2014}.
+Bernardini et al.
+used second order finite differences throughout.
+The Jim\'{e}nez group used dealiased Fourier in the two periodic directions and seventh order compact finite differences, with fourth-order consistency and extended spectral-like resolution \cite{Lele92}, for the wall-normal direction.
+The solver by Jim\'{e}nez' group is reported to switch from Chebyshev to finite differences if the resolution is above a certain threshold \cite{hoyas08} (reached around $Re_{\tau}=1000$).
+In other words, they attempt to use spectral methods for as large $Re_{\tau}$ as possible.
+Spectral methods are attractive for their accuracy, but also for their resolution properties, that are superior to those of any finite difference or spline method.
+As such, it is desirable to develop fully spectral solvers that can be used for large-scale turbulence simulations.
+
+In his seminal papers \cite{Shen94,Shen95}, Jie Shen describe how to make Legendre and Chebyshev basis functions that lead to sparse matrices, susceptible to fast ($\mathcal{O}(N)$) direct solvers.
+To the authors knowledge, the bases have not been used for large-scale channel flow simulations, and algorithms for the required direct solvers have not, until now, been devised.
+Shen's bases have been used for the Navier Stokes equations before, though.
+Bouchon et.al.
+\cite{Bouchon01} describe a spectral-Galerkin formulation very similar to the one used in this paper.
+However, they choose the Legendre basis over the Chebyshev, due to the simpler structure of the matrices and the lack of weight functions for the Legendre basis.
+Also, the channel simulations performed were rather small, and there is also the issue of fast transforms for the Legendre basis.
+
+In this paper we will describe and assess a channel flow solver based on Shen's Chebyshev basis \cite{Shen95}.
+We will give a proper description of the theoretical basis in Sec~\ref{sec:prelim}, the discretization of Navier-Stokes equations in Sec~\ref{sec:discretizationNS} and we will here describe necessary algorithms, including a new fast direct solver for the biharmonic problem that arise in Sec~\ref{sec:implementation}.
+We will finally show in Sec~\ref{sec:verification} that roundoff is not a major issue, and that the spectral-Galerkin method is in deed applicable to large-scale simulations, larger than what has been achieved for fully spectral channel simulations before.
 
 \section{Preliminaries}
+\label{sec:prelim}
 The Navier-Stokes equations used to describe turbulent flow in a doubly 
@@
 and $t$ are position and time and $\bm{\mathcal{H}}(\bm{x}, t) = \bm{u}\times 
-\bm{\omega}$ (where $\bm{\omega} = \nabla \times \bm{u}$)  represents the 
-nonlinear convection.  The constant dynamic viscosity is denoted as $\nu$ and 
-$P$ is a pressure modified to account for the kinetic energy, i.e., $P = p + 
-\bm{u} \cdot \bm{u}/2$, where $p$ the instantaneous pressure 
-normalized by a constant density. The computational domain is $\Omega=[-1, 
-1]\times [0, L_y] \times [0, L_z]$ with channel walls located at $x=\pm 1$ such 
-that no-slip applies at $ \bm{u}(\pm 1, y, z, t) = 0$. The walls are spanning 
-the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions 
-with periodic lengths $L_y$ and $L_z$, respectively.
+\bm{\omega}$, where $\bm{\omega} = \nabla \times \bm{u}$, represents 
+the nonlinearity.  The constant dynamic viscosity is denoted as $\nu$ 
+and $P(\bm{x}, t)$ is a pressure modified to account for both the driving force, $\beta(t)$, and the kinetic energy, i.e., $P = p + \bm{u} \cdot \bm{u}/2 + \beta y$, where $p$ is the instantaneous 
+pressure normalized by a constant density. The computational domain is 
+$\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$ with channel walls 
+located at $x=\pm 1$, such that no-slip applies at 
+$ \bm{u}(\pm 1, y, z, t) = 0$. The walls are spanning 
+the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions with periodic lengths $L_y$ and $L_z$, respectively.
 
@@
 \begin{align}
-&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_0, \phi_1,\ldots, \phi_{N_x}\} \label{eq:Tk}\\
-& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_0, \D{\phi}_1, \ldots, \D{\phi}_{N_x-2} \} \label{eq:phiD}\\
+&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_k\}_{k=0}^{N_x} \label{eq:Tk}\\
+& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_k\}_{k=0}^{N_x-2} \label{eq:phiD}\\
 & \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
-\frac{k+1}{k+3} T_{k+4}, & \N{V}_{N_x} &= \text{span} \{\N{\phi}_0, \N{\phi}_1, 
-\ldots, \N{\phi}_{N_x-4} \} \label{eq:phiN} 
+\frac{k+1}{k+3} T_{k+4}, & \N{V}_{N_x} &= \text{span} \{\N{\phi}_k\}_{k=0}^{N_x-4} \label{eq:phiN} 
 \end{align}
@@
 
-Three-dimensional basis functions and function spaces that are periodic in $y$ 
+Three-dimensional basis functions and function spaces, that are periodic in $y$ 
 and $z$ directions, can now be defined as
@@
 and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). We use 
-notation $K_N^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
-dimensions of $K_N$.
+notation $K^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
+dimensions of $K_N$. Likewise, ${K}^x(l)={K}_N(l, 0, 0), \D{K}^x(l)=\D{K}_N(l, 0, 0)$ and $\N{K}^x(l)=\N{K}_N(l, 0, 0)$ refer to the wavenumbers in the wall normal direction.
 
@@
 \end{align}
-where $x_i = h(i)$ represents either
+where $x_i = h(i)$ represents either Gauss-Lobatto or Gauss-Chebyshev points
 \begin{equation}
@@
 
-A three-dimensional scalar product in the weighted $[L^2_w(\Omega)]^3$ space is defined as
+A three-dimensional scalar product in the weighted $L^2_w(\Omega)$ space is defined as
 \begin{align}
@@
 periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction. In 
-this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, 
+this work we will make use of the discrete weighted $l^2_w(\Omega)$ space, 
 where quadrature is employed for integration. As such the scalar product 
@@
 \end{align}
-Here $h\N{\mathcal{S}}$ denotes the complete three dimensional scalar product 
+Here $h\N{\mathcal{S}}$ denotes the complete three-dimensional scalar product 
 and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
@@
 $c_{N_x}=1$ for the continuous scalar product $L^2_w(\Omega)$ (see, e.g., Eq. 
-(2.7) of Shen2), but for the discrete $l^2_w(\Omega)$ it follows that 
+(2.7) of \cite{Shen95}), but for the discrete $l^2_w(\Omega)$ it follows that 
 $c_{N_x}=2$ due to inexact quadrature for the highest wavenumber.}
@@
 complete transform. To find a transformation between physical $u$ and spectral 
-$\hat{u}$ we make use of Eq. (\ref{eq:u_solx}) directly and use the discrete 
+$\hat{u}$ we make use of Eq. (\ref{eq:u_solx}) directly in the first line of (\ref{eq:sst1}) and use extensively the discrete 
 orthogonality of the Fourier basis functions
@@
 \sum_{(q,r,s) \in \N{K}_N} \hat{u}(q,r,s,t) \N{\psi}(q,r,s), 
-\N{\psi}(l,m,n,t) \right>_w, \notag \\
+\N{\psi}(l,m,n) \right>_w, \notag \\
            &= h \sum_{q=0}^{N_x-4} \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) 
            \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
-           &= h \sum_{q=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t), 
+           &= h \sum_{q=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t). 
            \label{eq:sst2}
 \end{align}
-where $\N{B}_{lq} = (\N{\phi}_q, \N{\phi}_l)_w = 
+Here $\{\N{B}_{lq}\}_{l,q=0}^{N_x-4} = (\N{\phi}_q, \N{\phi}_l)_w = 
 \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
@@
 h\N{\mathcal{S}}(u)(l, m, n, t), \notag \\
- \hat{u}_{\bm{k}}(t) &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
+ \hat{u} &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
  \quad \forall \, \bm{k} \in \N{K}_N, \label{eq:fstB}
@@
 where $\N{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = 
-\N{\mathcal{T}}^{-1}(\N{\mathcal{T}}(u))$. Note that since $\N{B}$ assembles to 
-a pentadiagonal matrix the solution ($\N{B}^{-1}$) can be obtained very fast 
-and 
-the complete 
-transformation thus requires a fast Chebyshev transform ($O(N_x \log N_x)$) and 
-a fast linear algebraic solve ($O(N_x)$). Similar transforms $\mathcal{T}$ 
+\N{\mathcal{T}}^{-1}(\N{\mathcal{T}}(u))$. Note that, since $\N{B}$ assembles to 
+a pentadiagonal matrix, the solution ($\N{B}^{-1}$) can be obtained very fast and the complete 
+transformation thus requires a fast Chebyshev transform ($\mathcal{O}(N_x \log N_x)$) and 
+a fast linear algebraic solve ($\mathcal{O}(N_x)$). Similar transforms $\mathcal{T}$ 
 and $\D{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and 
@@
 $\mathcal{S}_l(\cdot) = (\cdot, T_l)_w$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
-\D{\phi}_l)_w$.
+\D{\phi}_l)_w$. See the Appendix for algorithms for all required transforms.
 
@@
 	\left<h_u, \N{\psi} \right>_w + \nu \left<\nabla^4u, \N{\psi}\right>_w 
-	&\forall \N{\psi} \in \N{V}_N, \label{eq:u1} \\
+	&\forall \,\N{\psi} \in \N{V}_N, \label{eq:u1} \\
 	\frac{\partial}{\partial t}\left<g, \D{\psi}\right>_w &= \left<h_g, 
 	\D{\psi}\right>_w + \nu 
-	\left<\nabla^2 g, \D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N ,
+	\left<\nabla^2 g, \D{\psi}\right>_w &\forall \, \D{\psi} \in \D{V}_N ,
 	\label{eq:g1} \\
 	\left<f, \D{\psi}\right>_w &= \left<\frac{\partial u}{\partial x}, 
-	\D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N. \label{eq:f1}
+	\D{\psi}\right>_w &\forall \, \D{\psi} \in \D{V}_N. \label{eq:f1}
 \end{align}
@@
 \left<f, \D{\psi}\right>_w &= \left<\frac{\partial v}{\partial y} + 
-\frac{\partial w}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
+\frac{\partial w}{\partial z}, \D{\psi}\right>_w \, &\forall \, \D{\psi} \in 
 \D{V}_N, \label{eq:f2} \\
 \left<g, \D{\psi}\right>_w &= \left<\frac{\partial w}{\partial y}  - 
-\frac{\partial v}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
+\frac{\partial v}{\partial z}, \D{\psi}\right>_w \, &\forall \, \D{\psi} \in 
 \D{V}_N, \label{eq:g2}
@@
 \hat{f}_{\bm{k}} &= \imath \underline{m}\, \hat{v}_{\bm{k}} + \imath 
-\underline{n}\, \hat{w}_{\bm{k}} &\forall \bm{k} \in \D{K}_N^0, \label{eq:f3} \\
+\underline{n}\, \hat{w}_{\bm{k}} &\forall \, \bm{k} \in \D{K}_N^0, \label{eq:f3} \\
 \hat{g}_{\bm{k}} &= \imath \underline{m}\, \hat{w}_{\bm{k}} - \imath 
-\underline{n}\, \hat{v}_{\bm{k}} & \forall \bm{k} \in \D{K}_N^0, \label{eq:g3}
+\underline{n}\, \hat{v}_{\bm{k}} & \forall \, \bm{k} \in \D{K}_N^0, \label{eq:g3}
 \end{align}
@@
 \frac{\partial }{\partial t} \D{B}_{lq}\hat{v}(q, 0, 0, t) &= 
-\D{B}_{lq}\hat{\mathcal{H}}_y(q, 0, 0, t) + \nu \D{A}_{lq} \hat{v}(q, 0, 0, t) 
-& \forall\, l \in [0, 1, 
-\ldots, N_x-2], \label{eq:v0}\\
+\D{B}_{lq}\hat{\mathcal{H}}_y(q, 0, 0, t) + \nu \D{A}_{lq} \hat{v}(q, 0, 0, t) - \left(\beta(t), \D{\phi}_l \right)_w
+& \forall\, l \in \D{K}^x, \label{eq:v0}\\
 \frac{\partial }{\partial t} \D{B}_{lq}\hat{w}(q, 0, 0, t) &= 
 \D{B}_{lq}\hat{\mathcal{H}}_z(q, 0, 0, t) + \nu \D{A}_{lq} \hat{w}(q, 0, 0, t) 
-& \forall\, l \in [0, 1, \ldots, N_x-2].\label{eq:w0}
-\end{align}
-
-Equations (\ref{eq:u1}-\ref{eq:f1}) are also solved in spectral space. 
-Inserting for (\ref{eq:u_solx}) and a similar expansion for $g$ in $\D{V}_N$, 
-the inner products required to solve Eqs. (\ref{eq:u1}-\ref{eq:f1}) are
+& \forall\, l \in \D{K}^x,\label{eq:w0}
+\end{align}
+where $\D{B}_{lq} = \left(\D{\phi}_q, \D{\phi}_l \right)_w$ and $\D{A}_{lq}=\left(\D{\phi}_q^{''}, \D{\phi}_l \right)_w$.
+Note that (\ref{eq:v0}) is the only place where the driving force, or the mean pressure gradient, $\beta$, enters, and, since it is constant in space, the term $\left(\beta(t), \D{\phi}_l \right)_w$ will only be non-zero for $l=0$.
+
+Equations (\ref{eq:u1}-\ref{eq:f1}) need to be rewritten in terms of one-dimensional scalar products. 
+Inserting the expansion (\ref{eq:u_solx}) for $u$ and a similar expansion for $g$ in $\D{V}_N$, the inner products required to solve Eqs. (\ref{eq:u1}-\ref{eq:f1}) are
 \begin{align}
@@
 h\left[-(\underline{m}^2+\underline{n}^2) 
-\N{\mathcal{S}}(\mathcal{H}_x) - \imath \underline{m}\left(\D{\phi}_q^{'}, 
+\left(\D{\phi}_q, 
+\N{\phi}_l \right)_w \hat{\mathcal{H}}_{x, q} - \imath \underline{m}\left(\D{\phi}_q^{'}, 
 \N{\phi}_l \right)_w \hat{\mathcal{H}}_{y, q} - \imath 
 \underline{n}\left(\D{\phi}_q^{'}, \N{\phi}_l \right)_w 
-\hat{\mathcal{H}}_{z, q}\right], 
+\hat{\mathcal{H}}_{z, q}\right], \label{eq:S_hv}
 \\
-\left< h_g, \D{\psi}_{\bm{k}} \right>_w &= h\left[ \imath \underline{m}\, \D{B} 
-\hat{\mathcal{H}}_z - \imath \underline{n}\, \D{B} \hat{\mathcal{H}}_{y} 
-\right],
+\left< h_g, \D{\psi}_{\bm{k}} \right>_w &= h\left[ \imath \underline{m}\, \left(\D{\phi}_q, \D{\phi}_l \right)_w 
+\hat{\mathcal{H}}_{z, q} - \imath \underline{n}\, \left(\D{\phi}_q, \D{\phi}_l \right)_w \hat{\mathcal{H}}_{y, q} 
+\right],\label{eq:S_hg}
 \end{align}
@@
 vector product $ \sum_{q=0}^{N_x-4}\left( \N{\phi}_q^{'}, \D{\phi}_l\right)_w 
-\hat{u}(q, {m}, {n}, t)$. The inner products are used to set up linear systems 
+\hat{u}(q, {m}, {n}, t)$, where $m$ and $n$ are given. The inner products are used to set up linear systems 
 of equations for the inhomogeneous wall normal direction. All inner products 
@@
 %\end{align}
-The mass matrices, stiffness matrices $\N{A}_{lq} = \left( \N{\phi}_q^{''}, 
+The required matrices $ \N{B}_{lq} = \left( \N{\phi}_q, 
+\N{\phi}_l\right)_w, \D{B}_{lq} = \left( \D{\phi}_q, 
+\D{\phi}_l\right)_w, \N{A}_{lq} = \left( \N{\phi}_q^{''}, 
 \N{\phi}_l\right)_w, 
 \D{A}_{lq} = \left( \D{\phi}_q^{''}, \D{\phi}_l\right)_w 
-$ and $\N{Q}_{lq} = \left(\N{\phi}^{''''}_q, \N{\phi}_l\right)_w$ are given by 
-Shen, whereas $\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w$ and 
-$\left(\D{\phi}_q^{'}, \N{\phi}_l\right)_w$ are given in the 
-appendix.
+, \D{C}_{lq} = \left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w$ and 
+$ \N{C}_{lq} = \left(\D{\phi}_q^{'}, \N{\phi}_l\right)_w$ are given in Table~\ref{tab:matrices}, whereas the matrix $\N{Q}_{lq} = \left(\N{\phi}^{''''}_q, \N{\phi}_l\right)_w$ is given below in Eqs (\ref{eq:N_Q})-(\ref{eq:rk}).
+
+The convection vector $\bm{\hat{\mathcal{H}}}$, needed to compute Eqs.
+(\ref{eq:S_hv}) and (\ref{eq:S_hg}), is found by projecting to the Dirichlet vector space $\D{V}_N^3$, which corresponds to transforming the nonlinear cross product over the entire mesh
+\begin{equation}
+\bm{\hat{\mathcal{H}}}_i = \mathcal{\D{T}}((\bm{u} \times \bm{\omega})_i) \quad \forall \, i=0,1,2. \label{eq:H_hat}
+\end{equation}
+The curl, $\bm{\omega} = (g, \partial_z u - \partial_x w, \partial_y u - \partial_x v)$, is computed by projecting each term to its appropriate space.
+To this end the two terms $\partial_x v$ and $\partial_x w$ are projected to $V_N$, whereas the remaining $\partial_y u$ and $\partial_z u$ are projected to $\N{V}_N$.
+The nonlinear term is also in need of dealiasing.
+In this work dealiasing has been performed with the 3/2-rule, see Canuto~et al.
+\cite{canuto1988}.
 
@@
 Discretization in time is performed using a constant time step $\triangle t$, 
-such that the time is represented discretely as $t_n = n \triangle t, \, n \in 
+such that the time is represented discretely as $t_{\kappa} = {\kappa} \triangle t, \, {\kappa} \in 
 \mathcal{Z}^{\ge}$. After 
 discretizing in time, equations (\ref{eq:u1}) and (\ref{eq:g1}) are transformed 
-to a matrix form that can be solved for $\hat{u}^{n+1}$ and $\hat{g}^{n+1}$ 
+to a matrix form that can be solved for $\hat{u}^{{\kappa}+1}$ and $\hat{g}^{{\kappa}+1}$ 
 given wavenumbers $m$ and $n$
 \begin{align}
-\N{H}\hat{u}^{n+1} & = \left(2\N{A} - 2\underline{z}^2\N{B} - \N{H} 
-\right)\hat{u}^{n} + \N{\mathcal{S}}(h_v^{n+1/2}) \triangle t & 
+\N{H}\hat{u}^{{\kappa}+1} & = \left(2\N{A} - 2\underline{z}^2\N{B} - \N{H} 
+\right)\hat{u}^{{\kappa}} + \N{\mathcal{S}}(h_v^{{\kappa}+1/2}) \triangle t & 
 \forall \, {\bm{k}} \in \N{K}_N, 
 \label{eq:ufin}\\ 
-\D{H} \hat{g}^{n+1} &= \left(2 \D{B}-\D{H}\right) 
-\hat{g}^{n} + \D{\mathcal{S}}(h_g^{n+1/2}) \triangle t &\forall \, 
+\D{H} \hat{g}^{{\kappa}+1} &= \left(2 \D{B}-\D{H}\right) 
+\hat{g}^{{\kappa}} + \D{\mathcal{S}}(h_g^{{\kappa}+1/2}) \triangle t &\forall \, 
 {\bm{k}} \in 
@@
 \underline{z}^2 \right) \N{A}_{lq} - \frac{2\underline{z}^2 + \nu \triangle t 
-\underline{z}^4}{2} \N{B}_{lq} &\forall m, n \in K_N^p, 
+\underline{z}^4}{2} \N{B}_{lq} &\forall \, m, n \in K^p, 
 \label{eq:Biharmonic_matrix} \\
@@
 t}{2}\D{A}_{lq} + (1 + \frac{\nu \triangle t \underline{z}^2}{2}) \D{B}_{lq} 
-&\forall m, n \in K_N^p, \label{eq:Helmholtz_matrix}
+&\forall \, m, n \in K^p, \label{eq:Helmholtz_matrix}
 \end{align}
@@
 terms are computed with a forward Adams-Bashforth projection, such that 
-$\N{\mathcal{S}}(h_v^{n+1/2}) = 
-\N{\mathcal{S}}(1.5h_v^{n} - 0.5 h_v^{n-1})$ and $\D{\mathcal{S}}(h_g^{n+1/2}) 
-= \D{\mathcal{S}}(1.5h_g^{n} - 0.5 h_g^{n-1})$. Equations 
+$\N{\mathcal{S}}(h_v^{{\kappa}+1/2}) = 
+\N{\mathcal{S}}(1.5h_v^{{\kappa}} - 0.5 h_v^{{\kappa}-1})$ and $\D{\mathcal{S}}(h_g^{{\kappa}+1/2}) 
+= \D{\mathcal{S}}(1.5h_g^{{\kappa}} - 0.5 h_g^{{\kappa}-1})$. Equations 
 (\ref{eq:v0}) and (\ref{eq:w0}) are also discretized in time using 
@@
 \begin{align}
-\left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{v}_0^{n+1} &= 
+\left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{v}_0^{{\kappa}+1} &= 
 \left(\frac{\nu \triangle t}{2}\D{A} + \D{B} 
-\right)\hat{v}^{n}_0 + \triangle t\D{B}\hat{\mathcal{H}}_{y,0}^{n+1/2} , 
+\right)\hat{v}^{{\kappa}}_0 + \triangle t\D{B}\hat{\mathcal{H}}_{y,0}^{{\kappa}+1/2} - \triangle t \left(\beta(t), \D{\phi}_l \right)_w , 
 \label{eq:v00} \\
-\left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{w}_0^{n+1} &= 
+\left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{w}_0^{{\kappa}+1} &= 
 \left(\frac{\nu \triangle t}{2}\D{A} + \D{B} 
-\right)\hat{w}^{n}_0 + \triangle t\D{B}\hat{\mathcal{H}}_{z,0}^{n+1/2}, 
+\right)\hat{w}^{{\kappa}}_0 + \triangle t\D{B}\hat{\mathcal{H}}_{z,0}^{{\kappa}+1/2}, 
 \label{eq:w00}
 \end{align}
-where we have used notation $\hat{v}^{n}_0 = \hat{v}(l, 0, 0, t_n)$, 
-$\hat{\mathcal{H}}^{n}_{y,0} = \hat{\mathcal{H}}_y(l, 0, 0, t_n)$ and similar 
-for $\hat{w}^n_0$ and $\hat{\mathcal{H}}^n_{z,0}$.
+where we have used notation $\hat{v}^{{\kappa}}_0 = \hat{v}(l, 0, 0, t_{\kappa})$, 
+$\hat{\mathcal{H}}^{{\kappa}}_{y,0} = \hat{\mathcal{H}}_y(l, 0, 0, t_{\kappa})$ and similar 
+for $\hat{w}^{\kappa}_0$ and $\hat{\mathcal{H}}^{\kappa}_{z,0}$.
 
@@
 \label{sec:implementation}
-The solution procedure for the numerical method described in 
-Sec.~\ref{sec:discretizationNS} is given in Algorithm \ref{alg:NS}
-
+An outline of the solution procedure used, for the numerical method described in Sec.~\ref{sec:discretizationNS}, is given in Algorithm \ref{alg:NS}.
 \begin{algorithm}
-	\caption{Solution procedure for Navier Stokes equations}
-    \label{alg:NS}
+	\caption{Solution procedure for Navier Stokes equations.     \label{alg:NS}}
     \begin{algorithmic}[1]
@@
 		\State \text{Compute LU decomposition of} $\N{H}(m,n), \D{H}(m,n) 
-		\,\forall \, m, n \in K_N^p$
-		\State $t=0; n=0$
+		\,\forall \, m, n \in K^p$
+		\State \text{Compute LU decomposition of coefficient matrices in} (\ref{eq:v00}) and (\ref{eq:w00})
+		\State $t=0; {\kappa}=0$
     	\While{$t<T$}
-	    	\State \text{Compute nonlinear convection} $\bm{\mathcal{H}}^n$
+	    	\State \text{Compute nonlinear convection} $\bm{\hat{\mathcal{H}}}^{\kappa}$
 	    	\State \text{Compute right hand sides of Eqs.~}(\ref{eq:ufin}), 
@@
 		    \State \text{Solve Eq.}(\ref{eq:ufin}) for 
-		    $\hat{u}^{n+1}_{\bm{k}} \, \forall \,\bm{k} \in \N{K}_N$
+		    $\hat{u}^{{\kappa}+1}_{\bm{k}} \, \forall \,\bm{k} \in \N{K}_N$
 		    \State \text{Solve Eqs.}(\ref{eq:gfin}) for 
-		    $\hat{g}^{n+1}_{\bm{k}} \, \forall\, \bm{k} \in \D{K}_N$
+		    $\hat{g}^{{\kappa}+1}_{\bm{k}} \, \forall\, \bm{k} \in \D{K}_N$
 		    \State \text{Solve Eqs.}(\ref{eq:f3}, \ref{eq:g3}) for  
@@
 		    \State \text{Solve Eqs.}(\ref{eq:v00}, \ref{eq:w00}) for 
-		    $\hat{v}_0^{n+1}, \hat{w}_0^{n+1}\, 
-		    \forall \, l \in [0, 1, \ldots, N_x-2] $
-		    \State $n \gets n + 1$
+		    $\hat{v}_0^{{\kappa}+1}, \hat{w}_0^{{\kappa}+1}\, 
+		    \forall \, l \in \D{K}^x$
+		    \State ${\kappa} \gets {\kappa} + 1$
 		    \State $t \gets t+\triangle t$
@@
 \end{algorithm}
-
 The major computational cost comes from  computing the nonlinear convection 
@@
 convection 
-term $\bm{\hat{\mathcal{H}}}$ is computed using a pseudo-spectral method with 
-the 3/2-rule for de-aliasing in the periodic directions (ref Canuto). The 
+term, $\bm{\hat{\mathcal{H}}}$, is computed using a pseudo-spectral method with 
+the 3/2-rule for dealiasing in the periodic directions, see \cite{canuto1988}. The 
 transforms are carried out using fast Fourier and fast Chebyshev transforms 
-that are all of order $O(N\log N)$ per one-dimensional transform of size 
-N. The Shen transforms make use of the Chebyshev transforms that are using 
-discrete cosine transforms. The forward scalar product is implemented as shown 
-in Algorithm \ref{alg:fst}. A complete transform is obtained through a scalar 
+that are all of order $\mathcal{O}(N_x\log N_x)$ per one-dimensional transform of size 
+$N_x$. The Shen transforms make use of the Chebyshev transforms that are using 
+discrete cosine transforms. The forward scalar products and transforms are implemented as shown 
+in Algorithm \ref{alg:fst} of the Appendix. A complete transform is obtained through a scalar 
 product followed by solving the linear system involving the mass 
@@
 pentadiagonal for $V_N, \D{V}_N$ and $\N{V}_N$ respectively, solving the linear 
-system is $O(N)$ and the complete transform still scales like $O(N\log N)$. 
-\begin{algorithm}
-	\caption{Forward scalar product for all spaces $V_N, \D{V}_N, \N{V}_N$. 
-	Here "dct" is the discrete cosine transform from SciPy.}
-	\label{alg:fst}
-	\begin{algorithmic}[1]
-	\Procedure{fastScalar}{}
-		\State \textbf{Input:} $\{f_j\}_{j=0}^{N}$
-		\State \textbf{Input:} points
-		\State \textbf{Input:} space
-		\If{ points = "Gauss-Lobatto"}
-			\State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=2}, 
-			\text{axis=0})\frac{\pi}{2 N}$
-		\Else{ points = "Gauss-Chebyshev"}
-		    \State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=1}, 
-		    \text{axis=0}) \frac{\pi}{2 (N-1)}$
-	    \EndIf
-	    \If{ space = $\D{V}_N$}
-	        \For{$k=0$ to $N-2$}
-		        \State $\hat{f}_k \gets w_k - w_{k+2}$
-	        \EndFor
-			\State \Return $\{\hat{f}_k\}_{k=0}^{N-2}$	        
-	    \ElsIf{ space = $\N{V}_N$}	                
-  	        \For{$k=0$ to $N-4$}
-       	        \State $\hat{f}_k \gets w_k - \frac{2(k+2)}{k+3} w_{k+2} + 
-       	        \frac{k+1}{k+3}w_{k+4}$
-   	        \EndFor
-   	        \State \Return $\{\hat{f}_k\}_{k=0}^{N-4}$
-   	    \Else{}
-   	        \For{$k = 0$ to $N$}
-     	        \State $f_k \gets \frac{2}{\pi} w_k $
-     	    \EndFor
-     	    \State $f_0 \gets f_0 / 2$
-     	    \If{ points = "Gauss-Lobatto"}
-	     	\State $f_N \gets f_N / 2$	     	
-     	    \EndIf
-			\State \Return $\{f_k\}_{k=0}^{N}$    
-	    \EndIf
-	\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-Shen \cite{shen1995} states that it is possible to solve Eq.~(\ref{eq:gfin}) 
+system is $\mathcal{O}(N_x)$ and the complete transform still scales like $\mathcal{O}(N_x\log N_x)$.
+
+Shen \cite{Shen95} states that it is possible to solve Eq.~(\ref{eq:gfin}) 
 with 
@@
 that the matrix 
-$\D{H}$ is of type upper Hessenberg and a solution
-procedure is simply to perform a direct LU decomposition (se, e.g., 
+$\D{H}$ decouples into odd and even components, leading to two matrices of type upper Hessenberg. A solution
+procedure is then simply to perform a direct LU decomposition (see, e.g., 
 \cite{stewart98}), which leads to a 
 lower triangular  
-matrix $L$ with only one subdiagonal and an upper triangular matrix $U$ which 
+matrix $L$ with only one subdiagonal and an upper triangular matrix $U$ that 
 is dense. However, each row in $U$ 
 contains at most three distinct values at $U_{k,k}, U_{k,k+2}$ and $U_{k,k+4}$, 
-and then $U_{k,j} = U_{k,k+4}\, \forall \, j = k+6, k+8, \ldots N$. 
+and then $U_{k,j} = U_{k,k+4}\, \forall \, j = k+6, k+8, \ldots N_x$. 
 Consequently 
 only three diagonals in $U$ need storage and the back solve can be traversed 
-very efficiently in $O(N)$. Note also that the system could be decoupled into 
-odd and even coefficients, which leads to two subsystem that may be trivially 
-solved 
-simultaneously in two threads. For optimal 
+very efficiently in $\mathcal{O}(N_x)$. Note also that the decoupling into 
+odd and even coefficients leads to two subsystem that may be trivially 
+solved simultaneously in two threads. For optimal 
 performance, the odd and even coefficient would then need to be stored 
-contiguously in computer memory, and not alternately, which would be the normal 
-way of storing the coefficients.
+contiguously in computer memory, and not alternately, which would be the normal way of storing the coefficients.
 
@@
 challenging to solve for efficiently, but 
-it is still possible to find an algorithm that is $O(N)$ for a system of $N$ 
+it is still possible to find an algorithm that is $\mathcal{O}(N_x)$ for a system of $N_x$ 
 unknowns. Note 
@@
 \begin{equation}
-\N{H} = c_0\N{Q} + c_1\N{A} + c_2 \N{B},
+\N{H} = \xi_0\N{Q} + \xi_1\N{A} + \xi_2 \N{B},
 \end{equation}
-where $c_0, c_1$ and $c_2$ are constants. The matrix $\N{A}_{kj}$ contains only 
+where $\xi_0, \xi_1$ and $\xi_2$ are constants. The matrix $\N{A}_{kj}$ contains only 
 three nonzero diagonals at $ j = k-2, k, k+2$,  whereas matrix $\N{B}_{kj}$ 
@@
 \begin{align}
+\label{eq:N_Q}
  \N{Q}_{kk} &= -4(k+1)(k+2)^2, \\
@@
 \end{align} 
-For each row in $\N{H}_{kj}$ we have that $\N{H}_{kj}=c_0\N{Q}_{kj}\, \forall j 
+For each row in $\N{H}_{kj}$ we have that $\N{H}_{kj}=\xi_0\N{Q}_{kj}\, \forall j 
 > k+4$. A straight forward direct LU decomposition of $\N{H}$ can be performed 
@@
 \begin{equation}
-\N{U}_{kj} = c_0(a_k q_j + b_k s_j), \quad j = k+6, k+8, \ldots, N_x-4, 
+\N{U}_{kj} = \xi_0(a_k q_j + b_k s_j), \quad j = k+6, k+8, \ldots, N_x-4, 
 \label{eq:ab}
 \end{equation}
-where $\bm{a}$ and $\bm{b}$ are two new vectors that can be computed 
-recursively from $\N{L}$ 
-\begin{align}
-a_k &= p_k, \, &\forall \, k= 0, 1, \notag \\
-b_k &= r_k, \, &\forall \, k= 0, 1, \notag\\
-a_k &= p_k - \N{L}_{k,k-2}a_{k-2}, \, &\forall \, k = 2, 3, \notag\\
-b_k &= r_k - \N{L}_{k,k-2}b_{k-2}, \, &\forall \, k = 2, 3, \notag\\
-a_k &= p_k - \N{L}_{k,k-2}a_{k-2} - \N{L}_{k,k-4}a_{k-4}, \, &\forall \, k = 4, 
-5, 
-\ldots, N, \notag\\
-b_k &= r_k - \N{L}_{k,k-2}b_{k-2} - \N{L}_{k,k-4}b_{k-4}, \, &\forall \, k = 4, 
-5, 
-\ldots, N.
-\end{align}
+where $\{a_k\}_{k=0}^{N_x-4}$ and $\{b_k\}_{k=0}^{N_x-4}$ are two new 
+vectors that can be computed recursively from $\N{L}$, as shown in Alg.~\ref{alg:ab}. 
+%\begin{align}
+%a_k &= p_k, \, &\forall \, k= 0, 1, \notag \\
+%b_k &= r_k, \, &\forall \, k= 0, 1, \notag\\
+%a_k &= p_k - \N{L}_{k,k-2}a_{k-2}, \, &\forall \, k = 2, 3, \notag\\
+%b_k &= r_k - \N{L}_{k,k-2}b_{k-2}, \, &\forall \, k = 2, 3, \notag\\
+%a_k &= p_k - \N{L}_{k,k-2}a_{k-2} - \N{L}_{k,k-4}a_{k-4}, \, &\forall \, k = 4, 
+%5, 
+%\ldots, N, \notag\\
+%b_k &= r_k - \N{L}_{k,k-2}b_{k-2} - \N{L}_{k,k-4}b_{k-4}, \, &\forall \, k = 4, 
+%5, 
+%\ldots, N.
+%\end{align}
 If $\bm{a}$ and $\bm{b}$ 
-are pre-computed 
+are pre-computed, 
 the total storage requirement for the complete LU decomposition is no more than 
@@
 one simple forward elimination and a back solve, where the back 
-solve is very fast ($O(N)$) because of the formulation (\ref{eq:ab}), leading 
+solve is very fast ($\mathcal{O}(N_x)$) because of the formulation (\ref{eq:ab}), leading 
 to a formula for the backwards substitution like (valid for $k \le N_x-8$)
@@
 u_k = \left(y_k - \N{U}_{k, k+2} u_{k+2} - \N{U}_{k, k+4} u_{k+4} - 
-c_0 a_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} q_j u_j - 
-c_0 b_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} 
+\xi_0 a_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} q_j u_j - 
+\xi_0 b_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} 
 s_j u_j\right)/\N{U}_{kk},
@@
 	\caption{LU decomposition of biharmonic operator $\N{H}$. Current algorithm 
-	is using dense storage of $\N{U}_{kj}$, but no more than the 3 diagonals 
+	is using dense storage of $\N{U}_{kj}$, but no more than the 3 items per row
 	need 
-	to be stored eventually.}
-	\label{alg:lu_biharmonic}
+	to be stored after the row is no longer needed in the for-loop below (as the loop is traversed we do not need $\N{U}_{m, j}$ for $m<k-4$). \label{alg:lu_biharmonic}}
 	\begin{algorithmic}[1]
@@
 			\State \textbf{Input:} $\{\N{H}_{kj}\}_{k, j = 0}^{N}$
-	        \State $\N{U}_{0j} = \N{H}_{0j} \quad \forall \, j \in 0, 2, 
-	        \ldots, N$
-	        \State $\N{U}_{1j} = \N{H}_{1j} \quad \forall \, j \in 1, 3, 
-	        \ldots, N$
-	        \State $\N{L}_{20} = \N{H}_{20}/\N{U}_{00}$
-	        \State $\N{L}_{31} = \N{H}_{31}/\N{U}_{11}$
-	        \State $\N{U}_{2j} = \N{H}_{2j} - \N{L}_{20} \N{U}_{0j} \quad 
-	        \forall \, j \in 2, 4, \ldots, N$
-	        \State $\N{U}_{3j} = \N{H}_{3j} - \N{L}_{31} \N{U}_{1j} \quad 
-	        \forall \, j \in 
-	        3, 5, \ldots, N$
+			\State Integer $M_e \gets \lfloor N/2 \rfloor$
+			\State Integer $M_o \gets \lfloor (N-1)/2 \rfloor$  
+	        \State $\{\hat{U}_{0,2j}\}_{j=0}^{M_e} \gets \{\hat{H}_{0,2j}\}_{j=0}^{M_e}$ 
+	        \State $\{\hat{U}_{1,2j+1}\}_{j=0}^{M_o} \gets \{\hat{H}_{1,2j+1}\}_{j=0}^{M_o}$ 
+	        \State $\N{L}_{20} \gets \N{H}_{20}/\N{U}_{00}$
+	        \State $\N{L}_{31} \gets \N{H}_{31}/\N{U}_{11}$
+	        \State $\{\hat{U}_{2,2j}\}_{j=1}^{M_e} \gets \{\N{H}_{2,2j}\}_{j=1}^{M_e} - \N{L}_{20} \{\N{U}_{0,2j}\}_{j=1}^{M_e}$
+	        \State $\{\hat{U}_{3,2j+1}\}_{j=1}^{M_o} \gets \{\N{H}_{3,2j+1}\}_{j=1}^{M_o} - \N{L}_{31} \{\N{U}_{1,2j+1}\}_{j=1}^{M_o}$
+
+	        %\State $\N{U}_{2j} = \N{H}_{2j} - \N{L}_{20} \N{U}_{0j} \quad 
+	        %\forall \, j \in 2, 4, \ldots, N$
+	        %\State $\N{U}_{3j} = \N{H}_{3j} - \N{L}_{31} \N{U}_{1j} \quad 
+	        %\forall \, j \in 
+	        %3, 5, \ldots, N$
 	        \For{$k = 4, 5, \ldots, N$}
-		        \State $\N{L}_{k,k-4} = \N{H}_{k, k-4} / \N{U}_{k-4, k-4}$
-		        \State $\N{U}_{k,j} = \N{H}_{k,j} - \N{L}_{k, k-4} 
-		        \N{U}_{k-4, j} 
-		        \quad \forall j = k-2, k, \ldots, N$
-	            \State $\N{L}_{k,k-2} = {\N{U}}_{k,k-2}/{\N{U}}_{k-2, k-2}$
-		        \State $\N{U}_{kj} = {\N{U}}_{kj} - \N{L}_{k, k-2} \N{U}_{k-2, 
-		        j} \quad 
-		        \forall j = k, k+2, \ldots, N$
+	            \State Integer $M \gets \lfloor \frac{N-k}{2}\rfloor$
+		        \State $\N{L}_{k,k-4} \gets \N{H}_{k, k-4} / \N{U}_{k-4, k-4}$
+		        \State $\{\N{U}_{k,k+2j}\}_{j=-1}^{M} \gets \{\N{H}_{k,k+2j}\}_{j=-1}^{M} - \N{L}_{k, k-4} 
+		        \{\N{U}_{k-4, k+2j}\}_{j=-1}^{M}$
+		        %\State $\N{U}_{k,j} = \N{H}_{k,j} - \N{L}_{k, k-4}  \N{U}_{k-4, j} \quad \forall j = k-2, k, \ldots, N$
+	            \State $\N{L}_{k,k-2} \gets {\N{U}}_{k,k-2}/{\N{U}}_{k-2, k-2}$
+	            \State $\{\N{U}_{k,k+2j}\}_{j=0}^{M} \gets \{{\N{U}}_{k,k+2j}\}_{j=0}^{M} - \N{L}_{k, k-2} \{\N{U}_{k-2,k+2j}\}_{j=0}^{M}$
+		        %\State $\N{U}_{kj} = {\N{U}}_{kj} - \N{L}_{k, k-2} \N{U}_{k-2, j} \quad \forall j = k, k+2, \ldots, N$
 	        \EndFor
@@
 
-%\begin{algorithm}
-%	\caption{Recursive formula to compute $\{a_k\}_{k=0}^{N_x-4}, 
-%	\{b_k\}_{k=0}^{N_x-4}$ in Eq. (\ref{eq:ab}). Note that the lower triangular 
-%	matrix $\{L_{kj}\}_{k,j=0}^{N_x-4}$ contains only two nonzero diagonals and 
-%	the parameters $p_k$ and $r_k$ are given in Eqs. (\ref{eq:pk}) and 
-%	(\ref{eq:rk}) respectively. }
-%    \label{alg:ab}
-%	\begin{algorithmic}[1]
-%		\Procedure{ReduceAB}{}
-%		\State \textbf{Input:} $ \{\N{L}_{kj}\}_{k,j=0}^{N_x-4}, c_0$
-%		\For{$k=0,1$}
-%		     \State $a_k \gets p_k$
-%		     \State $b_k \gets r_k$
-%		\EndFor
-%		\For{$k=2,3$}
-%		    \State $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2}$
-%		    \State $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}$
-%		\EndFor
-%		\For{$k=4, 5, \ldots, N_x-4$}
-%     		\State $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2} - \N{L}_{k, k-4} 
-%\,a_{k-4}$
-%	    	\State $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}- \N{L}_{k, k-4} 
-%\,b_{k-4}$
-%		\EndFor
-%		\For{$k=0, 1, \ldots, N_x-4$}
-%		    \State $a_k \gets a_k\cdot c_0$
-%		    \State $b_k \gets b_k \cdot c_0$
-%		\EndFor
-%		\State \Return $\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$
-%        \EndProcedure
-%\end{algorithmic}
-%\end{algorithm}
-
 \begin{algorithm}
-	\caption{Solve biharmonic Eq. (\ref{eq:ufin}) with pre-computed $\N{L}$, 
-	$\N{U}$ 
-	matrices. The parameters $q_j$ and $s_j$ are given in Eqs. (\ref{eq:pk}) 
-	and 
-	(\ref{eq:rk}) respectively. }
-	\label{alg:SolveBiharmonic}
+	\caption{Recursive formula to compute $\{a_k\}_{k=0}^{N_x-4}, 
+	\{b_k\}_{k=0}^{N_x-4}$ in Eq. (\ref{eq:ab}). Note that the lower triangular 
+	matrix $\{L_{kj}\}_{k,j=0}^{N_x-4}$ contains only two nonzero diagonals and 
+	the parameters $p_k$ and $r_k$ are given in Eqs. (\ref{eq:pk}) and 
+	(\ref{eq:rk}) respectively. \label{alg:ab}}
+	\begin{algorithmic}[1]
+		\Procedure{ReduceAB}{}
+		\State \textbf{Input:} $ \{\N{L}_{kj}\}_{k,j=0}^{N_x-4}, \xi_0$
+		\For{$k=0,1$}
+		     \State $a_k \gets p_k$
+		     \State $b_k \gets r_k$
+		\EndFor
+		\For{$k=2,3$}
+		    \State $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2}$
+		    \State $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}$
+		\EndFor
+		\For{$k=4, 5, \ldots, N_x-4$}
+     		\State $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2} - \N{L}_{k, k-4} 
+\,a_{k-4}$
+	    	\State $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}- \N{L}_{k, k-4} 
+\,b_{k-4}$
+		\EndFor
+
+		\State \Return $\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$
+        \EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+	\caption{Solve biharmonic Eq. (\ref{eq:ufin}) with pre-computed $\{\N{U}_{kj}\}_{k,j=0}^{N}, 
+		\{\N{L}_{kj}\}_{k,j=0}^{N}$ 
+	matrices, as well as the $\{ a_k\}_{k=0}^N$ and $\{ b_k\}_{k=0}^N$ vectors. The parameters $q_j$ and $s_j$ are given in Eqs. (\ref{eq:pk}) and 
+	(\ref{eq:rk}) respectively. 	\label{alg:SolveBiharmonic}}
 	\begin{algorithmic}[1]
@@
 		  \{\N{L}_{kj}\}_{k,j=0}^{N}, 
-		  \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N}, \{b_k\}_{k=0}^{N}$ 
+		  \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N}, \{b_k\}_{k=0}^{N}, \xi_0$ 
 		  \State // Solve $\N{L}_{kj}y_j=f_k$ by forward elimination 
@@
 		  substitution                    
-		  \State $u_N = y_N / \N{U}_{NN}$
-	      \State $u_{N-1} = y_{N-1} / \N{U}_{N-1,N-1}$
-	      \State $u_{N-2} = \left(y_{N-2} - \N{U}_{N-2, 
+		  \State $u_N \gets y_N / \N{U}_{NN}$
+	      \State $u_{N-1} \gets y_{N-1} / \N{U}_{N-1,N-1}$
+	      \State $u_{N-2} \gets \left(y_{N-2} - \N{U}_{N-2, 
 	      N}y_N\right)/\N{U}_{N-2,N-2}$
-	      \State $u_{N-3} = \left(y_{N-3} - \N{U}_{N-3, 
+	      \State $u_{N-3} \gets \left(y_{N-3} - \N{U}_{N-3, 
 	      N-1}y_{N-1}\right)/\N{U}_{N-3,N-3}$
 
-	      \State $q^o = q^e = s^o = s^e = 0$
+	      \State Float $q^o = q^e = s^o = s^e = 0.0$
 		  \For{$k = N-4$ \textbf{to} $0$ \textbf{step} $-1$}
@@
 		  \If{$k < N-5$}
-		  \State $j=k+6$
+		  \State $j \gets k+6$
 		  \If{$k$ is odd}
@@
 		    \State $s^o \gets s^o + u_{j}s_{j}$
-		    \State $u_k \gets u_k - q^o a_k - s^o b_k $	    
+		    \State $u_k \gets u_k - \xi_0q^o a_k - \xi_0s^o b_k $	    
 		  \Else
@@
 		    \State $s^e \gets s^e + u_{j}s_{j}$	  
-		    \State $u_k \gets u_k - q^e a_k - s^e b_k $
+		    \State $u_k \gets u_k - \xi_0q^e a_k - \xi_0s^e b_k $
 		  \EndIf
@@
 \section{Verification}
-We have described a Navier-Stokes solver for channel flows that scales no worse 
-than the fast transforms used to go between spectral and real space. All the 
-matrices that make up the linear systems can be inverted using sparse direct LU 
-decompositions that scales like $O(N)$. To verify and to further show that we 
-can use these algorithms for large scale simulations, we will first study some 
-simple 1D problems.
-
-We first consider the roundoff error of the matrix decomposition methods 
-described in Sec.~\ref{sec:implementation}. The roundoff error may be 
-approximated 
-by the following approach: Let $\{u_i\}_{i=0}^{N-4}$ be a uniformly distributed 
+\label{sec:verification}
+We have described a Navier-Stokes solver for channel flows applicable to large-scale simulations. The 
+matrices that make up the linear systems, that need to be solved in order to advance the Navier-Stokes equations, can be inverted using sparse direct LU 
+decompositions, and the solution to these systems may be obtained in $\approx \mathcal{O}(N_x)$. To verify and to further show that we 
+can use these algorithms for large-scale simulations, we will first study some simple 1D problems.
+
+Since we are describing a fully spectral solver, we should be concerned with roundoff errors. The roundoff error of the matrix decomposition methods 
+described for the biharmonic problem in Sec.~\ref{sec:implementation}, and similarly for the Helmholtz problem, may be estimated 
+by the following approach: Let $\{u_i\}_{i=0}^{N_x-4}$ be a uniformly distributed 
 random vector. Compute the matrix vector product $\bm{f}=\N{H}\bm{u}$ and 
 then solve $\N{H} \bm{v} = \bm{f}$, using Alg. \ref{alg:SolveBiharmonic}, for 
-$\{v_i\}_{i=0}^{i=N-4}$. The roundoff error of the matrix decomposition 
-algorithm may then be found as $\max_{0 \ge i \ge N-4}(|u_i-v_i|)/\max_{0 \ge i 
-\ge N-4}|u_i|$. For this experiment we assume that $\triangle t=0.0001$ 
-and $\nu=1/1200$, that are reasonable parameters for a turbulent channel flow 
-simulation at $\text{Re}_{\tau}=1200$. Three different wavenumbers 
-$z=(2, 200, 5600)$ are 
-also chosen as representative for simulations that are using up to 8000 
-wavenumbers in each periodic direction, which is approximately the resolution 
+$\{v_i\}_{i=0}^{N_x-4}$. The roundoff error of the matrix decomposition 
+algorithm may then be found as $\max_{0 \ge i \ge N_x-4}|u_i-v_i|/\max_{0 \ge i 
+\ge N_x-4}|u_i|$. For this experiment we assume that $\triangle t= 10^{-5}$ 
+and $\nu=1/5200$, that are reasonable parameters for a turbulent channel flow 
+simulation at $\text{Re}_{\tau}=5200$. Four different wavenumbers 
+$z=(0, 200, 1800, 5400)$ are 
+also chosen as representative for such large-scale simulations. The 1800 case corresponds to the highest wavenumber used for the resolution 
 of the largest known channel simulations, performed by Lee and 
-Moser~\cite{leemoser15}. The mesh in the wall normal direction 
-is varied from 64 to 4096 points, which covers almost three times higher 
-resolution than used by Lee and Moser \cite{leemoser15}. Results shown in the 
-first three rows of Table \ref{tab:roundoff} indicate that 
+Moser~\cite{leemoser15}, where maximum $\underline{z} = \sqrt{\underline{m}_{max}^2+\underline{n}_{max}^2} = \sqrt{(10240/8)^2 + (7680/6)^2} \approx 1800$. The mesh in the wall normal direction 
+is varied from 64 to 4096 points, which also covers almost three times higher 
+resolution than used by Lee and Moser. Results shown in the 
+first three rows of Table ~\ref{tab:roundoff} indicate that 
 roundoff errors are not very pronounced, even for problems as large as 
@@
 results are shown in the last three rows of Table \ref{tab:roundoff}. For 
-the Helmholtz matrix the roundoff errors are seen to be insignificant.
+the Helmholtz matrix roundoff errors are seen to be insignificant.
 
 \begin{table}
-	\caption{Roundoff error of the matrix decomposition method. The first three 
-	rows show results for Alg. (\ref{alg:SolveBiharmonic}) and matrix $\N{H}(z, 
+	\caption{ Roundoff error of the matrix decomposition method. The first three 
+	rows show results for Alg. (\ref{alg:SolveBiharmonic}) and matrix $\N{H}(\underline{z}, 
 	\triangle t, \nu)$, whereas the last three rows show results for matrix 
-	$\D{H}(z, \triangle t, \nu)$. Parameters 
-	are chosen as $\nu=1/1200$ and $\triangle t=0.0001$ and the numbers shown 
-	are for an average of 100 runs. }
-	\label{tab:roundoff}
+	$\D{H}(\underline{z}, \triangle t, \nu)$. Parameters 
+	are chosen as $\nu=1/5200$ and $\triangle t=10^{-5}$ and the numbers shown 
+	are for an average of 100 runs.	 \label{tab:roundoff}}
    \begin{tabular}{cccccccc}
@@
 	   \hline
-	$z$ & 64 & 128 & 256 & 512 & 1024 & 2048 & 4096 \\
+$\underline{z}$ & 64 & 128 & 256 & 512 & 1024 & 2048 & 4096 \\ 
 \hline
-0 & 2.66e-15 & 6.71e-15 & 1.73e-14 & 4.64e-14 & 1.18e-13 & 3.30e-13 & 9.21e-13 
-\\
-200 & 8.39e-15 & 1.78e-14 & 5.12e-14 & 1.56e-13 & 4.47e-13 & 1.18e-12 & 
-3.45e-12 \\
-5600 & 7.09e-15 & 2.08e-14 & 5.65e-14 & 1.43e-13 & 4.38e-13 & 1.01e-12 & 
-2.61e-12 \\
+0 & 3.27e-15  & 8.36e-15  & 2.10e-14  & 2.65e-14  & 2.88e-14  & 2.83e-14  & 3.29e-14  \\ 
+200 & 5.77e-14  & 8.50e-14  & 1.03e-13  & 1.08e-13  & 1.06e-13  & 1.06e-13  & 1.14e-13  \\ 
+1800 & 2.37e-13  & 1.65e-12  & 3.41e-12  & 3.33e-12  & 3.94e-12  & 3.79e-12  & 3.35e-12  \\ 
+5400 & 1.76e-13  & 1.84e-12  & 1.24e-11  & 1.57e-11  & 1.58e-11  & 1.55e-11  & 1.51e-11  \\ 
 \hline
-0 & 3.29e-15 & 6.07e-15 & 6.05e-15 & 6.50e-15 & 5.63e-15 & 5.43e-15 & 5.13e-15 
-\\
-200 & 3.03e-15 & 6.30e-15 & 6.14e-15 & 5.87e-15 & 5.67e-15 & 5.50e-15 & 
-5.39e-15 \\
-5600 & 3.35e-15 & 7.78e-15 & 8.86e-15 & 7.94e-15 & 7.18e-15 & 7.20e-15 & 
-7.03e-15
+0 & 3.26e-15  & 7.83e-15  & 1.89e-14  & 2.47e-14  & 2.16e-14  & 2.31e-14  & 1.94e-14  \\ 
+200 & 3.34e-15  & 8.63e-15  & 2.04e-14  & 2.54e-14  & 2.19e-14  & 2.01e-14  & 2.08e-14  \\ 
+1800 & 3.09e-15  & 8.23e-15  & 1.82e-14  & 2.27e-14  & 2.04e-14  & 2.18e-14  & 2.12e-14  \\ 
+5400 & 3.15e-15  & 7.77e-15  & 1.88e-14  & 2.25e-14  & 2.22e-14  & 2.10e-14  & 2.02e-14
 
@@
 for solving $\N{H} \bm{v} = \bm{f}$ and $\D{H} \bm{v} = \bm{f}$ on a complete 
-wavenumber mesh of size $(N_x,64,64)$. The numbers have been produced on a MacBook Pro laptop with the 2.8 GHz intel Core i7 processor. The average time for one single 
-solve is shown in Table~\ref{tab:timings}, where the numbers in parenthesis are used to indicate scaling (unity would be perfectly $O(N)$). The scaling is seen to be close to linear, and variations 
-are most likely due to how well datastructures fit in the cache. Naturally, the timings are not 100$\%$ reproducible, but the ability to solve large scale problems is evident.
+wavenumber mesh of size $(N_x,64,64)$.
+We use a MacBook Pro laptop with the 2.8 GHz intel Core i7 processor.
+The average time for one single solve is shown in Table~\ref{tab:timings}, where the numbers in parenthesis are used to indicate scaling (unity would be perfectly $\mathcal{O}(N_x)$).
+The scaling is seen to be close to linear, and variations
+are most likely due to how well datastructures fit in the cache. Naturally, the timings are not 100$\%$ reproducible, but the ability to solve large-scale problems is evident.
 \begin{table}
@@
 	(\ref{eq:gfin}) and Biharmonic 
-	(\ref{eq:ufin}) problems. The numbers in parenthesis are computed as $t_n / t_{n-1} / 2$, where $t_n$ is the current solvers timing in row $n$.}
-	\label{tab:timings}
+	(\ref{eq:ufin}) problems. The numbers in parenthesis are computed as $t_n / t_{n-1} / 2$, where $t_n$ is the current solvers timing in row $n$. 	\label{tab:timings}}
 	\begin{tabular}{ccc}
@@
 %8192 & 1.25e-04 (1.02) & 8.18e-05 (1.00) 
-Nx & Biharmonic & Helmholtz \\ 
+$N_x$ & Biharmonic & Helmholtz \\ 
 \hline
-64 & 8.09e-07 (0.00) & 4.51e-07 (0.00) \\ 
-128 & 1.57e-06 (0.97) & 9.41e-07 (1.04) \\ 
-256 & 2.76e-06 (0.88) & 1.97e-06 (1.05) \\ 
-512 & 4.96e-06 (0.90) & 4.22e-06 (1.07) \\ 
-1024 & 1.18e-05 (1.18) & 8.40e-06 (1.00) \\ 
-2048 & 2.37e-05 (1.01) & 1.79e-05 (1.07) \\ 
-4096 & 4.82e-05 (1.02) & 3.78e-05 (1.05) \\ 
-8192 & 9.72e-05 (1.01) & 7.91e-05 (1.05)
+64 & 6.03e-07 (0.00) & 4.48e-07 (0.00) \\ 
+128 & 1.20e-06 (0.99) & 9.00e-07 (1.01) \\ 
+256 & 2.30e-06 (0.96) & 1.81e-06 (1.01) \\ 
+512 & 4.44e-06 (0.96) & 3.59e-06 (0.99) \\ 
+1024 & 1.02e-05 (1.15) & 7.72e-06 (1.07) \\ 
+2048 & 2.03e-05 (0.99) & 1.68e-05 (1.09) \\ 
+4096 & 4.14e-05 (1.02) & 3.33e-05 (0.99) \\ 
+8192 & 8.25e-05 (1.00) & 6.92e-05 (1.04)
 	\end{tabular}
 \end{table}
+
+The complete Navier Stokes solver should scale like $\mathcal{O}(aN + bN\log N)$ since it is a combination of linear algebra routines $\mathcal{O}(N)$, fast transforms $\mathcal{O}(N \log N)$ and several elementwise array operations $\mathcal{O}(N)$.
+We show in Table \ref{tab:timings_solver} some timings for one single time step, increasing only the wall normal discretization.
+We show that the total time scales approximately as $N \log N$ for large $N$ and better than linearly for small $N$.
+This shift happens since at small $N$ the transforms are fast and linear operations, like elementwise multiplications, are more dominating.
+We also see that the time for assembly of the right hand side is approximately $\mathcal{O}(N \log N)$, whereas the time for solving the linear systems is very nearly linear.
+
+\begin{table}
+	\centering
+	\caption{Computing time for one time step using 32 wavenumbers in each periodic direction and $N_x$ points in the wall normal direction.
+The Assemble and Solve columns show the time required to assemble the right hand side (should be $\mathcal{O}(N_x \log N_x)$ due to transforms) and solve the linear systems, respectively.
+The number in parenthesis shows the scaling, and unity would mean $\mathcal{O}(N_X \log N_x)$ for Total and Assemble, whereas it would indicate $\mathcal{O}(N_x)$ for Solve.
+\label{tab:timings_solver}}
+	\begin{tabular}{cccc}	
+$N_x$ & Total ($\frac{t_k }{t_{k-1}} \frac{\log N-1}{2 \log N}$) & Assemble ($\frac{t_k }{t_{k-1}} \frac{\log N-1}{2 \log N}$) & Solve ($\frac{t_k}{t_{k-1}} \frac{1}{2}$) \\ 
+\hline
+32 & 0.010 (0.00) & 7.83e-03 (0.00) & 1.07e-03 (0.00) \\ 
+64 & 0.019 (0.86) & 1.53e-02 (0.88) & 1.98e-03 (0.92) \\ 
+128 & 0.040 (0.96) & 3.23e-02 (0.97) & 3.81e-03 (0.96) \\ 
+256 & 0.085 (0.98) & 6.77e-02 (0.98) & 7.58e-03 (1.00) \\ 
+512 & 0.185 (1.04) & 1.53e-01 (1.07) & 1.58e-02 (1.04) \\ 
+1024 & 0.384 (0.99) & 3.18e-01 (0.99) & 3.13e-02 (0.99) \\ 
+2048 & 0.816 (1.02) & 6.98e-01 (1.05) & 6.45e-02 (1.03) \\ 
+4096 & 1.736 (1.03) & 1.44e+00 (0.99) & 1.28e-01 (0.99)
+    \end{tabular}
+\end{table}
+
+The Orr-Sommerfeld equations are further used to establish that we are solving the right equations, and that the accuracy of the solver is second order in time.
+The Orr-Sommerfeld eigenvalue problem is first solved for $Re=8000$ using the biharmonic Shen basis (\ref{eq:phiN}).
+The leading eigenvalue/eigenvector pair is then added as a perturbation to a laminar, parabolic, base flow, leading to an initial field
+\begin{align}
+u(x,y,t) &= -\epsilon  \mathrm{Re}\{\xi(x) \exp (\imath (y-\lambda t))\}, \label{eq:OS1} \\
+v(x,y,t) &= 1-x^2 + \epsilon \mathrm{Re}\{\imath \xi(x) \exp(\imath(y-\lambda t)\}, \label{eq:OS2}
+\end{align}
+where $\xi(x)$ is the eigenvector, $\lambda$ the eigenvalue and $\epsilon \in \mathcal{R}^+$ is a small positive constant.
+Note that if $\epsilon$ is chosen as too big, then the initial/boundary value problem will not agree well with the linearized Navier-Stokes equations used in deriving the Orr-Sommerfeld equations, and the evolution from (\ref{eq:OS1}, \ref{eq:OS2}) will not follow the linearized equations.
+Likewise, if $\epsilon$ is chosen as too small, then roundoff errors will be too large, since the perturbation is added to a base flow of order unity.
+Since the evolution of the perturbation is known analytically, the departure from this exact linearized solution is used as a measure of the error of the current solver.
+We also compute the error in the perturbation energy by
+\begin{equation}
+E(t) =  \frac{\left<\tilde{\bm{u}}(\bm{x}, t), \tilde{\bm{u}}(\bm{x}, t) \right>}{\left<\tilde{\bm{u}}(\bm{x}, 0), \tilde{\bm{u}}(\bm{x}, 0) \right>}  - \exp(2 \mathrm{Im}\{\lambda\}t), \label{eq:OSenergy}
+\end{equation}
+where $\tilde{\bm{u}} = [u, v-(1-x^2)]$ represents the perturbation vector, and $\lambda=0.2470750602+\imath 2.664410371\cdot 10^{-3}$ is the eigenvalue.
+
+Table \ref{tab:epsilon} shows the perturbation error as a function of $\epsilon$, using a mesh of size $128 \times 8 \times 2$, $\triangle t=0.2$ and final time $t=5$.
+Since the perturbation is added to a base flow of order unity, we know that one digit will be lost in accuracy, for each row, moving down Table \ref{tab:epsilon}.
+Adding to this the normal roundoff, see Table \ref{tab:roundoff}, $\epsilon=10^{-8}$ should not be expected to yield more than 6-7 digits of accuracy, whereas the final row ($\epsilon=10^{-12}$) no more than 2-3.
+Still, we observe that the solution is nearly independent of $\epsilon$ for $10^{-6} < \epsilon < 10^{-8}$, and as such $\epsilon=10^{-7}$ is chosen as suitable for computing the temporal accuracy.
+\begin{table}
+	\centering
+	\caption{ Error in Orr-Sommerfeld perturbation as a function of perturbation magnitude $\epsilon$.
+The $L^2$-norm $ \| \bm{u}' \| = \sqrt{\left< \bm{u}', \bm{u}'\right>}$, where $\bm{u}'=\bm{u}-\bm{u}_{\mathrm{exact}}$ and energy error (\ref{eq:OSenergy}), for the most unstable Orr-Sommerfeld mode.
+Results are computed with a mesh of size $128 \times 8 \times 2$, $\triangle t=0.2$ and final time $t=5$.
+\label{tab:epsilon}}
+	\begin{tabular}{ccc}	
+		$-\log_{10}(\epsilon)$ & $\frac{1}{\epsilon}\| \bm{u}' \|(t=5)$ & $E(t=5)$ \\
+		\hline
+		4 & 1.234544e-02 & 2.441506e-04 \\ 
+		5 & 1.157535e-02 & 2.440520e-04 \\ 
+		6 & 1.156739e-02 & 2.440510e-04 \\ 
+		7 & 1.156732e-02 & 2.440510e-04 \\ 
+		8 & 1.156731e-02 & 2.440504e-04 \\ 
+		9 & 1.156732e-02 & 2.440498e-04 \\ 
+		10 & 1.156726e-02 & 2.440435e-04 \\ 
+		11 & 1.156497e-02 & 2.437132e-04 \\ 
+		12 & 1.155687e-02 & 2.435411e-04
+	\end{tabular}
+\end{table}
+
+Table \ref{tab:OS} shows results from using a mesh of size $128 \times 8 \times 2$, with final time $t=50$, corresponding to 2 flow-throughs, and a varying $\triangle t$.
+We observe, as expected, second order accuracy in the $L^2$-norm in time.
+The error in the accumulated perturbation energy is also found to be close to second order.
+
+\begin{table}
+	\centering
+	\caption{Error in Orr-Sommerfeld perturbation as a function of time step.
+The mesh size is $128 \times 8 \times 2$.
+The $L^2$-norm $ \| \bm{u}' \| = \sqrt{\left< \bm{u}', \bm{u}'\right>}$, where $\bm{u}'=\bm{u}-\bm{u}_{\mathrm{exact}}$, and energy error (\ref{eq:OSenergy}), for the most unstable Orr-Sommerfeld mode.
+The numbers in parenthesis are showing the order of the error, computed as $\log(\tau_i/\tau_{i-1})/\log(\triangle t_i/ \triangle t_{i-1})$, where $i$ is the row number of the table and $\tau$ represents the values of $\|\bm{u}'\|$ or $E(t=50)$.
+\label{tab:OS}}
+	\begin{tabular}{ccc}	
+ $\triangle t$ & $\| \bm{u}' \|(t=50)$ & $\int_{t=0}^{t=50}E(t) \mathrm{d}t$ \\
+ \hline
+
+1.0000e-01 & 2.9005e-09  (0.000) & 1.2775e-02 (0.000) \\
+6.6667e-02 & 1.2877e-09  (2.003) & 5.3068e-03 (2.167) \\
+5.0000e-02 & 7.2559e-10  (1.994) & 2.8939e-03 (2.108) \\
+4.0000e-02 & 4.6445e-10  (1.999) & 1.8138e-03 (2.094) \\
+3.3333e-02 & 3.2257e-10  (1.999) & 1.2418e-03 (2.078) \\
+2.8571e-02 & 2.3701e-10  (1.999) & 9.0307e-04 (2.066) \\
+2.5000e-02 & 1.8148e-10  (1.999) & 6.8608e-04 (2.058)
+	\end{tabular}
+\end{table}
+
+The ultimate test for this fully spectral Navier-Stokes solver is the pressure driven turbulent channel flow.
+We use the turbulent Reynolds number $Re_{\tau}=2000$, which has been simulated previously by Hoyas and Jim\'{e}nez \cite{hoyas06}, Lozano and Jim\'{e}nez \cite{Lozano2014} and Lee and Moser \cite{leemoser15}.
+Lozano and Jim\'{e}nez showed that a small box  of size $2 \times 2\pi \times \pi$ in $x, y,$ and $z$ directions, respectively, could be simulated with $633 \times 1024 \times 1024$ collocation points (compact finite differences in wall normal direction, Fourier in periodic) in order to capture all relevant one-point statistics found in much larger boxes.
+We use the same small computational box, with resolution of $512 \times 1024 \times 1024$, and expect the superior resolution properties of spectral methods to work in our favor.
+The solver has been implemented with a slab MPI decomposition, and simulations were run on the Shaheen II supercomputer at KAUST using 512 cores.
+With a time step of $2\cdot10^{-5}$, the solver used approximately 1 second per time step.
+After reaching steady state, statistics were sampled for 100,000 time steps (enough to converge first order statistics), and the average velocity is shown in Figure \ref{fig:U_mean} in a dashed line, alongside the results obtained by Hoyas and Jim\'{e}nez \cite{hoyas06} in a dotted line.
+The difference is hardly visible, showing that the fully spectral Shen-Fourier solver is reliable for $Re_{\tau} = 2000$.
+Furthermore, the results presented in Tables \ref{tab:roundoff}, \ref{tab:timings} and \ref{tab:timings_solver} indicate that the solver is more than likely to be reliable for even larger scales.
+
+\begin{figure}[t]
+	\begin{center}
+	\includegraphics[width=0.9\textwidth]{U_Re2000.eps}
+	\caption{Average velocity $\D{v}^+ = \D{v}/u_{\tau}$ in $y$-direction, for $Re_{\tau}=2000$. Our results (dashed) are compared with those of Hoyas and Jim\'{e}nez \cite{hoyas06} (dotted).}
+	\label{fig:U_mean}
+	\end{center}
+\end{figure}
+
+\section{Conclusions}
 
@@
 \section{Appendices}
+\label{sec:app}
 %\subsection*{Fast Shen transforms}
@@
 %\end{python}
+
+\subsection*{Fast scalar products and transforms}
+
+The algorithms for the fast scalar products and transforms, for the three spaces $V_N, \D{V}_N, \N{V}_N$, are given in Alg.
+~\ref{alg:fst}.
+The algorithms for the inverse transforms are given in Alg.~\ref{alg:ifst}.
+The algorithms are using the discrete cosine transforms (dct) of type 1, 2 or 3 depending on the choice of discretization points and direction.
+\begin{algorithm}
+	\caption{Forward scalar products and transforms for all spaces $V_N, \D{V}_N, \N{V}_N$. 
+		Here "dct" is the discrete cosine transform from SciPy.}
+		\label{alg:fst}
+		\begin{algorithmic}[1]
+			\Function{forwardScalar}{$\{f_j\}_{j=0}^{N}$, points, space}
+			\If{ points = "Gauss-Lobatto"}
+			\State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, \text{type=2}, 
+			\text{axis=0})\frac{\pi}{2 N}$
+			\Else{ points = "Gauss-Chebyshev"}
+			\State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=1}, 
+			\text{axis=0}) \frac{\pi}{2 (N-1)}$
+			\EndIf
+			\State $\{y\}_{k=0}^N \gets 0$
+			\If{ space = $\D{V}_N$}
+			\State $\{y_k\}_{k=0}^{N-2} \gets \{w_k\}_{k=0}^{N-2} - \{w_k\}_{k=2}^{N}$
+			\ElsIf{ space = $\N{V}_N$}	                
+			\For{$k=0$ to $N-4$}
+			\State ${y}_k \gets w_k - \frac{2(k+2)}{k+3} w_{k+2} + 
+			\frac{k+1}{k+3}w_{k+4}$
+			\EndFor
+			\ElsIf{space = $V_N$}
+			\State $\{y_k\}_{k=0}^N \gets \frac{2}{\pi} \{w_k\}_{k=0}^N $
+			\State $y_0 \gets y_0 / 2$
+			\If{ points = "Gauss-Lobatto"}
+			\State $y_N \gets y_N / 2$	     	
+			\EndIf
+			\EndIf
+			\State \Return $\{y_k\}_{k=0}^{N}$    
+			\EndFunction
+			\Function{forwardTransform}{$\{f_j\}_{j=0}^{N}$, points, space}
+			\State $\{y_k\}_{k=0}^{N} \gets$ forwardScalar($\{f_j\}_{j=0}^{N}$, points, space)
+			\If{space = $\D{V}_N$}
+			\State $B \gets \D{B}$
+			\ElsIf{space = $\N{V}_N$}
+			\State $B \gets \N{B}$
+			\ElsIf{space = $V_N$}
+			\State $B \gets B$
+			\EndIf
+			\State $\bm{y} \gets B^{-1} \bm{y}$
+			\State \Return $\{y_k\}_{k=0}^{N}$
+			\EndFunction
+
+		\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+	\caption{Inverse transforms for all spaces $V_N, \D{V}_N, \N{V}_N$. 
+		Here "dct" is the discrete cosine transform from SciPy.}
+	\label{alg:ifst}
+	\begin{algorithmic}[1]
+		\Function{inverseChebTransform}{$\{f\}_{j=0}^N$, points}
+		\If{points = "Gauss-Chebyshev"}
+		    \State $\{w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, type=3, axis=0)$
+		    \State $\{w_k\}_{k=0}^{N} \gets \frac{1}{2}\{w_k\}_{k=0}^{N} + \frac{1}{2}f_0$
+
+        \ElsIf{ points = "Gauss-Lobatto"}
+	        \State $\{w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, type=1, axis=0)$
+	        \State $\{w_k\}_{k=0}^{N} \gets \frac{1}{2}\{w_k\}_{k=0}^{N} + \frac{1}{2}f_0$
+	        \For{$k=0$ \textbf{to} $N$}
+	         \State $w_k \gets \frac{(-1)^k}{2} f_N$
+	        \EndFor
+	    \EndIf
+	    \State \Return{$\{w_k\}_{k=0}^{N}$}
+
+	    \EndFunction
+
+		\Function{inverseShenTransform}{$\{f_j\}_{j=0}^{N}$, points, space}
+
+		\If{ space = $\D{V}_N$}
+		\State $\{w_k\}_{k=0}^{N-2} \gets \{f_k\}_{k=0}^{N-2}$
+		\State $\{w_k\}_{k=2}^{N} \gets \{w_k\}_{k=2}^{N} - \{f_k\}_{k=0}^{N-2} $
+        \ElsIf{ space = $\N{V}_N$}	                
+		\State $\{w_k\}_{k=0}^{N-4} \gets \{f_k\}_{k=0}^{N-4}$
+		\For{$k=0$ \textbf{to} $N-4$}
+		\State $w_{k+2} \gets w_{k+2} - \frac{2(k+2)}{k+3} f_k$
+		\EndFor
+		\For{$k=0$ \textbf{to} $N-4$}
+		\State $w_{k+4} \gets w_{k+4} + \frac{k+1}{k+3} f_k$
+		\EndFor
+        \EndIf
+        \State $\{{w}_k\}_{k=0}^{N}$ = inverseChebTransform($\{w_k\}_{k=0}^N$, points)
+		\State \Return $\{{w}_k\}_{k=0}^{N}$
+		\EndFunction
+	\end{algorithmic}
+\end{algorithm}
+
 \subsection*{Matrices}
 
-\begin{equation}
- \{\D{A}_{kj}\}_{k,j=0}^{N_x-2} = -\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w = \begin{cases}
- 2\pi(k+1)(k+2), &j=k,\\
+Matrices representing one-dimensional scalar products are given in Table \ref{tab:matrices}.
+Note that the mass matrices are symmetrical, and as such only the upper triangular parts are described.
+Also note that the matrices are computed using quadrature, and not exact integration, which has some minor implications for Gauss-Lobatto (GL) points.
+This follows since for an exact $L^2_w$ scalar product we would have $c_{N_x}=1$ for GL as well as Gauss Chebyshev (GC), but for the $l^2_w$ scalar product used here $c_{N_x}=2$.
+This disagreement, that follows from inexact quadrature at the highest mode, is not reported by Shen \cite{Shen95}, but is reported by, e.g., Kopriva \cite{kopriva09}.
+This explains the inclusion of the $c_{k+4}$ term for matrix $\N{B}$, and the $c_{k+2}$ term for matrix $\D{B}$, that are missing from Lemma 2.1 and 3.1 of \cite{Shen95}.
+
+\begin{table}
+	\centering
+	\caption{Matrix notation and description for one dimensional scalar products.
+Note that $c_0=2$ and $c_k=1$ for $0<k<N_x$.
+For Gauss-Chebyshev points $c_{N_x}=1$, whereas $c_{N_x}=2$ for Gauss-Lobatto points.	\label{tab:matrices}}
+	\begin{tabular}{ccl}	
+		Notation & Scalar product & Description \\ 
+		\hline
+		$\{\D{A}_{kj}\}_{k,j=0}^{N_x-2}$ & $-\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w$ &
+$\begin{cases} 2\pi(k+1)(k+2), &j=k,\\
 4\pi(k+1), & j=k+2, k+4, k+6, \ldots, \\
- 0, &\text{otherwise}.
- \end{cases}
-\end{equation}
-
-\begin{equation}
- \{\N{A}_{kj}\}_{k,j=1}^{N_x-2} = -\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w = \begin{cases}
- 2\pi j^2(k+1)/(k+2), &j=k,\\
+0, &\text{otherwise}.\end{cases}$		\\
+
+$\{\N{A}_{kj}\}_{k,j=1}^{N_x-2}$ & $-\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w$ & $\begin{cases} 2\pi j^2(k+1)/(k+2), &j=k,\\
 4\pi j^2(k+1)/(k+2)^2, & j=k+2, k+4, k+6, \ldots, \\
- 0, &j>k, \text{ or } k+j \text{ odd}.
- \end{cases}
-\end{equation}
-
-\begin{equation}
- \{\N{C}_{kj}\}_{k,j=0}^{N_x-4, N_x-2} = \left(\D{\phi}^{'}_j, \N{\phi}_k 
- \right)_w = \begin{cases}
- -\pi(k+1), &j=k-1,\\
+0, &j>k, \text{ or } k+j \text{ odd} \end{cases}$ \\
+
+$\{\N{C}_{kj}\}_{k,j=0}^{N_x-4, N_x-2}$ & $\left(\D{\phi}^{'}_j, \N{\phi}_k 
+\right)_w$ & $\begin{cases} -\pi(k+1), &j=k-1,\\
 2\pi(k+1), & j=k+1, \\
 -\pi(k+1), & j=k+3, \\
- 0, &\text{otherwise}.
- \end{cases}
-\end{equation}
-
-\begin{equation}
- \{\D{C}_{kj}\}_{k,j=0}^{N_x-2, N_x-4} = \left(\N{\phi}^{'}_j, \D{\phi}_k 
- \right)_w = \begin{cases}
+0, &\text{otherwise} \end{cases}$ \\
+
+$\{\D{C}_{kj}\}_{k,j=0}^{N_x-2, N_x-4}$ & $\left(\N{\phi}^{'}_j, \D{\phi}_k 
+\right)_w$ & $\begin{cases}
 \pi (k-2)(k+1)/k, &j=k-3,\\
@@
 \pi(k+1), & j=k+1, \\
- 0, &\text{otherwise}.
- \end{cases}
-\end{equation}
-
-\begin{equation}
- \{\N{B}_{kj}\}_{k,j=0}^{N_x-4} = \{\N{B}_{jk}\}_{j,k=0}^{N_x-4} = (\N{\phi}_j, 
- \N{\phi}_k)_w = \begin{cases}
+0, &\text{otherwise}
+\end{cases}$ \\
+
+$\{\N{B}_{kj}\}_{k,j=0}^{N_x-4} = \{\N{B}_{jk}\}_{j,k=0}^{N_x-4}$ & $(\N{\phi}_j, \N{\phi}_k)_w$ & $\begin{cases}
 \frac{\pi}{2} \left(c_k + 4 \left(\frac{k+2}{k+3} \right)^2 + c_{k+4} 
 \left(\frac{k+1}{k+3}\right)^2    \right), &j=k,\\
- -\pi \left( \frac{k+2}{k+3} + \frac{(k+4)(k+1)}{(k+5)(k+3)} \right), &j=k+2,\\
- \frac{\pi}{2} \frac{k+1}{k+3} , & j=k+4, \\
- 0, &\text{otherwise}.
- \end{cases}
-\end{equation}
-
-\begin{equation}
- \{B_{kj}\}_{k,j=0}^{N_x} = (T_k, T_j)_w = \frac{c_k \pi}{2} \delta_{kj}
-\end{equation}
-
-For Gauss-Lobatto points we have $c_0=c_{N_x}=2$ and $c_k=1$ for $0<k<N_x$.
-For Gauss-Chebyshev points we have $c_0=2$ and $c_k=1$ for $k>0$.
-Note that for an $L^2_w$ scalar product we would have $c_{N_x}=1$ as well, but for the $l^2_w$ scalar product used here we get $c_{N_x}=2$.
-This disagreement, that follows from inexact quadrature at the highest mode, is not reported by Shen, but is reported by, e.g., Kopriva.
+-\pi \left( \frac{k+2}{k+3} + \frac{(k+4)(k+1)}{(k+5)(k+3)} \right), &j=k + 2,\\
+\frac{\pi}{2} \frac{k+1}{k+3} , & j=k + 4, \\
+0, &\text{otherwise}
+\end{cases} $\\
+
+$\{\D{B}_{kj}\}_{k,j=0}^{N_x-2} = \{\D{B}_{jk}\}_{j,k=0}^{N_x-2}$ & $(\D{\phi}_j, \D{\phi}_k)_w$ & $ \begin{cases} 
+\frac{\pi}{2} (c_k+c_{k+2}) &j=k, \\
+-\frac{\pi}{2} &j=k + 2, \\
+0, &\text{otherwise}
+\end{cases}$ \\
+
+$\{B_{kj}\}_{k,j=0}^{N_x} = \{B_{jk}\}_{j,k=0}^{N_x}$ & $(T_k, T_j)_w$ & $\frac{c_k \pi}{2} \delta_{kj}$
+
+	\end{tabular}
+\end{table}
 

### 2016-12-06 paper.tex
Algorithm2e

@@
-\documentclass[11pt, oneside]{elsarticle}
+\documentclass[preprint]{elsarticle}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{subfigure}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-
-\subfiglabelskip=0pt
-
-%\newcommand{\codetitlestyle}[1]{\small\textit{#1}\hspace{0.1cm}}
-\newcommand{\belowtitleskip}{2pt}%\smallskipamount}
-%\newcommand{\captionposition}{t}
+\usepackage[ruled,vlined]{algorithm2e}
+%\usepackage{algpseudocode}
 
@@
 \label{sec:implementation}
-An outline of the solution procedure used, for the numerical method described in Sec.~\ref{sec:discretizationNS}, is given in Algorithm \ref{alg:NS}.
+An outline of the solution procedure used, for the numerical method described 
+in Sec.~\ref{sec:discretizationNS}, is given in Algorithm \ref{alg:NS}. Note 
+that the Adams-Bashforth projection of convection reverts to a first order 
+forward Euler on the first time step. For simulations of turbulent flows this 
+is insignificant, but for accuracy tests, like the Orr-Sommerfeld instability, 
+a proper initialization would need to set both the two first time steps instead.
 \begin{algorithm}
-	\caption{Solution procedure for Navier Stokes equations.     \label{alg:NS}}
-    \begin{algorithmic}[1]
-    	\State \text{Initialize} $\bm{u}(\bm{x}, 0)$ and compute 
-    	$\hat{g}_{\bm{k}}(0)$ and $\hat{u}_{\bm{k}}(0)$
-		\State \text{Compute LU decomposition of} $\N{H}(m,n), \D{H}(m,n) 
-		\,\forall \, m, n \in K^p$
-		\State \text{Compute LU decomposition of coefficient matrices in} (\ref{eq:v00}) and (\ref{eq:w00})
-		\State $t=0; {\kappa}=0$
-    	\While{$t<T$}
-	    	\State \text{Compute nonlinear convection} $\bm{\hat{\mathcal{H}}}^{\kappa}$
-	    	\State \text{Compute right hand sides of Eqs.~}(\ref{eq:ufin}), 
-	    	(\ref{eq:gfin})
-		    \State \text{Solve Eq.}(\ref{eq:ufin}) for 
-		    $\hat{u}^{{\kappa}+1}_{\bm{k}} \, \forall \,\bm{k} \in \N{K}_N$
-		    \State \text{Solve Eqs.}(\ref{eq:gfin}) for 
-		    $\hat{g}^{{\kappa}+1}_{\bm{k}} \, \forall\, \bm{k} \in \D{K}_N$
-		    \State \text{Solve Eqs.}(\ref{eq:f3}, \ref{eq:g3}) for  
+	\caption{Solution procedure for Navier Stokes equations.    
+	 \label{alg:NS}}
+ 	\DontPrintSemicolon	
+    	Initialize $\bm{u}(\bm{x}, 0)$ and compute 
+    	$\hat{g}_{\bm{k}}(0)$ and $\hat{\bm{u}}_{\bm{k}}(0)$ \;
+		Compute LU decomposition of $\N{H}(m,n), \D{H}(m,n) 
+		\,\forall \, m, n \in K^p$ \;
+		Compute LU decomposition of coefficient matrices in (\ref{eq:v00}) and 
+		(\ref{eq:w00}) \;
+		$t \gets 0$\; 
+		${\kappa} \gets 0$ \;
+    	Compute nonlinear convection 
+    	$\bm{\hat{\mathcal{H}}}^{\kappa-1}(\hat{\bm{u}}_{\bm{k}}(0), 
+    	\hat{g}_{\bm{k}}(0))$ \;
+    	\While{$t<T$}{
+	    	Compute nonlinear convection $\bm{\hat{\mathcal{H}}}^{\kappa}$ \;
+	    	{Compute right hand sides of Eqs.~}(\ref{eq:ufin}), 
+	    	(\ref{eq:gfin}) \;
+		    {Solve Eq.}(\ref{eq:ufin}) for 
+		    $\hat{u}^{{\kappa}+1}_{\bm{k}} \, \forall \,\bm{k} \in \N{K}_N$ \;
+		    {Solve Eqs.}(\ref{eq:gfin}) for 
+		    $\hat{g}^{{\kappa}+1}_{\bm{k}} \, \forall\, \bm{k} \in \D{K}_N$ \;
+		    {Solve Eqs.}(\ref{eq:f3}, \ref{eq:g3}) for  
 		    $\hat{v}_{\bm{k}}, \hat{w}_{\bm{k}}\, \forall\, \bm{k} 
-		    \in \D{K}_N^0$
-		    \State \text{Solve Eqs.}(\ref{eq:v00}, \ref{eq:w00}) for 
+		    \in \D{K}_N^0$ \;
+		    {Solve Eqs.}(\ref{eq:v00}, \ref{eq:w00}) for 
 		    $\hat{v}_0^{{\kappa}+1}, \hat{w}_0^{{\kappa}+1}\, 
-		    \forall \, l \in \D{K}^x$
-		    \State ${\kappa} \gets {\kappa} + 1$
-		    \State $t \gets t+\triangle t$
-		    \State \text{Update to new time step}
-	    \EndWhile
-
-    \end{algorithmic}
+		    \forall \, l \in \D{K}^x$ \;
+		    ${\kappa} \gets {\kappa} + 1$ \;
+		    $t \gets t+\triangle t$ \;
+		    {Update to new time step} \;
+	    }
 \end{algorithm}
-The major computational cost comes from  computing the nonlinear convection 
-and solving for Eqs.~(\ref{eq:ufin}, \ref{eq:gfin}). The nonlinear 
-convection 
-term, $\bm{\hat{\mathcal{H}}}$, is computed using a pseudo-spectral method with 
-the 3/2-rule for dealiasing in the periodic directions, see \cite{canuto1988}. The 
+
+The major computational cost in Alg. \ref{alg:NS} comes from  computing the 
+nonlinear convection and solving for Eqs.~(\ref{eq:ufin}, \ref{eq:gfin}). The 
+nonlinear convection term, $\bm{\hat{\mathcal{H}}}$, is computed using a 
+pseudo-spectral method with the 3/2-rule for dealiasing in the periodic 
+directions, see \cite{canuto1988}. The 
 transforms are carried out using fast Fourier and fast Chebyshev transforms 
@@
 
+\SetKwProg{Fn}{Function}{}{}
+\SetKwFunction{reduceAB}{ReduceAB}
 \begin{algorithm}
-	\caption{LU decomposition of biharmonic operator $\N{H}$. Current algorithm 
-	is using dense storage of $\N{U}_{kj}$, but no more than the 3 items per row
-	need 
-	to be stored after the row is no longer needed in the for-loop below (as the loop is traversed we do not need $\N{U}_{m, j}$ for $m<k-4$). \label{alg:lu_biharmonic}}
-	\begin{algorithmic}[1]
-		\Procedure{LUbiharmonic}{}
-			\State \textbf{Input:} $\{\N{H}_{kj}\}_{k, j = 0}^{N}$
-			\State Integer $M_e \gets \lfloor N/2 \rfloor$
-			\State Integer $M_o \gets \lfloor (N-1)/2 \rfloor$  
-	        \State $\{\hat{U}_{0,2j}\}_{j=0}^{M_e} \gets \{\hat{H}_{0,2j}\}_{j=0}^{M_e}$ 
-	        \State $\{\hat{U}_{1,2j+1}\}_{j=0}^{M_o} \gets \{\hat{H}_{1,2j+1}\}_{j=0}^{M_o}$ 
-	        \State $\N{L}_{20} \gets \N{H}_{20}/\N{U}_{00}$
-	        \State $\N{L}_{31} \gets \N{H}_{31}/\N{U}_{11}$
-	        \State $\{\hat{U}_{2,2j}\}_{j=1}^{M_e} \gets \{\N{H}_{2,2j}\}_{j=1}^{M_e} - \N{L}_{20} \{\N{U}_{0,2j}\}_{j=1}^{M_e}$
-	        \State $\{\hat{U}_{3,2j+1}\}_{j=1}^{M_o} \gets \{\N{H}_{3,2j+1}\}_{j=1}^{M_o} - \N{L}_{31} \{\N{U}_{1,2j+1}\}_{j=1}^{M_o}$
-
-	        %\State $\N{U}_{2j} = \N{H}_{2j} - \N{L}_{20} \N{U}_{0j} \quad 
-	        %\forall \, j \in 2, 4, \ldots, N$
-	        %\State $\N{U}_{3j} = \N{H}_{3j} - \N{L}_{31} \N{U}_{1j} \quad 
-	        %\forall \, j \in 
-	        %3, 5, \ldots, N$
-	        \For{$k = 4, 5, \ldots, N$}
-	            \State Integer $M \gets \lfloor \frac{N-k}{2}\rfloor$
-		        \State $\N{L}_{k,k-4} \gets \N{H}_{k, k-4} / \N{U}_{k-4, k-4}$
-		        \State $\{\N{U}_{k,k+2j}\}_{j=-1}^{M} \gets \{\N{H}_{k,k+2j}\}_{j=-1}^{M} - \N{L}_{k, k-4} 
-		        \{\N{U}_{k-4, k+2j}\}_{j=-1}^{M}$
-		        %\State $\N{U}_{k,j} = \N{H}_{k,j} - \N{L}_{k, k-4}  \N{U}_{k-4, j} \quad \forall j = k-2, k, \ldots, N$
-	            \State $\N{L}_{k,k-2} \gets {\N{U}}_{k,k-2}/{\N{U}}_{k-2, k-2}$
-	            \State $\{\N{U}_{k,k+2j}\}_{j=0}^{M} \gets \{{\N{U}}_{k,k+2j}\}_{j=0}^{M} - \N{L}_{k, k-2} \{\N{U}_{k-2,k+2j}\}_{j=0}^{M}$
-		        %\State $\N{U}_{kj} = {\N{U}}_{kj} - \N{L}_{k, k-2} \N{U}_{k-2, j} \quad \forall j = k, k+2, \ldots, N$
-	        \EndFor
-	        \State \Return $\{\N{L}_{kj}\}_{k, j = 0}^{N}$, $\{\N{U}_{kj}\}_{k, 
-	        j = 
-	        0}^{N}$
-		\EndProcedure	
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
+	\DontPrintSemicolon
     \caption{Recursive formula to compute $\{a_k\}_{k=0}^{N_x-4}, 
@@
 	(\ref{eq:rk}) respectively. \label{alg:ab}}
-	\begin{algorithmic}[1]
-		\Procedure{ReduceAB}{}
-		\State \textbf{Input:} $ \{\N{L}_{kj}\}_{k,j=0}^{N_x-4}, \xi_0$
-		\For{$k=0,1$}
-		     \State $a_k \gets p_k$
-		     \State $b_k \gets r_k$
-		\EndFor
-		\For{$k=2,3$}
-		    \State $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2}$
-		    \State $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}$
-		\EndFor
-		\For{$k=4, 5, \ldots, N_x-4$}
-     		\State $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2} - \N{L}_{k, k-4} 
-\,a_{k-4}$
-	    	\State $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}- \N{L}_{k, k-4} 
-\,b_{k-4}$
-		\EndFor
-
-		\State \Return $\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$
-        \EndProcedure
-\end{algorithmic}
+		\Fn{\reduceAB{$ \{\N{L}_{kj}\}_{k,j=0}^{N_x-4}, \,\xi_0$}}{
+		\For{$k=0,1$}{
+		     $a_k \gets p_k$ \;
+		     $b_k \gets r_k$ \;
+		}
+		\For{$k=2,3$}{
+		    $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2}$ \;
+		    $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}$ \;
+		}
+		\For{$k=4, 5, \ldots, N_x-4$}{
+     		$a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2} - \N{L}_{k, k-4} \,a_{k-4}$ 
+     		\;
+	    	$b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}- \N{L}_{k, k-4} \,b_{k-4}$ 
+	    	\;
+		}
+		\Return $\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$ \;
+    }
 \end{algorithm}
 
+\SetKwFunction{LUSolveBiharmonic}{LUSolveBiharmonic}
 \begin{algorithm}
-	\caption{Solve biharmonic Eq. (\ref{eq:ufin}) with pre-computed $\{\N{U}_{kj}\}_{k,j=0}^{N}, 
+    \DontPrintSemicolon
+	\caption{Solve biharmonic Eq. (\ref{eq:ufin}) with pre-computed 
+	$\{\N{U}_{kj}\}_{k,j=0}^{N}, 
 		\{\N{L}_{kj}\}_{k,j=0}^{N}$ 
-	matrices, as well as the $\{ a_k\}_{k=0}^N$ and $\{ b_k\}_{k=0}^N$ vectors. The parameters $q_j$ and $s_j$ are given in Eqs. (\ref{eq:pk}) and 
+	matrices, as well as the $\{ a_k\}_{k=0}^N$ and $\{ b_k\}_{k=0}^N$ vectors. 
+	The parameters $q_j$ and $s_j$ are given in Eqs. (\ref{eq:pk}) and 
 	(\ref{eq:rk}) respectively. 	\label{alg:SolveBiharmonic}}
-	\begin{algorithmic}[1]
-		\Procedure{LUSolveBiharmonic}{}
-		  \State \textbf{Input:} $\{\N{U}_{kj}\}_{k,j=0}^{N}, 
+		\Fn{\LUSolveBiharmonic{$\{\N{U}_{kj}\}_{k,j=0}^{N}, 
                 \{\N{L}_{kj}\}_{k,j=0}^{N}, 
-		  \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N}, \{b_k\}_{k=0}^{N}, \xi_0$ 
-		  \State // Solve $\N{L}_{kj}y_j=f_k$ by forward elimination 
-		  \State $y_0 \gets f_0$
-		  \State $y_1 \gets f_1$
-		  \State $y_2 \gets f_2 - \N{L}_{20}y_{0}$
-		  \State $y_3 \gets f_3 - \N{L}_{31}y_{1}$
-		  \For{$k = 4$ \textbf{to} $N$}
-		  \State $y_k \gets f_k - \N{L}_{k,k-2}y_{k-2} - \N{L}_{k, k-4}y_{k-4}$
-		  \EndFor
-		  \State // Solve $\N{U}_{kj}u_j=y_k$ with back 
-		  substitution                    
-		  \State $u_N \gets y_N / \N{U}_{NN}$
-	      \State $u_{N-1} \gets y_{N-1} / \N{U}_{N-1,N-1}$
-	      \State $u_{N-2} \gets \left(y_{N-2} - \N{U}_{N-2, 
-	      N}y_N\right)/\N{U}_{N-2,N-2}$
-	      \State $u_{N-3} \gets \left(y_{N-3} - \N{U}_{N-3, 
-	      N-1}y_{N-1}\right)/\N{U}_{N-3,N-3}$
-
-	      \State Float $q^o = q^e = s^o = s^e = 0.0$
-		  \For{$k = N-4$ \textbf{to} $0$ \textbf{step} $-1$}
-		  \State $u_k \gets y_k - \N{U}_{k, k+2} u_{k+2} - \N{U}_{k, k+4} 
-		  u_{k+4} $ 
-		  \If{$k < N-5$}
-		  \State $j \gets k+6$
-		  \If{$k$ is odd}
-		    \State $q^o \gets q^o + u_{j}q_{j}$
-		    \State $s^o \gets s^o + u_{j}s_{j}$
-		    \State $u_k \gets u_k - \xi_0q^o a_k - \xi_0s^o b_k $	    
-		  \Else
-		    \State $q^e \gets q^e + u_{j}q_{j}$
-		    \State $s^e \gets s^e + u_{j}s_{j}$	  
-		    \State $u_k \gets u_k - \xi_0q^e a_k - \xi_0s^e b_k $
-		  \EndIf
-		  \EndIf
-		  \State $u_k \gets  u_k / \N{U}_{kk}$
-		  \EndFor
-		  \State \Return $\{u_k\}_{k=0}^{N}$
-
-		\EndProcedure
-	\end{algorithmic}
+                \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N}, \{b_k\}_{k=0}^{N}, 
+                \xi_0$}}{
+          \SetKwInOut{Output}{output}
+          \Output{$\{ \hat{u}_k\}_{k=0}^N$}
+		  \tcp*[l]{Solve $\N{L}_{kj}y_j=f_k$ by forward elimination}
+		  $y_0 \gets f_0$ \;
+		  $y_1 \gets f_1$ \;
+		  $y_2 \gets f_2 - \N{L}_{20}y_{0}$ \;
+		  $y_3 \gets f_3 - \N{L}_{31}y_{1}$ \;
+		  \For{$k = 4$ \textbf{to} $N$}{
+		  $y_k \gets f_k - \N{L}_{k,k-2}y_{k-2} - \N{L}_{k, k-4}y_{k-4}$ \;
+		  }
+		  \tcp*[l]{Solve $\N{U}_{kj}u_j=y_k$ with back 
+              substitution}                    
+		  $\hat{u}_N \gets y_N / \N{U}_{NN}$ \;
+	      $\hat{u}_{N-1} \gets y_{N-1} / \N{U}_{N-1,N-1}$ \;
+	      $\hat{u}_{N-2} \gets \left(y_{N-2} - \N{U}_{N-2, 
+	      N}y_N\right)/\N{U}_{N-2,N-2}$ \;
+	      $\hat{u}_{N-3} \gets \left(y_{N-3} - \N{U}_{N-3, 
+	      N-1}y_{N-1}\right)/\N{U}_{N-3,N-3}$ \;
+
+	      float $q^o = q^e = s^o = s^e = 0.0$ \;
+		  \For{$k = N-4$ \textbf{to} $0$ \textbf{step} $-1$}{
+		  $\hat{u}_k \gets y_k - \N{U}_{k, k+2} \hat{u}_{k+2} - \N{U}_{k, k+4} 
+		  \hat{u}_{k+4} $ \;
+		  \If{$k < N-5$}{
+		  $j \gets k+6$ \;
+		  \If{$k$ is odd}{
+		    $q^o \gets q^o + \hat{u}_{j}q_{j}$ \;
+		    $s^o \gets s^o + \hat{u}_{j}s_{j}$ \;
+		    $\hat{u}_k \gets \hat{u}_k - \xi_0q^o a_k - \xi_0s^o b_k $ \;
+        }
+		  \Else{
+		    $q^e \gets q^e + \hat{u}_{j}q_{j}$ \;
+		    $s^e \gets s^e + \hat{u}_{j}s_{j}$ \;	  
+		    $\hat{u}_k \gets \hat{u}_k - \xi_0q^e a_k - \xi_0s^e b_k $ \;
+		}
+		}
+		  $\hat{u}_k \gets  \hat{u}_k / \N{U}_{kk}$ \;
+		  }
+		  \Return $\{\hat{u}_k\}_{k=0}^{N}$ \;  
+		}
 \end{algorithm}
@@
 
+\bibliographystyle{model1-num-names}
 \bibliography{bib.bib}
@@
 The algorithms are using the discrete cosine transforms (dct) of type 1, 2 or 3 depending on the choice of discretization points and direction.
+\SetKwFunction{forwardScalar}{forwardScalar}
 \begin{algorithm}
-	\caption{Forward scalar products and transforms for all spaces $V_N, \D{V}_N, \N{V}_N$. 
-		Here "dct" is the discrete cosine transform from SciPy.}
-		\label{alg:fst}
-		\begin{algorithmic}[1]
-			\Function{forwardScalar}{$\{f_j\}_{j=0}^{N}$, points, space}
-			\If{ points = "Gauss-Lobatto"}
-			\State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, \text{type=2}, 
-			\text{axis=0})\frac{\pi}{2 N}$
-			\Else{ points = "Gauss-Chebyshev"}
-			\State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=1}, 
-			\text{axis=0}) \frac{\pi}{2 (N-1)}$
-			\EndIf
-			\State $\{y\}_{k=0}^N \gets 0$
-			\If{ space = $\D{V}_N$}
-			\State $\{y_k\}_{k=0}^{N-2} \gets \{w_k\}_{k=0}^{N-2} - \{w_k\}_{k=2}^{N}$
-			\ElsIf{ space = $\N{V}_N$}	                
-			\For{$k=0$ to $N-4$}
-			\State ${y}_k \gets w_k - \frac{2(k+2)}{k+3} w_{k+2} + 
-			\frac{k+1}{k+3}w_{k+4}$
-			\EndFor
-			\ElsIf{space = $V_N$}
-			\State $\{y_k\}_{k=0}^N \gets \frac{2}{\pi} \{w_k\}_{k=0}^N $
-			\State $y_0 \gets y_0 / 2$
-			\If{ points = "Gauss-Lobatto"}
-			\State $y_N \gets y_N / 2$	     	
-			\EndIf
-			\EndIf
-			\State \Return $\{y_k\}_{k=0}^{N}$    
-			\EndFunction
-			\Function{forwardTransform}{$\{f_j\}_{j=0}^{N}$, points, space}
-			\State $\{y_k\}_{k=0}^{N} \gets$ forwardScalar($\{f_j\}_{j=0}^{N}$, points, space)
-			\If{space = $\D{V}_N$}
-			\State $B \gets \D{B}$
-			\ElsIf{space = $\N{V}_N$}
-			\State $B \gets \N{B}$
-			\ElsIf{space = $V_N$}
-			\State $B \gets B$
-			\EndIf
-			\State $\bm{y} \gets B^{-1} \bm{y}$
-			\State \Return $\{y_k\}_{k=0}^{N}$
-			\EndFunction
-
-		\end{algorithmic}
+    \DontPrintSemicolon
+	\caption{Forward scalar products and transforms for all spaces $V_N, 
+	\D{V}_N, \N{V}_N$. 
+		Here "dct" is the discrete cosine transform from SciPy. \label{alg:fst}}
+			\Fn{\forwardScalar{$\{f_j\}_{j=0}^{N}$, points, space}}{
+			\If{ points = "Gauss-Lobatto"}{
+			$\{ w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, \text{type=2}, 
+			\text{axis=0})\frac{\pi}{2 N}$ \;
+            }
+			\ElseIf{ points = "Gauss-Chebyshev"}{
+			$\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=1}, 
+			\text{axis=0}) \frac{\pi}{2 (N-1)}$ \;
+		    }
+			$\{y\}_{k=0}^N \gets 0$ \;
+			\If{ space = $\D{V}_N$}{
+			$\{y_k\}_{k=0}^{N-2} \gets \{w_k\}_{k=0}^{N-2} - \{w_k\}_{k=2}^{N}$ 
+			\;
+            }
+			\ElseIf{ space = $\N{V}_N$}{
+			\For{$k=0$ to $N-4$}{
+			${y}_k \gets w_k - \frac{2(k+2)}{k+3} w_{k+2} + 
+			\frac{k+1}{k+3}w_{k+4}$ \;
+		    }
+            }
+			\ElseIf{space = $V_N$}{
+			$\{y_k\}_{k=0}^N \gets \frac{2}{\pi} \{w_k\}_{k=0}^N $ \;
+			$y_0 \gets y_0 / 2$ \;
+            }
+			\If{ points = "Gauss-Lobatto"}{
+			$y_N \gets y_N / 2$ \;   	
+		    }
+			\Return $\{y_k\}_{k=0}^{N}$    \;
+        }
+%			\EndFunction
+%			\Function{forwardTransform}{$\{f_j\}_{j=0}^{N}$, points, space}
+%			\State $\{y_k\}_{k=0}^{N} \gets$ forwardScalar($\{f_j\}_{j=0}^{N}$, 
+%points, space)
+%			\If{space = $\D{V}_N$}
+%			\State $B \gets \D{B}$
+%			\ElsIf{space = $\N{V}_N$}
+%			\State $B \gets \N{B}$
+%			\ElsIf{space = $V_N$}
+%			\State $B \gets B$
+%			\EndIf
+%			\State $\bm{y} \gets B^{-1} \bm{y}$
+%			\State \Return $\{y_k\}_{k=0}^{N}$
+%		}
 \end{algorithm}
-
-\begin{algorithm}
-	\caption{Inverse transforms for all spaces $V_N, \D{V}_N, \N{V}_N$. 
-		Here "dct" is the discrete cosine transform from SciPy.}
-	\label{alg:ifst}
-	\begin{algorithmic}[1]
-		\Function{inverseChebTransform}{$\{f\}_{j=0}^N$, points}
-		\If{points = "Gauss-Chebyshev"}
-		    \State $\{w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, type=3, axis=0)$
-		    \State $\{w_k\}_{k=0}^{N} \gets \frac{1}{2}\{w_k\}_{k=0}^{N} + \frac{1}{2}f_0$
-
-        \ElsIf{ points = "Gauss-Lobatto"}
-	        \State $\{w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, type=1, axis=0)$
-	        \State $\{w_k\}_{k=0}^{N} \gets \frac{1}{2}\{w_k\}_{k=0}^{N} + \frac{1}{2}f_0$
-	        \For{$k=0$ \textbf{to} $N$}
-	         \State $w_k \gets \frac{(-1)^k}{2} f_N$
-	        \EndFor
-	    \EndIf
-	    \State \Return{$\{w_k\}_{k=0}^{N}$}
-
-	    \EndFunction
-
-		\Function{inverseShenTransform}{$\{f_j\}_{j=0}^{N}$, points, space}
-
-		\If{ space = $\D{V}_N$}
-		\State $\{w_k\}_{k=0}^{N-2} \gets \{f_k\}_{k=0}^{N-2}$
-		\State $\{w_k\}_{k=2}^{N} \gets \{w_k\}_{k=2}^{N} - \{f_k\}_{k=0}^{N-2} $
-        \ElsIf{ space = $\N{V}_N$}	                
-		\State $\{w_k\}_{k=0}^{N-4} \gets \{f_k\}_{k=0}^{N-4}$
-		\For{$k=0$ \textbf{to} $N-4$}
-		\State $w_{k+2} \gets w_{k+2} - \frac{2(k+2)}{k+3} f_k$
-		\EndFor
-		\For{$k=0$ \textbf{to} $N-4$}
-		\State $w_{k+4} \gets w_{k+4} + \frac{k+1}{k+3} f_k$
-		\EndFor
-        \EndIf
-        \State $\{{w}_k\}_{k=0}^{N}$ = inverseChebTransform($\{w_k\}_{k=0}^N$, points)
-		\State \Return $\{{w}_k\}_{k=0}^{N}$
-		\EndFunction
-	\end{algorithmic}
-\end{algorithm}
+%
+%\begin{algorithm}
+%	\caption{Inverse transforms for all spaces $V_N, \D{V}_N, \N{V}_N$. 
+%		Here "dct" is the discrete cosine transform from SciPy.}
+%	\label{alg:ifst}
+%	\begin{algorithmic}[1]
+%		\Function{inverseChebTransform}{$\{f\}_{j=0}^N$, points}
+%		\If{points = "Gauss-Chebyshev"}
+%		    \State $\{w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, type=3, axis=0)$
+%		    \State $\{w_k\}_{k=0}^{N} \gets \frac{1}{2}\{w_k\}_{k=0}^{N} + \frac{1}{2}f_0$
+%		    
+%        \ElsIf{ points = "Gauss-Lobatto"}
+%	        \State $\{w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, type=1, axis=0)$
+%	        \State $\{w_k\}_{k=0}^{N} \gets \frac{1}{2}\{w_k\}_{k=0}^{N} + \frac{1}{2}f_0$
+%	        \For{$k=0$ \textbf{to} $N$}
+%	         \State $w_k \gets \frac{(-1)^k}{2} f_N$
+%	        \EndFor
+%	    \EndIf
+%	    \State \Return{$\{w_k\}_{k=0}^{N}$}
+%	    
+%	    \EndFunction
+%		
+%		\Function{inverseShenTransform}{$\{f_j\}_{j=0}^{N}$, points, space}
+%		
+%		\If{ space = $\D{V}_N$}
+%		\State $\{w_k\}_{k=0}^{N-2} \gets \{f_k\}_{k=0}^{N-2}$
+%		\State $\{w_k\}_{k=2}^{N} \gets \{w_k\}_{k=2}^{N} - \{f_k\}_{k=0}^{N-2} $
+%        \ElsIf{ space = $\N{V}_N$}	                
+%		\State $\{w_k\}_{k=0}^{N-4} \gets \{f_k\}_{k=0}^{N-4}$
+%		\For{$k=0$ \textbf{to} $N-4$}
+%		\State $w_{k+2} \gets w_{k+2} - \frac{2(k+2)}{k+3} f_k$
+%		\EndFor
+%		\For{$k=0$ \textbf{to} $N-4$}
+%		\State $w_{k+4} \gets w_{k+4} + \frac{k+1}{k+3} f_k$
+%		\EndFor
+%        \EndIf
+%        \State $\{{w}_k\}_{k=0}^{N}$ = inverseChebTransform($\{w_k\}_{k=0}^N$, points)
+%		\State \Return $\{{w}_k\}_{k=0}^{N}$
+%		\EndFunction
+%	\end{algorithmic}
+%\end{algorithm}
 

### 2016-12-09 paper.tex
Major commit

@@
 
+\SetKwProg{Fn}{Function}{}{}
+\DontPrintSemicolon
+\SetKwInOut{Output}{output}
+\SetKwInOut{Input}{input}
+\SetKw{KwStep}{ step }
 \newcommand{\mmo}[1]{({\bf mmo comment:} \emph{#1})}
-\bibliographystyle{plain}
+%\bliographystyle{plain}
+\bibliographystyle{model1-num-names}
 
@@
 
-\title{Assessment of a Shen-Fourier turbulent channel flow solver for large-scale simulations}
+\title{A Shen-Fourier turbulent channel flow solver for large-scale simulations}
 \author[mmo]{Mikael Mortensen}
@@
 Comput., 16, 1, 1995], that lead to stable and robust numerics, even at very large scale.
-New algorithms for direct solution of the linear systems are devised, and we show that the solver is accurate and applicable for very large simulations.
+New algorithms for direct solution of the linear systems are devised, and we show that the solver is accurate and applicable for high Reynolds numbers and very large simulations.
 Specifically, the solver is shown to reproduce first order statistics of a channel flow at $Re_{\tau}=2000$ by Hoyas and Jim\'{e}nez [Phys.
@@
 Isotropic and homogeneous turbulence are usually studied in triply periodic domains, which allows a spectral Fourier decomposition in all three spatial directions.
-The equations are solved in spectral space and nonlinearities are usually handled with pseudo-spectral techniques.
+The equations are most often solved in spectral space and nonlinearities are usually handled with pseudo-spectral techniques.
 An important feature is that the pressure can be trivially eliminated, which simplifies the Navier-Stokes equations substantially.
@@
 The first DNS channel solvers typically used a Chebyshev expansion for the wall-normal direction and, as such, were still able to obtain spectral accuracy in all three spatial directions.
-A Chebyshev-tau technique, that utilize the recurrence relations of the Chebyshev polynomials, is then used to approximate derivatives, and the coefficient matrices that appear (tridiagonal) can be inverted directly and efficiently.
+A Chebyshev-tau technique, that utilize the recurrence relations of the Chebyshev polynomials, was frequently  used to approximate derivatives, and the coefficient matrices that appeared (tridiagonal) could then be inverted directly and efficiently.
 A downside to the Chebyshev-tau method is usually quoted \cite{canuto1988} as numerical stability and roundoff errors, caused by the recurrence relation, and severe condition numbers of the coefficient matrices.
-For Chebyshev-tau methods the condition numbers have been reported to grow with size as $\mathcal{O}(N^8)$, for a discretization using N points in the wall-normal direction.
+For Chebyshev-tau methods the condition numbers have been reported to grow with size as $\mathcal{O}(N^8)$, for a discretization using $N$ points in the wall-normal direction.
 Discouraged by these numbers, all major recent channel flow simulations have found other, non-spectral, ways of discretizing the non-periodic direction.
@@
 Lee and Moser used seventh-order B-splines for the wall-normal direction.
-Other simulations of similar magnitude are performed by Hoyas and Jim\'{e}nez \cite{hoyas06, hoyas08}, Lozano-Duran and Jim\'{e}nez \cite{Lozano2014} and Bernardini, Pirozzoli and Orlandi \cite{bernardini2014}.
+Other simulations of similar magnitude have been performed by Hoyas and Jim\'{e}nez \cite{hoyas06, hoyas08}, Lozano-Duran and Jim\'{e}nez \cite{Lozano2014} and Bernardini, Pirozzoli and Orlandi \cite{bernardini2014}.
 Bernardini et al.
@@
 
-In his seminal papers \cite{Shen94,Shen95}, Jie Shen describe how to make Legendre and Chebyshev basis functions that lead to sparse matrices, susceptible to fast ($\mathcal{O}(N)$) direct solvers.
+In his seminal papers \cite{Shen94,Shen95}, Jie Shen describe how to construct Legendre and Chebyshev basis functions that lead to sparse matrices, susceptible to fast ($\mathcal{O}(N)$) direct solvers.
 To the authors knowledge, the bases have not been used for large-scale channel flow simulations, and algorithms for the required direct solvers have not, until now, been devised.
@@
 \cite{Bouchon01} describe a spectral-Galerkin formulation very similar to the one used in this paper.
-However, they choose the Legendre basis over the Chebyshev, due to the simpler structure of the matrices and the lack of weight functions for the Legendre basis.
-Also, the channel simulations performed were rather small, and there is also the issue of fast transforms for the Legendre basis.
+However, they choose the Legendre basis over the Chebyshev, neglecting the higher cost of transforms and arguing instead for the simpler structure of the matrices and the lack of weight functions for the Legendre basis.
+The cost of the Legendre transforms, though, makes the basis unsuitable for larger scales, and the simulations performed in \cite{Bouchon01} were rather small.
 
 In this paper we will describe and assess a channel flow solver based on Shen's Chebyshev basis \cite{Shen95}.
-We will give a proper description of the theoretical basis in Sec~\ref{sec:prelim}, the discretization of Navier-Stokes equations in Sec~\ref{sec:discretizationNS} and we will here describe necessary algorithms, including a new fast direct solver for the biharmonic problem that arise in Sec~\ref{sec:implementation}.
-We will finally show in Sec~\ref{sec:verification} that roundoff is not a major issue, and that the spectral-Galerkin method is in deed applicable to large-scale simulations, larger than what has been achieved for fully spectral channel simulations before.
+We will give a proper description of the theoretical basis in Sec~\ref{sec:prelim}, the discretization of Navier-Stokes equations in Sec~\ref{sec:discretizationNS} and we will describe necessary algorithms, including a new fast direct solver for the biharmonic problem that arise, in Sec~\ref{sec:implementation}.
+We will finally show in Sec~\ref{sec:verification} that roundoff is not a major issue, and that the Shen-Fourier spectral Galerkin method is in deed applicable to large-scale simulations, larger than what has been achieved for fully spectral channel simulations before.
 
@@
 \begin{align}
-&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_k\}_{k=0}^{N_x} \label{eq:Tk}\\
-& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_k\}_{k=0}^{N_x-2} \label{eq:phiD}\\
+&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_k\}_{k=0}^{N_x}, \label{eq:Tk}\\
+& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_k\}_{k=0}^{N_x-2}, \label{eq:phiD}\\
 & \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
-\frac{k+1}{k+3} T_{k+4}, & \N{V}_{N_x} &= \text{span} \{\N{\phi}_k\}_{k=0}^{N_x-4} \label{eq:phiN} 
+\frac{k+1}{k+3} T_{k+4}, & \N{V}_{N_x} &= \text{span} \{\N{\phi}_k\}_{k=0}^{N_x-4}, \label{eq:phiN} 
 \end{align}
@@
 The Navier Stokes equations (\ref{eq:NS}) are solved using a scheme 
-proposed by Kim, Moin and Moser (1987). This scheme is developed by taking the 
+proposed by Kim, Moin and Moser \cite{Kim87}. This scheme is developed by taking the 
 Laplacian of the wall normal momentum equation and the curl of the momentum 
@@
 \end{equation}
-The curl, $\bm{\omega} = (g, \partial_z u - \partial_x w, \partial_y u - \partial_x v)$, is computed by projecting each term to its appropriate space.
-To this end the two terms $\partial_x v$ and $\partial_x w$ are projected to $V_N$, whereas the remaining $\partial_y u$ and $\partial_z u$ are projected to $\N{V}_N$.
-The nonlinear term is also in need of dealiasing.
-In this work dealiasing has been performed with the 3/2-rule, see Canuto~et al.
-\cite{canuto1988}.
+The curl, $\bm{\omega} = (g, \partial_z u - \partial_x w, \partial_y u - 
+\partial_x v)$, is computed by projecting each term to its appropriate space. 
+To this end the two terms $\partial_x v$ and $\partial_x w$ are projected to 
+$V_N$, whereas the remaining $\partial_y u$ and $\partial_z u$ are projected to 
+$\N{V}_N$. Note that the nonlinear term is also generally in need of 
+dealiasing, at least for the two periodic directions.
 
@@
 An outline of the solution procedure used, for the numerical method described 
-in Sec.~\ref{sec:discretizationNS}, is given in Algorithm \ref{alg:NS}. Note 
+in Sec.~\ref{sec:discretizationNS}, is given in Algorithm 
+\ref{alg:NS}.\footnote{Note that throughout this paper we are using the 
+pseudocode conventions of Kopriva \cite{kopriva09}, with one minor difference. 
+A statement like $\{w\}_{k=0}^{N} \gets \{u\}_{k=0}^N$ will here be used to 
+represent a 
+vectorization, where components $u_k$, for $k=0,1,\ldots,N$, are copied from 
+$u_k$ to $w_k$. Similar conventions apply to matrices, e.g., 
+$\{U_{1,k}\}_{k=0}^N = \{V_{2,k}\}_{k=0}^N$ can be used for copying row 2 of 
+$V$ to row 1 of $U$.} Note 
 that the Adams-Bashforth projection of convection reverts to a first order 
@@
 is insignificant, but for accuracy tests, like the Orr-Sommerfeld instability, 
-a proper initialization would need to set both the two first time steps instead.
+a proper initialization would need to set both the two first time steps.
 \begin{algorithm}
@@
     	$\hat{g}_{\bm{k}}(0)$ and $\hat{\bm{u}}_{\bm{k}}(0)$ \;
+        Set parameters (e.g., mesh and viscosity) and end time $T$ \;
 		Compute LU decomposition of $\N{H}(m,n), \D{H}(m,n) 
@@
 nonlinear convection term, $\bm{\hat{\mathcal{H}}}$, is computed using a 
-pseudo-spectral method with the 3/2-rule for dealiasing in the periodic 
-directions, see \cite{canuto1988}. The 
+pseudo-spectral method, and dealiasing in the periodic directions can be either 
+the 3/2-rule \cite{canuto1988} or the 2/3-rule \cite{Orzag71}. The 
 transforms are carried out using fast Fourier and fast Chebyshev transforms 
-that are all of order $\mathcal{O}(N_x\log N_x)$ per one-dimensional transform of size 
-$N_x$. The Shen transforms make use of the Chebyshev transforms that are using 
+that are all of order $\mathcal{O}(N_x\log N_x)$ per one-dimensional transform 
+of size $N_x$. The Shen transforms make use of the Chebyshev transforms that in 
+turn are using 
 discrete cosine transforms. The forward scalar products and transforms are implemented as shown 
@@
 as shown in Alg. 
-\ref{alg:lu_biharmonic}, which leads to a lower triangular 
+\ref{alg:lu_biharmonic}, which leads to a lower 
+triangular 
 matrix $\N{L}_{kj}$ with two nonzero diagonals at $j=k-2$ and $k-4$ plus a 
@@
 
-\SetKwProg{Fn}{Function}{}{}
+\SetKwFunction{LUbiharmonic}{LUbiharmonic}
+\begin{algorithm}
+    \caption{LU decomposition of biharmonic operator $\N{H}$. Current algorithm 
+        is using dense storage of $\N{U}_{kj}$, but no more than the 3 items 
+        per row need 
+        to be stored after the row is no longer needed in the for-loop below 
+        (as the loop is traversed to row $k$, $\N{U}_{m, j}$ is no longer 
+        needed for $m<k-4$). 
+        \label{alg:lu_biharmonic}}
+        \Fn{\LUbiharmonic{$\{\N{H}_{kj}\}_{k, j = 0}^{N}$}}{
+        \Output{$\{\N{L}_{kj}\}_{k, j = 0}^{N}$, $\{\N{U}_{kj}\}_{k, j = 
+        0}^{N}$}
+        integer $M_e \gets \lfloor N/2 \rfloor$ \;
+        integer $M_o \gets \lfloor (N-1)/2 \rfloor$  \;
+        $\{\hat{U}_{0,2j}\}_{j=0}^{M_e} \gets \{\hat{H}_{0,2j}\}_{j=0}^{M_e}$ \;
+        $\{\hat{U}_{1,2j+1}\}_{j=0}^{M_o} \gets 
+        \{\hat{H}_{1,2j+1}\}_{j=0}^{M_o}$ \;
+        $\N{L}_{20} \gets \N{H}_{20}/\N{U}_{00}$ \;
+        $\N{L}_{31} \gets \N{H}_{31}/\N{U}_{11}$ \;
+        $\{\hat{U}_{2,2j}\}_{j=1}^{M_e} \gets \{\N{H}_{2,2j}\}_{j=1}^{M_e} - 
+        \N{L}_{20} \{\N{U}_{0,2j}\}_{j=1}^{M_e}$ \;
+        $\{\hat{U}_{3,2j+1}\}_{j=1}^{M_o} \gets \{\N{H}_{3,2j+1}\}_{j=1}^{M_o} 
+        - \N{L}_{31} \{\N{U}_{1,2j+1}\}_{j=1}^{M_o}$ \;
+
+        %\State $\N{U}_{2j} = \N{H}_{2j} - \N{L}_{20} \N{U}_{0j} \quad 
+        %\forall \, j \in 2, 4, \ldots, N$
+        %\State $\N{U}_{3j} = \N{H}_{3j} - \N{L}_{31} \N{U}_{1j} \quad 
+        %\forall \, j \in 
+        %3, 5, \ldots, N$
+        \For{$k = 4$ \KwTo $N$}{
+        integer $M \gets \lfloor \frac{N-k}{2}\rfloor$ \;
+        $\N{L}_{k,k-4} \gets \N{H}_{k, k-4} / \N{U}_{k-4, k-4}$ \;
+        $\{\N{U}_{k,k+2j}\}_{j=-1}^{M} \gets \{\N{H}_{k,k+2j}\}_{j=-1}^{M} - 
+        \N{L}_{k, k-4} 
+        \{\N{U}_{k-4, k+2j}\}_{j=-1}^{M}$ \;
+        %\State $\N{U}_{k,j} = \N{H}_{k,j} - \N{L}_{k, k-4}  \N{U}_{k-4, j} 
+        %\quad \forall j = k-2, k, \ldots, N$
+        $\N{L}_{k,k-2} \gets {\N{U}}_{k,k-2}/{\N{U}}_{k-2, k-2}$ \;
+        $\{\N{U}_{k,k+2j}\}_{j=0}^{M} \gets \{{\N{U}}_{k,k+2j}\}_{j=0}^{M} - 
+        \N{L}_{k, k-2} \{\N{U}_{k-2,k+2j}\}_{j=0}^{M}$ \;
+        %\State $\N{U}_{kj} = {\N{U}}_{kj} - \N{L}_{k, k-2} \N{U}_{k-2, j} 
+        %\quad \forall j = k, k+2, \ldots, N$
+        }
+        \Return $\{\N{L}_{kj}\}_{k, j = 0}^{N}$, $\{\N{U}_{kj}\}_{k, j = 0}^{N}$
+    }
+\end{algorithm}
+
 \SetKwFunction{reduceAB}{ReduceAB}
@@
 		\Fn{\reduceAB{$ \{\N{L}_{kj}\}_{k,j=0}^{N_x-4}, \,\xi_0$}}{
+        \Output{$\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$}  \;
 		\For{$k=0,1$}{
@@
 		}
-		\For{$k=4, 5, \ldots, N_x-4$}{
+		\For{$k=4$ \KwTo $N_x-4$}{
      		$a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2} - \N{L}_{k, k-4} \,a_{k-4}$ 
@@
                 \xi_0$}}{
-          \SetKwInOut{Output}{output}
           \Output{$\{ \hat{u}_k\}_{k=0}^N$}
@@
 		  $y_3 \gets f_3 - \N{L}_{31}y_{1}$ \;
-		  \For{$k = 4$ \textbf{to} $N$}{
+		  \For{$k = 4$ \KwTo $N$}{
 		  $y_k \gets f_k - \N{L}_{k,k-2}y_{k-2} - \N{L}_{k, k-4}y_{k-4}$ \;
@@
 
-	      float $q^o = q^e = s^o = s^e = 0.0$ \;
-		  \For{$k = N-4$ \textbf{to} $0$ \textbf{step} $-1$}{
+	      $q^o \gets 0.0$ \;
+          $q^e \gets 0.0$ \;
+          $s^o \gets 0.0$ \;
+          $s^e \gets 0.0$ \;
+		  \For{$k = N-4$ \KwTo $0$ \KwStep $-1$}{
 		  $\hat{u}_k \gets y_k - \N{U}_{k, k+2} \hat{u}_{k+2} - \N{U}_{k, k+4} 
@@
 
-The complete Navier Stokes solver should scale like $\mathcal{O}(aN + bN\log N)$ since it is a combination of linear algebra routines $\mathcal{O}(N)$, fast transforms $\mathcal{O}(N \log N)$ and several elementwise array operations $\mathcal{O}(N)$.
-We show in Table \ref{tab:timings_solver} some timings for one single time step, increasing only the wall normal discretization.
-We show that the total time scales approximately as $N \log N$ for large $N$ and better than linearly for small $N$.
-This shift happens since at small $N$ the transforms are fast and linear operations, like elementwise multiplications, are more dominating.
-We also see that the time for assembly of the right hand side is approximately $\mathcal{O}(N \log N)$, whereas the time for solving the linear systems is very nearly linear.
+The complete Navier Stokes solver should scale like $\mathcal{O}(N)$ and 
+$\mathcal{O}N\log N)$ since it is a combination of linear algebra routines 
+$\mathcal{O}(N)$, fast transforms $\mathcal{O}(N \log N)$ and several 
+elementwise array operations $\mathcal{O}(N)$.
+
+We show in Table \ref{tab:timings_solver} some timings for one single time 
+step, increasing only the wall normal discretization. We show that the total 
+time scales approximately as $N \log N$ for large $N$ and better than linearly 
+for small $N$. This shift happens since at small $N$ the transforms are fast 
+and linear operations, like elementwise multiplications, are more dominating. 
+We also see that the time for assembly of the right hand side is approximately 
+$\mathcal{O}(N \log N)$, whereas the time for solving the linear systems is 
+very nearly linear. Note that for fast transforms we are using 
+the FFTW library \cite{libfftw}, with planner effort set to FFTW$\_$MEASURE. 
+Timings in Table \ref{tab:timings_solver} are computed with dealiasing 
+performed with the 2/3-rule \cite{Orzag71}.
 
@@
 The $L^2$-norm $ \| \bm{u}' \| = \sqrt{\left< \bm{u}', \bm{u}'\right>}$, where $\bm{u}'=\bm{u}-\bm{u}_{\mathrm{exact}}$ and energy error (\ref{eq:OSenergy}), for the most unstable Orr-Sommerfeld mode.
-Results are computed with a mesh of size $128 \times 8 \times 2$, $\triangle t=0.2$ and final time $t=5$.
+Results are computed with a mesh of size $128 \times 8 \times 2$ or $512 \times 8 \times 2$, with $\triangle t=0.2$ and final time $t=5$.
 \label{tab:epsilon}}
-	\begin{tabular}{ccc}	
-		$-\log_{10}(\epsilon)$ & $\frac{1}{\epsilon}\| \bm{u}' \|(t=5)$ & $E(t=5)$ \\
+	\begin{tabular}{ccccc}	
+		$-\log_{10}(\epsilon)$ & \multicolumn{2}{c}{$\frac{1}{\epsilon}\| \bm{u}' \|(t=5)$} & \multicolumn{2}{c}{$E(t=5)$} \\
 		\hline
-		4 & 1.234544e-02 & 2.441506e-04 \\ 
-		5 & 1.157535e-02 & 2.440520e-04 \\ 
-		6 & 1.156739e-02 & 2.440510e-04 \\ 
-		7 & 1.156732e-02 & 2.440510e-04 \\ 
-		8 & 1.156731e-02 & 2.440504e-04 \\ 
-		9 & 1.156732e-02 & 2.440498e-04 \\ 
-		10 & 1.156726e-02 & 2.440435e-04 \\ 
-		11 & 1.156497e-02 & 2.437132e-04 \\ 
-		12 & 1.155687e-02 & 2.435411e-04
+		 & $N=128$ & $N=512$ & $N=128$ & $N=512$ \\
+		\hline
+4	&	1.23454361e-02	&	2.44150583e-04	&	1.23454363e-02	&	
+2.44150404e-04	\\
+5	&	1.15753545e-02	&	2.44052002e-04	&	1.15753547e-02	&	
+2.44051822e-04	\\
+6	&	1.15673948e-02	&	2.44051018e-04	&	1.15673950e-02	&	
+2.44050836e-04	\\
+7	&	1.15673151e-02	&	2.44051004e-04	&	1.15673154e-02	&	
+2.44050816e-04	\\
+8	&	1.15673142e-02	&	2.44050426e-04	&	1.15673156e-02	&	
+2.44050992e-04	\\
+9	&	1.15673192e-02	&	2.44049842e-04	&	1.15673185e-02	&	
+2.44050701e-04	\\
+10	&	1.15672575e-02	&	2.44043530e-04	&	1.15674820e-02	&	
+2.44052242e-04	\\
+11	&	1.15649705e-02	&	2.43713173e-04	&	1.15684397e-02	&	
+2.44010407e-04	\\
+12	&	1.15568665e-02	&	2.43541116e-04	&	1.15844457e-02	&	
+2.44202451e-04
 	\end{tabular}
@@
 \section{Conclusions}
+Spectral Navier Stokes solvers are attractive for their accuracy and resolution 
+properties, and are usually the preferred tools for performing Direct Numerical 
+Simulations of turbulent flows. A spectral discretization of the wall normal 
+direction in a plane channel is challenging due to roundoff errors from poorly 
+conditions coefficient matrices, and as such all recent large-scale channel DNS 
+have been conducted without spectral accuracy in the wall normal direction.
+
+In this paper a Navier Stokes solver for large-scale turbulent channel flows 
+has been described. In the wall normal direction the solver utilizes base 
+functions 
+suggested by J. Shen \cite{Shen95}, constructed from Chebyshev polynomials. The 
+solver uses Fourier decompositions in both periodic directions, and, as such, 
+it is fully spectral in all spatial directions. To show that this solver is 
+applicable to large-scale simulations at high Reynolds numbers,  we have shown 
+that the roundoff errors are small, even for simulations twice the size of the 
+largest simulations known to date \cite{leemoser15}.
+The computational cost of the solver is shown to scale with the expected 
+$\mathcal{O}(N \log N)$, for a mesh of size $N \times N \times N$, due to fast 
+transforms. Direct 
+solvers have been devised for Helmholtz and biharmonic coefficient matrices, 
+and these have been shown to scale close to 
+$\mathcal{O}(N)$, with insignificant roundoff errors. The fully spectral solver 
+has been applied to a turbulent 
+channel flow at $Re_{\tau}=2000$, and the first order statistics are shown to agree well with the literature.
+
+\section*{Acknowledgements}
+This research is a part of the 4DSpace Strategic Research Initiative at the 
+University of Oslo.
 
@@
 
-\bibliographystyle{model1-num-names}
 \bibliography{bib.bib}
@@
 
-The algorithms for the fast scalar products and transforms, for the three spaces $V_N, \D{V}_N, \N{V}_N$, are given in Alg.
-~\ref{alg:fst}.
-The algorithms for the inverse transforms are given in Alg.~\ref{alg:ifst}.
-The algorithms are using the discrete cosine transforms (dct) of type 1, 2 or 3 depending on the choice of discretization points and direction.
+Algorithms for fast scalar products and transforms, for the three 
+spaces $V_N, \D{V}_N, \N{V}_N$, are given in Alg. ~\ref{alg:fst}. The 
+algorithms for the inverse transforms are given in Alg.~\ref{alg:ifst}. The 
+algorithms are using the discrete cosine transforms (dct) of type 1, 2 or 3 
+depending on the choice of discretization points and direction.
 \SetKwFunction{forwardScalar}{forwardScalar}
+\SetKwFunction{forwardTransform}{forwardTransform}
 \begin{algorithm}
@@
 			\Fn{\forwardScalar{$\{f_j\}_{j=0}^{N}$, points, space}}{
+
+            \Output{$\{y_k\}_{k=0}^{N}$}
+
 			\If{ points = "Gauss-Lobatto"}{
@@
 			\ElseIf{ space = $\N{V}_N$}{
-			\For{$k=0$ to $N-4$}{
+			\For{$k=0$ \KwTo $N-4$}{
 			${y}_k \gets w_k - \frac{2(k+2)}{k+3} w_{k+2} + 
@@
 			$y_0 \gets y_0 / 2$ \;
-            }
 			\If{ points = "Gauss-Lobatto"}{
@@
 		    }
+            }
 			\Return $\{y_k\}_{k=0}^{N}$    \;
             }
-%			\EndFunction
-%			\Function{forwardTransform}{$\{f_j\}_{j=0}^{N}$, points, space}
-%			\State $\{y_k\}_{k=0}^{N} \gets$ forwardScalar($\{f_j\}_{j=0}^{N}$, 
-%points, space)
-%			\If{space = $\D{V}_N$}
-%			\State $B \gets \D{B}$
-%			\ElsIf{space = $\N{V}_N$}
-%			\State $B \gets \N{B}$
-%			\ElsIf{space = $V_N$}
-%			\State $B \gets B$
-%			\EndIf
-%			\State $\bm{y} \gets B^{-1} \bm{y}$
-%			\State \Return $\{y_k\}_{k=0}^{N}$
-%		}
+			\Fn{\forwardTransform{$\{f_j\}_{j=0}^{N}$, points, space}}{
+			\Output{$\{y_k\}_{k=0}^{N}$}
+            $\{y_k\}_{k=0}^{N} \gets$ forwardScalar($\{f_j\}_{j=0}^{N}$, 
+points, space) \;
+			\If{space = $\D{V}_N$}{
+			$\{y_k\}_{k=0}^{N-2} \gets \{\D{B}_{kj}^{-1}\}_{k,j=0}^{N-2}\{y_k\}_{k=0}^{N-2}$ \;
+            }
+			\ElseIf{space = $\N{V}_N$}{
+			$\{y_k\}_{k=0}^{N-4} \gets \{\N{B}_{kj}^{-1}\}_{k,j=0}^{N-4}\{y_k\}_{k=0}^{N-4}$ \;
+            }
+			\ElseIf{space = $V_N$}{
+			$\{y_k\}_{k=0}^{N} \gets \{{B}_{kj}^{-1}\}_{k,j=0}^{N}\{y_k\}_{k=0}^{N}$ \;
+            }
+			\Return $\{y_k\}_{k=0}^{N}$
+		}
 \end{algorithm}
-%
-%\begin{algorithm}
-%	\caption{Inverse transforms for all spaces $V_N, \D{V}_N, \N{V}_N$. 
-%		Here "dct" is the discrete cosine transform from SciPy.}
-%	\label{alg:ifst}
-%	\begin{algorithmic}[1]
-%		\Function{inverseChebTransform}{$\{f\}_{j=0}^N$, points}
-%		\If{points = "Gauss-Chebyshev"}
-%		    \State $\{w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, type=3, axis=0)$
-%		    \State $\{w_k\}_{k=0}^{N} \gets \frac{1}{2}\{w_k\}_{k=0}^{N} + \frac{1}{2}f_0$
-%		    
-%        \ElsIf{ points = "Gauss-Lobatto"}
-%	        \State $\{w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, type=1, axis=0)$
-%	        \State $\{w_k\}_{k=0}^{N} \gets \frac{1}{2}\{w_k\}_{k=0}^{N} + \frac{1}{2}f_0$
-%	        \For{$k=0$ \textbf{to} $N$}
-%	         \State $w_k \gets \frac{(-1)^k}{2} f_N$
-%	        \EndFor
-%	    \EndIf
-%	    \State \Return{$\{w_k\}_{k=0}^{N}$}
-%	    
-%	    \EndFunction
-%		
-%		\Function{inverseShenTransform}{$\{f_j\}_{j=0}^{N}$, points, space}
-%		
-%		\If{ space = $\D{V}_N$}
-%		\State $\{w_k\}_{k=0}^{N-2} \gets \{f_k\}_{k=0}^{N-2}$
-%		\State $\{w_k\}_{k=2}^{N} \gets \{w_k\}_{k=2}^{N} - \{f_k\}_{k=0}^{N-2} $
-%        \ElsIf{ space = $\N{V}_N$}	                
-%		\State $\{w_k\}_{k=0}^{N-4} \gets \{f_k\}_{k=0}^{N-4}$
-%		\For{$k=0$ \textbf{to} $N-4$}
-%		\State $w_{k+2} \gets w_{k+2} - \frac{2(k+2)}{k+3} f_k$
-%		\EndFor
-%		\For{$k=0$ \textbf{to} $N-4$}
-%		\State $w_{k+4} \gets w_{k+4} + \frac{k+1}{k+3} f_k$
-%		\EndFor
-%        \EndIf
-%        \State $\{{w}_k\}_{k=0}^{N}$ = inverseChebTransform($\{w_k\}_{k=0}^N$, points)
-%		\State \Return $\{{w}_k\}_{k=0}^{N}$
-%		\EndFunction
-%	\end{algorithmic}
-%\end{algorithm}
+
+\SetKwFunction{inverseChebTransform}{inverseChebTransform}
+\SetKwFunction{inverseShenTransform}{inverseShenTransform}
+\begin{algorithm}
+	\caption{Inverse transforms for all spaces $V_N, \D{V}_N, \N{V}_N$. 
+		Here "dct" is the discrete cosine transform from SciPy.\label{alg:ifst}}
+		\Fn {\inverseChebTransform{$\{f\}_{j=0}^N$, points}}{
+        \Output{$\{y_k\}_{k=0}^{N}$}    
+        \If{points = "Gauss-Chebyshev"}{
+		    $\{y_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, \text{type=3}, 
+		    \text{axis=0})$ \;
+		    $\{y_k\}_{k=0}^{N} \gets \frac{1}{2}\{y_k\}_{k=0}^{N} + 
+		    \frac{1}{2}f_0$ \;
+		}
+        \ElseIf{ points = "Gauss-Lobatto"}{
+	        $\{y_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, \text{type=1}, 
+	        \text{axis=0})$ \;
+	        $\{y_k\}_{k=0}^{N} \gets \frac{1}{2}\{y_k\}_{k=0}^{N} + 
+	        \frac{1}{2}f_0$ \;
+	        \For{$k=0$ \KwTo $N$}{
+	         $y_k \gets y_k + \frac{(-1)^k}{2} f_N$ \;
+    	    }
+	    }
+	    \Return{$\{y_k\}_{k=0}^{N}$}
+    }
+
+	    \Fn{\inverseShenTransform{$\{f_j\}_{j=0}^{N}$, points, space}}{
+		\Output{$\{y_k\}_{k=0}^{N}$}
+        $\{z_k\}_{k=0}^{N} \gets 0$ \;
+		\If{ space = $\D{V}_N$}{
+		$\{z_k\}_{k=0}^{N-2} \gets \{f_k\}_{k=0}^{N-2}$ \;
+		$\{z_k\}_{k=2}^{N} \gets \{z_k\}_{k=2}^{N} - \{f_k\}_{k=0}^{N-2} $ \;
+        }
+        \ElseIf{ space = $\N{V}_N$}{	       
+		$\{z_k\}_{k=0}^{N-4} \gets \{f_k\}_{k=0}^{N-4}$ \;
+		\For{$k=0$ \KwTo $N-4$}{
+		$z_{k+2} \gets z_{k+2} - \frac{2(k+2)}{k+3} f_k$ \;
+	    }
+		\For{$k=0$ \KwTo $N-4$}{
+		$z_{k+4} \gets z_{k+4} + \frac{k+1}{k+3} f_k$ \;
+	    }
+        }
+        $\{y_k\}_{k=0}^{N} \gets $ inverseChebTransform($\{z_k\}_{k=0}^N$, 
+        points) 
+        \;
+		\Return $\{y_k\}_{k=0}^{N}$ \;
+    }
+\end{algorithm}
 

### 2017-01-04 paper.tex
Work

@@
 \newcommand{\N}[1]{\check{#1}}
-\newcommand{\D}[1]{\overline{#1}}
+\newcommand{\D}[1]{\bar{#1}}
 
@@
 
-\title{A Shen-Fourier turbulent channel flow solver for large-scale simulations}
+\title{A spectral-Galerkin turbulent channel flow solver for large-scale simulations}
 \author[mmo]{Mikael Mortensen}
@@
 Sci.
-Comput., 16, 1, 1995], that lead to stable and robust numerics, even at very large scale.
-New algorithms for direct solution of the linear systems are devised, and we show that the solver is accurate and applicable for high Reynolds numbers and very large simulations.
-Specifically, the solver is shown to reproduce first order statistics of a channel flow at $Re_{\tau}=2000$ by Hoyas and Jim\'{e}nez [Phys.
-Fluids, 18(1), 2006], which is higher than has been achieved in a channel by a fully spectral method before.
+Comput., 16, 1, 1995], that lead to stable and robust numerical schemes, even at very large scale.
+New algorithms for direct solution of the linear systems are devised, and we show that the solver is accurate and applicable for very high Reynolds numbers.
+Specifically, the solver is shown to reproduce the first order statistics of Hoyas and Jim\'{e}nez [Phys.
+Fluids, 18(1), 2006], for a channel flow at $Re_{\tau}=2000$, which is higher than has been reported by a fully spectral method before.
+The solver is available through the open source project spectralDNS [https://github.com/spectralDNS].
 \end{abstract}
 \begin{keyword}
-DNS \sep Fourier \sep Chebyshev
+DNS \sep Fourier \sep Chebyshev \sep Biharmonic \sep Helmholtz \sep turbulence
 \end{keyword}
@@
 \section{Introduction}
-Direct Numerical Simulations (DNS) of turbulent flows is a very important research tool, utilized across a range of scientific communities.
+Direct Numerical Simulations (DNS) of turbulent flows is a very important research tool, utilized across a range of scientific communities \cite{Moin98}.
 DNS is used extensively to validate statistical models, and to further our understanding of complex mechanisms taking place inside turbulent flows.
@@
 The most commonly known DNS use simple geometries, because turbulence physics may then most easily be isolated and studied.
-Isotropic and homogeneous turbulence are usually studied in triply periodic domains, which allows a spectral Fourier decomposition in all three spatial directions.
-The equations are most often solved in spectral space and nonlinearities are usually handled with pseudo-spectral techniques.
-An important feature is that the pressure can be trivially eliminated, which simplifies the Navier-Stokes equations substantially.
-
-Inhomogeneous, wall bounded, flows, successfully studied with DNS are, e.g., the pipe flow, free shear layers and channel flow between two parallel plates (both Couette and Poiseuille).
-In this paper we will constrain ourselves to the pressure driven Poiseuille channel flow.
-In a channel there are two periodic directions that can be handled with Fourier expansions, whereas the wall normal direction is non-periodic and requires a different type of discretization.
+Isotropic and homogeneous turbulence are usually studied numerically in triply periodic domains, which allows for a spectral Fourier decomposition in all three spatial directions.
+Spectral methods are often favored in DNS due to their superior accuracy and resolution properties.
+One example is given in the DNS review of Moin and Mahesh \cite{Moin98}, who report that, for similar accuracy in first derivatives, a second-order finite difference scheme requires approximately 5.5 times more points than Fourier, in each spatial direction, whereas for a 6'th order Pad\'{e} scheme the factor is about 1.6.
+
+In this paper we will consider the pressure driven turbulent channel flow, where there are two periodic directions that can be handled with Fourier expansions, and a non-periodic wall-normal direction that requires a different type of discretization.
 There are many challenges associated with this inhomogeneity not faced by the pure Fourier solvers, but the first problem at hand is the discretization.
-The first DNS channel solvers typically used a Chebyshev expansion for the wall-normal direction and, as such, were still able to obtain spectral accuracy in all three spatial directions.
-A Chebyshev-tau technique, that utilize the recurrence relations of the Chebyshev polynomials, was frequently  used to approximate derivatives, and the coefficient matrices that appeared (tridiagonal) could then be inverted directly and efficiently.
-A downside to the Chebyshev-tau method is usually quoted \cite{canuto1988} as numerical stability and roundoff errors, caused by the recurrence relation, and severe condition numbers of the coefficient matrices.
+Early DNS channel solvers, see, e.g., \cite{Moin80, Kleiser80, Kim87}, typically used a Chebyshev expansion for the wall-normal direction and, as such, were still able to obtain spectral accuracy in all three spatial directions.
+A Chebyshev-tau technique, that utilize the recurrence relations of the Chebyshev polynomials, was used to approximate derivatives, and the coefficient matrices that appeared (tridiagonal) could then be inverted directly and efficiently \cite{Kim87}.
+A downside to the Chebyshev-tau method is usually quoted \cite{canuto1988} as numerical instability and roundoff errors, caused by the recurrence relation, and severe condition numbers of the coefficient matrices.
 For Chebyshev-tau methods the condition numbers have been reported to grow with size as $\mathcal{O}(N^8)$, for a discretization using $N$ points in the wall-normal direction.
@@
 
-The largest known channel simulations to date have been performed by Lee and Moser \cite{leemoser15}, where for $Re_{\tau}\approx 5200$ they used a computational box of resolution $[10240 \times 1530 \times 7680]$ for the streamwise, wall-normal and spanwise directions, respectively.
+The largest known channel simulations to date have been performed by Lee and Moser \cite{leemoser15}, where, for $Re_{\tau}\approx 5200$, they used a computational box of resolution $[10240 \times 1530 \times 7680]$ for the streamwise, wall-normal and spanwise directions, respectively.
 Lee and Moser used seventh-order B-splines for the wall-normal direction.
@@
 Bernardini et al.
-used second order finite differences throughout.
-The Jim\'{e}nez group used dealiased Fourier in the two periodic directions and seventh order compact finite differences, with fourth-order consistency and extended spectral-like resolution \cite{Lele92}, for the wall-normal direction.
+used second-order finite differences throughout.
+The Jim\'{e}nez group used dealiased Fourier in the two periodic directions and seventh-order compact finite differences, with fourth-order consistency and extended spectral-like resolution \cite{Lele92}, for the wall-normal direction.
 The solver by Jim\'{e}nez' group is reported to switch from Chebyshev to finite differences if the resolution is above a certain threshold \cite{hoyas08} (reached around $Re_{\tau}=1000$).
-In other words, they attempt to use spectral methods for as large $Re_{\tau}$ as possible.
-Spectral methods are attractive for their accuracy, but also for their resolution properties, that are superior to those of any finite difference or spline method.
+In other words, they attempt to use fully spectral methods for as large $Re_{\tau}$ as possible.
+As previously mentioned, spectral methods are attractive for their accuracy and resolution properties, that are superior to those of any finite difference or spline method.
 As such, it is desirable to develop fully spectral solvers that can be used for large-scale turbulence simulations.
 
-In his seminal papers \cite{Shen94,Shen95}, Jie Shen describe how to construct Legendre and Chebyshev basis functions that lead to sparse matrices, susceptible to fast ($\mathcal{O}(N)$) direct solvers.
+In his seminal papers \cite{Shen94,Shen95}, Jie Shen describe how to construct Legendre and Chebyshev basis functions that lead to sparse matrices, susceptible to very fast direct solvers.
 To the authors knowledge, the bases have not been used for large-scale channel flow simulations, and algorithms for the required direct solvers have not, until now, been devised.
@@
 \cite{Bouchon01} describe a spectral-Galerkin formulation very similar to the one used in this paper.
-However, they choose the Legendre basis over the Chebyshev, neglecting the higher cost of transforms and arguing instead for the simpler structure of the matrices and the lack of weight functions for the Legendre basis.
+However, they choose the Legendre basis over Chebyshev, neglecting the higher cost of transforms and arguing instead for the simpler structure of the matrices and the lack of weight functions for the Legendre basis.
 The cost of the Legendre transforms, though, makes the basis unsuitable for larger scales, and the simulations performed in \cite{Bouchon01} were rather small.
 
-In this paper we will describe and assess a channel flow solver based on Shen's Chebyshev basis \cite{Shen95}.
+In this paper we will describe and assess a spectral-Galerkin channel flow solver based on Fourier and Shen's Chebyshev basis \cite{Shen95}.
 We will give a proper description of the theoretical basis in Sec~\ref{sec:prelim}, the discretization of Navier-Stokes equations in Sec~\ref{sec:discretizationNS} and we will describe necessary algorithms, including a new fast direct solver for the biharmonic problem that arise, in Sec~\ref{sec:implementation}.
-We will finally show in Sec~\ref{sec:verification} that roundoff is not a major issue, and that the Shen-Fourier spectral Galerkin method is in deed applicable to large-scale simulations, larger than what has been achieved for fully spectral channel simulations before.
+We will finally show in, Sec~\ref{sec:verification}, that roundoff is not a major issue, and that the Shen-Fourier spectral-Galerkin method is in deed applicable to large-scale simulations, larger than what has been reported for fully spectral channel simulations before.
+The solver has been implemented in the open source code spectralDNS \cite{spectralDNS}, where the bulk of the solver is written in high-level Python \cite{python}, with critical parts of the code migrated to Cython \cite{cython} for efficiency.
 
@@
 \label{sec:prelim}
-The Navier-Stokes equations used to describe turbulent flow in a doubly 
-periodic channel can be written in rotational form as
+The Navier-Stokes equations, used to describe turbulent flow in a doubly 
+periodic channel, can be written in rotational form as
 \begin{align}
@@
 where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ 
-and $t$ are position and time and $\bm{\mathcal{H}}(\bm{x}, t) = \bm{u}\times 
-\bm{\omega}$, where $\bm{\omega} = \nabla \times \bm{u}$, represents 
-the nonlinearity.  The constant dynamic viscosity is denoted as $\nu$ 
+and $t$ are position and time, and the nonlinearity $\bm{\mathcal{H}}(\bm{x}, t) = (\mathcal{H}_x, \mathcal{H}_y, \mathcal{H}_z) = \bm{u}\times 
+\bm{\omega}$, where $\bm{\omega} = \nabla \times \bm{u}$.  The constant dynamic viscosity is denoted as $\nu$ 
 and $P(\bm{x}, t)$ is a pressure modified to account for both the driving force, $\beta(t)$, and the kinetic energy, i.e., $P = p + \bm{u} \cdot \bm{u}/2 + \beta y$, where $p$ is the instantaneous 
 pressure normalized by a constant density. The computational domain is 
-$\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$ with channel walls 
+$\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$, with channel walls 
 located at $x=\pm 1$, such that no-slip applies at 
@@
 
-The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall normal direction.
-Three different sets of basis functions and function spaces are relevant for the wall normal direction
+The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall-normal direction.
+Three different sets of basis functions and function spaces are relevant for the wall-normal direction
 \begin{align}
@@
 & \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
-\frac{k+1}{k+3} T_{k+4}, & \N{V}_{N_x} &= \text{span} \{\N{\phi}_k\}_{k=0}^{N_x-4}, \label{eq:phiN} 
+\frac{k+1}{k+3} T_{k+4}(x), & \N{V}_{N_x} &= \text{span} \{\N{\phi}_k\}_{k=0}^{N_x-4}, \label{eq:phiN} 
 \end{align}
@@
 basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were 
-suggested by Shen, and satisfy, respectively, the boundary conditions 
-$\D{\phi}_k(\pm 1) = 0$, $\N{\phi}_k(\pm 1)=0$ and $\N{\phi}'_k(\pm 1)=0$. The 
-last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in 
-V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1) = v'(\pm 
-1) = 0 \}$.
+suggested by Shen \cite{Shen95}, and satisfy, respectively, the boundary conditions 
+$\D{\phi}_k(\pm 1) = 0$, $\N{\phi}_k(\pm 1)=0$ and $\N{\phi}'_k(\pm 1)=0$. 
+%The last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1) = v'(\pm 1) = 0 \}$.
 
@@
 \begin{align}
-K_N = \Big\{\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / &(l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
- &(l, m, n) \in  [0, \ldots, N_x] \times [-\frac{N_y}{2},\ldots, 
- \frac{N_y}{2}-1] \times[-\frac{N_z}{2},\ldots, \frac{N_z}{2}-1] \Big\} 
+K_N = \Big\{&\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / (l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
+ &(l, m, n) \in  [0,1, \ldots, N_x] \times  [-\frac{N_y}{2},-\frac{N_y}{2}+1,\ldots, 
+ \frac{N_y}{2}-1] \times [-\frac{N_z}{2},-\frac{N_z}{2},+1,\ldots, \frac{N_z}{2}-1] \Big\} 
  \label{eq:wavenumbermesh}
@@
 notation $K^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
-dimensions of $K_N$. Likewise, ${K}^x(l)={K}_N(l, 0, 0), \D{K}^x(l)=\D{K}_N(l, 0, 0)$ and $\N{K}^x(l)=\N{K}_N(l, 0, 0)$ refer to the wavenumbers in the wall normal direction.
+dimensions of $K_N$. Likewise, ${K}^x(l)={K}_N(l, 0, 0), \D{K}^x(l)=\D{K}_N(l, 0, 0)$ and $\N{K}^x(l)=\N{K}_N(l, 0, 0)$ refer to the wavenumbers, for the different bases, in the wall-normal direction.
 
@@
   X_N = \Big\{ \bm{x} \in \mathcal{R}^3/&(x_i, y_j, z_k) \in \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
-  &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\} \label{eq:Xn}
-\end{align}
-where $x_i = h(i)$ represents either Gauss-Lobatto or Gauss-Chebyshev points
+  &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\}, \label{eq:Xn}
+\end{align}
+where $x_i = h(i)$ represents
 \begin{equation}
  h(i) = \begin{cases}
-   \cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Lobatto points}, \\
-   \cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Chebyshev points}. \\
+   \cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Chebyshev-Gauss-Lobatto points}, \\
+   \cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Chebyshev-Gauss points}. \\
  \end{cases}
@@
 mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_solx}) may be 
-evaluated on the mesh (\ref{eq:Xn}) using fast Fourier and fast Shen inverse 
-transforms as
+evaluated on the mesh (\ref{eq:Xn}) using fast inverse transforms as
 \begin{align}
@@
 \underline{m} 
-y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , 
+y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} 
 \quad \forall \, \bm{x} \in X_N, \notag \\
   u &= \N{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y 
-  (\N{\mathcal{S}}^{-1}_x (\hat{u}))), \quad \forall \, \bm{x} \in X_N.  
+  (\N{\mathcal{S}}^{-1}_x (\hat{u}))) \quad \forall \, \bm{x} \in X_N.  
   \label{eq:ifft} 
@@
 Here $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
-transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represent inverse 
-Fourier transforms along directions $y$ and $z$ respectively, and the inverse 
-Shen transform $\N{\mathcal{S}}_{x}^{-1}$ is performed along the wall normal 
-$x$ direction. Note that the transforms are applied sequentially to the entire 
-computational mesh. The inverse Shen transform may be computed using fast 
+transform, and $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represent inverse discrete
+Fourier transforms along directions $y$ and $z$ respectively. For simplicity, we introduce here a special notation called the inverse 
+\emph{Shen} transform, $\N{\mathcal{S}}^{-1}_x$, that is used to transform coefficients from spectral to physical space in a series expansion that is using either one of the bases in (\ref{eq:Tk}, \ref{eq:phiD}, \ref{eq:phiN}).
+The inverse Shen transform, $\mathcal{S}^{-1}_x$, is performed along the wall-normal
+$x$ direction, and it may be computed using fast 
 Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, 
-\ref{eq:phiN}). The inverse transforms are slightly different for the three 
+\ref{eq:phiN}), see Alg~\ref{alg:ifst} in the Appendix. The transforms are slightly different for the three 
 bases and 
-${\mathcal{S}}^{-1}, \D{\mathcal{S}}^{-1}$ and $\N{\mathcal{S}}^{-1}$ are used 
+${\mathcal{S}}_x^{-1}, \D{\mathcal{S}}_x^{-1}$ and $\N{\mathcal{S}}_x^{-1}$ are used 
 to distinguish between them with obvious notation.
@@
 this work we will make use of the discrete weighted $l^2_w(\Omega)$ space, 
-where quadrature is employed for integration. As such the scalar product 
+where quadrature is employed for the integration. As such, the scalar product 
 corresponds to sums that may be computed using fast transforms. Using 
-$\N{\psi}_{\bm{k}}$ for $v$ we obtain
+$\N{\psi}_{\bm{k}}$ for $v$ we redefine the scalar product with quadrature as
 \begin{align}
@@
   &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
-  (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
+  (\mathcal{F}_{{n}}(u))) \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
 \end{align}
@@
 $y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
-\N{\phi}_l)_w$ is used to represent the Shen scalar product in the 
+\N{\phi}_l)_w$ is used to represent the forward \emph{Shen} scalar product in the 
 $x$-direction. The weights, $w_i$, required for the Shen transforms are 
@@
        \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
-       Gauss-Lobatto points},\\
+       Chebyshev-Gauss-Lobatto points},\\
        \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for 
-       Gauss-Chebyshev points},      
+       Chebyshev-Gauss points},      
  \end{cases}
 \end{equation}
-where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.\footnote{Note that 
-$c_{N_x}=1$ for the continuous scalar product $L^2_w(\Omega)$ (see, e.g., Eq. 
-(2.7) of \cite{Shen95}), but for the discrete $l^2_w(\Omega)$ it follows that 
-$c_{N_x}=2$ due to inexact quadrature for the highest wavenumber.}
+where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$, see Sec.
+1.11 of \cite{kopriva09}.%\footnote{Note that $c_{N_x}=1$ for the continuous scalar product $L^2_w(\Omega)$ (see, e.g., Eq.
+(2.7) of \cite{Shen95}), but for the discrete $l^2_w(\Omega)$ it follows that $c_{N_x}=2$ due to inexact quadrature for the highest wavenumber.}
 
@@
 complete transform. To find a transformation between physical $u$ and spectral 
-$\hat{u}$ we make use of Eq. (\ref{eq:u_solx}) directly in the first line of (\ref{eq:sst1}) and use extensively the discrete 
+$\hat{u}$, we make use of Eq. (\ref{eq:u_solx}) directly on the left hand side of (\ref{eq:sst1}) and use extensively the discrete 
 orthogonality of the Fourier basis functions
@@
            \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
-           &= h \sum_{q=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t). 
+           &= h \sum_{q=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t) \quad \forall \, \bm{k} \in \N{K}_N. 
            \label{eq:sst2}
@@
 \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
-pentadiagonal mass matrix. The complete transformation is now obtained by 
+banded mass matrix with only 5 nonzero diagonals, see Table~\ref{tab:matrices}. The complete transformation is now obtained by 
 setting Eq. (\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. 
@@
 h\N{\mathcal{S}}(u)(l, m, n, t), \notag \\
- \hat{u} &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
+ \hat{u} &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u) 
  \quad \forall \, \bm{k} \in \N{K}_N, \label{eq:fstB}
@@
 \N{\mathcal{T}}^{-1}(\N{\mathcal{T}}(u))$. Note that, since $\N{B}$ assembles to 
-a pentadiagonal matrix, the solution ($\N{B}^{-1}$) can be obtained very fast and the complete 
+a pentadiagonal matrix for its decoupled odd and even coefficients, the solution ($\N{B}^{-1}$) can be obtained very fast and the complete 
 transformation thus requires a fast Chebyshev transform ($\mathcal{O}(N_x \log N_x)$) and 
@@
 (\ref{eq:phiD}), using mass matrices $B_{lq}=(T_q, T_l)_w$ (diagonal) and 
-$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_w$ (tridiagonal) and scalar products 
+$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_w$ (three non-zero diagonals) and scalar products 
 $\mathcal{S}_l(\cdot) = (\cdot, T_l)_w$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
-\D{\phi}_l)_w$. See the Appendix for algorithms for all required transforms.
+\D{\phi}_l)_w$. See Alg.~\ref{alg:fst} and \ref{alg:ifst} in the Appendix for algorithms for all required transforms.
 
@@
 proposed by Kim, Moin and Moser \cite{Kim87}. This scheme is developed by taking the 
-Laplacian of the wall normal momentum equation and the curl of the momentum 
+Laplacian of the wall-normal momentum equation and the curl of the momentum 
 equation. Following elimination of the pressure, the equations to solve are 
@@
 \begin{align}
-	\frac{\partial }{\partial t} \left< \nabla^2 u, \N{\psi}\right>_w &= 
-	\left<h_u, \N{\psi} \right>_w + \nu \left<\nabla^4u, \N{\psi}\right>_w 
-	&\forall \,\N{\psi} \in \N{V}_N, \label{eq:u1} \\
-	\frac{\partial}{\partial t}\left<g, \D{\psi}\right>_w &= \left<h_g, 
-	\D{\psi}\right>_w + \nu 
-	\left<\nabla^2 g, \D{\psi}\right>_w &\forall \, \D{\psi} \in \D{V}_N ,
+	\frac{\partial }{\partial t} \left< \nabla^2 u, \N{\psi}_{\bm{k}}\right>_w &= 
+	\left<h_u, \N{\psi}_{\bm{k}} \right>_w + \nu \left<\nabla^4u, \N{\psi}_{\bm{k}}\right>_w 
+	&\forall \,\N{\psi}_{\bm{k}} \in \N{V}_N, \label{eq:u1} \\
+	\frac{\partial}{\partial t}\left<g, \D{\psi}_{\bm{k}}\right>_w &= \left<h_g, 
+	\D{\psi}_{\bm{k}}\right>_w + \nu 
+	\left<\nabla^2 g, \D{\psi}_{\bm{k}}\right>_w &\forall \, \D{\psi}_{\bm{k}} \in \D{V}_N ,
 	\label{eq:g1} \\
-	\left<f, \D{\psi}\right>_w &= \left<\frac{\partial u}{\partial x}, 
-	\D{\psi}\right>_w &\forall \, \D{\psi} \in \D{V}_N. \label{eq:f1}
+	\left<f, \D{\psi}_{\bm{k}}\right>_w &= \left<\frac{\partial u}{\partial x}, 
+	\D{\psi}_{\bm{k}}\right>_w &\forall \, \D{\psi}_{\bm{k}} \in \D{V}_N. \label{eq:f1}
 \end{align}
@@
 \begin{align}
-\left<f, \D{\psi}\right>_w &= \left<\frac{\partial v}{\partial y} + 
-\frac{\partial w}{\partial z}, \D{\psi}\right>_w \, &\forall \, \D{\psi} \in 
+\left<f, \D{\psi}_{\bm{k}}\right>_w &= \left<\frac{\partial v}{\partial y} + 
+\frac{\partial w}{\partial z}, \D{\psi}_{\bm{k}}\right>_w \, &\forall \, \D{\psi}_{\bm{k}} \in 
 \D{V}_N, \label{eq:f2} \\
-\left<g, \D{\psi}\right>_w &= \left<\frac{\partial w}{\partial y}  - 
-\frac{\partial v}{\partial z}, \D{\psi}\right>_w \, &\forall \, \D{\psi} \in 
+\left<g, \D{\psi}_{\bm{k}}\right>_w &= \left<\frac{\partial w}{\partial y}  - 
+\frac{\partial v}{\partial z}, \D{\psi}_{\bm{k}}\right>_w \, &\forall \, \D{\psi}_{\bm{k}} \in 
 \D{V}_N, \label{eq:g2}
@@
 \frac{\partial }{\partial t} \D{B}_{lq}\hat{v}(q, 0, 0, t) &= 
-\D{B}_{lq}\hat{\mathcal{H}}_y(q, 0, 0, t) + \nu \D{A}_{lq} \hat{v}(q, 0, 0, t) - \left(\beta(t), \D{\phi}_l \right)_w
+\D{B}_{lq}\hat{\mathcal{H}}_y(q, 0, 0, t) - \nu \D{A}_{lq} \hat{v}(q, 0, 0, t) - \left(\beta(t), \D{\phi}_l \right)_w
 & \forall\, l \in \D{K}^x, \label{eq:v0}\\
 \frac{\partial }{\partial t} \D{B}_{lq}\hat{w}(q, 0, 0, t) &= 
-\D{B}_{lq}\hat{\mathcal{H}}_z(q, 0, 0, t) + \nu \D{A}_{lq} \hat{w}(q, 0, 0, t) 
+\D{B}_{lq}\hat{\mathcal{H}}_z(q, 0, 0, t) - \nu \D{A}_{lq} \hat{w}(q, 0, 0, t) 
 & \forall\, l \in \D{K}^x,\label{eq:w0}
 \end{align}
-where $\D{B}_{lq} = \left(\D{\phi}_q, \D{\phi}_l \right)_w$ and $\D{A}_{lq}=\left(\D{\phi}_q^{''}, \D{\phi}_l \right)_w$.
-Note that (\ref{eq:v0}) is the only place where the driving force, or the mean pressure gradient, $\beta$, enters, and, since it is constant in space, the term $\left(\beta(t), \D{\phi}_l \right)_w$ will only be non-zero for $l=0$.
+where $\D{A}_{lq}=-(\D{\phi}_q^{''}, \D{\phi}_l )_w$.
+Note that (\ref{eq:v0}) is the only place where the driving force, or the mean pressure gradient, $\beta$, enters, and, since it is constant in space, the term $(\beta(t), \D{\phi}_l )_w$ will only be non-zero for $l=0$.
 
@@
 act along the first dimension of the transformed variables, i.e.,  $ 
-\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q$ is short for the matrix 
-vector product $ \sum_{q=0}^{N_x-4}\left( \N{\phi}_q^{'}, \D{\phi}_l\right)_w 
+(\N{\phi}_q^{'}, \D{\phi}_l)_w \hat{u}_q$ is short for the matrix 
+vector product $ \sum_{q=0}^{N_x-4}( \N{\phi}_q^{'}, \D{\phi}_l)_w 
 \hat{u}(q, {m}, {n}, t)$, where $m$ and $n$ are given. The inner products are used to set up linear systems 
-of equations for the inhomogeneous wall normal direction. All inner products 
-$\left(\cdot, \cdot\right)_w$ can be represented by sparse matrices.
+of equations for the inhomogeneous wall-normal direction. All scalar products 
+$(\cdot, \cdot)_w$ can be represented by sparse matrices.
 %\begin{align}
@@
 %\end{align}
-The required matrices $ \N{B}_{lq} = \left( \N{\phi}_q, 
-\N{\phi}_l\right)_w, \D{B}_{lq} = \left( \D{\phi}_q, 
-\D{\phi}_l\right)_w, \N{A}_{lq} = \left( \N{\phi}_q^{''}, 
-\N{\phi}_l\right)_w, 
-\D{A}_{lq} = \left( \D{\phi}_q^{''}, \D{\phi}_l\right)_w 
-, \D{C}_{lq} = \left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w$ and 
-$ \N{C}_{lq} = \left(\D{\phi}_q^{'}, \N{\phi}_l\right)_w$ are given in Table~\ref{tab:matrices}, whereas the matrix $\N{Q}_{lq} = \left(\N{\phi}^{''''}_q, \N{\phi}_l\right)_w$ is given below in Eqs (\ref{eq:N_Q})-(\ref{eq:rk}).
+The required matrices $ \N{B}_{lq} = ( \N{\phi}_q, 
+\N{\phi}_l)_w, \D{B}_{lq} = ( \D{\phi}_q, 
+\D{\phi}_l)_w 
+, \N{M}_{lq} = ( \D{\phi}_q, 
+\N{\phi}_l)_w, \D{C}_{lq} = (\N{\phi}_q^{'}, \D{\phi}_l)_w$ and 
+$ \N{C}_{lq} = (\D{\phi}_q^{'}, \N{\phi}_l)_w$ are given in Table~\ref{tab:matrices}, $\N{A}_{lq} = -( \N{\phi}_q^{''}, 
+\N{\phi}_l)_w, \D{A}_{lq} = -( \D{\phi}_q^{''}, \D{\phi}_l)_w$ are given in \cite{Shen95}, whereas the matrix $\N{Q}_{lq} = (\N{\phi}^{''''}_q, \N{\phi}_l)_w$ is given below in Eqs (\ref{eq:N_Q})-(\ref{eq:rk}).
+Note that the matrices are computed using quadrature, which has some implications for Chebyshev-Gauss-Lobatto (GL) points, where the rows of the highest modes differ from those presented in Lemma 2.1 and 3.1 of \cite{Shen95}.
+This disagreement, that follows from inexact quadrature at the highest mode, explains the inclusion of the $c_{k+4}$ term for matrix $\N{B}$ and the $c_{k+2}$ term for matrix $\D{B}$.
+Matrix $B=(T_k, T_j)_w$ is also different in the highest mode from Eq.
+2.7 of \cite{Shen95}, but agrees with Eqs.
+1.135 and 1.136 of \cite{kopriva09}.
 
@@
 \begin{equation}
-\bm{\hat{\mathcal{H}}}_i = \mathcal{\D{T}}((\bm{u} \times \bm{\omega})_i) \quad \forall \, i=0,1,2. \label{eq:H_hat}
+\bm{\hat{\mathcal{H}}}_i = \mathcal{\D{T}}((\bm{u} \times \bm{\omega})_i) \quad \forall \, i=x,y,z. \label{eq:H_hat}
 \end{equation}
 The curl, $\bm{\omega} = (g, \partial_z u - \partial_x w, \partial_y u - 
-\partial_x v)$, is computed by projecting each term to its appropriate space. 
-To this end the two terms $\partial_x v$ and $\partial_x w$ are projected to 
-$V_N$, whereas the remaining $\partial_y u$ and $\partial_z u$ are projected to 
-$\N{V}_N$. Note that the nonlinear term is also generally in need of 
+\partial_x v)$, is computed by projecting each term to its appropriate space, before transforming back to physical space. 
+To this end, the two terms $\partial_x v$ and $\partial_x w$ are projected to 
+$V_N$ (requires $C_{kj} = (\D{\phi}_j^{'}, T_k)_w$, see Tab. \ref{tab:matrices}), whereas the remaining $\partial_y u$ and $\partial_z u$ are projected to 
+$\N{V}_N$:
+\begin{align}
+ \partial_x v(\bm{x}) &= \mathcal{T}^{-1}(B^{-1}C\hat{v}) &\forall \, \bm{x} \in X_N, \\
+ \partial_x w(\bm{x}) &= \mathcal{T}^{-1}(B^{-1}C\hat{w}) &\forall \, \bm{x} \in X_N, \\
+ \partial_y u(\bm{x}) &= \N{\mathcal{T}}^{-1}(\imath \underline{m}\hat{u}) &\forall \, \bm{x} \in X_N, \\
+ \partial_z u(\bm{x}) &= \N{\mathcal{T}}^{-1}(\imath \underline{n}\hat{u}) &\forall \, \bm{x} \in X_N.
+\end{align}
+Note that the nonlinear term is also generally in need of 
 dealiasing, at least for the two periodic directions.
@@
 nonlinear term by Moser, Kim and Mansour (1999). We have implemented both 
-approaches, but for simplicity only the first one is described here. 
+approaches in \cite{spectralDNS}, but for simplicity only the first one is described here. 
 Discretization in time is performed using a constant time step $\triangle t$, 
@@
 where $\underline{z}^2 = \underline{m}^2 + \underline{n}^2$. The convective 
-terms are computed with a forward Adams-Bashforth projection, such that 
+terms are computed with the recipes given in Eqs. (\ref{eq:S_hv}, \ref{eq:S_hg}), discretized in time with a forward Adams-Bashforth projection, such that 
 $\N{\mathcal{S}}(h_v^{{\kappa}+1/2}) = 
@@
 \label{sec:implementation}
-An outline of the solution procedure used, for the numerical method described 
+An outline of the solution procedure, used for the numerical method described 
 in Sec.~\ref{sec:discretizationNS}, is given in Algorithm 
 \ref{alg:NS}.\footnote{Note that throughout this paper we are using the 
-pseudocode conventions of Kopriva \cite{kopriva09}, with one minor difference. 
-A statement like $\{w\}_{k=0}^{N} \gets \{u\}_{k=0}^N$ will here be used to 
-represent a 
-vectorization, where components $u_k$, for $k=0,1,\ldots,N$, are copied from 
+pseudocode conventions of Kopriva \cite{kopriva09}, with some minor differences. 
+A vectorization statement like $\{w_k\}_{k=0}^{N} \gets \{u_k\}_{k=0}^N$ indicates that components $u_k$, for $k=0,1,\ldots,N$, are copied from 
 $u_k$ to $w_k$. Similar conventions apply to matrices, e.g., 
 $\{U_{1,k}\}_{k=0}^N = \{V_{2,k}\}_{k=0}^N$ can be used for copying row 2 of 
-$V$ to row 1 of $U$.} Note 
+$V$ to row 1 of $U$. Broadcasting is also implied, here meaning that for $\{w\}_{k=0}^{N} \gets c$, where $c$ is a scalar, all elements of $w_k$ from $k=0$ to $N$ gets the scalar value $c$.} Note 
 that the Adams-Bashforth projection of convection reverts to a first order 
@@
 is insignificant, but for accuracy tests, like the Orr-Sommerfeld instability, 
-a proper initialization would need to set both the two first time steps.
+a proper initialization will be required to set both the two first time steps.
 \begin{algorithm}
@@
 pseudo-spectral method, and dealiasing in the periodic directions can be either 
-the 3/2-rule \cite{canuto1988} or the 2/3-rule \cite{Orzag71}. The 
-transforms are carried out using fast Fourier and fast Chebyshev transforms 
-that are all of order $\mathcal{O}(N_x\log N_x)$ per one-dimensional transform 
-of size $N_x$. The Shen transforms make use of the Chebyshev transforms that in 
-turn are using 
-discrete cosine transforms. The forward scalar products and transforms are implemented as shown 
-in Algorithm \ref{alg:fst} of the Appendix. A complete transform is obtained through a scalar 
-product followed by solving the linear system involving the mass 
-matrix, just like shown in Eq. (\ref{eq:fstB}) for space $\N{V}_N$ and similar 
-for $\D{V}_N$ and $V_N$. Since the mass matrices are diagonal, tridiagonal and 
-pentadiagonal for $V_N, \D{V}_N$ and $\N{V}_N$ respectively, solving the linear 
-system is $\mathcal{O}(N_x)$ and the complete transform still scales like $\mathcal{O}(N_x\log N_x)$.
-
-Shen \cite{Shen95} states that it is possible to solve Eq.~(\ref{eq:gfin}) 
-with 
-essentially the same number of operations as a pentadiagonal solver. We note 
-that the matrix 
-$\D{H}$ decouples into odd and even components, leading to two matrices of type upper Hessenberg. A solution
-procedure is then simply to perform a direct LU decomposition (see, e.g., 
-\cite{stewart98}), which leads to a 
-lower triangular  
-matrix $L$ with only one subdiagonal and an upper triangular matrix $U$ that 
-is dense. However, each row in $U$ 
-contains at most three distinct values at $U_{k,k}, U_{k,k+2}$ and $U_{k,k+4}$, 
-and then $U_{k,j} = U_{k,k+4}\, \forall \, j = k+6, k+8, \ldots N_x$. 
+the 3/2-rule \cite{canuto1988} or the 2/3-rule \cite{Orzag71}.
+
+Shen \cite{Shen95} states that it is possible to solve Eq.~(\ref{eq:gfin}) directly 
+with essentially the same number of operations as a pentadiagonal solver. To this end, note that the matrix 
+$\D{H}$ decouples into odd and even components, leading to two matrices of type upper Hessenberg. A direct LU decomposition (see, e.g., \cite{stewart98}), leads to a lower triangular  
+matrix $\D{L}$ with only one subdiagonal and an upper triangular matrix $\D{U}$ that is dense. However, each row in $\D{U}$ 
+contains at most three distinct values at $\D{U}_{k,k}, \D{U}_{k,k+2}$ and $\D{U}_{k,k+4}$, 
+and then $\D{U}_{k,j} = \D{U}_{k,k+4}\, \forall \, j = k+6, k+8, \ldots N_x-2$. 
 Consequently 
-only three diagonals in $U$ need storage and the back solve can be traversed 
+only three diagonals in $\D{U}$ need storage and the back solve can be traversed 
 very efficiently in $\mathcal{O}(N_x)$. Note also that the decoupling into 
@@
 
-The biharmonic matrix $\N{H}$ in Eq.~(\ref{eq:Biharmonic_matrix}) is more 
-challenging to solve for efficiently, but 
-it is still possible to find an algorithm that is $\mathcal{O}(N_x)$ for a system of $N_x$ 
-unknowns. Note 
-that $\N{H}$ is the sum of three matrices $\N{Q}, \N{A}$ and $\N{B}$ and can be 
+The biharmonic problem in Eq.~(\ref{eq:ufin}) is more challenging to solve for efficiently, but 
+it is still possible, as suggested yet not devised by Shen \cite{Shen95}, to find an algorithm that is $\mathcal{O}(N_x)$ for a system of $N_x$ unknowns. Note 
+that $\N{H}$ is the sum of three matrices $\N{Q}, \N{A}$ and $\N{B}$, and can be 
 written as
@@
 \end{equation}
-where $\xi_0, \xi_1$ and $\xi_2$ are constants. The matrix $\N{A}_{kj}$ contains only 
+where $\xi_0, \xi_1$ and $\xi_2$ are constants (depending on $m$ and $n$). The matrix $\N{A}_{kj}$ contains only 
 three nonzero diagonals at $ j = k-2, k, k+2$,  whereas matrix $\N{B}_{kj}$ 
@@
 nonzero diagonals at $ j = k-4, k-2, k, k+2, k+4$. The nonzero elements of the 
-upper triangular $\N{Q}_{kj}$ matrix are given as
+upper triangular $\N{Q}_{kj}$ matrix are given as \cite{Shen95}
 \begin{align}
@@
 show stopper for an efficient back solve. However, we note that $\N{U}_{kj}$ 
-contains only three distinct 
-diagonals at $j=k, k+2$ and $k+4$, whereas the remaining part of $\N{U}_{kj}$ 
+contains three distinct 
+diagonals at $j=k, k+2$ and $k+4$, whereas the remaining part of $\N{U}_{kj}$, for $k\le N_x-10$, 
 can be expressed as
@@
 \end{equation}
-where $\{a_k\}_{k=0}^{N_x-4}$ and $\{b_k\}_{k=0}^{N_x-4}$ are two new 
+where $\{a_k\}_{k=0}^{N_x-10}$ and $\{b_k\}_{k=0}^{N_x-10}$ are two new 
 vectors that can be computed recursively from $\N{L}$, as shown in Alg.~\ref{alg:ab}. 
@@
 the total storage requirement for the complete LU decomposition is no more than 
-$7N$, i.e., two diagonals for $\N{L}$, three diagonals for $\N{U}$ plus 
-$\bm{a}$ and 
+$7N_x$, i.e., two diagonals for $\N{L}$, three diagonals for $\N{U}$ plus $\bm{a}$ and 
 $\bm{b}$. The solution of the complete system can be obtained very quickly with 
 one simple forward elimination and a back solve, where the back 
-solve is very fast ($\mathcal{O}(N_x)$) because of the formulation (\ref{eq:ab}), leading 
-to a formula for the backwards substitution like (valid for $k \le N_x-8$)
+solve is very fast ($\mathcal{O}(N_x)$) because of (\ref{eq:ab}), leading 
+to a formula for the backwards substitution of $\N{U}_{kj} \hat{u}_j = y_k$ like (valid for $k \le N_x-10$)
 \begin{equation}
-u_k = \left(y_k - \N{U}_{k, k+2} u_{k+2} - \N{U}_{k, k+4} u_{k+4} - 
-\xi_0 a_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} q_j u_j - 
+\hat{u}_k = \left(y_k - \N{U}_{k, k+2} \hat{u}_{k+2} - \N{U}_{k, k+4} \hat{u}_{k+4} - 
+\xi_0 a_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} q_j \hat{u}_j - 
 \xi_0 b_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} 
-s_j u_j\right)/\N{U}_{kk},
+s_j \hat{u}_j\right)/\N{U}_{kk},
 \end{equation}
@@
 	\DontPrintSemicolon
-    \caption{Recursive formula to compute $\{a_k\}_{k=0}^{N_x-4}, 
-	\{b_k\}_{k=0}^{N_x-4}$ in Eq. (\ref{eq:ab}). Note that the lower triangular 
-	matrix $\{L_{kj}\}_{k,j=0}^{N_x-4}$ contains only two nonzero diagonals and 
+    \caption{Recursive formula to compute $\{a_k\}_{k=0}^{N_x-10}, 
+	\{b_k\}_{k=0}^{N_x-10}$ in Eq. (\ref{eq:ab}). Note that the lower triangular 
+	matrix $\{L_{kj}\}_{k,j=0}^{N_x-4}$ contains only two nonzero subdiagonals and 
 	the parameters $p_k$ and $r_k$ are given in Eqs. (\ref{eq:pk}) and 
 	(\ref{eq:rk}) respectively. \label{alg:ab}}
-		\Fn{\reduceAB{$ \{\N{L}_{kj}\}_{k,j=0}^{N_x-4}, \,\xi_0$}}{
-        \Output{$\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$}  \;
-		\For{$k=0,1$}{
+		\Fn{\reduceAB{$ \{\N{L}_{kj}\}_{k,j=0}^{N_x-4}$}}{
+        \Output{$\{a_k\}_{k=0}^{N_x-10}, \{b_k\}_{k=0}^{N_x-10}$}
+
+        \For{$k=0$ \KwTo $1$}{
         	$a_k \gets p_k$ \;
         	$b_k \gets r_k$ \;
-		}
-		\For{$k=2,3$}{
-		    $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2}$ \;
-		    $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}$ \;
-		}
-		\For{$k=4$ \KwTo $N_x-4$}{
+        	$a_{k+2} \gets p_{k+2} - \N{L}_{k+2,k}\,a_{k}$ \;
+        	$b_{k+2} \gets r_{k+2} - \N{L}_{k+2,k}\,b_{k}$ \;
+        }
+
+		\For{$k=4$ \KwTo $N_x-10$}{
      		$a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2} - \N{L}_{k, k-4} \,a_{k-4}$ 
@@
 		}
-		\Return $\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$ \;
+		\Return $\{a_k\}_{k=0}^{N_x-10}, \{b_k\}_{k=0}^{N_x-10}$ \;
     }
@@
 		\{\N{L}_{kj}\}_{k,j=0}^{N}$ 
-	matrices, as well as the $\{ a_k\}_{k=0}^N$ and $\{ b_k\}_{k=0}^N$ vectors. 
+	matrices, as well as the $\{ a_k\}_{k=0}^{N-6}$ and $\{ b_k\}_{k=0}^{N-6}$ vectors. 
 	The parameters $q_j$ and $s_j$ are given in Eqs. (\ref{eq:pk}) and 
@@
                 \{\N{L}_{kj}\}_{k,j=0}^{N}, 
-                \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N}, \{b_k\}_{k=0}^{N}, 
+                \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N-6}, \{b_k\}_{k=0}^{N-6}, 
                 \xi_0$}}{
@@
 		  \tcp*[l]{Solve $\N{L}_{kj}y_j=f_k$ by forward elimination}
-		  $y_0 \gets f_0$ \;
-		  $y_1 \gets f_1$ \;
-		  $y_2 \gets f_2 - \N{L}_{20}y_{0}$ \;
-		  $y_3 \gets f_3 - \N{L}_{31}y_{1}$ \;
+          \For{$k=0$ \KwTo $1$}
+		  {$y_k \gets f_k$ \;
+		   $y_{k+2} \gets f_{k+2} - \N{L}_{k+2,k} y_{k}$} 
 		  \For{$k = 4$ \KwTo $N$}{
@@
 		  }
-		  \tcp*[l]{Solve $\N{U}_{kj}u_j=y_k$ with back 
+		  \tcp*[l]{Solve $\N{U}_{kj}\hat{u}_j=y_k$ with back 
               substitution}      
-		  $\hat{u}_N \gets y_N / \N{U}_{NN}$ \;
-	      $\hat{u}_{N-1} \gets y_{N-1} / \N{U}_{N-1,N-1}$ \;
-	      $\hat{u}_{N-2} \gets \left(y_{N-2} - \N{U}_{N-2, 
-	      N}y_N\right)/\N{U}_{N-2,N-2}$ \;
-	      $\hat{u}_{N-3} \gets \left(y_{N-3} - \N{U}_{N-3, 
-	      N-1}y_{N-1}\right)/\N{U}_{N-3,N-3}$ \;
+          \For{$k = N$ \KwTo $N-1$ \KwStep $-1$}{
+            $\hat{u}_k \gets y_k / \N{U}_{k,k}$ \;
+            $\hat{u}_{k-2} \gets \left(y_{k-2} - \N{U}_{k-2, 
+            	k}\hat{u}_k\right)/\N{U}_{k-2,k-2}$ \;
+          }
 
@@
 \label{sec:verification}
-We have described a Navier-Stokes solver for channel flows applicable to large-scale simulations. The 
-matrices that make up the linear systems, that need to be solved in order to advance the Navier-Stokes equations, can be inverted using sparse direct LU 
-decompositions, and the solution to these systems may be obtained in $\approx \mathcal{O}(N_x)$. To verify and to further show that we 
-can use these algorithms for large-scale simulations, we will first study some simple 1D problems.
-
-Since we are describing a fully spectral solver, we should be concerned with roundoff errors. The roundoff error of the matrix decomposition methods 
-described for the biharmonic problem in Sec.~\ref{sec:implementation}, and similarly for the Helmholtz problem, may be estimated 
+We have in previous sections described a Navier-Stokes solver for channel flows applicable to large-scale simulations.
+We have devised algorithms for fast direct sparse solvers, scalar products, and for the necessary fast transforms between physical and spectral space.
+The complete solver has been implemented in the open source code spectralDNS~\cite{spectralDNS}.
+For MPI we are using a slab decomposition, as described by Mortensen \cite{Mortensen2015}, but parallel scalability is not considered here since all algorithms described in previous sections are executed in serial.
+In this section we simply verify the implementation, and, to show that these algorithms can be used for large-scale simulations, we will first study some simple 1D problems.
+
+Roundoff errors are often causing problems for spectral methods.
+The roundoff error of the linear solver described for the biharmonic problem in Sec.~\ref{sec:implementation}, and similarly for the Helmholtz problem, may be estimated
 by the following approach: Let $\{u_i\}_{i=0}^{N_x-4}$ be a uniformly distributed 
-random vector. Compute the matrix vector product $\bm{f}=\N{H}\bm{u}$ and 
+random vector of double precision in the interval $(0, 1)$. Compute the matrix vector product $\bm{f}=\N{H}\bm{u}$ and 
 then solve $\N{H} \bm{v} = \bm{f}$, using Alg. \ref{alg:SolveBiharmonic}, for 
-$\{v_i\}_{i=0}^{N_x-4}$. The roundoff error of the matrix decomposition 
-algorithm may then be found as $\max_{0 \ge i \ge N_x-4}|u_i-v_i|/\max_{0 \ge i 
-\ge N_x-4}|u_i|$. For this experiment we assume that $\triangle t= 10^{-5}$ 
+$\{v_i\}_{i=0}^{N_x-4}$. The roundoff error of the solution 
+algorithm may then be found as $\max \{|u_i-v_i|\}_{i=0}^{N_x-4} / \max \{|u_i|\}_{i=0}^{N_x-4}$. For this experiment we use double precision and assume that $\triangle t= 10^{-5}$ 
 and $\nu=1/5200$, that are reasonable parameters for a turbulent channel flow 
-simulation at $\text{Re}_{\tau}=5200$. Four different wavenumbers 
-$z=(0, 200, 1800, 5400)$ are 
+simulation at ${Re}_{\tau}=5200$. Four different wavenumbers 
+$\underline{z}=(0, 200, 1800, 5400)$ are 
 also chosen as representative for such large-scale simulations. The 1800 case corresponds to the highest wavenumber used for the resolution 
 of the largest known channel simulations, performed by Lee and 
-Moser~\cite{leemoser15}, where maximum $\underline{z} = \sqrt{\underline{m}_{max}^2+\underline{n}_{max}^2} = \sqrt{(10240/8)^2 + (7680/6)^2} \approx 1800$. The mesh in the wall normal direction 
+Moser~\cite{leemoser15}, where maximum $\underline{z} = \sqrt{\underline{m}_{max}^2+\underline{n}_{max}^2} = \sqrt{(10240/8)^2 + (7680/6)^2} \approx 1800$. The mesh in the wall-normal direction 
 is varied from 64 to 4096 points, which also covers almost three times higher 
@@
 the Helmholtz matrix roundoff errors are seen to be insignificant.
+Note that Chebyshev-Gauss points are used for the computations in Table~\ref{tab:roundoff} and that Chebyshev-Gauss-Lobatto points give very similar (slightly better) results.
 
 \begin{table}
-	\caption{ Roundoff error of the matrix decomposition method. The first three 
+	\caption{ Roundoff error of the Biharmonic and Helmholtz linear algebra solvers. The first three 
 	rows show results for Alg. (\ref{alg:SolveBiharmonic}) and matrix $\N{H}(\underline{z}, 
@@
 The scaling is seen to be close to linear, and variations
-are most likely due to how well datastructures fit in the cache. Naturally, the timings are not 100$\%$ reproducible, but the ability to solve large-scale problems is evident.
+are most likely due to how well datastructures fit in the cache. The ability to solve large-scale problems is evident.
 \begin{table}
@@
 \hline
-64 & 6.03e-07 (0.00) & 4.48e-07 (0.00) \\ 
-128 & 1.20e-06 (0.99) & 9.00e-07 (1.01) \\ 
-256 & 2.30e-06 (0.96) & 1.81e-06 (1.01) \\ 
-512 & 4.44e-06 (0.96) & 3.59e-06 (0.99) \\ 
-1024 & 1.02e-05 (1.15) & 7.72e-06 (1.07) \\ 
-2048 & 2.03e-05 (0.99) & 1.68e-05 (1.09) \\ 
-4096 & 4.14e-05 (1.02) & 3.33e-05 (0.99) \\ 
-8192 & 8.25e-05 (1.00) & 6.92e-05 (1.04)
+64 & 5.15e-07 (0.00) & 4.01e-07 (0.00) \\ 
+128 & 1.05e-06 (1.02) & 8.03e-07 (1.00) \\ 
+256 & 2.13e-06 (1.01) & 1.65e-06 (1.03) \\ 
+512 & 4.39e-06 (1.03) & 3.39e-06 (1.03) \\ 
+1024 & 8.67e-06 (0.99) & 6.98e-06 (1.03) \\ 
+2048 & 1.75e-05 (1.01) & 1.39e-05 (0.99) \\ 
+4096 & 3.96e-05 (1.13) & 3.03e-05 (1.09) \\ 
+8192 & 7.95e-05 (1.00) & 6.27e-05 (1.03) \\
 	\end{tabular}
@@
 
-The complete Navier Stokes solver should scale like $\mathcal{O}(N)$ and 
-$\mathcal{O}N\log N)$ since it is a combination of linear algebra routines 
-$\mathcal{O}(N)$, fast transforms $\mathcal{O}(N \log N)$ and several 
-elementwise array operations $\mathcal{O}(N)$.
-
-We show in Table \ref{tab:timings_solver} some timings for one single time 
-step, increasing only the wall normal discretization. We show that the total 
-time scales approximately as $N \log N$ for large $N$ and better than linearly 
-for small $N$. This shift happens since at small $N$ the transforms are fast 
-and linear operations, like elementwise multiplications, are more dominating. 
+The Orr-Sommerfeld equation (see, e.g., \cite{Orszag_1971}) is further used to establish that we are solving the right equations, and that the accuracy of the solver is second-order in time.
+The Orr-Sommerfeld eigenvalue problem is first solved for $Re=8000$, using the biharmonic Shen basis (\ref{eq:phiN}).
+The leading eigenvalue/eigenvector pair is then added as a perturbation to a laminar, parabolic, base flow, leading to an initial field
+\begin{align}
+u(x,y,t) &= -\epsilon  \mathrm{Re}\{\imath \xi(x) \exp (\imath (y-\lambda t))\}, \label{eq:OS1} \\
+v(x,y,t) &= 1-x^2 + \epsilon \mathrm{Re}\{\xi'(x) \exp(\imath(y-\lambda t)\}, \label{eq:OS2}
+\end{align}
+where $\xi(x)$ is the eigenvector, $\lambda=0.2470750602+\imath 2.664410371\cdot 10^{-3}$ the eigenvalue and $\epsilon \in \mathcal{R}^+$ is a small positive constant.
+Note that if $\epsilon$ is chosen as too big, then the initial/boundary value problem will not agree well with the linearized Navier-Stokes equations used in deriving the Orr-Sommerfeld equations, due to second-order terms.
+Likewise, if $\epsilon$ is chosen as too small, then roundoff errors will be too large, since the perturbation is added to a base flow of order unity.
+It turns our that $\epsilon=10^{-7}$ is a good compromise, and that a mesh of size $128 \times 8 \times 2$ is accurate enough to isolate temporal errors.
+Since the evolution of the perturbation is known analytically, the $L^2$-norm of the deviation from this exact linearized solution is used as a measure of the error.
+We also compute the error in the perturbation energy by
+\begin{equation}
+E(t) =  \frac{\left<\tilde{\bm{u}}(\bm{x}, t), \tilde{\bm{u}}(\bm{x}, t) \right>}{\left<\tilde{\bm{u}}(\bm{x}, 0), \tilde{\bm{u}}(\bm{x}, 0) \right>}  - \exp(2 \mathrm{Im}\{\lambda\}t), \label{eq:OSenergy}
+\end{equation}
+where $\tilde{\bm{u}} = [u, v-(1-x^2)]$ represents the perturbation vector. 
+%Table \ref{tab:epsilon} shows the perturbation error as a function of $\epsilon$, using a mesh of size $128 \times 8 \times 2$ or $512 \times 8 \times 2$, $\triangle t=0.2$ and final time $t=5$.
+Since the perturbation is added to a base flow of order unity, we know that one digit will be lost in accuracy, for each row, moving down Table \ref{tab:epsilon}.
+Adding to this the normal roundoff, see Table \ref{tab:roundoff}, $\epsilon=10^{-8}$ should not be expected to yield more than 6-7 digits of accuracy, whereas the final row ($\epsilon=10^{-12}$) no more than 2-3.
+Still, we observe that the solution is nearly independent of $\epsilon$ for $10^{-6} < \epsilon < 10^{-8}$, and as such $\epsilon=10^{-7}$ is chosen as suitable for computing the temporal accuracy.
+Note also that we can assume spatial discretization errors are negligible, as the case with $N_x=512$ gives more or less the same result as $N_x=128$.
+%\begin{table}
+%	\centering
+%	\caption{ Error in Orr-Sommerfeld perturbation as a function of perturbation magnitude $\epsilon$.
+The $L^2$-norm $ \| \bm{u}' \| = \sqrt{\left< \bm{u}', \bm{u}'\right>}$, where $\bm{u}'=\bm{u}-\bm{u}_{\mathrm{exact}}$ and energy error (\ref{eq:OSenergy}), for the most unstable Orr-Sommerfeld mode.
+Results are computed with a mesh of size $128 \times 8 \times 2$ or $512 \times 8 \times 2$, with $\triangle t=0.2$ and final time $t=5$.
+\label{tab:epsilon}}
+%	\begin{tabular}{ccccc}	
+%		$-\log_{10}(\epsilon)$ & \multicolumn{2}{c}{$\frac{1}{\epsilon}\| \bm{u}' \|(t=5)$} & \multicolumn{2}{c}{$E(t=5)$} \\
+%		\hline
+%		 & $N_x=128$ & $N_X=512$ & $N_x=128$ & $N_X=512$ \\
+%		\hline
+%4	&	1.23454361e-02	&	1.23454363e-02	&	2.44150583e-04	&	
+%2.44150404e-04	\\
+%5	&	1.15753545e-02	&	1.15753547e-02	&	2.44052002e-04	&	
+%2.44051822e-04	\\
+%6	&	1.15673948e-02	&	1.15673950e-02	&	2.44051018e-04	&	
+%2.44050836e-04	\\
+%7	&	1.15673151e-02	&	1.15673154e-02	&	2.44051004e-04	&	
+%2.44050816e-04	\\
+%8	&	1.15673142e-02	&	1.15673156e-02	&	2.44050426e-04	&	
+%2.44050992e-04	\\
+%9	&	1.15673192e-02	&	1.15673185e-02	&	2.44049842e-04	&	
+%2.44050701e-04	\\
+%10	&	1.15672575e-02	&	1.15674820e-02	&	2.44043530e-04	&	
+%2.44052242e-04	\\
+%11	&	1.15649705e-02	&	1.15684397e-02	&	2.43713173e-04	&	
+%2.44010407e-04	\\
+%12	&	1.15568665e-02	&	1.15844457e-02	&	2.43541116e-04	&	
+%2.44202451e-04
+%	\end{tabular}
+%\end{table}
+Table \ref{tab:OS} shows the computed errors using a final time of $t=50$, corresponding to 2 flow-throughs, and a varying $\triangle t$.
+We observe, as expected, second-order accuracy in the $L^2$-norm in time.
+The error in the accumulated perturbation energy is also found to be close to second order.
+Table \ref{tab:OS_spatial} shows the computed error as a function of spatial discretization, where the time step is kept constant at $\triangle t=10^{-3}$ for 50 steps.
+The spectral decay of the error, with increasing $N_x$, is evident, and there is little difference between using either Chebyshev-Gauss (GC) or Chebyshev-Gauss-Lobatto (GL) points.
+Note that there is no difference at all between GC and GL for the results shown Table \ref{tab:OS}.
+
+\begin{table}
+	\centering
+	\caption{Error in the Orr-Sommerfeld perturbation at $Re=8000$, for the most unstable eigenmode, as a function of time step.
+The mesh size is $128 \times 8 \times 2$.
+The $L^2$-norm $ \| \bm{u}' \| = \sqrt{\left< \bm{u}', \bm{u}'\right>}$, where $\bm{u}'=\bm{u}-\bm{u}_{\mathrm{exact}}$, and the energy error is computed as Eq.
+(\ref{eq:OSenergy}).
+The numbers in parenthesis are showing the order of the error, computed as $\log(\tau_i/\tau_{i-1})/\log(\triangle t_i/ \triangle t_{i-1})$, where $i$ is the row number of the table and $\tau$ represents the values of $\|\bm{u}'\|$ or $E(t=50)$.
+\label{tab:OS}}
+	\begin{tabular}{ccc}	
+ $\triangle t$ & $\| \bm{u}' \|(t=50)$ & $\int_{t=0}^{t=50}E(t) \mathrm{d}t$ \\
+ \hline
+
+1.0000e-01 & 2.9005e-09  (0.000) & 1.2775e-02 (0.000) \\
+6.6667e-02 & 1.2877e-09  (2.003) & 5.3068e-03 (2.167) \\
+5.0000e-02 & 7.2559e-10  (1.994) & 2.8939e-03 (2.108) \\
+4.0000e-02 & 4.6445e-10  (1.999) & 1.8138e-03 (2.094) \\
+3.3333e-02 & 3.2257e-10  (1.999) & 1.2418e-03 (2.078) \\
+2.8571e-02 & 2.3701e-10  (1.999) & 9.0307e-04 (2.066) \\
+2.5000e-02 & 1.8148e-10  (1.999) & 6.8608e-04 (2.058)
+	\end{tabular}
+\end{table}
+
+\begin{table}
+	\centering
+	\caption{Error in the Orr-Sommerfeld perturbation at $Re=8000$, for the most unstable eigenmode, as a function of $N_x$.
+The time step $\triangle t = 10^{-3}$ and the end time is $t=0.05$.
+The $L^2$-norm $ \| \bm{u}' \| = \sqrt{\left< \bm{u}', \bm{u}'\right>}$, where $\bm{u}'=\bm{u}-\bm{u}_{\mathrm{exact}}$, and the energy error is computed as Eq.
+(\ref{eq:OSenergy}).
+GC and GL refer to Chebyshev-Gauss and Chebyshev-Gauss-Lobatto points, respectively.
+\label{tab:OS_spatial}}
+	\begin{tabular}{ccccc}	
+		$N_x$ & \multicolumn{2}{c}{${\| \bm{u}' \|(t=0.05)}$} & \multicolumn{2}{c}{$\int_{t=0}^{t=0.05}E(t) \mathrm{d}t$} \\
+		\hline
+			 & GC & GL & GC & GL \\
+		\hline
+16	&	3.23081791e-01	&	4.76365364e-01	&	3.35105681e-02	&	3.10940983e-03	\\
+32	&	5.71635963e-03	&	6.77845098e-03	&	2.57328408e-04	&	5.02990613e-05	\\
+64	&	6.99681587e-08	&	8.22331688e-08	&	1.75070691e-09	&	8.04535105e-10	\\
+128	&	5.06389229e-08	&	4.81745061e-08	&	1.01924047e-09	&	1.53428292e-09	\\
+256	&	4.89447535e-08	&	4.83040873e-08	&	8.40445491e-10	&	3.77475917e-09	
+	\end{tabular}
+\end{table}
+
+The final test for this fully spectral Navier-Stokes solver is the pressure driven turbulent channel flow.
+First, to measure scaling of the complete 3-dimensional solver, we use a rather small turbulent Reynolds number of $Re_{\tau}=590$ and a small computational box of size $N_x \times 32 \times 32$.
+Since the resolutions in $y$ and $z$ directions are kept constant, the complete Navier Stokes solver should ultimately scale like and
+$\mathcal{O}(N_x\log N_x)$, since it is a combination of linear algebra routines 
+$\mathcal{O}(N_x)$, fast transforms $\mathcal{O}(N_x \log N_x)$ and several elementwise array operations $\mathcal{O}(N_x)$. We show in Table \ref{tab:timings_solver} some timings for one single time 
+step, increasing only the wall-normal discretization. We see that the total 
+time scales approximately as $N_x \log N_x$ for large $N_x$ and linearly 
+for small $N_x$. This shift happens since at small $N_x$ the transforms are fast and linear operations, like elementwise multiplications, are more dominating. 
 We also see that the time for assembly of the right hand side is approximately 
-$\mathcal{O}(N \log N)$, whereas the time for solving the linear systems is 
+$\mathcal{O}(N_x \log N_x)$, whereas the time for solving the linear systems is 
 very nearly linear. Note that for fast transforms we are using 
 the FFTW library \cite{libfftw}, with planner effort set to FFTW$\_$MEASURE. 
-Timings in Table \ref{tab:timings_solver} are computed with dealiasing 
+The results in Table \ref{tab:timings_solver} are computed with GC points and dealiasing is
 performed with the 2/3-rule \cite{Orzag71}.
@@
 	\centering
-	\caption{Computing time for one time step using 32 wavenumbers in each periodic direction and $N_x$ points in the wall normal direction.
+	\caption{Computing time for one time step using 32 wavenumbers in each periodic direction and $N_x$ points in the wall-normal direction.
 The Assemble and Solve columns show the time required to assemble the right hand side (should be $\mathcal{O}(N_x \log N_x)$ due to transforms) and solve the linear systems, respectively.
-The number in parenthesis shows the scaling, and unity would mean $\mathcal{O}(N_X \log N_x)$ for Total and Assemble, whereas it would indicate $\mathcal{O}(N_x)$ for Solve.
+The number in parenthesis shows the scaling, and unity would mean $\mathcal{O}(N_x \log N_x)$ for Total and Assemble, whereas it would indicate $\mathcal{O}(N_x)$ for Solve.
 \label{tab:timings_solver}}
 	\begin{tabular}{cccc}	
-$N_x$ & Total ($\frac{t_k }{t_{k-1}} \frac{\log N-1}{2 \log N}$) & Assemble ($\frac{t_k }{t_{k-1}} \frac{\log N-1}{2 \log N}$) & Solve ($\frac{t_k}{t_{k-1}} \frac{1}{2}$) \\ 
+		$N_x$ & Total ($\frac{t_k }{t_{k-1}} \frac{\log N_x-1}{2 \log N_x}$) & Assemble ($\frac{t_k }{t_{k-1}} \frac{\log N_x-1}{2 \log N_x}$) & Solve ($\frac{t_k}{t_{k-1}} \frac{1}{2}$) \\ 
 		\hline
@@
 
-The Orr-Sommerfeld equations are further used to establish that we are solving the right equations, and that the accuracy of the solver is second order in time.
-The Orr-Sommerfeld eigenvalue problem is first solved for $Re=8000$ using the biharmonic Shen basis (\ref{eq:phiN}).
-The leading eigenvalue/eigenvector pair is then added as a perturbation to a laminar, parabolic, base flow, leading to an initial field
-\begin{align}
-u(x,y,t) &= -\epsilon  \mathrm{Re}\{\xi(x) \exp (\imath (y-\lambda t))\}, \label{eq:OS1} \\
-v(x,y,t) &= 1-x^2 + \epsilon \mathrm{Re}\{\imath \xi(x) \exp(\imath(y-\lambda t)\}, \label{eq:OS2}
-\end{align}
-where $\xi(x)$ is the eigenvector, $\lambda$ the eigenvalue and $\epsilon \in \mathcal{R}^+$ is a small positive constant.
-Note that if $\epsilon$ is chosen as too big, then the initial/boundary value problem will not agree well with the linearized Navier-Stokes equations used in deriving the Orr-Sommerfeld equations, and the evolution from (\ref{eq:OS1}, \ref{eq:OS2}) will not follow the linearized equations.
-Likewise, if $\epsilon$ is chosen as too small, then roundoff errors will be too large, since the perturbation is added to a base flow of order unity.
-Since the evolution of the perturbation is known analytically, the departure from this exact linearized solution is used as a measure of the error of the current solver.
-We also compute the error in the perturbation energy by
-\begin{equation}
-E(t) =  \frac{\left<\tilde{\bm{u}}(\bm{x}, t), \tilde{\bm{u}}(\bm{x}, t) \right>}{\left<\tilde{\bm{u}}(\bm{x}, 0), \tilde{\bm{u}}(\bm{x}, 0) \right>}  - \exp(2 \mathrm{Im}\{\lambda\}t), \label{eq:OSenergy}
-\end{equation}
-where $\tilde{\bm{u}} = [u, v-(1-x^2)]$ represents the perturbation vector, and $\lambda=0.2470750602+\imath 2.664410371\cdot 10^{-3}$ is the eigenvalue.
-
-Table \ref{tab:epsilon} shows the perturbation error as a function of $\epsilon$, using a mesh of size $128 \times 8 \times 2$, $\triangle t=0.2$ and final time $t=5$.
-Since the perturbation is added to a base flow of order unity, we know that one digit will be lost in accuracy, for each row, moving down Table \ref{tab:epsilon}.
-Adding to this the normal roundoff, see Table \ref{tab:roundoff}, $\epsilon=10^{-8}$ should not be expected to yield more than 6-7 digits of accuracy, whereas the final row ($\epsilon=10^{-12}$) no more than 2-3.
-Still, we observe that the solution is nearly independent of $\epsilon$ for $10^{-6} < \epsilon < 10^{-8}$, and as such $\epsilon=10^{-7}$ is chosen as suitable for computing the temporal accuracy.
-\begin{table}
-	\centering
-	\caption{ Error in Orr-Sommerfeld perturbation as a function of perturbation magnitude $\epsilon$.
-The $L^2$-norm $ \| \bm{u}' \| = \sqrt{\left< \bm{u}', \bm{u}'\right>}$, where $\bm{u}'=\bm{u}-\bm{u}_{\mathrm{exact}}$ and energy error (\ref{eq:OSenergy}), for the most unstable Orr-Sommerfeld mode.
-Results are computed with a mesh of size $128 \times 8 \times 2$ or $512 \times 8 \times 2$, with $\triangle t=0.2$ and final time $t=5$.
-\label{tab:epsilon}}
-	\begin{tabular}{ccccc}	
-		$-\log_{10}(\epsilon)$ & \multicolumn{2}{c}{$\frac{1}{\epsilon}\| \bm{u}' \|(t=5)$} & \multicolumn{2}{c}{$E(t=5)$} \\
-		\hline
-		 & $N=128$ & $N=512$ & $N=128$ & $N=512$ \\
-		\hline
-4	&	1.23454361e-02	&	2.44150583e-04	&	1.23454363e-02	&	
-2.44150404e-04	\\
-5	&	1.15753545e-02	&	2.44052002e-04	&	1.15753547e-02	&	
-2.44051822e-04	\\
-6	&	1.15673948e-02	&	2.44051018e-04	&	1.15673950e-02	&	
-2.44050836e-04	\\
-7	&	1.15673151e-02	&	2.44051004e-04	&	1.15673154e-02	&	
-2.44050816e-04	\\
-8	&	1.15673142e-02	&	2.44050426e-04	&	1.15673156e-02	&	
-2.44050992e-04	\\
-9	&	1.15673192e-02	&	2.44049842e-04	&	1.15673185e-02	&	
-2.44050701e-04	\\
-10	&	1.15672575e-02	&	2.44043530e-04	&	1.15674820e-02	&	
-2.44052242e-04	\\
-11	&	1.15649705e-02	&	2.43713173e-04	&	1.15684397e-02	&	
-2.44010407e-04	\\
-12	&	1.15568665e-02	&	2.43541116e-04	&	1.15844457e-02	&	
-2.44202451e-04
-	\end{tabular}
-\end{table}
-
-Table \ref{tab:OS} shows results from using a mesh of size $128 \times 8 \times 2$, with final time $t=50$, corresponding to 2 flow-throughs, and a varying $\triangle t$.
-We observe, as expected, second order accuracy in the $L^2$-norm in time.
-The error in the accumulated perturbation energy is also found to be close to second order.
-
-\begin{table}
-	\centering
-	\caption{Error in Orr-Sommerfeld perturbation as a function of time step.
-The mesh size is $128 \times 8 \times 2$.
-The $L^2$-norm $ \| \bm{u}' \| = \sqrt{\left< \bm{u}', \bm{u}'\right>}$, where $\bm{u}'=\bm{u}-\bm{u}_{\mathrm{exact}}$, and energy error (\ref{eq:OSenergy}), for the most unstable Orr-Sommerfeld mode.
-The numbers in parenthesis are showing the order of the error, computed as $\log(\tau_i/\tau_{i-1})/\log(\triangle t_i/ \triangle t_{i-1})$, where $i$ is the row number of the table and $\tau$ represents the values of $\|\bm{u}'\|$ or $E(t=50)$.
-\label{tab:OS}}
-	\begin{tabular}{ccc}	
- $\triangle t$ & $\| \bm{u}' \|(t=50)$ & $\int_{t=0}^{t=50}E(t) \mathrm{d}t$ \\
- \hline
-
-1.0000e-01 & 2.9005e-09  (0.000) & 1.2775e-02 (0.000) \\
-6.6667e-02 & 1.2877e-09  (2.003) & 5.3068e-03 (2.167) \\
-5.0000e-02 & 7.2559e-10  (1.994) & 2.8939e-03 (2.108) \\
-4.0000e-02 & 4.6445e-10  (1.999) & 1.8138e-03 (2.094) \\
-3.3333e-02 & 3.2257e-10  (1.999) & 1.2418e-03 (2.078) \\
-2.8571e-02 & 2.3701e-10  (1.999) & 9.0307e-04 (2.066) \\
-2.5000e-02 & 1.8148e-10  (1.999) & 6.8608e-04 (2.058)
-	\end{tabular}
-\end{table}
-
-The ultimate test for this fully spectral Navier-Stokes solver is the pressure driven turbulent channel flow.
-We use the turbulent Reynolds number $Re_{\tau}=2000$, which has been simulated previously by Hoyas and Jim\'{e}nez \cite{hoyas06}, Lozano and Jim\'{e}nez \cite{Lozano2014} and Lee and Moser \cite{leemoser15}.
-Lozano and Jim\'{e}nez showed that a small box  of size $2 \times 2\pi \times \pi$ in $x, y,$ and $z$ directions, respectively, could be simulated with $633 \times 1024 \times 1024$ collocation points (compact finite differences in wall normal direction, Fourier in periodic) in order to capture all relevant one-point statistics found in much larger boxes.
+For the ultimate test of a fully turbulent flow, we use the turbulent Reynolds number $Re_{\tau}=2000$, which has been simulated previously by Hoyas and Jim\'{e}nez \cite{hoyas06}, Lozano and Jim\'{e}nez \cite{Lozano2014} and Lee and Moser \cite{leemoser15}.
+Lozano and Jim\'{e}nez showed that a small box  of size $2 \times 2\pi \times \pi$ in $x, y,$ and $z$ directions, respectively, could be simulated with $633 \times 1024 \times 1024$ collocation points (compact finite differences in wall-normal direction, Fourier in periodic) in order to capture all relevant one-point statistics found in much larger boxes.
 We use the same small computational box, with resolution of $512 \times 1024 \times 1024$, and expect the superior resolution properties of spectral methods to work in our favor.
-The solver has been implemented with a slab MPI decomposition, and simulations were run on the Shaheen II supercomputer at KAUST using 512 cores.
+Simulations were run on the Shaheen II supercomputer at KAUST, using 512 cores.
 With a time step of $2\cdot10^{-5}$, the solver used approximately 1 second per time step.
 After reaching steady state, statistics were sampled for 100,000 time steps (enough to converge first order statistics), and the average velocity is shown in Figure \ref{fig:U_mean} in a dashed line, alongside the results obtained by Hoyas and Jim\'{e}nez \cite{hoyas06} in a dotted line.
-The difference is hardly visible, showing that the fully spectral Shen-Fourier solver is reliable for $Re_{\tau} = 2000$.
-Furthermore, the results presented in Tables \ref{tab:roundoff}, \ref{tab:timings} and \ref{tab:timings_solver} indicate that the solver is more than likely to be reliable for even larger scales.
+The difference is hardly visible, showing that the fully spectral Shen-Fourier solver is reliable and robust for $Re_{\tau} = 2000$.
+Furthermore, the results presented in Tables \ref{tab:roundoff}, \ref{tab:timings} and \ref{tab:timings_solver} indicate that the solver is more than likely to be reliable for even higher Reynolds numbers and larger scales.
 
@@
 properties, and are usually the preferred tools for performing Direct Numerical 
-Simulations of turbulent flows. A spectral discretization of the wall normal 
-direction in a plane channel is challenging due to roundoff errors from poorly 
-conditions coefficient matrices, and as such all recent large-scale channel DNS 
-have been conducted without spectral accuracy in the wall normal direction.
-
-In this paper a Navier Stokes solver for large-scale turbulent channel flows 
-has been described. In the wall normal direction the solver utilizes base 
+Simulations (DNS) of turbulent flows. A spectral discretization of the wall-normal 
+direction in a plane channel is challenging, due to roundoff errors, and as such recent large-scale channel DNS 
+have been conducted without spectral accuracy in the wall-normal direction.
+
+In this paper a spectral-Galerkin Navier Stokes solver for large-scale turbulent channel flows 
+has been described. In the wall-normal direction, the solver utilizes base 
 functions 
@@
 The computational cost of the solver is shown to scale with the expected 
-$\mathcal{O}(N \log N)$, for a mesh of size $N \times N \times N$, due to fast 
+$\mathcal{O}(N_x \log N_x)$, for a mesh of size $N_x \times N_y \times N_z$ ($N_y$ and $N_z$ kept constant), due to fast 
 transforms. Direct 
@@
 and these have been shown to scale close to 
-$\mathcal{O}(N)$, with insignificant roundoff errors. The fully spectral solver 
+$\mathcal{O}(N_x)$, with insignificant roundoff errors. The fully spectral solver 
 has been applied to a turbulent 
 channel flow at $Re_{\tau}=2000$, and the first order statistics are shown to agree well with the literature.
+The solver has been implemented as part of the open source project spectralDNS \cite{spectralDNS}.
 
 \section*{Acknowledgements}
-This research is a part of the 4DSpace Strategic Research Initiative at the 
-University of Oslo.
+This research is a part of the 4DSpace Strategic Research Initiative at the University of Oslo.
+I am also grateful to David Ketcheson and the KAUST Supercomputing Laboratory for providing access to Shaheen II.
 
@@
 
-\section{Appendices}
+\section*{Appendix}
 \label{sec:app}
-%\subsection*{Fast Shen transforms}
-%The fast forward and inverse Chebyshev transforms are used by the Shen 
-%transforms. The code below computes both scalar products and complete 
-%transforms for both 
-%
-%\begin{python}
-%from numpy import *
-%from scipy.fftpack import dct
-%
-%def fct(f, transform="GL"):
-%    """Fast Chebyshev transform."""
-%    N = f.shape[0]
-%    f_hat = zeros(N)
-%    if transform == "GL":
-%        fk[:] = dct(fj, type=2, axis=0)            
-%        fk /= N
-%        fk[0] /= 2
-%                
-%    elif transform == "GC":
-%        fk[:] = dct(fj, type=1, axis=0)/(N-1)
-%        fk[0] /= 2
-%        fk[-1] /= 2
-%            
-%    return fk
-%
-%def ifct(fk, transform="GL"):
-%    """Inverse fast Chebyshev transform."""
-%    if transform == "GL":
-%        f = 0.5*dct(fk, type=3, axis=0)
-%        f += 0.5*fk[0]
-%    
-%    elif transform == "GC":
-%        f = 0.5*dct(fk, type=1, axis=0)
-%        f += 0.5*fk[0]
-%        f[::2] += 0.5*fk[-1]
-%        f[1::2] -= 0.5*fk[-1]
-%
-%    return f
-%
-%def fastChebScalar(f, transform="GL"):
-%    """Fast Chebyshev scalar product."""
-%    N = f.shape[0]
-%    if transform == "GL":
-%        fk = dct(fj, type=2, axis=0)*pi/(2*N)
-%        
-%    elif transform == "GC":
-%        fk = dct(fj, type=1, axis=0)*pi/(2*(N-1))
-%    return fk
-%
-%def fastShenScalar(f, transform="GL"):
-%    """Fast Shen scalar product"""                
-%    fk = fastChebScalar(f, transform)
-%    fk[:-2] -= fk[2:]
-%    return fk
-%
-%def ifst(fk):
-%    """Fast inverse Shen transform
-%    """
-%    N = len(fj)
-%    wk = np.zeros(N, dtype=fk.dtype)        
-%    wk[:-2] = fk[:-2] 
-%    wk[2:] -= fk[:-2] 
-%    return ifct(wk)
-%    
-%
-%def fst(fj, transform="GL"):
-%    """Fast Shen transform"""
-%    fk = fastShenScalar(fj, transform)
-%    
-%    N = fj.shape[0]
-%    if transform == "GL":
-%        ck = np.ones(N-2); ck[0] = 2
-%        
-%    elif transform == "GC":
-%        ck = np.ones(N-2); ck[0] = 2; ck[-1] = 2
-%        
-%    a = np.ones(N-4)*(-np.pi/2)
-%    b = np.pi/2*(ck+1)
-%    c = a.copy()
-%    fk[:-2] = TDMA_1D(a, b, c, fk[:-2])
-%        
-%    return fk
-%
-%def fastShenScalar(f, transform="GL"):
-%    """Fast Shen scalar product Neumann basis"""        
-%    k  = wavenumbers(f.shape)
-%    fk = fastChebScalar(f, transform)
-%    fk[:-2] -= ((k/(k+2))**2) * fk[2:]
-%    return fk
-%
-%def ifst(fk):
-%    """Fast inverse Shen scalar transform Neumann basis"""
-%    k = self.wavenumbers(fk.shape[0])
-%    wk = np.zeros(fk.shape[0])    
-%    wk[1:-2] = fk[1:-2]
-%    wk[3:] -= (k[1:]/(k[1:]+2))**2*fk[1:-2]
-%    f = ifct(wk, transform)
-%    return f
-%        
-%def fst(f, transform="GL"):
-%    """Fast Shen transform Neumann basis"""
-%    fk = fastShenScalar(f, transform)
-%    N = fj.shape[0]
-%    k = wavenumbers(N)
-%    ck = np.ones(N-3)
-%    if transform == "GC": ck[-1] = 2
-%    a = np.ones(N-5)*(-np.pi/2)*(k[1:-2]/(k[1:-2]+2))**2
-%    b = np.pi/2*(1+ck*(k[1:]/(k[1:]+2))**4)
-%    c = a.copy()
-%    fk[1:-2] = TDMA_1D(a, b, c, fk[1:-2])
-%    return fk
-%
-%# Cython implementation of TDMA    
-%import numpy as np
-%cimport cython
-%cimport numpy as np
-%#cython: boundscheck=False
-%#cython: wraparound=False
-%
-%ctypedef fused T:
-%    np.float64_t
-%    np.complex128_t
-%
-%def TDMA_1D(np.ndarray[np.float64_t, ndim=1] a, 
-%            np.ndarray[np.float64_t, ndim=1] b, 
-%            np.ndarray[np.float64_t, ndim=1] c, 
-%            np.ndarray[T, ndim=1] d):
-%    cdef:
-%        unsigned int n = b.shape[0]
-%        unsigned int m = a.shape[0]
-%        unsigned int k = n - m
-%        int i
-%        
-%    for i in range(m):
-%        d[i + k] -= d[i] * a[i] / b[i]
-%        b[i + k] -= c[i] * a[i] / b[i]
-%    for i in range(m - 1, -1, -1):
-%        d[i] -= d[i + k] * c[i] / b[i + k]
-%    for i in range(n):
-%        d[i] /= b[i]
-%        
-%    return d
-%    
-%
-%\end{python}
-
-\subsection*{Fast scalar products and transforms}
-
+
+\paragraph*{Fast scalar products and transforms}
 Algorithms for fast scalar products and transforms, for the three 
@@
 depending on the choice of discretization points and direction.
+
+\paragraph*{Matrices}
+Matrices representing one-dimensional scalar products are given in Table \ref{tab:matrices}.
+Note that the mass matrices are symmetrical, and as such only the upper triangular parts are described.
+Also note that the matrices are computed using quadrature, and not exact integration, which has some minor implications for Chebyshev-Gauss-Lobatto (GL) points.
+This follows since for an exact $L^2_w$ scalar product we would have $c_{N_x}=1$ for GL as well as Chebyshev-Gauss (GC), but for the $l^2_w$ scalar product used here $c_{N_x}=2$.
+This disagreement, that follows from inexact quadrature at the highest mode, explains the inclusion of the $c_{k+4}$ term for matrix $\N{B}$, and the $c_{k+2}$ term for matrix $\D{B}$.
+These are missing from Lemma 2.1 and 3.1 of \cite{Shen95}, because Shen here considers only the exact $L^2_w$ scalar product.
+
 \SetKwFunction{forwardScalar}{forwardScalar}
@@
 
-			\If{ points = "Gauss-Lobatto"}{
-			$\{ w_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, \text{type=2}, 
+			\If{ points = "Chebyshev-Gauss-Lobatto"}{
+			$\{ w_k\}_{k=0}^{N} \gets \text{dct}(\{f_j\}_{j=0}^{N}, \text{type=2}, 
 			\text{axis=0})\frac{\pi}{2 N}$ \;
             }
-			\ElseIf{ points = "Gauss-Chebyshev"}{
-			$\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=1}, 
+			\ElseIf{ points = "Chebyshev-Gauss"}{
+			$\{ w_k\}_{k=0}^{N} \gets \text{dct}(\{f_j\}_{j=0}^{N}, \text{type=1}, 
 			\text{axis=0}) \frac{\pi}{2 (N-1)}$ \;
@@
 			$y_0 \gets y_0 / 2$ \;
-			\If{ points = "Gauss-Lobatto"}{
+			\If{ points = "Chebyshev-Gauss-Lobatto"}{
 			$y_N \gets y_N / 2$ \;   	
@@
             $\{y_k\}_{k=0}^{N} \gets$ forwardScalar($\{f_j\}_{j=0}^{N}$, 
-points, space) \;
+\emph{points}, \emph{space}) \;
 			\If{space = $\D{V}_N$}{
@@
 \end{algorithm}
-
 \SetKwFunction{inverseChebTransform}{inverseChebTransform}
@@
 		Here "dct" is the discrete cosine transform from SciPy.\label{alg:ifst}}
-		\Fn {\inverseChebTransform{$\{f\}_{j=0}^N$, points}}{
-        \Output{$\{y_k\}_{k=0}^{N}$}    
-        \If{points = "Gauss-Chebyshev"}{
-		    $\{y_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, \text{type=3}, 
+		\Fn {\inverseChebTransform{$\{f\}_{k=0}^N$, points}}{
+        \Output{$\{y_j\}_{j=0}^{N}$}    
+        \If{points = "Chebyshev-Gauss"}{
+		    $\{y_j\}_{j=0}^{N} \gets \text{dct}(\{f\}_{k=0}^N, \text{type=3}, 
 		    \text{axis=0})$ \;
-		    $\{y_k\}_{k=0}^{N} \gets \frac{1}{2}\{y_k\}_{k=0}^{N} + 
+		    $\{y_j\}_{j=0}^{N} \gets \frac{1}{2}\{y_j\}_{j=0}^{N} + 
 		    \frac{1}{2}f_0$ \;
 		}
-        \ElseIf{ points = "Gauss-Lobatto"}{
-	        $\{y_k\}_{k=0}^{N} \gets \text{dct}(\bm{f}, \text{type=1}, 
+        \ElseIf{ points = "Chebyshev-Gauss-Lobatto"}{
+	        $\{y_j\}_{j=0}^{N} \gets \text{dct}(\{f\}_{k=0}^N, \text{type=1}, 
 	        \text{axis=0})$ \;
-	        $\{y_k\}_{k=0}^{N} \gets \frac{1}{2}\{y_k\}_{k=0}^{N} + 
+	        $\{y_j\}_{j=0}^{N} \gets \frac{1}{2}\{y_k\}_{k=0}^{N} + 
 	        \frac{1}{2}f_0$ \;
-	        \For{$k=0$ \KwTo $N$}{
-	         $y_k \gets y_k + \frac{(-1)^k}{2} f_N$ \;
-    	    }
-	    }
-	    \Return{$\{y_k\}_{k=0}^{N}$}
-    }
-
-	    \Fn{\inverseShenTransform{$\{f_j\}_{j=0}^{N}$, points, space}}{
-		\Output{$\{y_k\}_{k=0}^{N}$}
+	        \For{$j=0$ \KwTo $N$}{
+	         $y_j \gets y_j + \frac{(-1)^j}{2} f_N$ \;
+    	    }
+	    }
+	    \Return{$\{y_j\}_{j=0}^{N}$}
+    }
+
+	    \Fn{\inverseShenTransform{$\{f_k\}_{k=0}^{N}$, points, space}}{
+		\Output{$\{y_j\}_{j=0}^{N}$}
         $\{z_k\}_{k=0}^{N} \gets 0$ \;
@@
         }
-        $\{y_k\}_{k=0}^{N} \gets $ inverseChebTransform($\{z_k\}_{k=0}^N$, 
-        points) 
+        $\{y_j\}_{j=0}^{N} \gets $ inverseChebTransform($\{z_k\}_{k=0}^N$, 
+        \emph{points}) 
         \;
-		\Return $\{y_k\}_{k=0}^{N}$ \;
+		\Return $\{y_j\}_{j=0}^{N}$ \;
     }
 \end{algorithm}
-
-\subsection*{Matrices}
-
-Matrices representing one-dimensional scalar products are given in Table \ref{tab:matrices}.
-Note that the mass matrices are symmetrical, and as such only the upper triangular parts are described.
-Also note that the matrices are computed using quadrature, and not exact integration, which has some minor implications for Gauss-Lobatto (GL) points.
-This follows since for an exact $L^2_w$ scalar product we would have $c_{N_x}=1$ for GL as well as Gauss Chebyshev (GC), but for the $l^2_w$ scalar product used here $c_{N_x}=2$.
-This disagreement, that follows from inexact quadrature at the highest mode, is not reported by Shen \cite{Shen95}, but is reported by, e.g., Kopriva \cite{kopriva09}.
-This explains the inclusion of the $c_{k+4}$ term for matrix $\N{B}$, and the $c_{k+2}$ term for matrix $\D{B}$, that are missing from Lemma 2.1 and 3.1 of \cite{Shen95}.
-
 \begin{table}
@@
 Note that $c_0=2$ and $c_k=1$ for $0<k<N_x$.
-For Gauss-Chebyshev points $c_{N_x}=1$, whereas $c_{N_x}=2$ for Gauss-Lobatto points.	\label{tab:matrices}}
+For Chebyshev-Gauss points $c_{N_x}=1$, whereas $c_{N_x}=2$ for Chebyshev-Gauss-Lobatto points.	\label{tab:matrices}}
 	\begin{tabular}{ccl}	
@@
 		\hline
-		$\{\D{A}_{kj}\}_{k,j=0}^{N_x-2}$ & $-\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w$ &
-$\begin{cases} 2\pi(k+1)(k+2), &j=k,\\
-4\pi(k+1), & j=k+2, k+4, k+6, \ldots, \\
-0, &\text{otherwise}.\end{cases}$		\\
-
-$\{\N{A}_{kj}\}_{k,j=1}^{N_x-2}$ & $-\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w$ & $\begin{cases} 2\pi j^2(k+1)/(k+2), &j=k,\\
-4\pi j^2(k+1)/(k+2)^2, & j=k+2, k+4, k+6, \ldots, \\
-0, &j>k, \text{ or } k+j \text{ odd} \end{cases}$ \\
+%		$\{\D{A}_{kj}\}_{k,j=0}^{N_x-2}$ & $-\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w$ &
+%$\begin{cases} 2\pi(k+1)(k+2), &j=k,\\
+%4\pi(k+1), & j=k+2, k+4, k+6, \ldots, \\
+%0, &\text{otherwise}.\end{cases}$		\\
+%
+%$\{\N{A}_{kj}\}_{k,j=1}^{N_x-2}$ & $-\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w$ & $\begin{cases} 2\pi j^2(k+1)/(k+2), &j=k,\\
+%4\pi j^2(k+1)/(k+2)^2, & j=k+2, k+4, k+6, \ldots, \\
+%0, &j>k, \text{ or } k+j \text{ odd.} \end{cases}$ \\
 
@@
 -\pi(k+1), & j=k+3, \\
-0, &\text{otherwise} \end{cases}$ \\
+0, &\text{otherwise.} \end{cases}$ \\
 
@@
 \pi (k-2)(k+1)/k, &j=k-3,\\
--2 \pi(k+1)^2/(k+2), & j=k-1, \\
+-2 \pi \frac{(k+1)^2}{k+2}, & j=k-1, \\
 \pi(k+1), & j=k+1, \\
-0, &\text{otherwise}
+0, &\text{otherwise.}
+\end{cases}$ \\
+
+$\{{C}_{kj}\}_{k,j=0}^{N_x, N_x-2}$ & $\left(\D{\phi}^{'}_j, T_k 
+\right)_w$ & $\begin{cases}
+-\pi (k+1), &j=k-1,\\
+-2 \pi, & j=k+1,k+3, \ldots, N_x-2 \\
+0, &\text{otherwise.}
+\end{cases}$ \\
+
+$\{\N{M}_{kj}\}_{k,j=0}^{N_x-4, N_x-2}$  & $ \left(\D{\phi}_j, \N{\phi}_k \right)_w $ & $\begin{cases}
+-\pi/2 & j=k-2, \\
+\frac{\pi}{2} \left(c_k+2 \frac{k+2}{k+3} \right) & j=k, \\
+-\frac{\pi}{2} \left(2\frac{k+2}{k+3} + c_{j+2} \frac{k+1}{k+3}\right) & j=k+2 \\
+\frac{\pi}{2}\frac{k+1}{k+3} & j=k+4, \\
+0, &\text{otherwise.}
 \end{cases}$ \\
@@
 \frac{\pi}{2} \frac{k+1}{k+3} , & j=k + 4, \\
-0, &\text{otherwise}
+0, &\text{otherwise.}
 \end{cases} $\\
@@
 -\frac{\pi}{2} &j=k + 2, \\
-0, &\text{otherwise}
+0, &\text{otherwise.}
 \end{cases}$ \\
 
-$\{B_{kj}\}_{k,j=0}^{N_x} = \{B_{jk}\}_{j,k=0}^{N_x}$ & $(T_k, T_j)_w$ & $\frac{c_k \pi}{2} \delta_{kj}$
+$\{B_{kj}\}_{k,j=0}^{N_x} = \{B_{jk}\}_{j,k=0}^{N_x}$ & $(T_k, T_j)_w$ & $\frac{c_k \pi}{2} \delta_{kj}.$
 

### 2017-01-13 paper.tex
Tikz figure
Submitted

@@
 
+\usepackage{tikz}
+\usepackage{pgfplots}
+
 %\newcommand{\mmo}[1]{\emph{#1}}
+\interfootnotelinepenalty=10000
+\newcommand*{\Scale}[2][4]{\scalebox{#1}{$#2$}}%
+\newcommand{\half}{\Scale[0.6]{\tfrac{1}{2}}}
+\newcommand{\nhalf}{\Scale[0.6]{-\tfrac{1}{2}}}
 
@@
 Comput., 16, 1, 1995], that lead to stable and robust numerical schemes, even at very large scale.
-New algorithms for direct solution of the linear systems are devised, and we show that the solver is accurate and applicable for very high Reynolds numbers.
+New and fast algorithms for the direct solution of the linear systems are devised, and algorithms and matrices for all required scalar products and transforms are provided.
+We validate the solver for very high Reynolds numbers.
 Specifically, the solver is shown to reproduce the first order statistics of Hoyas and Jim\'{e}nez [Phys.
-Fluids, 18(1), 2006], for a channel flow at $Re_{\tau}=2000$, which is higher than has been reported by a fully spectral method before.
+Fluids, 18(1), 2006], for a channel flow at $Re_{\tau}=2000$.
 The solver is available through the open source project spectralDNS [https://github.com/spectralDNS].
@@
 The solver by Jim\'{e}nez' group is reported to switch from Chebyshev to finite differences if the resolution is above a certain threshold \cite{hoyas08} (reached around $Re_{\tau}=1000$).
-In other words, they attempt to use fully spectral methods for as large $Re_{\tau}$ as possible.
+In other words, they attempt to use a fully spectral discretization for as large $Re_{\tau}$ as possible.
 As previously mentioned, spectral methods are attractive for their accuracy and resolution properties, that are superior to those of any finite difference or spline method.
@@
 In his seminal papers \cite{Shen94,Shen95}, Jie Shen describe how to construct Legendre and Chebyshev basis functions that lead to sparse matrices, susceptible to very fast direct solvers.
-To the authors knowledge, the bases have not been used for large-scale channel flow simulations, and algorithms for the required direct solvers have not, until now, been devised.
+To the author's knowledge, the bases have not been used for large-scale channel flow simulations, and algorithms for the required direct solvers have not, until now, been devised.
 Shen's bases have been used for the Navier Stokes equations before, though.
@@
 \cite{Bouchon01} describe a spectral-Galerkin formulation very similar to the one used in this paper.
-However, they choose the Legendre basis over Chebyshev, neglecting the higher cost of transforms and arguing instead for the simpler structure of the matrices and the lack of weight functions for the Legendre basis.
-The cost of the Legendre transforms, though, makes the basis unsuitable for larger scales, and the simulations performed in \cite{Bouchon01} were rather small.
+However, they choose the Legendre basis over Chebyshev, which has some consequences when aiming for large-scale, because fast transforms are required in moving from spectral to physical space, and back again.
+For Fourier and Chebyshev bases, the Fast Fourier Transforms (FFTs of $\mathcal{O} (N \log N)$) apply directly.
+However, until recently, the discrete Legendre transforms required $\mathcal{O}(N^2)$ operations.
+This scaling has now been improved by several authors, as recently reviewed by Hale and Townsend \cite{Hale14, Hale2015}, but the methods are still not quite on par with the FFTs.
+For example, Hale and Townsend describe an $\mathcal{O}(N (\log N)^2 / \log \log N)$ algorithm, using intermediate fast transforms from Legendre to Chebyshev coefficients.
 
 In this paper we will describe and assess a spectral-Galerkin channel flow solver based on Fourier and Shen's Chebyshev basis \cite{Shen95}.
-We will give a proper description of the theoretical basis in Sec~\ref{sec:prelim}, the discretization of Navier-Stokes equations in Sec~\ref{sec:discretizationNS} and we will describe necessary algorithms, including a new fast direct solver for the biharmonic problem that arise, in Sec~\ref{sec:implementation}.
-We will finally show in, Sec~\ref{sec:verification}, that roundoff is not a major issue, and that the Shen-Fourier spectral-Galerkin method is in deed applicable to large-scale simulations, larger than what has been reported for fully spectral channel simulations before.
-The solver has been implemented in the open source code spectralDNS \cite{spectralDNS}, where the bulk of the solver is written in high-level Python \cite{python}, with critical parts of the code migrated to Cython \cite{cython} for efficiency.
-
-\section{Preliminaries}
+The solver will consist of parts that scale at worst as $N \log N$, for a 1D problem of size $N$, and as such as $N^3 \log N$ for a 3D box of size $N^3$.
+We will give a proper description of the theoretical basis in Sec~\ref{sec:prelim}, the discretization of Navier-Stokes equations in Sec~\ref{sec:discretizationNS}, and we will describe necessary algorithms, including a new fast direct solver for the biharmonic problem that arise, in Sec~\ref{sec:implementation}.
+We will finally show, in Sec~\ref{sec:verification}, that roundoff is not a major issue, and that the Shen-Fourier spectral-Galerkin method is in deed applicable to large-scale simulations.
+The solver has been implemented in the open source code spectralDNS \cite{spectralDNS}, where the bulk of the code is written in high-level Python \cite{python}, with critical parts migrated to Cython \cite{cython} for efficiency.
+
+\section{Basis functions and fast transforms}
 \label{sec:prelim}
@@
  \frac{\partial \bm{u}}{\partial t}   &= \bm{\mathcal{H}} + \nu 
- \nabla^2 \bm{u} - \nabla{P}, \notag \\
+ \nabla^2 \bm{u} - \nabla{\tilde{p}}, \notag \\
  \nabla \cdot \bm{u} &= 0, \label{eq:NS}
@@
 \bm{\omega}$, where $\bm{\omega} = \nabla \times \bm{u}$.  The constant dynamic viscosity is denoted as $\nu$ 
-and $P(\bm{x}, t)$ is a pressure modified to account for both the driving force, $\beta(t)$, and the kinetic energy, i.e., $P = p + \bm{u} \cdot \bm{u}/2 + \beta y$, where $p$ is the instantaneous 
+and $\tilde{p}(\bm{x}, t)$ is a pressure modified to account for both the driving force, $\beta(t)$, and the kinetic energy, i.e., $\tilde{p} = p + \bm{u} \cdot \bm{u}/2 + \beta y$, where $p$ is the instantaneous
 pressure normalized by a constant density. The computational domain is 
@@
 
-The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall-normal direction.
-Three different sets of basis functions and function spaces are relevant for the wall-normal direction
-\begin{align}
-&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_k\}_{k=0}^{N_x}, \label{eq:Tk}\\
-& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_k\}_{k=0}^{N_x-2}, \label{eq:phiD}\\
-& \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
-\frac{k+1}{k+3} T_{k+4}(x), & \N{V}_{N_x} &= \text{span} \{\N{\phi}_k\}_{k=0}^{N_x-4}, \label{eq:phiN} 
-\end{align}
-where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind. The 
-basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were 
-suggested by Shen \cite{Shen95}, and satisfy, respectively, the boundary conditions 
-$\D{\phi}_k(\pm 1) = 0$, $\N{\phi}_k(\pm 1)=0$ and $\N{\phi}'_k(\pm 1)=0$. 
-%The last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1) = v'(\pm 1) = 0 \}$.
-
-Three-dimensional basis functions and function spaces, that are periodic in $y$ 
-and $z$ directions, can now be defined as
-\begin{align}
-  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
-  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
-  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
-\end{align}
-where $\imath=\sqrt{-1}$ and 
-\begin{align}
-K_N = \Big\{&\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / (l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
- &(l, m, n) \in  [0,1, \ldots, N_x] \times  [-\frac{N_y}{2},-\frac{N_y}{2}+1,\ldots, 
- \frac{N_y}{2}-1] \times [-\frac{N_z}{2},-\frac{N_z}{2},+1,\ldots, \frac{N_z}{2}-1] \Big\} 
- \label{eq:wavenumbermesh}
-\end{align}
-is the wavenumber mesh for $V_N$. The other two wavenumber meshes $\D{K}_N$ and 
-$\N{K}_N$ are both equal to $K_N$ except from in index $l$ that 
-ranges $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
-and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). We use 
-notation $K^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
-dimensions of $K_N$. Likewise, ${K}^x(l)={K}_N(l, 0, 0), \D{K}^x(l)=\D{K}_N(l, 0, 0)$ and $\N{K}^x(l)=\N{K}_N(l, 0, 0)$ refer to the wavenumbers, for the different bases, in the wall-normal direction.
-
-The computational mesh in real space can be created using $N = (N_x, N_y, 
-N_z)$ intervals, where the two periodic directions use uniform intervals. The 
-computational mesh is given as
-\begin{align}
-  X_N = \Big\{ \bm{x} \in \mathcal{R}^3/&(x_i, y_j, z_k) \in \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
+The domain $\Omega$ is discretized using $N = (N_x, N_y, N_z)$ intervals, where the two periodic directions use uniform intervals. The computational mesh is given as
+\begin{align}
+X_N = \Big\{ \bm{x} \in \mathbb{R}^3 | &(x_i, y_j, z_k) = \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
 &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\}, \label{eq:Xn}
@@
 
-For the Navier Stokes equations we look for solutions of the velocity 
+The spectral-Galerkin method makes use of a three-dimensional scalar product in the weighted $L^2_{\sigma}(\Omega)$ space, that is defined as
+\begin{align}
+\left<u, \upsilon \right>_{\sigma} &= \int_{\Omega} {u(\bm{x}) \upsilon^*(\bm{x})} \sigma(\bm{x})\,dxdydz, 
+\end{align}
+where $\upsilon^*$ is the complex conjugate of the test function $\upsilon$ and the weights $\sigma$ are unity for 
+periodic directions and  $\sigma(x)=1/\sqrt{1-x^2}$ for the inhomogeneous direction. In 
+this work we will make use of the discrete weighted $l^2_{\sigma}(\Omega)$ space, 
+where quadrature is employed for the integration. As such, we redefine the scalar product as
+\begin{equation}
+\left<u, \upsilon\right>_{\sigma} = \sum_{i=0}^{N_x}\sum_{j=0}^{N_y-1} \sum_{k=0}^{N_z-1} u(x_i, y_j, z_k) \upsilon^*(x_i, y_j, z_k) \sigma(x_i), \label{eq:quadrature}
+\end{equation}
+that is more amendable to the fast transforms that will be defined later in this section.
+
+The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall-normal direction.
+Three different sets of basis functions and function spaces are relevant for the wall-normal direction
+\begin{align}
+&  \phi_k(x) = T_k(x), & W_{N_x} &= \text{span}\{\phi_k\}_{k=0}^{N_x}, \label{eq:Tk}\\
+& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{W}_{N_x} &= \text{span} \{ \D{\phi}_k\}_{k=0}^{N_x-2}, \label{eq:phiD}\\
+& \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
+\frac{k+1}{k+3} T_{k+4}(x), & \N{W}_{N_x} &= \text{span} \{\N{\phi}_k\}_{k=0}^{N_x-4}, \label{eq:phiN} 
+\end{align}
+where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind. The 
+basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were 
+suggested by Shen \cite{Shen95}, and satisfy, respectively, the boundary conditions 
+$\D{\phi}_k(\pm 1) = 0$, $\N{\phi}_k(\pm 1)=0$ and $\N{\phi}'_k(\pm 1)=0$. 
+%The last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1) = v'(\pm 1) = 0 \}$.
+
+Three-dimensional basis functions and function spaces, that are periodic in $y$ 
+and $z$ directions, can now be defined as
+\begin{align}
+  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
+  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
+  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
+\end{align}
+where $\imath=\sqrt{-1}$. 
+%\begin{align}
+%K_N = \Big\{&\bm{k} \in \mathbb{Z}\times\mathbb{R}^2 / (l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
+% &(l, m, n) \in  [0,1, \ldots, N_x] \times  [-\frac{N_y}{2},-\frac{N_y}{2}+1,\ldots, 
+% \frac{N_y}{2}-1] \times [-\frac{N_z}{2},-\frac{N_z}{2},+1,\ldots, \frac{N_z}{2}-1] \Big\} 
+% \label{eq:wavenumbermesh}
+%\end{align}
+%is the wavenumber mesh for $V_N$. The other two wavenumber meshes $\D{K}_N$ and 
+%$\N{K}_N$ are both equal to $K_N$ except from in index $l$ that 
+%ranges $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
+%and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). We use 
+%notation $K^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
+%dimensions of $K_N$.
+Likewise, ${K}^x(l)={K}_N(l, 0, 0), \D{K}^x(l)=\D{K}_N(l, 0, 0)$ and $\N{K}^x(l)=\N{K}_N(l, 0, 0)$ refer to the wavenumbers, for the different bases, in the wall-normal direction.
+The wavenumbermesh, $K_N$, for space $V_{N}$, is defined by the Cartesian product of wavenumbers from the two periodic and the inhomogeneous wall-normal direction: $K_N(l,m,n)= K^x(l) \times K^p(m,n)$, where
+\begin{align}
+K^p = \Big\{&(\underline{m}, \underline{n}) = \left(\frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
+&(m, n) \in  [-\frac{N_y}{2},-\frac{N_y}{2}+1,\ldots, 
+\frac{N_y}{2}-1] \times [-\frac{N_z}{2},-\frac{N_z}{2},+1,\ldots, \frac{N_z}{2}-1] \Big\} 
+\label{eq:wavenumbermeshp}
+\end{align}
+and $K^x(l) = \{l \in \mathbb{Z} | l=0,1, \ldots, N_x\}$.
+The two remaining wavenumber meshes, $\D{K}_{N}$ and $\N{K}_{N}$, differ from $K_N$ only in the range of the first index sets, $\D{K}^x$ and $\N{K}^x$, ending in $N_x-2$ and $N_x-4$, respectively (see Eqs.
+(\ref{eq:phiD}) and (\ref{eq:phiN})).
+
+In the spectral-Galerkin method we look for solutions of the velocity 
 components of the form
@@
 convenience 
-and compliance with the definition used for the inverse discrete Fourier 
+and compliance with the definition used later for the inverse discrete Fourier 
 transform. Note that from now on we will simply use the notation $\hat{u}$ for 
@@
 
-If the expansion coefficients $\hat{u}$  are known for the entire wavenumber 
-mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_solx}) may be 
-evaluated on the mesh (\ref{eq:Xn}) using fast inverse transforms as
+For an efficient method it is crucial to be able to compute $u$ quickly from $\hat{u}$, or, vice versa, to compute $\hat{u}$ quickly from $u$.
+This may be achieved using the fast transform methods to be defined next.
+Consider first how to compute $u$ from the known expansion coefficients $\hat{u}$. Writing out the summation terms, the expression (\ref{eq:u_solx}) may be 
+evaluated on the entire mesh (\ref{eq:Xn}) as
 \begin{align}
@@
   u &= \N{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y 
-  (\N{\mathcal{S}}^{-1}_x (\hat{u}))) \quad \forall \, \bm{x} \in X_N.  
+  (\N{\mathcal{S}}^{-1}_x (\hat{u}))),
   \label{eq:ifft} 
 \end{align}
-Here $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
-transform, and $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represent inverse discrete
-Fourier transforms along directions $y$ and $z$ respectively. For simplicity, we introduce here a special notation called the inverse 
-\emph{Shen} transform, $\N{\mathcal{S}}^{-1}_x$, that is used to transform coefficients from spectral to physical space in a series expansion that is using either one of the bases in (\ref{eq:Tk}, \ref{eq:phiD}, \ref{eq:phiN}).
+where $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
+transform in space $\N{V}_N$, and $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represent inverse discrete
+Fourier transforms along directions $y$ and $z$ respectively. For simplicity, we have introduce here a special notation called the inverse 
+\emph{Shen} transform, ${\mathcal{S}}^{-1}_x$, that is used to transform coefficients from spectral to physical space in a series expansion that is using either one of the bases in (\ref{eq:Tk}, \ref{eq:phiD}, \ref{eq:phiN}).
 The inverse Shen transform, $\mathcal{S}^{-1}_x$, is performed along the wall-normal
@@
 to distinguish between them with obvious notation.
-
-A three-dimensional scalar product in the weighted $L^2_w(\Omega)$ space is defined as
-\begin{align}
- \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, 
-\end{align}
-where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for 
-periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction. In 
-this work we will make use of the discrete weighted $l^2_w(\Omega)$ space, 
-where quadrature is employed for the integration. As such, the scalar product 
-corresponds to sums that may be computed using fast transforms. Using 
-$\N{\psi}_{\bm{k}}$ for $v$ we redefine the scalar product with quadrature as
-\begin{align}
- \left< u, \N{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i 
+Similarly, $\D{\mathcal{T}}^{-1}$ and ${\mathcal{T}}^{-1}$ define the inverse transforms for spaces $\D{V}_N$ and $V_N$.
+Note that a transform in  any one direction is performed over all indices in the other two directions.
+For example, for the transform in the $x$-direction we have
+\begin{equation}
+u(x_i,m,n,t) = \N{\mathcal{S}}_x^{-1}(\hat{u}(l,m,n,t)) \quad \forall \, i \in [0,1,\ldots, N_x] \text{ and }  m,n \in {K}^p,
+\end{equation}
+and similar for the other two directions.
+
+Fast transforms may also be used in the scalar product defined in Eq. (\ref{eq:quadrature}), using basis $\N{\psi}_{\bm{k}}$ for $\upsilon$
+\begin{align}
+ \left< u, \N{\psi}_{\bm{k}} \right>_{\sigma} &= h \underbrace{\sum_i 
  \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath 
  \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} 
- }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) w_i}_{\N{\mathcal{S}}_l},   \notag \\
+ }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) \sigma(x_i)}_{\N{\mathcal{S}}_l} \quad \forall \, \bm{k} \in \N{K}_N,   \notag \\
   &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
-  (\mathcal{F}_{{n}}(u))) \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
-\end{align}
-Here $h\N{\mathcal{S}}$ denotes the complete three-dimensional scalar product 
+  (\mathcal{F}_{{n}}(u))). \label{eq:sst1}
+\end{align}
+Here $h\N{\mathcal{S}}(\cdot)$ denotes the complete three-dimensional scalar product 
 and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
@@
 $y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
-\N{\phi}_l)_w$ is used to represent the forward \emph{Shen} scalar product in the 
-$x$-direction. The weights, $w_i$, required for the Shen transforms are 
-defined as
+\N{\phi}_l)_{\sigma}$ is used to represent the forward \emph{Shen} scalar product in the 
+$x$-direction. The weights, $\sigma(x_i)$, required for the Shen transforms, are defined as
 \begin{equation}
- w_i = \begin{cases}
+ \sigma(x_i) = \begin{cases}
        \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
@@
 where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$, see Sec.
-1.11 of \cite{kopriva09}.%\footnote{Note that $c_{N_x}=1$ for the continuous scalar product $L^2_w(\Omega)$ (see, e.g., Eq.
-(2.7) of \cite{Shen95}), but for the discrete $l^2_w(\Omega)$ it follows that $c_{N_x}=2$ due to inexact quadrature for the highest wavenumber.}
+1.11 of \cite{kopriva09}.%\footnote{Note that $c_{N_x}=1$ for the continuous scalar product $L^2_{\sigma}(\Omega)$ (see, e.g., Eq.
+(2.7) of \cite{Shen95}), but for the discrete $l^2_{\sigma}(\Omega)$ it follows that $c_{N_x}=2$ due to inexact quadrature for the highest wavenumber.}
 
 The scalar product  $h\N{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a 
-complete transform. To find a transformation between physical $u$ and spectral 
+complete transform. To find a transformation from physical $u$ to spectral 
 $\hat{u}$, we make use of Eq. (\ref{eq:u_solx}) directly on the left hand side of (\ref{eq:sst1}) and use extensively the discrete 
@@
 \begin{align}
-%\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
+%\left<u, \N{\psi}_{\bm{k}}\right>_{\sigma} &= \frac{1}{N_yN_z}\left< 
 %\sum_{\bm{m}=(q,r,s) \in \N{K}_N} \hat{u}_{\bm{m}} \N{\psi}_{\bm{m}}, 
-%\N{\psi}_{\bm{k}} \right>_w, \notag \\
-\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
+%\N{\psi}_{\bm{k}} \right>_{\sigma}, \notag \\
+\left<u, \N{\psi}_{\bm{k}}\right>_{\sigma} &= \frac{1}{N_yN_z}\left< 
 \sum_{(q,r,s) \in \N{K}_N} \hat{u}(q,r,s,t) \N{\psi}(q,r,s), 
-\N{\psi}(l,m,n) \right>_w, \notag \\
+\N{\psi}(l,m,n) \right>_{\sigma} \quad \forall \, \bm{k} \in \N{K}_N, \notag \\
            &= h \sum_{q=0}^{N_x-4} \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) 
-           \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
-           &= h \sum_{q=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t) \quad \forall \, \bm{k} \in \N{K}_N. 
+           \N{\phi}_{l}(x_i) \sigma(x_i) \, \hat{u}(q, {m}, {n}, t), \notag \\
+           &= h \sum_{q=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t). 
            \label{eq:sst2}
 \end{align}
-Here $\{\N{B}_{lq}\}_{l,q=0}^{N_x-4} = (\N{\phi}_q, \N{\phi}_l)_w = 
-\sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
+Here $\N{B}_{lq} = (\N{\phi}_q, \N{\phi}_l)_{\sigma} = 
+\sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) \sigma(x_i)$ are the components of a 
 banded mass matrix with only 5 nonzero diagonals, see Table~\ref{tab:matrices}. The complete transformation is now obtained by 
@@
 h\sum_{q=0}^{N_x-4} \N{B}_{lq}\, \hat{u}(q, {m}, {n}, t) &= 
-h\N{\mathcal{S}}(u)(l, m, n, t), \notag \\
- \hat{u} &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u) 
- \quad \forall \, \bm{k} \in \N{K}_N, \label{eq:fstB}
+h\N{\mathcal{S}}(u)(l, m, n, t)  \quad \forall \, \bm{k} \in \N{K}_N, \notag \\
+ \hat{u} &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), \label{eq:fstB}
 \end{align}
@@
 transformation thus requires a fast Chebyshev transform ($\mathcal{O}(N_x \log N_x)$) and 
-a fast linear algebraic solve ($\mathcal{O}(N_x)$). Similar transforms $\mathcal{T}$ 
+a fast linear algebraic solve ($\mathcal{O}(N_x)$) for the wall-normal direction, given $m$ and $n$. Similar transforms $\mathcal{T}$ 
 and $\D{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and 
-(\ref{eq:phiD}), using mass matrices $B_{lq}=(T_q, T_l)_w$ (diagonal) and 
-$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_w$ (three non-zero diagonals) and scalar products 
-$\mathcal{S}_l(\cdot) = (\cdot, T_l)_w$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
-\D{\phi}_l)_w$. See Alg.~\ref{alg:fst} and \ref{alg:ifst} in the Appendix for algorithms for all required transforms.
+(\ref{eq:phiD}), using mass matrices with components $B_{lq}=(T_q, T_l)_{\sigma}$ (diagonal) and 
+$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_{\sigma}$ (three non-zero diagonals) and scalar products 
+$\mathcal{S}_l(\cdot) = (\cdot, T_l)_{\sigma}$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
+\D{\phi}_l)_{\sigma}$. See Alg.~\ref{alg:fst} and \ref{alg:ifst} in the Appendix for algorithms for all required transforms.
 
@@
 from the continuity equation. The boundary conditions for Eq. (\ref{eq:g}) are 
-$g(\pm 1) = 0$. The variational formulation becomes: Find ${u} \in 
-\N{V}_N$, ${g} \in \D{V}_N$ and ${f} \in \D{V}_N$ such that
-\begin{align}
-	\frac{\partial }{\partial t} \left< \nabla^2 u, \N{\psi}_{\bm{k}}\right>_w &= 
-	\left<h_u, \N{\psi}_{\bm{k}} \right>_w + \nu \left<\nabla^4u, \N{\psi}_{\bm{k}}\right>_w 
-	&\forall \,\N{\psi}_{\bm{k}} \in \N{V}_N, \label{eq:u1} \\
-	\frac{\partial}{\partial t}\left<g, \D{\psi}_{\bm{k}}\right>_w &= \left<h_g, 
-	\D{\psi}_{\bm{k}}\right>_w + \nu 
-	\left<\nabla^2 g, \D{\psi}_{\bm{k}}\right>_w &\forall \, \D{\psi}_{\bm{k}} \in \D{V}_N ,
+$g(\pm 1) = 0$.
+
+We consider the spectral-Galerkin discretization of (\ref{eq:biharmonic}), (\ref{eq:g}) and (\ref{eq:f}), using a central difference in time, with Crank-Nicolson for the linear terms and an Adams-Bashforth method for the nonlinear terms.
+To this end, the discretization in time is performed with a constant time step $\triangle t>0$,
+such that time is represented discretely as $t_{\kappa} = {\kappa} \triangle t, \, {\kappa} \in [0, 1, \ldots]$, and variables with a time superscript, like $u^{\kappa} = u(t_{\kappa})$.
+We get the variational formulation: with $u^0(\bm{x}), u^1(\bm{x}), g^0(\bm{x})$ and $g^1(\bm{x})$ given, for all $\kappa > 1$ find ${u}^{\kappa+1} \in \N{V}_N$, ${g}^{\kappa+1} \in \D{V}_N$ and ${f}^{\kappa+1} \in \D{V}_N$ such that
+\begin{align}
+  \left<  \frac{\nabla^2(u^{\kappa+1}-u^{\kappa})}{\triangle t}, \N{\psi}_{\bm{k}}\right>_{\sigma} &= 
+\left<h_u^{\kappa+\half}, \N{\psi}_{\bm{k}} \right>_{\sigma} + {\nu} \left<\nabla^4 u^{\kappa+\half}, \N{\psi}_{\bm{k}}\right>_{\sigma} 
+&\forall \,\psi_{\bm{k}} \in \N{V}_N, \label{eq:u1} \\
+  \left<\frac{g^{\kappa+1}- g^{\kappa}}{\triangle t}, \D{\psi}_{\bm{k}}\right>_{\sigma} &= \left<h_g^{\kappa + \half}, 
+\D{\psi}_{\bm{k}}\right>_{\sigma} + {\nu} 
+\left<\nabla^2 g^{\kappa+\half}, \D{\psi}_{\bm{k}}\right>_{\sigma} &\forall \, \psi_{\bm{k}} \in \D{V}_N ,
 \label{eq:g1} \\
-	\left<f, \D{\psi}_{\bm{k}}\right>_w &= \left<\frac{\partial u}{\partial x}, 
-	\D{\psi}_{\bm{k}}\right>_w &\forall \, \D{\psi}_{\bm{k}} \in \D{V}_N. \label{eq:f1}
-\end{align}
-The two remaining velocity components are computed by projection to the 
+	\left<f^{\kappa+1}, \D{\psi}_{\bm{k}}\right>_{\sigma} &= \left<\frac{\partial u^{\kappa+1}}{\partial x}, 
+	\D{\psi}_{\bm{k}}\right>_{\sigma} &\forall \, \psi_{\bm{k}} \in \D{V}_N. \label{eq:f1}
+\end{align}
+Here superscript $\kappa + \half$ is used to represent Crank-Nicolson for linear terms (e.g., $u^{\kappa+\half}=0.5(u^{\kappa+1}+u^{\kappa})$) and Adams-Bashforth for nonlinear (e.g., $h_u^{\kappa + \half} = 1.5h_u^{\kappa} - 0.5 h_u^{\kappa-1}$).
+
+With ${f}^{\kappa+1}$ and ${g}^{\kappa+1}$ known, the two remaining velocity components are then computed by projection to the 
 Dirichlet space $\D{V}_N$: Find 
-${v} \in \D{V}_N$ and $w \in \D{V}_N$ such that
-\begin{align}
-\left<f, \D{\psi}_{\bm{k}}\right>_w &= \left<\frac{\partial v}{\partial y} + 
-\frac{\partial w}{\partial z}, \D{\psi}_{\bm{k}}\right>_w \, &\forall \, \D{\psi}_{\bm{k}} \in 
+${v}^{\kappa+1} \in \D{V}_N$ and $w^{\kappa+1} \in \D{V}_N$ such that
+\begin{align}
+\left<f^{\kappa+1}, \D{\psi}_{\bm{k}}\right>_{\sigma} &= \left<\frac{\partial v^{\kappa+1}}{\partial y} + 
+\frac{\partial w^{\kappa+1}}{\partial z}, \D{\psi}_{\bm{k}}\right>_{\sigma} \, &\forall \, \D{\psi}_{\bm{k}} \in 
 \D{V}_N, \label{eq:f2} \\
-\left<g, \D{\psi}_{\bm{k}}\right>_w &= \left<\frac{\partial w}{\partial y}  - 
-\frac{\partial v}{\partial z}, \D{\psi}_{\bm{k}}\right>_w \, &\forall \, \D{\psi}_{\bm{k}} \in 
+\left<g^{\kappa+1}, \D{\psi}_{\bm{k}}\right>_{\sigma} &= \left<\frac{\partial w^{\kappa+1}}{\partial y}  - 
+\frac{\partial v^{\kappa+1}}{\partial z}, \D{\psi}_{\bm{k}}\right>_{\sigma} \, &\forall \, \D{\psi}_{\bm{k}} \in 
 \D{V}_N, \label{eq:g2}
@@
 directions. Written in spectral space Eqs. (\ref{eq:f2}) and (\ref{eq:g2}) 
-become simply
-\begin{align}
-\hat{f}_{\bm{k}} &= \imath \underline{m}\, \hat{v}_{\bm{k}} + \imath 
-\underline{n}\, \hat{w}_{\bm{k}} &\forall \, \bm{k} \in \D{K}_N^0, \label{eq:f3} \\
-\hat{g}_{\bm{k}} &= \imath \underline{m}\, \hat{w}_{\bm{k}} - \imath 
-\underline{n}\, \hat{v}_{\bm{k}} & \forall \, \bm{k} \in \D{K}_N^0, \label{eq:g3}
+become simply the algebraic expressions
+\begin{align}
+\hat{f}_{\bm{k}}^{\kappa+1} &= \imath \underline{m}\, \hat{v}_{\bm{k}}^{\kappa+1} + \imath 
+\underline{n}\, \hat{w}_{\bm{k}}^{\kappa+1} &\forall \, \bm{k} \in \D{K}_N^0, \label{eq:f3} \\
+\hat{g}_{\bm{k}}^{\kappa+1} &= \imath \underline{m}\, \hat{w}_{\bm{k}}^{\kappa+1} - \imath 
+\underline{n}\, \hat{v}_{\bm{k}}^{\kappa+1} & \forall \, \bm{k} \in \D{K}_N^0, \label{eq:g3}
 \end{align}
 where $\D{K}_N^0$ is used to denote that these equations can be solved for all 
-wavenumbers except $m=n=0$. For $m=n=0$ we solve instead the 
-corresponding transformed momentum equations in $y$ and $z$ directions:
-\begin{align}
-\frac{\partial }{\partial t} \D{B}_{lq}\hat{v}(q, 0, 0, t) &= 
-\D{B}_{lq}\hat{\mathcal{H}}_y(q, 0, 0, t) - \nu \D{A}_{lq} \hat{v}(q, 0, 0, t) - \left(\beta(t), \D{\phi}_l \right)_w
-& \forall\, l \in \D{K}^x, \label{eq:v0}\\
-\frac{\partial }{\partial t} \D{B}_{lq}\hat{w}(q, 0, 0, t) &= 
-\D{B}_{lq}\hat{\mathcal{H}}_z(q, 0, 0, t) - \nu \D{A}_{lq} \hat{w}(q, 0, 0, t) 
-& \forall\, l \in \D{K}^x,\label{eq:w0}
-\end{align}
-where $\D{A}_{lq}=-(\D{\phi}_q^{''}, \D{\phi}_l )_w$.
-Note that (\ref{eq:v0}) is the only place where the driving force, or the mean pressure gradient, $\beta$, enters, and, since it is constant in space, the term $(\beta(t), \D{\phi}_l )_w$ will only be non-zero for $l=0$.
-
-Equations (\ref{eq:u1}-\ref{eq:f1}) need to be rewritten in terms of one-dimensional scalar products. 
-Inserting the expansion (\ref{eq:u_solx}) for $u$ and a similar expansion for $g$ in $\D{V}_N$, the inner products required to solve Eqs. (\ref{eq:u1}-\ref{eq:f1}) are
-\begin{align}
-\left<\nabla^4u, \N{\psi}_{\bm{k}}\right>_w &= h\left[ \left( 
+wavenumbers except $m=n=0$. For $m=n=0$ we solve instead the momentum equations in $y$ and $z$ directions (see Eq. (\ref{eq:NS})):
+\begin{align}
+\left< \frac{v^{\kappa+1} - v^{\kappa}}{\triangle t}, \psi_{\bm{k}} \right>_{\sigma} &= \left< \mathcal{H}_{y}^{\kappa+\half}, \psi_{\bm{k}} \right>_{\sigma} + \nu \left< \nabla^2 v^{\kappa+\half},  \psi_{\bm{k}} \right>_{\sigma} -\left< \beta^{\kappa+1}, \psi_{\bm{k}} \right>_{\sigma} & \forall \, \bm{k} \in \D{K}^x \times K^p(0,0), \label{eq:v0} \\
+\left< \frac{w^{\kappa+1} - w^{\kappa}}{\triangle t}, \psi_{\bm{k}} \right>_{\sigma} &= \left< \mathcal{H}_{z}^{\kappa+\half}, \psi_{\bm{k}} \right>_{\sigma} + \nu \left< \nabla^2 w^{\kappa+\half},  \psi_{\bm{k}} \right>_{\sigma}  & \forall \, \bm{k} \in \D{K}^x \times K^p(0,0) \label{eq:w0}.
+\end{align}
+%\begin{align}
+%\frac{1}{\triangle t}\D{B}_{lq}(\hat{v}^{\kappa+1}_{q} - \hat{v}^{\kappa}_{q}) &= 
+%\D{B}_{lq}\hat{\mathcal{H}}_{y,q}^{\kappa+1/2} - \nu \D{A}_{lq} \hat{v}^{\kappa+1/2}_q - \left(\beta^{\kappa+1}, \D{\phi}_l \right)_{\sigma}
+%& \forall\, l \in \D{K}^x, m=n=0, \label{eq:v0}\\
+%\frac{1}{\triangle t}\D{B}_{lq}(\hat{w}^{\kappa+1}_{q} - \hat{w}^{\kappa}_{q}) &= 
+%\D{B}_{lq}\hat{\mathcal{H}}_{z,q}^{\kappa+1/2} - \nu \D{A}_{lq} \hat{w}^{\kappa+1/2}_q 
+%& \forall\, l \in \D{K}^x, m=n=0 \label{eq:w0}
+%\end{align}
+Note that the regular pressure is eliminated since $m=n=0$, and that (\ref{eq:v0}) is the only place where the driving force, or the mean pressure gradient, $\beta$, enters the equations.
+Also, since $\beta$ is constant in space, the term $<\beta^{\kappa+1}, \D{\psi}_{\bm{k}} >_{\sigma}$ will only be non-zero for $l=m=n=0$.
+
+The final step of the method is to rewrite all equations on matrix form, using one-dimensional scalar products.
+Inserting the expansion (\ref{eq:u_solx}) for $u$ in $\N{V}_N$, and similar for $g, h_u$ and $h_g$ in $\D{V}_N$, the inner products required to solve Eqs.
+(\ref{eq:u1}-\ref{eq:f1}) are
+\begin{align}
+\left<\nabla^4u, \N{\psi}_{\bm{k}}\right>_{\sigma} &= h\left[ \left( 
 \N{\phi}_q^{''''}, 
-\N{\phi}_l\right)_w -2(\underline{m}^2+\underline{n}^2) \left( \N{\phi}_q^{''}, 
-\N{\phi}_l\right)_w + (\underline{m}^2+\underline{n}^2)^2\left( \N{\phi}_q, 
-\N{\phi}_l\right)_w  \right] \hat{u}_q, \\
-\left< \nabla^2 u, \N{\psi}_{\bm{k}}\right>_w &= h\left[\left( \N{\phi}_q^{''}, 
-\N{\phi}_l\right)_w - (\underline{m}^2+\underline{n}^2)\left( \N{\phi}_q, 
-\N{\phi}_l \right)_w \right] \hat{u}_q, \\
-\left< \nabla^2 g, \D{\psi}_{\bm{k}}\right>_w &= h\left[\left( \D{\phi}_q^{''}, 
-\D{\phi}_l\right)_w - (\underline{m}^2+\underline{n}^2)\left( \D{\phi}_q, 
-\D{\phi}_l \right)_w \right] \hat{g}_q, \\
-\left<\frac{\partial u}{\partial x}, \D{\psi}_{\bm{k}}\right>_w &=
-h\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q, \\
-\left<h_u, \N{\psi}_{\bm{k}} \right>_w &= 
+\N{\phi}_l\right)_{\sigma} -2(\underline{m}^2+\underline{n}^2) \left( \N{\phi}_q^{''}, 
+\N{\phi}_l\right)_{\sigma} + (\underline{m}^2+\underline{n}^2)^2\left( \N{\phi}_q, 
+\N{\phi}_l\right)_{\sigma}  \right] \hat{u}_q, \\
+\left< \nabla^2 u, \N{\psi}_{\bm{k}}\right>_{\sigma} &= h\left[\left( \N{\phi}_q^{''}, 
+\N{\phi}_l\right)_{\sigma} - (\underline{m}^2+\underline{n}^2)\left( \N{\phi}_q, 
+\N{\phi}_l \right)_{\sigma} \right] \hat{u}_q, \\
+\left< \nabla^2 g, \D{\psi}_{\bm{k}}\right>_{\sigma} &= h\left[\left( \D{\phi}_q^{''}, 
+\D{\phi}_l\right)_{\sigma} - (\underline{m}^2+\underline{n}^2)\left( \D{\phi}_q, 
+\D{\phi}_l \right)_{\sigma} \right] \hat{g}_q, \\
+\left<\frac{\partial u}{\partial x}, \D{\psi}_{\bm{k}}\right>_{\sigma} &=
+h\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_{\sigma} \hat{u}_q, \\
+\left<h_u, \N{\psi}_{\bm{k}} \right>_{\sigma} &= 
 h\left[-(\underline{m}^2+\underline{n}^2) 
 \left(\D{\phi}_q, 
-\N{\phi}_l \right)_w \hat{\mathcal{H}}_{x, q} - \imath \underline{m}\left(\D{\phi}_q^{'}, 
-\N{\phi}_l \right)_w \hat{\mathcal{H}}_{y, q} - \imath 
-\underline{n}\left(\D{\phi}_q^{'}, \N{\phi}_l \right)_w 
+\N{\phi}_l \right)_{\sigma} \hat{\mathcal{H}}_{x, q} - \imath \underline{m}\left(\D{\phi}_q^{'}, 
+\N{\phi}_l \right)_{\sigma} \hat{\mathcal{H}}_{y, q} - \imath 
+\underline{n}\left(\D{\phi}_q^{'}, \N{\phi}_l \right)_{\sigma} 
 \hat{\mathcal{H}}_{z, q}\right], \label{eq:S_hv}
 \\
-\left< h_g, \D{\psi}_{\bm{k}} \right>_w &= h\left[ \imath \underline{m}\, \left(\D{\phi}_q, \D{\phi}_l \right)_w 
-\hat{\mathcal{H}}_{z, q} - \imath \underline{n}\, \left(\D{\phi}_q, \D{\phi}_l \right)_w \hat{\mathcal{H}}_{y, q} 
+\left< h_g, \D{\psi}_{\bm{k}} \right>_{\sigma} &= h\left[ \imath \underline{m}\, \left(\D{\phi}_q, \D{\phi}_l \right)_{\sigma} 
+\hat{\mathcal{H}}_{z, q} - \imath \underline{n}\, \left(\D{\phi}_q, \D{\phi}_l \right)_{\sigma} \hat{\mathcal{H}}_{y, q} 
 \right],\label{eq:S_hg}
 \end{align}
-where for brevity in notation (as before) it is understood that the matrices 
+where for brevity in notation (as before) it is understood that the scalar products
 act along the first dimension of the transformed variables, i.e.,  $ 
-(\N{\phi}_q^{'}, \D{\phi}_l)_w \hat{u}_q$ is short for the matrix 
-vector product $ \sum_{q=0}^{N_x-4}( \N{\phi}_q^{'}, \D{\phi}_l)_w 
-\hat{u}(q, {m}, {n}, t)$, where $m$ and $n$ are given. The inner products are used to set up linear systems 
+(\N{\phi}_q^{'}, \D{\phi}_l)_{\sigma} \hat{u}_q$ is short for the matrix 
+vector product $ \sum_{q=0}^{N_x-4}( \N{\phi}_q^{'}, \D{\phi}_l)_{\sigma} 
+\hat{u}(q, {m}, {n}, t)$, for all $m$ and $n$. The scalar products are used to set up linear systems 
 of equations for the inhomogeneous wall-normal direction. All scalar products 
-$(\cdot, \cdot)_w$ can be represented by sparse matrices.
+$(\cdot, \cdot)_{\sigma}$ can be represented by sparse matrices.
 %\begin{align}
-%\frac{\partial }{\partial t} \left( \N{\psi}_j^{''}, \N{\psi}_k\right)_w 
-%\hat{u}_j &=  \left(h_u, \N{\psi}_k \right)_w + \nu \left(\N{\psi}^{''''}_j, 
-%\N{\psi}_k\right)_w \hat{u}_j, \, &\forall \N{\psi}_k \in \N{V}_N, \\
-%\frac{\partial}{\partial t}\left(\D{\psi}_j, \D{\psi}_k\right)_w \hat{g}_j &= 
-%\left(h_g, \D{\psi}_k\right)_w + \nu \left(\D{\psi}_j^{''}, 
-%\D{\psi}_k\right)_w 
+%\frac{\partial }{\partial t} \left( \N{\psi}_j^{''}, \N{\psi}_k\right)_{\sigma} 
+%\hat{u}_j &=  \left(h_u, \N{\psi}_k \right)_{\sigma} + \nu \left(\N{\psi}^{''''}_j, 
+%\N{\psi}_k\right)_{\sigma} \hat{u}_j, \, &\forall \N{\psi}_k \in \N{V}_N, \\
+%\frac{\partial}{\partial t}\left(\D{\psi}_j, \D{\psi}_k\right)_{\sigma} \hat{g}_j &= 
+%\left(h_g, \D{\psi}_k\right)_{\sigma} + \nu \left(\D{\psi}_j^{''}, 
+%\D{\psi}_k\right)_{\sigma} 
 %\hat{g}_j, \, &\forall \D{\psi}_k \in \D{V}_N \\
-%\left(\D{\psi}_j, \D{\psi}_k\right)_w \hat{f}_j &= \left(\N{\psi}_j^{'}, 
-%\D{\psi}_k\right)_w \hat{u}_j, \, &\forall \D{\psi}_k \in \D{V}_N.
+%\left(\D{\psi}_j, \D{\psi}_k\right)_{\sigma} \hat{f}_j &= \left(\N{\psi}_j^{'}, 
+%\D{\psi}_k\right)_{\sigma} \hat{u}_j, \, &\forall \D{\psi}_k \in \D{V}_N.
 %\end{align}
-The required matrices $ \N{B}_{lq} = ( \N{\phi}_q, 
-\N{\phi}_l)_w, \D{B}_{lq} = ( \D{\phi}_q, 
-\D{\phi}_l)_w 
-, \N{M}_{lq} = ( \D{\phi}_q, 
-\N{\phi}_l)_w, \D{C}_{lq} = (\N{\phi}_q^{'}, \D{\phi}_l)_w$ and 
-$ \N{C}_{lq} = (\D{\phi}_q^{'}, \N{\phi}_l)_w$ are given in Table~\ref{tab:matrices}, $\N{A}_{lq} = -( \N{\phi}_q^{''}, 
-\N{\phi}_l)_w, \D{A}_{lq} = -( \D{\phi}_q^{''}, \D{\phi}_l)_w$ are given in \cite{Shen95}, whereas the matrix $\N{Q}_{lq} = (\N{\phi}^{''''}_q, \N{\phi}_l)_w$ is given below in Eqs (\ref{eq:N_Q})-(\ref{eq:rk}).
+The required matrices with components $ \N{B}_{lq} = ( \N{\phi}_q, 
+\N{\phi}_l)_{\sigma}, \D{B}_{lq} = ( \D{\phi}_q, 
+\D{\phi}_l)_{\sigma}, \N{M}_{lq} = ( \D{\phi}_q, 
+\N{\phi}_l)_{\sigma}, \D{C}_{lq} = (\N{\phi}_q^{'}, \D{\phi}_l)_{\sigma}$ and 
+$ \N{C}_{lq} = (\D{\phi}_q^{'}, \N{\phi}_l)_{\sigma}$ are given in Table~\ref{tab:matrices}, $\N{A}_{lq}= -( \N{\phi}_q^{''}, 
+\N{\phi}_l)_{\sigma}, \D{A}_{lq}= -( \D{\phi}_q^{''}, \D{\phi}_l)_{\sigma}$ are given in \cite{Shen95}, whereas $\N{Q}_{lq} = (\N{\phi}^{''''}_q, \N{\phi}_l)_{\sigma}$ is given below in Eqs (\ref{eq:N_Q})-(\ref{eq:rk}).
 Note that the matrices are computed using quadrature, which has some implications for Chebyshev-Gauss-Lobatto (GL) points, where the rows of the highest modes differ from those presented in Lemma 2.1 and 3.1 of \cite{Shen95}.
 This disagreement, that follows from inexact quadrature at the highest mode, explains the inclusion of the $c_{k+4}$ term for matrix $\N{B}$ and the $c_{k+2}$ term for matrix $\D{B}$.
-Matrix $B=(T_k, T_j)_w$ is also different in the highest mode from Eq.
+The matrix with components $B_{lq}=(T_q, T_l)_{\sigma}$ is also different in the highest mode from Eq.
 2.7 of \cite{Shen95}, but agrees with Eqs.
@@
 
-The convection vector $\bm{\hat{\mathcal{H}}}$, needed to compute Eqs.
-(\ref{eq:S_hv}) and (\ref{eq:S_hg}), is found by projecting to the Dirichlet vector space $\D{V}_N^3$, which corresponds to transforming the nonlinear cross product over the entire mesh
+Assembling all scalar products, the final matrix form of Eqs.
+(\ref{eq:u1}), (\ref{eq:g1}) and (\ref{eq:f1}), that can be used to solve for $\hat{u}^{{\kappa}+1}$, $\hat{g}^{{\kappa}+1}$ and $\hat{f}^{\kappa+1}$,
+given wavenumbers $m$ and $n$, are now found as
+\begin{align}
+\N{H}\hat{u}^{{\kappa}+1} & = \left(2\N{A} - 2\underline{z}^2\N{B} - \N{H} 
+\right)\hat{u}^{{\kappa}} + \triangle t \N{\mathcal{S}}(h_v^{{\kappa}+1/2})  & 
+\forall \, {\bm{k}} \in \N{K}_N, 
+\label{eq:ufin}\\ 
+\D{H} \hat{g}^{{\kappa}+1} &= \left(2 \D{B}-\D{H}\right) 
+\hat{g}^{{\kappa}} + \triangle t \D{\mathcal{S}}(h_g^{{\kappa}+1/2})  &\forall \, 
+{\bm{k}} \in 
+\D{K}_N, \label{eq:gfin} \\
+\D{B} \hat{f}^{\kappa+1} &= \D{C}\hat{u} &\forall \, \bm{k} \in \D{K_N}. \label{eq:ffin}
+\end{align}
+The coefficient matrices are given as
+\begin{align}
+\N{H}(m, n) &= -\frac{\nu \triangle t}{2}\N{Q} + \left( 1 + \nu 
+\triangle t 
+\underline{z}^2 \right) \N{A} - \frac{2\underline{z}^2 + \nu \triangle t 
+\underline{z}^4}{2} \N{B} &\forall \, m, n \in K^p, 
+\label{eq:Biharmonic_matrix} \\
+\D{H}(m, n) &= -\frac{\nu \triangle 
+t}{2}\D{A} + (1 + \frac{\nu \triangle t \underline{z}^2}{2}) \D{B} 
+&\forall \, m, n \in K^p, \label{eq:Helmholtz_matrix}
+\end{align}
+where $\underline{z}^2 = \underline{m}^2 + \underline{n}^2$. Equations 
+(\ref{eq:v0}) and (\ref{eq:w0}) are also written on matrix form as
+\begin{align}
+\D{H}\hat{v}^{{\kappa}+1} &= 
+\left(2\D{B} - \D{H} \right)\hat{v}^{{\kappa}} + \triangle t\D{B}\hat{\mathcal{H}}_{y}^{{\kappa}+\half} - \triangle t \D{\mathcal{S}}(\beta^{\kappa+1}) & \forall \, l \in \D{K}^x, m=n=0, 
+\label{eq:v00} \\
+\D{H}\hat{w}^{{\kappa}+1} &= 
+\left(2\D{B} - \D{H} \right)\hat{w}^{{\kappa}} + \triangle t\D{B}\hat{\mathcal{H}}_{z}^{{\kappa}+\half} & \forall \, l \in \D{K}^x, m=n=0. 
+\label{eq:w00}
+\end{align}
+
+Finally, the nonlinear
+terms are computed with the recipes given in Eqs.
+(\ref{eq:S_hv}, \ref{eq:S_hg}).
+To this end the required vector $\bm{\hat{\mathcal{H}}}$ is found by projecting to the Dirichlet vector space $\D{V}_N^3$, which corresponds to transforming the nonlinear cross product, evaluated in real space, over the entire mesh
 \begin{equation}
-\bm{\hat{\mathcal{H}}}_i = \mathcal{\D{T}}((\bm{u} \times \bm{\omega})_i) \quad \forall \, i=x,y,z. \label{eq:H_hat}
+{\hat{\mathcal{H}}}_i = \mathcal{\D{T}}((\bm{u} \times \bm{\omega})_i) \quad \forall \, i=x,y,z. \label{eq:H_hat}
 \end{equation}
 The curl, $\bm{\omega} = (g, \partial_z u - \partial_x w, \partial_y u - 
-\partial_x v)$, is computed by projecting each term to its appropriate space, before transforming back to physical space. 
+\partial_x v)$, is computed by projecting each individual term to its appropriate spectral space, before transforming back to physical space. 
 To this end, the two terms $\partial_x v$ and $\partial_x w$ are projected to 
-$V_N$ (requires $C_{kj} = (\D{\phi}_j^{'}, T_k)_w$, see Tab. \ref{tab:matrices}), whereas the remaining $\partial_y u$ and $\partial_z u$ are projected to 
-$\N{V}_N$:
+$V_N$ (requires $C_{lq} = (\D{\phi}_j^{'}, T_k)_{\sigma}$, see Tab. \ref{tab:matrices}), whereas the remaining $\partial_y u$ and $\partial_z u$ are projected to 
+$\N{V}_N$. With compact notation using the transforms, we obtain
 \begin{align}
@@
 \end{align}
+Here, with a slight abuse of notation, the left hand side is simply representing the respective expressions evaluated on the quadrature points in the real mesh $X_N$. 
 Note that the nonlinear term is also generally in need of 
-dealiasing, at least for the two periodic directions.
-
-It remains to discretize the equations in time. Kim, Moin and Mansour use 
-Crank-Nicolson for the diffusive terms and Adams-Bashforth for the nonlinear 
-convective terms, whereas a third order Runge-Kutta method is used for the 
-nonlinear term by Moser, Kim and Mansour (1999). We have implemented both 
-approaches in \cite{spectralDNS}, but for simplicity only the first one is described here. 
-Discretization in time is performed using a constant time step $\triangle t$, 
-such that the time is represented discretely as $t_{\kappa} = {\kappa} \triangle t, \, {\kappa} \in 
-\mathcal{Z}^{\ge}$. After 
-discretizing in time, equations (\ref{eq:u1}) and (\ref{eq:g1}) are transformed 
-to a matrix form that can be solved for $\hat{u}^{{\kappa}+1}$ and $\hat{g}^{{\kappa}+1}$ 
-given wavenumbers $m$ and $n$
-\begin{align}
-\N{H}\hat{u}^{{\kappa}+1} & = \left(2\N{A} - 2\underline{z}^2\N{B} - \N{H} 
-\right)\hat{u}^{{\kappa}} + \N{\mathcal{S}}(h_v^{{\kappa}+1/2}) \triangle t & 
-\forall \, {\bm{k}} \in \N{K}_N, 
-\label{eq:ufin}\\ 
-\D{H} \hat{g}^{{\kappa}+1} &= \left(2 \D{B}-\D{H}\right) 
-\hat{g}^{{\kappa}} + \D{\mathcal{S}}(h_g^{{\kappa}+1/2}) \triangle t &\forall \, 
-{\bm{k}} \in 
-\D{K}_N. \label{eq:gfin}
-\end{align}
-The coefficient matrices are given as
-\begin{align}
-\N{H}_{lq}(m, n) &= -\frac{\nu \triangle t}{2}\N{Q}_{lq} + \left( 1 + \nu 
-\triangle t 
-\underline{z}^2 \right) \N{A}_{lq} - \frac{2\underline{z}^2 + \nu \triangle t 
-\underline{z}^4}{2} \N{B}_{lq} &\forall \, m, n \in K^p, 
-\label{eq:Biharmonic_matrix} \\
-\D{H}_{lq}(m, n) &= -\frac{\nu \triangle 
-t}{2}\D{A}_{lq} + (1 + \frac{\nu \triangle t \underline{z}^2}{2}) \D{B}_{lq} 
-&\forall \, m, n \in K^p, \label{eq:Helmholtz_matrix}
-\end{align}
-where $\underline{z}^2 = \underline{m}^2 + \underline{n}^2$. The convective 
-terms are computed with the recipes given in Eqs. (\ref{eq:S_hv}, \ref{eq:S_hg}), discretized in time with a forward Adams-Bashforth projection, such that 
-$\N{\mathcal{S}}(h_v^{{\kappa}+1/2}) = 
-\N{\mathcal{S}}(1.5h_v^{{\kappa}} - 0.5 h_v^{{\kappa}-1})$ and $\D{\mathcal{S}}(h_g^{{\kappa}+1/2}) 
-= \D{\mathcal{S}}(1.5h_g^{{\kappa}} - 0.5 h_g^{{\kappa}-1})$. Equations 
-(\ref{eq:v0}) and (\ref{eq:w0}) are also discretized in time using 
-Crank-Nicolson and Adams-Bashforth, and we obtain
-\begin{align}
-\left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{v}_0^{{\kappa}+1} &= 
-\left(\frac{\nu \triangle t}{2}\D{A} + \D{B} 
-\right)\hat{v}^{{\kappa}}_0 + \triangle t\D{B}\hat{\mathcal{H}}_{y,0}^{{\kappa}+1/2} - \triangle t \left(\beta(t), \D{\phi}_l \right)_w , 
-\label{eq:v00} \\
-\left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{w}_0^{{\kappa}+1} &= 
-\left(\frac{\nu \triangle t}{2}\D{A} + \D{B} 
-\right)\hat{w}^{{\kappa}}_0 + \triangle t\D{B}\hat{\mathcal{H}}_{z,0}^{{\kappa}+1/2}, 
-\label{eq:w00}
-\end{align}
-where we have used notation $\hat{v}^{{\kappa}}_0 = \hat{v}(l, 0, 0, t_{\kappa})$, 
-$\hat{\mathcal{H}}^{{\kappa}}_{y,0} = \hat{\mathcal{H}}_y(l, 0, 0, t_{\kappa})$ and similar 
-for $\hat{w}^{\kappa}_0$ and $\hat{\mathcal{H}}^{\kappa}_{z,0}$.
+dealiasing, at least for the two periodic directions, but this is not discussed further in this paper.
 
@@
 $\{U_{1,k}\}_{k=0}^N = \{V_{2,k}\}_{k=0}^N$ can be used for copying row 2 of 
-$V$ to row 1 of $U$. Broadcasting is also implied, here meaning that for $\{w\}_{k=0}^{N} \gets c$, where $c$ is a scalar, all elements of $w_k$ from $k=0$ to $N$ gets the scalar value $c$.} Note 
-that the Adams-Bashforth projection of convection reverts to a first order 
-forward Euler on the first time step. For simulations of turbulent flows this 
-is insignificant, but for accuracy tests, like the Orr-Sommerfeld instability, 
-a proper initialization will be required to set both the two first time steps.
+$V$ to row 1 of $U$.
+Broadcasting is also implied, here meaning that for $\{w\}_{k=0}^{N} \gets c$, where $c$ is a scalar, all elements of $w_k$ from $k=0$ to $N$ gets the scalar value $c$.} The major computational cost in Alg.
+\ref{alg:NS} comes from  computing the
+nonlinear convection and solving for Eqs.~(\ref{eq:ufin}, \ref{eq:gfin}). We consider in this section the linear solvers.
+
 \begin{algorithm}
@@
  	\DontPrintSemicolon	
-    	Initialize $\bm{u}(\bm{x}, 0)$ and compute 
-    	$\hat{g}_{\bm{k}}(0)$ and $\hat{\bm{u}}_{\bm{k}}(0)$ \;
+    	Initialize $\bm{u}^0(\bm{x})$ and $\bm{u}^1(\bm{x})$ \;
+    	Compute $\hat{\bm{u}}^0_{\bm{k}},  \hat{\bm{u}}^1_{\bm{k}}, \hat{g}^0_{\bm{k}}$ and $\hat{g}^1_{\bm{k}}$ \;
         Set parameters (e.g., mesh and viscosity) and end time $T$ \;
-		Compute LU decomposition of $\N{H}(m,n), \D{H}(m,n) 
+		Compute LU decompositions of $\N{H}(m,n), \D{H}(m,n) 
 		\,\forall \, m, n \in K^p$ \;
-		Compute LU decomposition of coefficient matrices in (\ref{eq:v00}) and 
-		(\ref{eq:w00}) \;
-		$t \gets 0$\; 
-		${\kappa} \gets 0$ \;
+        $t \gets \triangle t$\; 
+		${\kappa} \gets 1$ \;
     	Compute nonlinear convection 
-    	$\bm{\hat{\mathcal{H}}}^{\kappa-1}(\hat{\bm{u}}_{\bm{k}}(0), 
-    	\hat{g}_{\bm{k}}(0))$ \;
+    	$\bm{\hat{\mathcal{H}}}_{\bm{k}}^{0}(\hat{\bm{u}}^0_{\bm{k}}, 
+    	\hat{g}^0_{\bm{k}})$ \;
     	\While{$t<T$}{
-	    	Compute nonlinear convection $\bm{\hat{\mathcal{H}}}^{\kappa}$ \;
+	    	Compute nonlinear convection $\bm{\hat{\mathcal{H}}}_{\bm{k}}^{\kappa}$ \;
 	    	{Compute right hand sides of Eqs.~}(\ref{eq:ufin}), 
@@
 		    $\hat{g}^{{\kappa}+1}_{\bm{k}} \, \forall\, \bm{k} \in \D{K}_N$ \;
+		    {Solve Eqs.}(\ref{eq:ffin}) for 
+		    $\hat{f}^{{\kappa}+1}_{\bm{k}} \, \forall\, \bm{k} \in \D{K}_N$ \;
 		    {Solve Eqs.}(\ref{eq:f3}, \ref{eq:g3}) for  
@@
 		    {Solve Eqs.}(\ref{eq:v00}, \ref{eq:w00}) for 
-		    $\hat{v}_0^{{\kappa}+1}, \hat{w}_0^{{\kappa}+1}\, 
-		    \forall \, l \in \D{K}^x$ \;
+		    $\hat{v}^{{\kappa}+1}, \hat{w}^{{\kappa}+1}\, 
+		    \forall \, l \in \D{K}^x, m=n=0$ \;
 		    ${\kappa} \gets {\kappa} + 1$ \;
@@
 
-The major computational cost in Alg. \ref{alg:NS} comes from  computing the 
-nonlinear convection and solving for Eqs.~(\ref{eq:ufin}, \ref{eq:gfin}). The 
-nonlinear convection term, $\bm{\hat{\mathcal{H}}}$, is computed using a 
-pseudo-spectral method, and dealiasing in the periodic directions can be either 
-the 3/2-rule \cite{canuto1988} or the 2/3-rule \cite{Orzag71}.
-
-Shen \cite{Shen95} states that it is possible to solve Eq.~(\ref{eq:gfin}) directly 
+Shen \cite{Shen95} writes that it is possible to solve Eq.~(\ref{eq:gfin}) directly 
 with essentially the same number of operations as a pentadiagonal solver. To this end, note that the matrix 
@@
 \end{equation}
-where $\xi_0, \xi_1$ and $\xi_2$ are constants (depending on $m$ and $n$). The matrix $\N{A}_{kj}$ contains only 
-three nonzero diagonals at $ j = k-2, k, k+2$,  whereas matrix $\N{B}_{kj}$ 
+where $\xi_0, \xi_1$ and $\xi_2$ are constants (depending on $m$ and $n$). The matrix $\{\N{A}_{kj}\}_{k,j=0}^{N_x-4}$ contains only 
+three nonzero diagonals at $ j = k-2, k, k+2$,  whereas matrix $\{\N{B}_{kj}\}_{k,j=0}^{N_x-4}$ 
 contains five 
 nonzero diagonals at $ j = k-4, k-2, k, k+2, k+4$. The nonzero elements of the 
-upper triangular $\N{Q}_{kj}$ matrix are given as \cite{Shen95}
+upper triangular $\{\N{Q}_{kj}\}_{k,j=0}^{N_x-4}$ matrix are given as \cite{Shen95}
 \begin{align}
@@
 \end{align} 
-For each row in $\N{H}_{kj}$ we have that $\N{H}_{kj}=\xi_0\N{Q}_{kj}\, \forall j 
-> k+4$. A straight forward direct LU decomposition of $\N{H}$ can be performed 
-as shown in Alg. 
+A straight forward direct LU decomposition (without pivoting) of $\N{H}$ can be performed as shown in Alg. 
 \ref{alg:lu_biharmonic}, which leads to a lower 
 triangular 
-matrix $\N{L}_{kj}$ with two nonzero diagonals at $j=k-2$ and $k-4$ plus a 
-unity 
-main diagonal. The upper triangular matrix $\N{U}_{kj}$ is dense and as such a 
-show stopper for an efficient back solve. However, we note that $\N{U}_{kj}$ 
-contains three distinct 
-diagonals at $j=k, k+2$ and $k+4$, whereas the remaining part of $\N{U}_{kj}$, for $k\le N_x-10$, 
-can be expressed as
+matrix $\{\N{L}_{kj}\}_{k,j=0}^{N_x-4}$ with two nonzero diagonals at $j=k-2$ and $k-4$ plus a unity 
+main diagonal. The upper triangular matrix $\N{U}$ is dense and as such a 
+show stopper for an efficient back solve. However, we note that $\{\N{U}_{kj}\}_{k,j=0}^{N_x-4}$ 
+contains three distinct diagonals at $j=k, k+2$ and $k+4$, whereas the remaining part can be expressed as
 \begin{equation}
-\N{U}_{kj} = \xi_0(a_k q_j + b_k s_j), \quad j = k+6, k+8, \ldots, N_x-4, 
+\N{U}_{kj} = \xi_0(a_k q_j + b_k s_j), \quad j = k+6, k+8, \ldots, N_x-4 \text{ and } k\le N_x-10, 
 \label{eq:ab}
@@
 %\end{align}
-If $\bm{a}$ and $\bm{b}$ 
+If $a$ and $b$
 are pre-computed, 
-the total storage requirement for the complete LU decomposition is no more than 
-$7N_x$, i.e., two diagonals for $\N{L}$, three diagonals for $\N{U}$ plus $\bm{a}$ and 
-$\bm{b}$. The solution of the complete system can be obtained very quickly with 
+the total storage requirement for the complete LU decomposition is less than 
+$7N_x$, since there are two diagonals for $\N{L}$, three for $\N{U}$ plus ${a}$ and 
+${b}$. The solution of the complete system can be obtained very quickly with 
 one simple forward elimination and a back solve, where the back 
@@
         per row need 
-        to be stored after the row is no longer needed in the for-loop below 
+        to be stored ($\N{U}_{kk}, \N{U}_{kk+2}$ and $\N{U}_{kk+4}$) after the row is no longer needed in the for-loop below 
         (as the loop is traversed to row $k$, $\N{U}_{m, j}$ is no longer 
@@
         $\N{L}_{k,k-2} \gets {\N{U}}_{k,k-2}/{\N{U}}_{k-2, k-2}$ \;
-        $\{\N{U}_{k,k+2j}\}_{j=0}^{M} \gets \{{\N{U}}_{k,k+2j}\}_{j=0}^{M} - 
-        \N{L}_{k, k-2} \{\N{U}_{k-2,k+2j}\}_{j=0}^{M}$ \;
+        $\{\N{U}_{k,k+2j}\}_{j=-1}^{M} \gets \{{\N{U}}_{k,k+2j}\}_{j=-1}^{M} - 
+        \N{L}_{k, k-2} \{\N{U}_{k-2,k+2j}\}_{j=-1}^{M}$ \;
         %\State $\N{U}_{kj} = {\N{U}}_{kj} - \N{L}_{k, k-2} \N{U}_{k-2, j} 
@@
 	The parameters $q_j$ and $s_j$ are given in Eqs. (\ref{eq:pk}) and 
-	(\ref{eq:rk}) respectively. 	\label{alg:SolveBiharmonic}}
-		\Fn{\LUSolveBiharmonic{$\{\N{U}_{kj}\}_{k,j=0}^{N}, 
+	(\ref{eq:rk}) respectively. The vector $\{f_k\}_{k=0}^{N}$ contains the right hand side and $\{ \hat{u}_k\}_{k=0}^N$ is the solution.	\label{alg:SolveBiharmonic}}
+		\Fn{\LUSolveBiharmonic{$\{f_k\}_{k=0}^{N}, \{\N{U}_{kj}\}_{k,j=0}^{N}, 
                 \{\N{L}_{kj}\}_{k,j=0}^{N}, 
-                \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N-6}, \{b_k\}_{k=0}^{N-6}, 
+                \{a_k\}_{k=0}^{N-6}, \{b_k\}_{k=0}^{N-6}, 
                 \xi_0$}}{
@@
 
-\section{Verification}
+\section{Verification and validation}
 \label{sec:verification}
@@
 The complete solver has been implemented in the open source code spectralDNS~\cite{spectralDNS}.
-For MPI we are using a slab decomposition, as described by Mortensen \cite{Mortensen2015}, but parallel scalability is not considered here since all algorithms described in previous sections are executed in serial.
-In this section we simply verify the implementation, and, to show that these algorithms can be used for large-scale simulations, we will first study some simple 1D problems.
+For MPI we are using a slab decomposition, as described by Mortensen \cite{Mortensen2016}, but parallel scalability is not considered here since all algorithms described in previous sections are executed in serial.
+In this section we will simply verify the implementation, and validate the method for large-scale simulations.
+To this end we will first study some simple 1D problems.
 
@@
 by the following approach: Let $\{u_i\}_{i=0}^{N_x-4}$ be a uniformly distributed 
-random vector of double precision in the interval $(0, 1)$. Compute the matrix vector product $\bm{f}=\N{H}\bm{u}$ and 
-then solve $\N{H} \bm{v} = \bm{f}$, using Alg. \ref{alg:SolveBiharmonic}, for 
+random vector of double precision in the interval $(0, 1)$. Compute the matrix vector product ${f}=\N{H}{u}$ and 
+then solve $\N{H} {v} = {f}$, using Algorithms \ref{alg:lu_biharmonic}, \ref{alg:ab} and \ref{alg:SolveBiharmonic}, for 
 $\{v_i\}_{i=0}^{N_x-4}$. The roundoff error of the solution 
@@
 	are for an average of 100 runs.	 \label{tab:roundoff}}
+\centering
    \begin{tabular}{cccccccc}
@@
 To show that the solvers are scaling with size, we investigate the execution time 
-for solving $\N{H} \bm{v} = \bm{f}$ and $\D{H} \bm{v} = \bm{f}$ on a complete 
+for solving $\N{H} {v} = {f}$ and $\D{H} {v} = {f}$ on a complete 
 wavenumber mesh of size $(N_x,64,64)$.
@@
 \begin{align}
-u(x,y,t) &= -\epsilon  \mathrm{Re}\{\imath \xi(x) \exp (\imath (y-\lambda t))\}, \label{eq:OS1} \\
-v(x,y,t) &= 1-x^2 + \epsilon \mathrm{Re}\{\xi'(x) \exp(\imath(y-\lambda t)\}, \label{eq:OS2}
-\end{align}
-where $\xi(x)$ is the eigenvector, $\lambda=0.2470750602+\imath 2.664410371\cdot 10^{-3}$ the eigenvalue and $\epsilon \in \mathcal{R}^+$ is a small positive constant.
+u(x,y,t) &= -\epsilon  \Re\{\imath \xi(x) \exp (\imath (y-\lambda t))\}, \label{eq:OS1} \\
+v(x,y,t) &= 1-x^2 + \epsilon \Re\{\xi'(x) \exp(\imath(y-\lambda t)\}, \label{eq:OS2}
+\end{align}
+where $\xi(x)$ is the eigenvector, $\lambda=0.2470750602+\imath 2.664410371\cdot 10^{-3}$ the eigenvalue and $\epsilon \in \mathbb{R}^+$ is a small positive constant.
 Note that if $\epsilon$ is chosen as too big, then the initial/boundary value problem will not agree well with the linearized Navier-Stokes equations used in deriving the Orr-Sommerfeld equations, due to second-order terms.
-Likewise, if $\epsilon$ is chosen as too small, then roundoff errors will be too large, since the perturbation is added to a base flow of order unity.
+Likewise, if $\epsilon$ is chosen as too small, then roundoff errors will be too large, since the perturbation is added to a base flow with magnitude of order unity.
 It turns our that $\epsilon=10^{-7}$ is a good compromise, and that a mesh of size $128 \times 8 \times 2$ is accurate enough to isolate temporal errors.
@@
 \begin{equation}
-E(t) =  \frac{\left<\tilde{\bm{u}}(\bm{x}, t), \tilde{\bm{u}}(\bm{x}, t) \right>}{\left<\tilde{\bm{u}}(\bm{x}, 0), \tilde{\bm{u}}(\bm{x}, 0) \right>}  - \exp(2 \mathrm{Im}\{\lambda\}t), \label{eq:OSenergy}
+E(t) =  \frac{\left<\tilde{\bm{u}}(\bm{x}, t), \tilde{\bm{u}}(\bm{x}, t) \right>}{\left<\tilde{\bm{u}}(\bm{x}, 0), \tilde{\bm{u}}(\bm{x}, 0) \right>}  - \exp(2 \Im\{\lambda\}t), \label{eq:OSenergy}
 \end{equation}
@@
 The mesh size is $128 \times 8 \times 2$.
-The $L^2$-norm $ \| \bm{u}' \| = \sqrt{\left< \bm{u}', \bm{u}'\right>}$, where $\bm{u}'=\bm{u}-\bm{u}_{\mathrm{exact}}$, and the energy error is computed as Eq.
+The $L^2$-norm $ \| \bar{\bm{u}} \| = \sqrt{\left< \bar{\bm{u}}, \bar{\bm{u}}\right>}$, where $\bar{\bm{u}}=\bm{u}-\bm{u}_{\mathrm{exact}}$, and the energy error is computed as Eq.
 (\ref{eq:OSenergy}).
-The numbers in parenthesis are showing the order of the error, computed as $\log(\tau_i/\tau_{i-1})/\log(\triangle t_i/ \triangle t_{i-1})$, where $i$ is the row number of the table and $\tau$ represents the values of $\|\bm{u}'\|$ or $E(t=50)$.
+The numbers in parenthesis are showing the order of the error, computed as $\log(\tau_i/\tau_{i-1})/\log(\triangle t_i/ \triangle t_{i-1})$, where $i$ is the row number of the table and $\tau$ represents the values of $\|\bar{\bm{u}}\|$ or $E(t=50)$.
 \label{tab:OS}}
 	\begin{tabular}{ccc}	
- $\triangle t$ & $\| \bm{u}' \|(t=50)$ & $\int_{t=0}^{t=50}E(t) \mathrm{d}t$ \\
+ $\triangle t$ & $\| \bar{\bm{u}} \|(t=50)$ & $\int_{t=0}^{t=50}E(t) \mathrm{d}t$ \\
  \hline
@@
 The time step $\triangle t = 10^{-3}$ and the end time is $t=0.05$.
-The $L^2$-norm $ \| \bm{u}' \| = \sqrt{\left< \bm{u}', \bm{u}'\right>}$, where $\bm{u}'=\bm{u}-\bm{u}_{\mathrm{exact}}$, and the energy error is computed as Eq.
+The $L^2$-norm $ \| \bar{\bm{u}} \| = \sqrt{\left< \bar{\bm{u}}, \bar{\bm{u}}\right>}$, where $\bar{\bm{u}}=\bm{u}-\bm{u}_{\mathrm{exact}}$, and the energy error is computed as Eq.
 (\ref{eq:OSenergy}).
@@
 	\begin{tabular}{ccccc}	
-		$N_x$ & \multicolumn{2}{c}{${\| \bm{u}' \|(t=0.05)}$} & \multicolumn{2}{c}{$\int_{t=0}^{t=0.05}E(t) \mathrm{d}t$} \\
+		$N_x$ & \multicolumn{2}{c}{$\frac{1}{\epsilon}{\| \bar{\bm{u}} \|(t=0.05)}$} & \multicolumn{2}{c}{$\int_{t=0}^{t=0.05}E(t) \mathrm{d}t$} \\
 		\hline
@@
 We use the same small computational box, with resolution of $512 \times 1024 \times 1024$, and expect the superior resolution properties of spectral methods to work in our favor.
+Furthermore, the driving force $\beta$ was dynamically adjusted in time to keep the flux through the channel constant.
 Simulations were run on the Shaheen II supercomputer at KAUST, using 512 cores.
@@
 
+%\begin{figure}[t]
+%	\begin{center}
+%	\includegraphics[width=0.9\textwidth]{U_Re2000.eps}
+%	\caption{Average velocity $\D{v}^+ = \D{v}/u_{\tau}$ in $y$-direction, for $Re_{\tau}=2000$. Our results (dashed) are compared with those of Hoyas and Jim\'{e}nez \cite{hoyas06} (dotted).}
+%	\label{fig:U_mean}
+%	\end{center}
+%\end{figure}
 \begin{figure}[t]
 	\begin{center}
-	\includegraphics[width=0.9\textwidth]{U_Re2000.eps}
+		%
+		\begin{tikzpicture}[xscale=1.2]
+		\begin{axis}[
+		xlabel=$x$,
+		ylabel=$\bar{v}^{+}$,
+		legend cell align=left,
+		legend pos=south east,
+		legend style={fill=none, draw=none},
+		]
+		\addplot[dashed, color=blue] table [x=x, y=u]{my_data.txt};
+		\addlegendentry{Current}
+		% not compat
+		\addplot[dotted, color=blue] table [x=x, y=u]{hoyas_data5.txt};
+		\addlegendentry{Hoyas \& Jim{\' e}nez}
+		\end{axis}
+		\end{tikzpicture}
+		%
 		\caption{Average velocity $\D{v}^+ = \D{v}/u_{\tau}$ in $y$-direction, for $Re_{\tau}=2000$. Our results (dashed) are compared with those of Hoyas and Jim\'{e}nez \cite{hoyas06} (dotted).}
@@
 \section{Conclusions}
-Spectral Navier Stokes solvers are attractive for their accuracy and resolution 
-properties, and are usually the preferred tools for performing Direct Numerical 
-Simulations (DNS) of turbulent flows. A spectral discretization of the wall-normal 
-direction in a plane channel is challenging, due to roundoff errors, and as such recent large-scale channel DNS 
-have been conducted without spectral accuracy in the wall-normal direction.
-
-In this paper a spectral-Galerkin Navier Stokes solver for large-scale turbulent channel flows 
+In this paper a spectral-Galerkin Navier Stokes solver amenable to large-scale turbulent channel flows 
 has been described. In the wall-normal direction, the solver utilizes base 
@@
 solver uses Fourier decompositions in both periodic directions, and, as such, 
-it is fully spectral in all spatial directions. To show that this solver is 
-applicable to large-scale simulations at high Reynolds numbers,  we have shown 
+it is fully spectral in all spatial directions. To validate the method for large-scale simulations at high Reynolds numbers, we have shown 
 that the roundoff errors are small, even for simulations twice the size of the 
@@
 and these have been shown to scale close to 
-$\mathcal{O}(N_x)$, with insignificant roundoff errors. The fully spectral solver 
+$\mathcal{O}(N_x)$, with insignificant roundoff errors.
+Furthermore, all required scalar product matrices and fast transforms between spectral and physical space have been described.
+The fully spectral solver
 has been applied to a turbulent 
@@
 %t$ from initial conditions $u(\bm{x}, 0) = u^0(\bm{x})$ at $t=0$, such that 
-%$t_n = n\triangle t$, $n \in \mathcal{Z}$. For turbulent channel flows initial 
+%$t_n = n\triangle t$, $n \in \mathbb{Z}$. For turbulent channel flows initial 
 %conditions are simply a perturbed state used to get a turbulent flow going. 
@@
 %\begin{multline}
-% \frac{1}{\triangle t}\left<\bm{u}^*-\bm{u}^{n-1}, \bm{\D{\psi}}\right>_w + 
-%\left<\bm{N}^{n-1/2}, \bm{\D{\psi}} \right>_w   = \nu \left<\nabla^2 
-%\bm{u}^{n-1/2}, \bm{\D{\psi}} \right>_w \\ - \left<\nabla{p}^{n-3/2}, 
-%\bm{\D{\psi}} \right>_w + \left<\bm{f}, \bm{\D{\psi}} \right>_w \,\,  \forall 
+% \frac{1}{\triangle t}\left<\bm{u}^*-\bm{u}^{n-1}, \bm{\D{\psi}}\right>_{\sigma} + 
+%\left<\bm{N}^{n-1/2}, \bm{\D{\psi}} \right>_{\sigma}   = \nu \left<\nabla^2 
+%\bm{u}^{n-1/2}, \bm{\D{\psi}} \right>_{\sigma} \\ - \left<\nabla{p}^{n-3/2}, 
+%\bm{\D{\psi}} \right>_{\sigma} + \left<\bm{f}, \bm{\D{\psi}} \right>_{\sigma} \,\,  \forall 
 %\, \bm{\D{\psi}} \in \D{V}_N^3. \label{eq:tentative}
@@
 %\begin{equation}
-%\left< \nabla^2 p^*, \N{\psi} \right>_w = \frac{1}{\triangle t} \left<\nabla 
-%\cdot \bm{u}^*, \N{\psi} \right>_w\,\, \forall \, \N{\psi} \in \N{V}_N. 
+%\left< \nabla^2 p^*, \N{\psi} \right>_{\sigma} = \frac{1}{\triangle t} \left<\nabla 
+%\cdot \bm{u}^*, \N{\psi} \right>_{\sigma}\,\, \forall \, \N{\psi} \in \N{V}_N. 
 %\label{eq:pcorr}
@@
 %\begin{align}
-%\left< \bm{u}^n, \bm{\D{\psi}}\right>_w &= \left< \bm{u}^*, 
-%\bm{\D{\psi}}\right>_w - \triangle t \left< \nabla p^*, \bm{\D{\psi}} 
-%\right>_w 
+%\left< \bm{u}^n, \bm{\D{\psi}}\right>_{\sigma} &= \left< \bm{u}^*, 
+%\bm{\D{\psi}}\right>_{\sigma} - \triangle t \left< \nabla p^*, \bm{\D{\psi}} 
+%\right>_{\sigma} 
 %\,\,  &\forall \, \bm{\D{\psi}} \in \D{V}_N^3, \label{eq:u_update} \\
-%\left<p^{n-1/2}, \N{\psi}\right>_w &= \left<p^{n-3/2}, \N{\psi}\right>_w + 
-%\left<p^*, \N{\psi}\right>_w \,\, &\forall \, \N{\psi} \in 
+%\left<p^{n-1/2}, \N{\psi}\right>_{\sigma} &= \left<p^{n-3/2}, \N{\psi}\right>_{\sigma} + 
+%\left<p^*, \N{\psi}\right>_{\sigma} \,\, &\forall \, \N{\psi} \in 
 %\N{V}_N,\label{eq:pres_update}
@@
 %\begin{align}
-% \left< u_i^* - u_i^{n-1}, \D{\psi}\right>_w &= 
+% \left< u_i^* - u_i^{n-1}, \D{\psi}\right>_{\sigma} &= 
 %h\D{B}\left({\hat{u}}_i^*-{\hat{u}}_i^{n-1} \right) \quad &i \in (0,1,2), \\ 
-% \left<\nabla^2 {u}_i^{n-1/2}, \D{\psi}\right>_w &= -h  \left( \D{A} 
+% \left<\nabla^2 {u}_i^{n-1/2}, \D{\psi}\right>_{\sigma} &= -h  \left( \D{A} 
 %+(\underline{m}^2+\underline{n}^2)\D{B}\right) \hat{u}_i^{n-1/2} \quad &i \in 
 %(0,1,2), \\ 
-% \left< f_i, \D{\psi}\right>_w &= h\D{\mathcal{S}}(f_i) &i \in (0,1,2), \\
-% \left<N_i^{n-1/2}, \D{\psi} \right>_w &= h\D{\mathcal{S}}(N_i^{n-1/2}) &i \in 
+% \left< f_i, \D{\psi}\right>_{\sigma} &= h\D{\mathcal{S}}(f_i) &i \in (0,1,2), \\
+% \left<N_i^{n-1/2}, \D{\psi} \right>_{\sigma} &= h\D{\mathcal{S}}(N_i^{n-1/2}) &i \in 
 %(0,1,2), \\  
-% \left< \nabla p^{n-3/2}, \bm{\D{\psi}} \right>_w &= h\left[\grave{C},\, 
+% \left< \nabla p^{n-3/2}, \bm{\D{\psi}} \right>_{\sigma} &= h\left[\grave{C},\, 
 %\imath \underline{m} \grave{B},\, \imath \underline{n} \grave{B} \right] 
 %\hat{p}^{n-3/2}, \\ 
-% \left< \nabla^2 p^*, \N{\psi} \right>_w &= -h\left(\N{A} +(\underline{m}^2 + 
+% \left< \nabla^2 p^*, \N{\psi} \right>_{\sigma} &= -h\left(\N{A} +(\underline{m}^2 + 
 %\underline{n}^2)\N{B} \right) \hat{p}^*, \\ 
-% \left<\nabla \cdot \bm{u}^*, \N{\psi} \right>_w &= h\left( \tilde{C} 
+% \left<\nabla \cdot \bm{u}^*, \N{\psi} \right>_{\sigma} &= h\left( \tilde{C} 
 %\hat{u}^* + \imath \underline{m} \tilde{B} \hat{v}^* + \imath \underline{n} 
@@
 %\end{align}
-%where $\D{A}_{kj} = -\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w, 
-%\N{A}_{kj} = -\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w,  \grave{C}_{kj} 
-%= \left(\N{\phi}^{'}_j, \D{\phi}_k \right)_w, \tilde{C}_{kj} = 
+%where $\D{A}_{kj} = -\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_{\sigma}, 
+%\N{A}_{kj} = -\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_{\sigma},  \grave{C}_{kj} 
+%= \left(\N{\phi}^{'}_j, \D{\phi}_k \right)_{\sigma}, \tilde{C}_{kj} = 
 %(\D{\phi}^{'}_j, 
-%\N{\phi}_k)_w, \N{B}_{kj} = \left( \N{\phi}_j, \N{\phi}_k \right)_w, 
-%\grave{B}_{kj} = (\N{\phi}_j, \D{\phi}_k)_w$ and $\tilde{B}_{kj} = 
+%\N{\phi}_k)_{\sigma}, \N{B}_{kj} = \left( \N{\phi}_j, \N{\phi}_k \right)_{\sigma}, 
+%\grave{B}_{kj} = (\N{\phi}_j, \D{\phi}_k)_{\sigma}$ and $\tilde{B}_{kj} = 
 %(\D{\phi}_j, 
-%\N{\phi}_k)_w$ are all more or less sparse matrices. Arranging the governing 
+%\N{\phi}_k)_{\sigma}$ are all more or less sparse matrices. Arranging the governing 
 %equations on matrix form we obtain for the three tentative velocity components 
@@
 %\end{align}
-%where $\D{C}_{kj} = (\D{\phi}_j^{'}, \D{\phi}_k)_w$,  $\acute{C_{kj}} = 
-%(\D{\phi}_j^{'}, T_k)_w$ and $B=(T_j, T_k)_w$. Partial derivatives with 
+%where $\D{C}_{kj} = (\D{\phi}_j^{'}, \D{\phi}_k)_{\sigma}$,  $\acute{C_{kj}} = 
+%(\D{\phi}_j^{'}, T_k)_{\sigma}$ and $B=(T_j, T_k)_{\sigma}$. Partial derivatives with 
 %respect 
@@
 Algorithms for fast scalar products and transforms, for the three 
-spaces $V_N, \D{V}_N, \N{V}_N$, are given in Alg. ~\ref{alg:fst}. The 
+spaces $W_N, \D{W}_N, \N{W}_N$, see Eqs. (\ref{eq:Tk}, \ref{eq:phiD}) and (\ref{eq:phiN}), are given in Alg. ~\ref{alg:fst}. The 
 algorithms for the inverse transforms are given in Alg.~\ref{alg:ifst}. The 
@@
 Also note that the matrices are computed using quadrature, and not exact integration, which has some minor implications for Chebyshev-Gauss-Lobatto (GL) points.
-This follows since for an exact $L^2_w$ scalar product we would have $c_{N_x}=1$ for GL as well as Chebyshev-Gauss (GC), but for the $l^2_w$ scalar product used here $c_{N_x}=2$.
+This follows since for an exact $L^2_{\sigma}$ scalar product we would have $c_{N_x}=1$ for GL as well as Chebyshev-Gauss (GC), but for the $l^2_{\sigma}$ scalar product used here $c_{N_x}=2$.
 This disagreement, that follows from inexact quadrature at the highest mode, explains the inclusion of the $c_{k+4}$ term for matrix $\N{B}$, and the $c_{k+2}$ term for matrix $\D{B}$.
-These are missing from Lemma 2.1 and 3.1 of \cite{Shen95}, because Shen here considers only the exact $L^2_w$ scalar product.
+These are missing from Lemma 2.1 and 3.1 of \cite{Shen95}, because Shen here considers only the exact $L^2_{\sigma}$ scalar product.
 
@@
     \DontPrintSemicolon
-	\caption{Forward scalar products and transforms for all spaces $V_N, 
-	\D{V}_N, \N{V}_N$. 
-		Here "dct" is the discrete cosine transform from SciPy. \label{alg:fst}}
+	\caption{Forward scalar products and transforms for all spaces $W_N, 
+	\D{W}_N, \N{W}_N$. 
+		Here "dct" is the discrete cosine transform from SciPy. The matrices $B$, $\D{B}$ and $\N{B}$ are given in Table \ref{tab:matrices}.\label{alg:fst}}
 			\Fn{\forwardScalar{$\{f_j\}_{j=0}^{N}$, points, space}}{
@@
 
-			\If{ points = "Chebyshev-Gauss-Lobatto"}{
+			\If{ points = "Chebyshev-Gauss"}{
 			$\{ w_k\}_{k=0}^{N} \gets \text{dct}(\{f_j\}_{j=0}^{N}, \text{type=2}, 
@@
             }
-			\ElseIf{ points = "Chebyshev-Gauss"}{
+			\ElseIf{ points = "Chebyshev-Gauss-Lobatto"}{
 			$\{ w_k\}_{k=0}^{N} \gets \text{dct}(\{f_j\}_{j=0}^{N}, \text{type=1}, 
@@
 			$\{y\}_{k=0}^N \gets 0$ \;
-			\If{ space = $\D{V}_N$}{
+			\If{ space = $\D{W}_N$}{
 			$\{y_k\}_{k=0}^{N-2} \gets \{w_k\}_{k=0}^{N-2} - \{w_k\}_{k=2}^{N}$ 
@@
             }
-			\ElseIf{ space = $\N{V}_N$}{
+			\ElseIf{ space = $\N{W}_N$}{
 			\For{$k=0$ \KwTo $N-4$}{
@@
             }
-			\ElseIf{space = $V_N$}{
+			\ElseIf{space = $W_N$}{
 			$\{y_k\}_{k=0}^N \gets \frac{2}{\pi} \{w_k\}_{k=0}^N $ \;
@@
 \emph{points}, \emph{space}) \;
-			\If{space = $\D{V}_N$}{
+			\If{space = $\D{W}_N$}{
 			$\{y_k\}_{k=0}^{N-2} \gets \{\D{B}_{kj}^{-1}\}_{k,j=0}^{N-2}\{y_k\}_{k=0}^{N-2}$ \;
             }
-			\ElseIf{space = $\N{V}_N$}{
+			\ElseIf{space = $\N{W}_N$}{
 			$\{y_k\}_{k=0}^{N-4} \gets \{\N{B}_{kj}^{-1}\}_{k,j=0}^{N-4}\{y_k\}_{k=0}^{N-4}$ \;
             }
-			\ElseIf{space = $V_N$}{
+			\ElseIf{space = $W_N$}{
 			$\{y_k\}_{k=0}^{N} \gets \{{B}_{kj}^{-1}\}_{k,j=0}^{N}\{y_k\}_{k=0}^{N}$ \;
@@
 \begin{algorithm}
-	\caption{Inverse transforms for all spaces $V_N, \D{V}_N, \N{V}_N$. 
+	\caption{Inverse transforms for all spaces $W_N, \D{W}_N, \N{W}_N$. 
 		Here "dct" is the discrete cosine transform from SciPy.\label{alg:ifst}}
@@
         $\{z_k\}_{k=0}^{N} \gets 0$ \;
-		\If{ space = $\D{V}_N$}{
+		\If{ space = $\D{W}_N$}{
 		$\{z_k\}_{k=0}^{N-2} \gets \{f_k\}_{k=0}^{N-2}$ \;
@@
         }
-        \ElseIf{ space = $\N{V}_N$}{	       
+        \ElseIf{ space = $\N{W}_N$}{	       
 		$\{z_k\}_{k=0}^{N-4} \gets \{f_k\}_{k=0}^{N-4}$ \;
@@
 Note that $c_0=2$ and $c_k=1$ for $0<k<N_x$.
-For Chebyshev-Gauss points $c_{N_x}=1$, whereas $c_{N_x}=2$ for Chebyshev-Gauss-Lobatto points.	\label{tab:matrices}}
+For Chebyshev-Gauss points $c_{N_x}=1$, whereas $c_{N_x}=2$ for Chebyshev-Gauss-Lobatto points.
+Also note that only the upper diagonals are described for the symmetric matrices.	\label{tab:matrices}}
 	\begin{tabular}{ccl}	
@@
 		\hline
-%		$\{\D{A}_{kj}\}_{k,j=0}^{N_x-2}$ & $-\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w$ &
+%		$\{\D{A}_{kj}\}_{k,j=0}^{N_x-2}$ & $-\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_{\sigma}$ &
 %$\begin{cases} 2\pi(k+1)(k+2), &j=k,\\
@@
 %
-%$\{\N{A}_{kj}\}_{k,j=1}^{N_x-2}$ & $-\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w$ & $\begin{cases} 2\pi j^2(k+1)/(k+2), &j=k,\\
+%$\{\N{A}_{kj}\}_{k,j=1}^{N_x-2}$ & $-\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_{\sigma}$ & $\begin{cases} 2\pi j^2(k+1)/(k+2), &j=k,\\
 %4\pi j^2(k+1)/(k+2)^2, & j=k+2, k+4, k+6, \ldots, \\
@@
 $\{\N{C}_{kj}\}_{k,j=0}^{N_x-4, N_x-2}$ & $\left(\D{\phi}^{'}_j, \N{\phi}_k 
-\right)_w$ & $\begin{cases} -\pi(k+1), &j=k-1,\\
+\right)_{\sigma}$ & $\begin{cases} -\pi(k+1), &j=k-1,\\
 2\pi(k+1), & j=k+1, \\
@@
 $\{\D{C}_{kj}\}_{k,j=0}^{N_x-2, N_x-4}$ & $\left(\N{\phi}^{'}_j, \D{\phi}_k 
-\right)_w$ & $\begin{cases}
+\right)_{\sigma}$ & $\begin{cases}
 \pi (k-2)(k+1)/k, &j=k-3,\\
@@
 $\{{C}_{kj}\}_{k,j=0}^{N_x, N_x-2}$ & $\left(\D{\phi}^{'}_j, T_k 
-\right)_w$ & $\begin{cases}
+\right)_{\sigma}$ & $\begin{cases}
 -\pi (k+1), &j=k-1,\\
@@
 
-$\{\N{M}_{kj}\}_{k,j=0}^{N_x-4, N_x-2}$  & $ \left(\D{\phi}_j, \N{\phi}_k \right)_w $ & $\begin{cases}
+$\{\N{M}_{kj}\}_{k,j=0}^{N_x-4, N_x-2}$  & $ \left(\D{\phi}_j, \N{\phi}_k \right)_{\sigma} $ & $\begin{cases}
 -\pi/2 & j=k-2, \\
@@
 
-$\{\N{B}_{kj}\}_{k,j=0}^{N_x-4} = \{\N{B}_{jk}\}_{j,k=0}^{N_x-4}$ & $(\N{\phi}_j, \N{\phi}_k)_w$ & $\begin{cases}
+$\{\N{B}_{kj}\}_{k,j=0}^{N_x-4}$ & $(\N{\phi}_j, \N{\phi}_k)_{\sigma}$ & $\begin{cases}
 \frac{\pi}{2} \left(c_k + 4 \left(\frac{k+2}{k+3} \right)^2 + c_{k+4} 
@@
 
-$\{\D{B}_{kj}\}_{k,j=0}^{N_x-2} = \{\D{B}_{jk}\}_{j,k=0}^{N_x-2}$ & $(\D{\phi}_j, \D{\phi}_k)_w$ & $ \begin{cases} 
+$\{\D{B}_{kj}\}_{k,j=0}^{N_x-2} $ & $(\D{\phi}_j, \D{\phi}_k)_{\sigma}$ & $ \begin{cases} 
 \frac{\pi}{2} (c_k+c_{k+2}) &j=k, \\
@@
 
-$\{B_{kj}\}_{k,j=0}^{N_x} = \{B_{jk}\}_{j,k=0}^{N_x}$ & $(T_k, T_j)_w$ & $\frac{c_k \pi}{2} \delta_{kj}.$
+$\{B_{kj}\}_{k,j=0}^{N_x} $ & $(T_k, T_j)_{\sigma}$ & $\frac{c_k \pi}{2} \delta_{kj}.$
 

### 2017-08-02 paper.tex
Abstract

@@
 \begin{abstract}
-A fully (pseudo-)spectral solver for direct numerical simulations of large-scale turbulent channel flows is described.
-The solver utilizes the Chebyshev base functions suggested by J.
+In this paper we describe in detail the implementation of a spectral Galerkin method intended for large scale turbulent channel flow simulations with the spectral Galerkin method.
+This modified Chebyshev basis was suggested some time ago by J.
 Shen [SIAM J.
 Sci.
-Comput., 16, 1, 1995], that lead to stable and robust numerical schemes, even at very large scale.
-New and fast algorithms for the direct solution of the linear systems are devised, and algorithms and matrices for all required scalar products and transforms are provided.
+Comput., 16, 1, 1995], but has until now not been fully taken advantage of in the study of turbulence physics.
+What is great about this basis, that makes it ideal for the turbulent channel flow, is both its accuracy (fully spectral in all spatial directions, when combined with Fourier bases in the periodic directions) and its efficiency (all algorithms may be executed with $\mathcal{O}(N\log N)$ or better scaling).
+But what sets it apart from other known fully spectral channel flow solvers is the robustness and low roundoff, caused by well conditioned coefficient and inner product matrices.
+
+For the implementation of this new solver, we devise new and fast ($\mathcal{O}(N\log N)$) algorithms for the direct solution of the linear systems, as well as algorithms and matrices for the required inner products and fast transforms.
 We validate the solver for very high Reynolds numbers.
@@
 The solver is available through the open source project spectralDNS [https://github.com/spectralDNS].
+
 \end{abstract}
@@
 
-In this paper we will describe and assess a spectral-Galerkin channel flow solver based on Fourier and Shen's Chebyshev basis \cite{Shen95}.
+In this paper we will describe and assess a spectral-Galerkin channel flow solver based on tensorproduct spaces of Fourier and Shen's Chebyshev basis \cite{Shen95}.
 The solver will consist of parts that scale at worst as $N \log N$, for a 1D problem of size $N$, and as such as $N^3 \log N$ for a 3D box of size $N^3$.
